### 2015-01-09 paper/sheaf_lics.tex

%% bare_conf.tex
%% V1.4a
%% 2014/09/17
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8a or later) with an IEEE
%% conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_conf_compsoc.tex,
%%                    bare_jrnl_compsoc.tex, bare_jrnl_transmag.tex
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices and paper sizes can       ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[conference]{IEEEtran}
% Some Computer Society conferences also require the compsoc mode option,
% but others use the standard conference format.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/tex-archive/info/epslatex/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
\usepackage[cmex10]{amsmath}
\usepackage{amssymb,amsthm,amsfonts}

\newtheorem{thm}{Theorem}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{defi}[thm]{Definition}
\newtheorem{lem}[thm]{Lemma}

% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Do not use the stfloats baselinefloat ability as IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/dblfloatfix/

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/url/
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\usepackage{wasysym}
\usepackage[all]{xy}
\def\dar[#1]{\ar@<2pt>[#1]\ar@<-2pt>[#1]}
\DeclareMathOperator{\Type}{Type}
\DeclareMathOperator{\HProp}{HProp}
\newcommand{\modal}{\ensuremath{\ocircle}}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Sheafification in HoTT}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Kevin Quirin and Nicolas Tabareau}
\IEEEauthorblockA{INRIA \\ \'Ecole des Mines de Nantes \\ Nantes, France}}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
Abstract
\end{abstract}

% no keywords

\section{Introduction}
\label{sec:intro}

\section{Homotopy Type Theory}
\label{sec:homotopy-type-theoryhott}

\subsection{Homotopy Type Theory}
\label{ssec:hott}

We refer to~\cite{hottbook}.

\begin{itemize}
\item

We use the subobjects classification seen in~\cite{sets_in_hott}~: for
all type $B$, $\Type_n$ classifies subobjects of $B$ with $n$-truncated
fibers ; we have an equivalence
$$\sum_{A:\Type} \sum_{f:A \to B} \prod_{b\in B}
\mathrm{IsTrunc}~n~\mathrm{fib}_f(b) \simeq (B \to \Type_n)$$
and a pullback (with $\Type_n^\bullet$ the universe of pointed
$n$-truncated types)~:
$$
\xymatrix{
  A \ar[r]^{t_f} \ar[d]_f & \Type_n^\bullet \ar[d]^{\mathrm{pr}_1}\\
  B \ar[r]_{\chi_f} & \Type_n
}
$$
forall $f$ with $n$-truncated fibers, with 
$$t_f = \lambda a,~(\mathrm{fib}_f(f(a)),(a,\mathrm{idpath})).$$
\end{itemize}

\subsection{Left exact Modalities}
\label{sec:lexmod}

We use the following definition of truncated left exact modalities~:
\begin{defi}
  Let $p$ be a truncation index. Then a left exact modality is the
  data of
  \begin{itemize}
  \item A predicate $P~:~\Type_p \to \HProp$
  \item For every $p$-truncated type $A$, a $p$-truncated type
    $\modal A$ such that $P(\modal A)$
  \item For every $p$-truncated type $A$, a map $\eta_A~:~A \to
    \modal A$
  \end{itemize}
  such that
  \begin{itemize}
  \item For every $p$-truncated types $A$ and $B$, if $P(B)$ then
    $$\left\{
      \begin{array}{rcl}
        (\modal A \to B) & \to & (A \to B) \\
        f & \mapsto & f \circ \eta_A
      \end{array} \right.$$
    is an equivalence.
  \item For all $A~:~\Type_p$ and $B~:~A \to \Type_p$ such that $P(A)$
    and $\Pi_{x:A} P(B x)$, then $P\left( \Sigma_{x:A} B(x)\right)$
  \item For all $A~:~\Type_p$ and $x,y~:~A$, if $\modal A$ is
    contractible, then $\modal (x=y)$ is contractible.
  \end{itemize}
\end{defi}

All theorem in~\cite{hottbook}, chapter 7.7 remains true + fibers
preservation :

for all $n$-truncated types,
  and all map $f~:~X \to Y$, the modalisation of fiber of $f$ above any element $y:Y$
  is the fiber of $\modal f$ above $\eta y$~:
  $$\modal \{x:X ~\&~ f~ x = y\} = \{x:\modal X ~\&~ \modal f ~x = \eta_Y
  y\}.$$

  We want moreover that the following diagram commute
  $$\xymatrix{
    \{x:X ~\&~ f~ x = y\} \ar[r]^\eta \ar[d]_\gamma & \modal \{x:X ~\&~ f~ x = y\} \ar@{=}[dl]\\
    \{x:\modal X ~\&~ \modal f ~x = \eta_Y y\} &
  }$$
  where $\Pi_1 \circ \gamma = \eta_X$, and $\Pi_2 \circ \gamma$ is the
  usual modalisation of paths.

\subsection{Generalized pullbacks}
\label{sec:gener-pb}

\begin{itemize}
\item definition
\item giraud axiom
\end{itemize}

\section{Sheaves}

We try to define sheaves inductively~: we define sheaves in $\HProp$
as the classical types ({\em ie} implied by their double negation), and
we define sheaves on $\Type_{n+1}$ {\em via} usual definition of sheaves,
as seen in~maclanemoerdijk and sheaves on $\Type_n$. Of course,
the definition must be cumulative.

We will call $\mathcal S_i$ the universe of sheaves on $\Type_i$.

\subsection{Definitions and forst properties}
\label{sec:def}

\begin{itemize}
\item Density
\item Separated Type / Sheaf / closure : a subobject is always dense
  into its closure.
\item $\Omega_n^\modal$ is a sheaf.
\item Sheaves are stable by dependent products.
\end{itemize}

\subsection{Construction of sheafification}
\label{sec:sheafification}

We mimic the construction in~\cite{maclanemoerdijk}
Two steps~: separation, sheafification.

\subsubsection{For h-propositions}
\label{sec:h-propositions}

\begin{itemize}
\item Any left exact modality on $\HProp$.
\item We use $\lnot\lnot$
\end{itemize}
\subsubsection{From Type to Separated Type}
\label{sec:from-type-separated}

\begin{itemize}
\item Definition of the separation reflective subuniverse
  $(\square,~\mu)$.

  $\square$ + $\square T$ is separated for all $(n+1)$-truncated
    type $T$ 
    \begin{thm}
      A $(n+1)$-truncated type $T$ with an embedding $f : T \to U$
      into a separated $(n+1)$-truncated type $U$ is itself separated.
    \end{thm}

\item If $a,b~:~T~: \Type_{n+1}$ such that $\mu_T a = \mu_T b$, then
  $\modal (a=b)$.

\item Cech nerve of $\mu$ is the diagram of closure of diagonals,
  hence their colimits are equal. Since colimits are universal, so is $\square$.

\item $(\square,~\mu)$ is a modality.

\end{itemize}

\subsubsection{From Separated Type to Sheaf}
\label{sec:separated-to-sheaf}

\begin{itemize}
\item 
\begin{thm}
  Any type $A~:~\Type_{n+1}$ closed in a sheaf is itself a sheaf.
\end{thm}

\item Definition of sheafification reflextive subuniverse $(\star,~\nu)$
\end{itemize}

\section{Future Works}
\label{sec:future-works}

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively. 
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.

\section{Conclusion}
The conclusion goes here.

% conference papers do not normally have an appendix

% use section* for acknowledgment
\section*{Acknowledgment}

The authors would like to thank...

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{3}

\bibitem[RS13]{sets_in_hott}
Egbert Rijke and Bas Spitters.
\newblock Sets in homotopy type theory.
\newblock 2013.

\bibitem[{Uni}13]{hottbook}
{Univalent Foundations Project}.
\newblock {\em Homotopy Type Theory: Univalent Foundations for Mathematics}.
\newblock 2013.

\bibitem[MM92]{maclanemoerdijk}
Saunders MacLane and Ieke Moerdijk.
\newblock {\em Sheaves in Geometry and Logic}.
\newblock Springer-Verlag, 1992.

\end{thebibliography}

% that's all folks
\end{document}

### 2015-01-10 paper/sheaf_lics.tex
paper

@@
 
+\usepackage{enumerate}
 \usepackage{wasysym}
@@
 \DeclareMathOperator{\HProp}{HProp}
+\DeclareMathOperator{\nat}{nat}
+\DeclareMathOperator{\Unit}{Unit}
 \newcommand{\modal}{\ensuremath{\ocircle}}
@@
 
+Plan of the paper~:
+\begin{itemize}
+\item In section~\ref{sec:hott}, we define the foundations on which
+  our worked is based.
+\item In section~\ref{sec:sheaves}, we explain the construction of the
+  sheafification functor.
+\item Finally, we discuss related works and present future works.
+\end{itemize}
 \section{Homotopy Type Theory}
-\label{sec:homotopy-type-theoryhott}
+\label{sec:hott}
 
@@
 We refer to~\cite{hottbook}.
+Truncation levels ?
+
+If $x$ is an object of a product or a sigma type, we denote the $i$-th
+projection of $x$ by either $\pi_i x$ or $x_i$.
 
@@
   data of
-  \begin{itemize}
-  \item A predicate $P~:~\Type_p \to \HProp$
+  \begin{enumerate}[(i)]
+  \item A predicate $P:\Type_p \to \HProp$
   \item For every $p$-truncated type $A$, a $p$-truncated type
     $\modal A$ such that $P(\modal A)$
-  \item For every $p$-truncated type $A$, a map $\eta_A~:~A \to
+  \item For every $p$-truncated type $A$, a map $\eta_A:A \to
     \modal A$
-  \end{itemize}
+  \end{enumerate}
   such that
-  \begin{itemize}
+  \begin{enumerate}[(i)]
+    \setcounter{enumi}{3}
   \item For every $p$-truncated types $A$ and $B$, if $P(B)$ then
@@
     is an equivalence.
-  \item For all $A~:~\Type_p$ and $B~:~A \to \Type_p$ such that $P(A)$
-    and $\Pi_{x:A} P(B x)$, then $P\left( \Sigma_{x:A} B(x)\right)$
-  \item For all $A~:~\Type_p$ and $x,y~:~A$, if $\modal A$ is
+  \item For all $A:\Type_p$ and $B:A \to \Type_p$ such that $P(A)$
+    and $\Pi_{x:A} P(B x)$, then $P\left( \sum_{x:A} B(x)\right)$
+  \item For all $A:\Type_p$ and $x,y:A$, if $\modal A$ is
     contractible, then $\modal (x=y)$ is contractible.
-  \end{itemize}
+  \end{enumerate}
+  Conditions (i) to (iv) define a {\em reflective subuniverse}, (i) to
+  (v) a {\em modality}.
 \end{defi}
 
-All theorem in~\cite{hottbook}, chapter 7.7 remains true + fibers
-preservation :
+Since basic operations (dependent products, products, sigma types) are
+stable for truncation levels, all theorem in~\cite{hottbook}, chapter
+7.7 remains true.
 
-for all $n$-truncated types,
-  and all map $f~:~X \to Y$, the modalisation of fiber of $f$ above any element $y:Y$
-  is the fiber of $\modal f$ above $\eta y$~:
-  $$\modal \{x:X ~\&~ f~ x = y\} = \{x:\modal X ~\&~ \modal f ~x = \eta_Y
-  y\}.$$
+Moreover, left exactness implies in particular fibers preservations~:
+\begin{prop}
+For any $n$-truncated types $X$ and $Y$,
+and any map $f:X \to Y$, the modalisation of fiber of $f$ above any element $y:Y$
+is the fiber of $\modal f$ above $\eta_Y y$~:
+$$\modal \left(\sum_{x:X}  (f x = y)\right) = \sum_{x:\modal X} (\modal f x = \eta_Y y).$$
 
-  We want moreover that the following diagram commute
+Moreover the following diagram commute
 $$\xymatrix{
-    \{x:X ~\&~ f~ x = y\} \ar[r]^\eta \ar[d]_\gamma & \modal \{x:X ~\&~ f~ x = y\} \ar@{=}[dl]\\
-    \{x:\modal X ~\&~ \modal f ~x = \eta_Y y\} &
+  \sum_{x:X} (f x = y) \ar[r]^\eta \ar[d]_\gamma & \modal \left(\sum_{x:X}  (f x = y)\right) \ar@{=}[dl]\\
+  \sum_{x:\modal X} (\modal f x = \eta_Y y) &
 }$$
-  where $\Pi_1 \circ \gamma = \eta_X$, and $\Pi_2 \circ \gamma$ is the
+where $\pi_1 \circ \gamma = \eta_X$, and $\pi_2 \circ \gamma$ is the
 usual modalisation of paths.
-
+\end{prop}
 \subsection{Generalized pullbacks}
@@
 
-\begin{itemize}
-\item definition
-\item giraud axiom
-\end{itemize}
+\begin{defi}
+  Let $f:X \to Y$ be a map, and $p:\nat$. The $p$-pullback of $f$,
+  noted $X\times_Y \cdots\times_Y X$ is
+  the limit of the diagram
+  $$\xymatrix{
+    X \ar[rd]^f & \cdots & X \ar[ld]_f \\
+      &    Y   &
+  }$$
+  with $p$ copies of $X$. The $0$-pullback of $f$ is $\Unit$.
+\end{defi}
+
+In homotopy type theory, we have 
+$$X\times_Y \cdots\times_Y X = \sum_{x:X^n} (f x_1 = f x_2) \land
+\cdots \land (f x_{n-1} = f x_n).$$
+\begin{defi}
+  If $f:X \to Y$ is a map, then the {\em \v{C}ech nerve} of $f$ is the
+  diagram
+  $$C(f) := \cdots
+    X \times_Y X \times_Y X
+   \underset{\longrightarrow}{\stackrel{\longrightarrow}{\longrightarrow}}
+   X \times_Y X 
+\stackrel{\longrightarrow}{\longrightarrow} X$$
+with canonical projections.
+\end{defi}
+
+The Giraud axiom asserts that, if $f:X\to Y$ is a surjection, then the
+colimit of $C(f)$ is $Y$. See~\cite{lurie}.
 
 \section{Sheaves}
+\label{sec:sheaves}
 
-We try to define sheaves inductively~: we define sheaves in $\HProp$
-as the classical types ({\em ie} implied by their double negation), and
+We define sheaves by induction on the homotopical truncation level of types
+~: we define sheaves in $\HProp$ as the classical types ({\em ie}
+implied by their double negation), and
 we define sheaves on $\Type_{n+1}$ {\em via} usual definition of sheaves,
-as seen in~maclanemoerdijk and sheaves on $\Type_n$. Of course,
+as seen in~\cite{maclanemoerdijk} and sheaves on $\Type_n$. Of course,
 the definition must be cumulative.
@@
 
-\subsection{Definitions and forst properties}
+\subsection{Definitions and first properties}
 \label{sec:def}
@@
 \end{itemize}
+
+\begin{defi}
+  Let $E$ be a type. The {\em closure} of a subobject of $E$ classified by $\chi$
+  is the subobject of $E$ classified by $\modal \circ \chi$.
+
+  The subobject of $E$ classified by $\chi$ is said {\em closed in
+    $E$} if its closure is itself : $\chi = \modal \circ \chi$.
+\end{defi}
+
+\begin{defi}
+  Let $E$ be a type, and $\chi:E \to \Type_n$. The subobject $A$ of $E$
+  classified by $\chi$ is {\em dense} in $E$ if its $\modal$-closure
+  is $E$ seen as a subobject of $E$, {\em ie}
+  $$\forall e:E,~ \left(\sum_{e':E} e=e'\right) \simeq (\modal (\chi~e)).$$ 
+  Moreover, for all $x:A$ we need the following coherence diagram to
+  commute
+  $$\xymatrix{
+    \sum_{e':A} x = e' \ar@{=}[r] \ar[d]_\iota & (\chi~x) \ar[d]^{\eta_{(\chi~x)}}\\
+    \sum_{e':E} x = e' \ar@{=}[r] & \modal (\chi~x)
+  }$$
+  where $\iota: x \mapsto ({x_1}_1 ; x_2)$.
+\end{defi}
+
+It follows from fibers preservation that any $n$-subobject of a type seen as a $n$-subobject
+of its closure is closed.
 
@@
 
-\item If $a,b~:~T~: \Type_{n+1}$ such that $\mu_T a = \mu_T b$, then
+\item If $a,b:T:\Type_{n+1}$ such that $\mu_T a = \mu_T b$, then
   $\modal (a=b)$.
@@
 \begin{thm}
-  Any type $A~:~\Type_{n+1}$ closed in a sheaf is itself a sheaf.
+  Any type $A:\Type_{n+1}$ closed in a sheaf is itself a sheaf.
 \end{thm}
 
-\item Definition of sheafification reflextive subuniverse $(\star,~\nu)$
+\item Definition of sheafification reflextive subuniverse
+  $(\star,~\nu)$
+\item left exactness
+\item cumulativity
 \end{itemize}
@@
 \label{sec:future-works}
+
+Prove that the Giraud axiom holds in homotopy type theory.
 
@@
 % (used to reserve space for the reference number labels box)
-\begin{thebibliography}{3}
+\begin{thebibliography}{4}
 
@@
 
+\bibitem[HTT]{lurie}
+Jacob Lurie
+\newblock{\em Higher Topos Theory}
+
+\bibitem[lumsdaine]{lumsdaine}
+Lumsdaine
+\newblock{Diagrams and limits}
+
 \end{thebibliography}

### 2015-01-11 paper/sheaf_lics.tex
paper

@@
-%% bare_conf.tex
-%% V1.4a
-%% 2014/09/17
-%% by Michael Shell
-%% See:
-%% http://www.michaelshell.org/
-%% for current contact information.
-%%
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.8a or later) with an IEEE
-%% conference paper.
-%%
-%% Support sites:
-%% http://www.michaelshell.org/tex/ieeetran/
-%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
-%% and
-%% http://www.ieee.org/
-
-%%*************************************************************************
-%% Legal Notice:
-%% This code is offered as-is without any warranty either expressed or
-%% implied; without even the implied warranty of MERCHANTABILITY or
-%% FITNESS FOR A PARTICULAR PURPOSE! 
-%% User assumes all risk.
-%% In no event shall IEEE or any contributor to this code be liable for
-%% any damages or losses, including, but not limited to, incidental,
-%% consequential, or any other damages, resulting from the use or misuse
-%% of any information contained here.
-%%
-%% All comments are the opinions of their respective authors and are not
-%% necessarily endorsed by the IEEE.
-%%
-%% This work is distributed under the LaTeX Project Public License (LPPL)
-%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
-%% distributed and modified. A copy of the LPPL, version 1.3, is included
-%% in the base LaTeX documentation of all distributions of LaTeX released
-%% 2003/12/01 or later.
-%% Retain all contribution notices and credits.
-%% ** Modified files should be clearly indicated as such, including  **
-%% ** renaming them and changing author support contact information. **
-%%
-%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
-%%                    bare_conf.tex, bare_jrnl.tex, bare_conf_compsoc.tex,
-%%                    bare_jrnl_compsoc.tex, bare_jrnl_transmag.tex
-%%*************************************************************************
-
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. IEEE's font choices and paper sizes can       ***
-% *** trigger bugs that do not appear when using other class files.       ***                          ***
-% The testflow support page is at:
-% http://www.michaelshell.org/tex/testflow/
-
 \documentclass[conference]{IEEEtran}
-% Some Computer Society conferences also require the compsoc mode option,
-% but others use the standard conference format.
-%
-% If IEEEtran.cls has not been installed into the LaTeX system files,
-% manually specify the path to it like:
-% \documentclass[conference]{../sty/IEEEtran}
-
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
-
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
-
-% *** CITATION PACKAGES ***
-%
-%\usepackage{cite}
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off
-% such as if a citation ever needs to be enclosed in parenthesis.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 5.0 (2009-03-20) and later if using hyperref.sty.
-% The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
-% The documentation is contained in the cite.sty file itself.
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
-  % \usepackage[pdftex]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../pdf/}{../jpeg/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-\else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
-  % \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.eps}
-\fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation
-% can be obtained at: 
-% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found at:
-% http://www.ctan.org/tex-archive/info/epslatex/
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
-
 % *** MATH PACKAGES ***
@@
 \newtheorem{lem}[thm]{Lemma}
-
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics. If using
-% it, be sure to load this package with the cmex10 option to ensure that
-% only type 1 fonts will utilized at all point sizes. Without this option,
-% it is possible that some math symbols, particularly those within
-% footnotes, will be rendered in bitmap form which will result in a
-% document that can not be IEEE Xplore compliant!
-%
-% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/
-
-% *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
-% There is also a support site at:
-% http://algorithms.berlios.de/index.html
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/
-
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/tools/
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-
-% *** SUBFIGURE PACKAGES ***
-%\ifCLASSOPTIONcompsoc
-%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-%\else
-%  \usepackage[caption=false,font=footnotesize]{subfig}
-%\fi
-% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
-% for subfigure.sty, the latter of which is no longer maintained and is
-% incompatible with some LaTeX packages including fixltx2e. However,
-% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
-% which will override IEEEtran.cls' handling of captions and this will result
-% in non-IEEE style figure/table captions. To prevent this problem, be sure
-% and invoke subfig.sty's "caption=false" package option (available since
-% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
-% handling of captions.
-% Note that the Computer Society format requires a larger sans serif font
-% than the serif footnote size font used in traditional IEEE formatting
-% and thus the need to invoke different subfig.sty package options depending
-% on whether compsoc mode has been enabled.
-%
-% The latest version and documentation of subfig.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
-
-% *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure. The latest version and documentation can be found at:
-% http://www.ctan.org/tex-archive/macros/latex/base/
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
-% Do not use the stfloats baselinefloat ability as IEEE does not allow
-% \baselineskip to stretch. Authors submitting work to the IEEE should note
-% that IEEE rarely uses double column equations and that authors should try
-% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
-% packages (also by Sigitas Tolusis) as IEEE does not format its papers in
-% such ways.
-% Do not attempt to use stfloats with fixltx2e as they are incompatible.
-% Instead, use Morten Hogholm'a dblfloatfix which combines the features
-% of both fixltx2e and stfloats:
-%
-% \usepackage{dblfloatfix}
-% The latest version can be found at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/dblfloatfix/
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-%\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/url/
-% Basically, \url{my_url_here}.
 
@@
 \DeclareMathOperator{\Unit}{Unit}
+\DeclareMathOperator{\im}{Im}
+\DeclareMathOperator{\id}{id}
 \newcommand{\modal}{\ensuremath{\ocircle}}
@@
 We refer to~\cite{hottbook}.
-Truncation levels ?
+Truncation levels + Im ?
 
@@
   \item For all $A:\Type_p$ and $B:A \to \Type_p$ such that $P(A)$
-    and $\Pi_{x:A} P(B x)$, then $P\left( \sum_{x:A} B(x)\right)$
+    and $\prod_{x:A} P(B x)$, then $P\left( \sum_{x:A} B(x)\right)$
   \item For all $A:\Type_p$ and $x,y:A$, if $\modal A$ is
@@
 \end{defi}
+
+The type of all $p$-types such that $P$ will be noted $\Type^\modal_p$
 
@@
 
-\begin{itemize}
-\item Density
-\item Separated Type / Sheaf / closure : a subobject is always dense
-  into its closure.
-\item $\Omega_n^\modal$ is a sheaf.
-\item Sheaves are stable by dependent products.
-\end{itemize}
+In this section, we suppose given a truncation index $n\geqslant -1$,
+and a left exact modality $\modal$ on $\Type_n$.
 
@@
 
+\begin{defi}[Restriction]
+  For any $A, E:\Type$ and $F:\Type_{n+1}$, we define~:
+  \begin{itemize}
+  \item if $A \hookrightarrow E$, $M_E^\chi$ is the map sending an
+    arrow $f~:~E\to F$ to $\restriction f A$ the restriction of $f$ to $A$.
+  \item if $A \to E$ with $n$-truncated fibers, $\Phi_E^\chi$ is the
+    map sending $f~:~E \to f$ to $\restriction f A$.
+  \end{itemize}
+\end{defi}
+
+\begin{defi}[Separated Type]
+  A type $F$ in $\Type_{n+1}$ is {\em separated} if for all type $E$, and
+  all dense subobject of $E$ classified by $\chi~:~E \to \Type_n$,
+  $\Phi_E^\chi$ is a monomorphism. In other words, the dotted arrow,
+  if exists, is unique.
+
+  $$\xymatrix{
+    \sum_{e:E} (\chi~ e) \ar[r]^f \ar[d]_{\pi_1} & F \\
+    E \ar@{-->}[ru]&
+  }$$
+\end{defi}
+
+\begin{defi}[Sheaves]
+  A type $F$ of $\Type_{n+1}$ is a {\em $(n+1)$-sheaf} if it is
+  separated, and for all type $E$ and all dense subobject of $E$
+  classified by $\chi~:~E \to \Type_{-1}$, $M_E^\chi$ is an
+  equivalence. In other words, the dotted arrow exists and is unique.
+
+  $$\xymatrix{
+    \sum_{e:E} (\chi~ e) \ar[r]^f \ar[d]_{\pi_1} & F \\
+    E \ar@{-->}[ru]&
+  }$$
+\end{defi}
+
+From these definitions, one can show that
+\begin{prop}\label{prop:sheaf_prop}
+  \begin{itemize}
+  \item $\Type_n^\modal$ is a sheaf.
+  \item If $A:\Type_{n+1}$ and $B:A \to \Type_{n+1}$ such that for all
+    $a:A$, $(B~a)$ is a sheaf, then $\prod_{a:A}B~a$ is a sheaf.
+  \end{itemize}
+\end{prop}
+
 \subsection{Construction of sheafification}
@@
 We mimic the construction in~\cite{maclanemoerdijk}
-Two steps~: separation, sheafification.
 
@@
 
-\begin{itemize}
-\item Any left exact modality on $\HProp$.
-\item We use $\lnot\lnot$
-\end{itemize}
+For the case $n=-1$, one can take any left exact modality on
+$\HProp$. Here we took the $\lnot\lnot$ modality~:
+$$\forall P:\HProp,~\modal P = \lnot\lnot P.$$
+One can easily show that it defines a left exact modality.
+
 \subsubsection{From Type to Separated Type}
 \label{sec:from-type-separated}
+In this section, we suppose given a truncation index $n\geqslant -1$,
+and a left exact modality $\modal$ on $\Type_n$, compatible with the
+modality on $\HProp$.
+
+Let $T$ be in $\Type_{n+1}$. We define $\square T$ as the image of
+$\modal^T \circ \{\cdot\}_T$, as in
+$$\xymatrix{
+  T \ar[r]^{\{\cdot\}_T} \ar[d]_{\mu_T} & \Type_n^T \ar[d]^{\modal^T} \\
+  \square T \ar[r]& \left( \Type_n^\modal \right)^T
+}, $$
+where $\{\cdot\}_T$ is the singleton map $\lambda (t:T),~\lambda
+(t':T),~t=t'$.
+In type theory words, 
+$$\square T = \im (\lambda~t:T,~\lambda~ t',~ \modal (t = t')).$$
+
+\begin{prop}
+  For all $T:\Type_{n+1}$, $\square T$ is separated.  
+\end{prop}
+
+\begin{proof}
+We will use the following lemma~:
+\begin{lem}
+  A $(n+1)$-truncated type $T$ with an embedding $f : T \to U$
+  into a separated $(n+1)$-truncated type $U$ is itself separated.
+\end{lem}
+As $\square T$ embeds in $\left( \Type_n^\modal \right)^T$, we only
+have to show that the latter is separated. But it is the case with
+both parts proposition~\ref{prop:sheaf_prop}.
+\end{proof}
+
+We now need to show that $\square$ is universal. The following
+lemma is central in the proof~:
+\begin{lem}
+  Let $T:\Type_{n+1}$. Then $\square T$ is the colimit of the closed
+  diagonal diagram
+  $$\cdots \overline{\Delta_3} \underset{\longrightarrow}{\stackrel{\longrightarrow}{\longrightarrow}}
+   \overline{\Delta_2}
+\stackrel{\longrightarrow}{\longrightarrow} \overline{\Delta_1},$$
+where $\Delta_k$ is the $k$-pullback of $\id : T \to T$.
+\end{lem}
+\begin{proof}
+  As $\mu_T$ is an embedding, we know by Giraud axiom that $\square T$ is the colimit
+  of $C(\mu_T)$. If we can show that $C(\mu_T) = C(\id)$, the the
+  result will follow.
+
+  Let $k:\nat$, let's show that 
+\begin{align*}
+  &\sum_{t:T^k} (\mu_T t_1 = \mu_T t_2 \land \cdots
+    \land \mu_T t_{k-1} \\
+  = &\mu_T t_k) = \sum_{t:T^k} \modal (t_1 = t_2 \land \cdots
+  \land t_{k-1} = t_k)
+\end{align*}
+By induction on $k$, and the preservation of products by $\modal$, it
+suffices to show that for all $a,b:T$, $\modal (a=b) = (\mu_T a =
+\mu_T b)$, which is true by univalence.
+If $U$ is a separated type, we can now easily define an inverse to
+$\lambda\,(f:\square T \to f),~x \circ \mu_T$.
+\end{proof}
 
 \begin{itemize}
-\item Definition of the separation reflective subuniverse
-  $(\square,~\mu)$.
-
-  $\square$ + $\square T$ is separated for all $(n+1)$-truncated
-    type $T$ 
-    \begin{thm}
-      A $(n+1)$-truncated type $T$ with an embedding $f : T \to U$
-      into a separated $(n+1)$-truncated type $U$ is itself separated.
-    \end{thm}
-
-\item If $a,b:T:\Type_{n+1}$ such that $\mu_T a = \mu_T b$, then
-  $\modal (a=b)$.
-
-\item Cech nerve of $\mu$ is the diagram of closure of diagonals,
-  hence their colimits are equal. Since colimits are universal, so is $\square$.
 

### 2015-01-12 paper/sheaf_lics.tex
extend is_classical_IsTrunc to -2

@@
 
-\subsection{Definitions and first properties}
-\label{sec:def}
+% \subsection{Definitions and first properties}
+% \label{sec:def}
 
@@
 
-\subsection{Construction of sheafification}
+\section{Construction of sheafification}
 \label{sec:sheafification}
@@
 
-\subsubsection{For h-propositions}
-\label{sec:h-propositions}
+\subsection{For h-propositions}
+\label{ssec:h-propositions}
 
@@
 $\HProp$. Here we took the $\lnot\lnot$ modality~:
-$$\forall P:\HProp,~\modal P = \lnot\lnot P.$$
+$$\forall P:\HProp,~\modal_j P = \lnot\lnot P.$$
 One can easily show that it defines a left exact modality.
 
-\subsubsection{From Type to Separated Type}
-\label{sec:from-type-separated}
+\subsection{From Type to Separated Type}
+\label{ssec:from-type-separated}
 In this section, we suppose given a truncation index $n\geqslant -1$,
@@
 In type theory words, 
-$$\square T = \im (\lambda~t:T,~\lambda~ t',~ \modal (t = t')).$$
+\begin{align*}
+\square T &= \im (\lambda~t:T,~\lambda~ t',~ \modal (t = t')) \\
+          &= \sum_{u:T \to \Type_n^\modal} \left\| \sum_{a:X} 
+            (\lambda t,~\modal (a=t)) = u\right\|
+\end{align*}
 
@@
 \mu_T b)$, which is true by univalence.
-If $U$ is a separated type, we can now easily define an inverse to
-$\lambda\,(f:\square T \to f),~x \circ \mu_T$.
 \end{proof}
 
-\begin{itemize}
+Now that $\square T$ is a colimit, we can easily define an inverse to
+$\lambda\,(f:\square T \to U),~x \circ \mu_T$ for any separated type
+$U$, using the universal property of the colimit.
 
-\item $(\square,~\mu)$ is a modality.
+We now have to prove point {\it (v)} in the definition of
+modality. Let $A:\Type_{n+1}$ be a sheaf and $B:A \to \Type_{n+1}$ be
+a sheaf family. We want to show that $\sum_{x:A} (Bx)$ is
+separated. Let $E$ be a type, and $\sum_{e:E} (\chi\,e)$ a dense
+subobject of E.
 
-\end{itemize}
+Let $f,g$ be two maps from $\sum_{e:E} (\chi\,e)$ to $\sum_{x:A}
+(Bx)$, equal when precomposed with $\pi_1$.
 
-\subsubsection{From Separated Type to Sheaf}
-\label{sec:separated-to-sheaf}
+As $A$ is separated, for any $x:E$, the first components of $f x$ and
+$g x$ are equal. For the second components, notice that for all $x:E$,
+$\sum_{y:E} x = y$ has a dense subobject, $\sum_{y:\sum_{e:E} (\chi\,
+  e)} x=y_1$. 
+Using the separation property with these type, one can show the second
+components of $f x$ and $g x$ are equals too, with respect to the
+first path.
 
-\begin{itemize}
-\item 
-\begin{thm}
-  Any type $A:\Type_{n+1}$ closed in a sheaf is itself a sheaf.
-\end{thm}
+\subsection{From Separated Type to Sheaf}
+\label{ssec:separated-to-sheaf}
 
-\item Definition of sheafification reflextive subuniverse
-  $(\star,~\nu)$
-\item left exactness
-\item cumulativity
-\end{itemize}
+If $T$ is already a separated type, the following lemma alllows us
+to build a sheaf~:
+\begin{lem}
+  Let $T:\Type_{n+1}$ be separated, and $U$ be a sheaf. If $T$ embeds
+  in $U$, and is closed in $U$, then $T$ is a sheaf.
+\end{lem}
 
+As any separated type $T$ embeds in $\left(\Type_n^\modal\right)^T$,
+it suffices to take the closure of $T$ to get a sheaf.
+
+\subsection{Summary}
+\label{ssec:summary}
+
+If $T:\Type_{n+1}$, the sheafification of $T$ is
+$$\star T = \sum_{u:T \to \Type_n^\modal} \lnot\lnot\left\| \sum_{a:X} 
+            (\lambda t,~\modal (a=t)) = u\right\|$$
+
+We already know that $\star$ defines a modality. It remains to show
+that this modality is left exact.
+
+The last thing to prove is the cumulativity of our sheafifications~:
+\begin{prop}
+  If $T:\Type_n$, then $\modal T = \star \widehat T$ (where $\widehat
+  T$ is $T$ seen as a $\Type_{n+1}$.
+\end{prop}
 \section{Future Works}
@@
 Prove that the Giraud axiom holds in homotopy type theory.
+
+Extend the sheafification to all types, even if non truncated.
 

### 2015-01-12 paper/sheaf_lics.tex
paper

@@
 \def\dar[#1]{\ar@<2pt>[#1]\ar@<-2pt>[#1]}
+\def\tar[#1]{\ar@<2pt>[#1]\ar@<0pt>[#1]\ar@<-2pt>[#1]}
 \DeclareMathOperator{\Type}{Type}
@@
 \DeclareMathOperator{\id}{id}
+\DeclareMathOperator{\Contr}{Contr}
 \newcommand{\modal}{\ensuremath{\ocircle}}
@@
   diagram
-  $$C(f) := \cdots
-    X \times_Y X \times_Y X
-   \underset{\longrightarrow}{\stackrel{\longrightarrow}{\longrightarrow}}
-   X \times_Y X 
-\stackrel{\longrightarrow}{\longrightarrow} X$$
+  $$\xymatrix{
+    C(f) := \cdots~ X \times_Y X \times_Y X \tar[r] & X \times_Y X \dar[r] & X
+  }$$
 with canonical projections.
@@
   diagonal diagram
-  $$\cdots \overline{\Delta_3} \underset{\longrightarrow}{\stackrel{\longrightarrow}{\longrightarrow}}
-   \overline{\Delta_2}
-\stackrel{\longrightarrow}{\longrightarrow} \overline{\Delta_1},$$
+  $$\xymatrix{ 
+    \cdots \overline{\Delta_3} \tar[r] & \overline{\Delta_2} \dar[r] &
+    \overline{\Delta_1}
+  }$$
 where $\Delta_k$ is the $k$-pullback of $\id : T \to T$.
@@
 that this modality is left exact.
+
+Let $T:\Type_{n+1}$ such that $\star X$ is contractible. Let
+$x,y:T$. We want to show that $\star (x=y)$ is contractible.
+
+As $\star T$ is contractible, we have a center $u_0 : T \to
+\Type_n^\modal$ with an hypothesis $H_0 :  \lnot\lnot\left\| \sum_{a:X} 
+            (\lambda t,~\modal (a=t)) = u_0\right\|$, and a
+          contractibility condition $c : \prod_{t:\star T} x = (u_0,H_0)$.
+As $\Contr (\star (x=y))$ is a $\HProp$-sheaf, we can change $H_0$
+into a $a:T$ and a proof $u_0 = (\lambda t,~\modal (a=t))$.
+Finally, $c$ type becomes $\prod_{v:\star T} v_1 =  (\lambda t,~\modal (a=t))$.
 
@@
 Egbert Rijke and Bas Spitters.
-\newblock Sets in homotopy type theory.
+\newblock {\em Sets in homotopy type theory}.
 \newblock 2013.
@@
 \bibitem[lumsdaine]{lumsdaine}
-Lumsdaine
-\newblock{Diagrams and limits}
+Peter LeFanu Lumsdaine, Jeremy Avigad and Chris Kapulkin
+\newblock{\em Homotopy limits in type theory}
 

### 2015-01-13 paper/sheaf_lics.tex

@@
 \end{prop}
+
+Another useful properties of modalities is the following~:
+\begin{prop}
+  If $P:\HProp$, then $\modal \widehat P : \HProp$, where $\widehat P$
+  is $P$ seen as a $p$-type.
+\end{prop}
+
+A left exact modality leads to a new theory with new principles. A
+basic example is the {\em open modality of $P$} for a $P:\HProp$, $\modal_P T = P \to T$. 
+When one works in the subuniverse defined by this modality, defining
+an inhabitant of a type $T$ is actually an inhabitant of $P \to
+T$. Thus, being in this subuniverse is just adding the axiom $P$ to
+the theory. Sadly, adding an axiom using the modality does not add any
+computational content to $P$, thus it is not completely satisfactory.
+
+Another example is the $\lnot\lnot$-modality, $\modal_{\lnot\lnot} T = (T
+\to \bot) \to \bot$. Although we might think it enables us to use
+classical reasonings in the corresponding reflective subuniverse, the
+main problem is that every type in the new universe is an $\HProp$,
+which is very restrictive.
+
 \subsection{Generalized pullbacks}
@@
 
-We define sheaves by induction on the homotopical truncation level of types
-~: we define sheaves in $\HProp$ as the classical types ({\em ie}
-implied by their double negation), and
-we define sheaves on $\Type_{n+1}$ {\em via} usual definition of sheaves,
-as seen in~\cite{maclanemoerdijk} and sheaves on $\Type_n$. Of course,
-the definition must be cumulative.
+We will define sheaves by induction on the homotopical truncation level of types
+~: the base case will be any left exact modality on $\HProp$, and we
+give in this section some definitions useful for th inductive case.
 
@@
 
+Note that these definitions are almost the same as the ones
+in~\cite{maclanemoerdijk}. The main difference is that {\em separated}
+is defined for $n$-subobjects,
+while {\em sheaf} only for $-1$-subobject.
+
 From these definitions, one can show that
@@
 
+There again, the separation step has the same definition as
+in~\cite{maclanemoerdijk}, using $\Type_n^\modal$ instead of the
+$j$-subobject classifier.
+
 \begin{prop}
@@
 \end{lem}
+
+This lemma is an adaptation of the sheafification process
+in~\cite{maclanemoerdijk}, where they consider only the kernel pair of
+$\mu_T$ instead of the \v{C}ech nerve.
+
 \begin{proof}

### 2015-01-13 paper/sheaf_lics.tex
abstract of sheaf_lics.pdf

--- 2015-01-12 paper/sheaf_lics.tex
@@
 \documentclass[conference]{IEEEtran}
+
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
+
 % *** MATH PACKAGES ***
@@
 \newtheorem{lem}[thm]{Lemma}
+
+\newcommand{\mynote}[2]{
+    \fbox{\bfseries\sffamily\scriptsize#1}
+    {\small$\blacktriangleright$\textsf{\emph{#2}}$\blacktriangleleft$}~}
+
+\newcommand\kq[1]{\mynote{KQ}{#1}}
+\newcommand\nt[1]{\mynote{NT}{#1}}
 
@@
 % Do not put math or special symbols in the title.
-\title{Sheafification in HoTT}
+\title{Lawvere-Tierney Sheafification\\ in Homotopy Type Theory}
 
@@
 \author{\IEEEauthorblockN{Kevin Quirin and Nicolas Tabareau}
-\IEEEauthorblockA{INRIA \\ \'Ecole des Mines de Nantes \\ Nantes, France}}
+\IEEEauthorblockA{Inria \\ \'Ecole des Mines de Nantes \\ Nantes, France}}
 
@@
 \begin{abstract}
-Abstract
+
+  Sheafification is a popular tool in (higher) topos theory which
+  allows to extend the internal logic of a topos with new
+  principles. One of the most famous applications is the possibility to
+  transform a topos into a boolean topos using the dense topology, which
+  corresponds in essence to Gödel's double negation translation. 
+  % 
+  The same construction has not been considered in Martin-Löf type
+  theory because of a mismatch between topos and type theory. This
+  mismatched has been fixed recently by considering homotopy type
+  theory, an extension of Martin-Löf type theory with new
+  principles inspired by category theory and homotopy theory.
+  % 
+  In this paper, we give a computer checked construction of
+  Lawvere-Tierney sheafification in homotopy type theory, which allows
+  in particular to give a meaning to the (propositional) law of
+  excluded middle inside homotopy type theory, without axiom.
+
 \end{abstract}
@@
 \label{sec:intro}
+
+\begin{itemize}
+\item 
+  links between HoTT and higher topos theory 
+\item 
+  modalities in HoTT
+\item 
+  Giraud's axiom
+\item
+  previous attempts to extend Gödel's double negation to 
+  Martin-Löf type theory
+\item 
+  compare to the work of Parigot on $\lambda\mu$-calculus (Classical proofs as programs), and Coquand 
+  (A realization of the negative interpretation of the Axiom of
+  Choice)
+\item
+  work of Hugo Herbelin, A Constructive Proof of Dependent Choice,
+  Compatible with Classical Logic
+\item
+  Barthe (CPS Translating Inductive and Coinductive Types)
+  ``No translation is possible along the same lines for small $\Sigma$-types
+  and sum types with dependent case.''
+\item
+
+\end{itemize}
 
@@
 \label{ssec:hott}
+
+\nt{
+  Il faut au minimum introduire proprement les n-Types, le squash
+  sur les props 
+}
 
@@
 \label{sec:lexmod}
+
+\nt{
+To be checked:
+  Condition iv of Thm 7.7.5 of HoTT
+}
 

### 2015-01-13 paper/sheaf_lics.tex
abstract updated

@@
 \usepackage[T1]{fontenc}
+\usepackage{xspace}
 
@@
 \newcommand\nt[1]{\mynote{NT}{#1}}
+
+\newcommand{\ie}{i.e,\xspace}
+\newcommand{\eg}{e.g,\xspace}
 
@@
   allows to extend the internal logic of a topos with new
-  principles. One of the most famous applications is the possibility to
-  transform a topos into a boolean topos using the dense topology, which
-  corresponds in essence to Gödel's double negation translation. 
+  principles. One of its most famous applications is the possibility
+  to transform a topos into a boolean (classical) topos using the
+  dense topology, which corresponds in essence to Gödel's double
+  negation translation.
   % 
   The same construction has not been considered in Martin-Löf type
-  theory because of a mismatch between topos and type theory. This
+  theory because of a mismatch between topos theory and type theory. This
   mismatched has been fixed recently by considering homotopy type
   theory, an extension of Martin-Löf type theory with new
-  principles inspired by category theory and homotopy theory.
+  principles inspired by category theory and homotopy theory, and
+  which corresponds closely to higher toposes.
   % 
-  In this paper, we give a computer checked construction of
+  In this paper, we give a computer-checked construction of
   Lawvere-Tierney sheafification in homotopy type theory, which allows
   in particular to give a meaning to the (propositional) law of
-  excluded middle inside homotopy type theory, without axiom.
+  excluded middle inside homotopy type theory, without axiom, being
+  compatible with the full type-theoretic axiom of choice.
 

### 2015-01-13 paper/sheaf_lics.tex
add open modality, references to MM

@@
 \usepackage[all]{xy}
-\def\dar[#1]{\ar@<2pt>[#1]\ar@<-2pt>[#1]}
+\def\dar[#1]{\ar@<2pt>[#1]\ar@<-2pt>[#1]} %double arrows in xy
+\def\tar[#1]{\ar@<2pt>[#1]\ar@<0pt>[#1]\ar@<-2pt>[#1]} %triple arrows in xy
 \DeclareMathOperator{\Type}{Type}
@@
 
+Another useful property is that we can extend property {\it (iii)} to
+dependent product.
+
 Moreover, left exactness implies in particular fibers preservations~:
@@
 The Giraud axiom asserts that, if $f:X\to Y$ is a surjection, then the
-colimit of $C(f)$ is $Y$. See~\cite{lurie}.
+colimit of $C(f)$ is $Y$. Viewing homotopy type theory as a
+type-theoristic version of higher topos theory, it seems that this
+axiom can be proved~\cite{lurie}, although we admit it in this paper.
 
@@
 
-We define sheaves by induction on the homotopical truncation level of types
-~: we define sheaves in $\HProp$ as the classical types ({\em ie}
-implied by their double negation), and
-we define sheaves on $\Type_{n+1}$ {\em via} usual definition of sheaves,
-as seen in~\cite{maclanemoerdijk} and sheaves on $\Type_n$. Of course,
-the definition must be cumulative.
-
-We will call $\mathcal S_i$ the universe of sheaves on $\Type_i$.
+We will define sheaves by induction on the homotopical truncation level of types
+~: the base case will be any left exact modality on $\HProp$, and we
+give in this section some definitions useful for the inductive case.
 

### 2015-01-13 paper/sheaf_lics.tex
add trunc

@@
 \DeclareMathOperator{\HProp}{HProp}
+\DeclareMathOperator{\IsHProp}{IsHProp}
 \DeclareMathOperator{\nat}{nat}
@@
 \DeclareMathOperator{\Contr}{Contr}
+\DeclareMathOperator{\IsContr}{IsContr}
+\newcommand{\IsType}[1]{\mathrm{Is-}#1\mathrm{-type}}
 \newcommand{\modal}{\ensuremath{\ocircle}}
@@
 
-We refer to~\cite{hottbook}.
-Truncation levels + Im ?
+Our work is mainly based on the stratification by $\Type_n$~:
+\begin{defi}
+  $\IsType n$ is defined by induction on $n\geqslant -2$~:
+  \begin{itemize}
+  \item $\IsType {(-2)} X$ if $X$ is a contractible type, {\em ie} $X$
+    is pointed by $c:X$, and every other point in $X$ is connected to $c$.
+  \item $\IsType {(n+1)} X = \prod_{x,y:X} \IsType n (x=y).$
+  \end{itemize}
+  Then, $\Type_n = \sum_{X:\Type} \IsType n X$.
+\end{defi}
+For a type $X$, to be in $\Type_n$ means that path spaces of $X$ are
+trivial after $n+1$ iteration.
+We define some shortcuts
+\begin{itemize}
+\item $\IsContr := \IsType {(-2)}$ and $\Contr := \Type_{-2}$
+\item $\IsHProp := \IsType {(-1)}$ and $\HProp := \Type_{-1}$
+\end{itemize}
 
-If $x$ is an object of a product or a sigma type, we denote the $i$-th
-projection of $x$ by either $\pi_i x$ or $x_i$.
+Using higher inductive types, one can define for any type $X$ $\|X\| : \HProp$
+generated by 
+\begin{itemize}
+\item a function $|\cdot|_X : X \to \|X\|$,
+\item paths $x=y$ for all $x,y:\|X\|$.
+\end{itemize}
+$\|X\|$ is called the {\em (propositional) truncation of $X$}. The recursion principle asserts~:
+\begin{lem}
+  For any $A:\Type$ and $B:\HProp$, if $f:A \to B$ then there is an
+  induced $g:\|A\|\to B$ such that $g(|a|)= f(a)$ for any $a:A$.
+\end{lem}
+We refer to~\cite{hottbook} for more details.
 
-\begin{itemize}
-\item
-
-We use the subobjects classification seen in~\cite{sets_in_hott}~: for
+Working in $\Type_n$ allows to use the subobjects classification as in~\cite{sets_in_hott}~: for
 all type $B$, $\Type_n$ classifies subobjects of $B$ with $n$-truncated
 fibers ; we have an equivalence
-$$\sum_{A:\Type} \sum_{f:A \to B} \prod_{b\in B}
-\mathrm{IsTrunc}~n~\mathrm{fib}_f(b) \simeq (B \to \Type_n)$$
+$$\left(\sum_{A:\Type} \sum_{f:A \to B} \prod_{b\in B}
+\IsType n~\mathrm{fib}_f(b)\right) \simeq (B \to \Type_n)$$
 and a pullback (with $\Type_n^\bullet$ the universe of pointed
@@
 \xymatrix{
-  A \ar[r]^{t_f} \ar[d]_f & \Type_n^\bullet \ar[d]^{\mathrm{pr}_1}\\
+  A \ar[r]^{t_f} \ar[d]_f & \Type_n^\bullet \ar[d]^{\pi_1}\\
   B \ar[r]_{\chi_f} & \Type_n
@@
 $$t_f = \lambda a,~(\mathrm{fib}_f(f(a)),(a,\mathrm{idpath})).$$
-\end{itemize}
 
@@
 7.7 remains true.
+In particular, if $X:\Type_p$ verifies $P(X)$, then for any $x,y:X$, $P(x=y)$.
 
@@
 
-Another useful properties of modalities is the following~:
+As we want, a left exact modality preserves $\HProp$~:
 \begin{prop}
@@
 When one works in the subuniverse defined by this modality, defining
-an inhabitant of a type $T$ is actually an inhabitant of $P \to
+an inhabitant of a type $T$ is actually giving an inhabitant of $P \to
 T$. Thus, being in this subuniverse is just adding the axiom $P$ to
-the theory. Sadly, adding an axiom using the modality does not add any
-computational content to $P$, thus it is not completely satisfactory.
+the theory. Sadly, adding an axiom this way does not add any
+computational content to $P$, thus is not completely satisfactory.
 
@@
 which is very restrictive. 
+We will use this modality only in the $\HProp$ layer, and try to
+extend it to every $\Type_n$ without collapsing every type.
 
@@
 colimit of $C(f)$ is $Y$. Viewing homotopy type theory as a
-type-theoristic version of higher topos theory, it seems that this
-axiom can be proved~\cite{lurie}, although we admit it in this paper.
+type-theoristic version of higher topos theory, \cite{lurie} suggests that this
+axiom can be proved, although we admit it in this paper.
 
@@
     &\sum_{t:T^k} (\mu_T t_1 = \mu_T t_2 \land \cdots
-    \land \mu_T t_{k-1} \\
-  = &\mu_T t_k) = \sum_{t:T^k} \modal (t_1 = t_2 \land \cdots
+      \land \mu_T t_{k-1} \mu_T t_k)\\
+    &= \sum_{t:T^k} \modal (t_1 = t_2 \land \cdots
       \land t_{k-1} = t_k)

### 2015-01-13 paper/sheaf_lics.tex
explicitize proof of (o(a=b)=(mu a = mu b))

@@
 \newtheorem{lem}[thm]{Lemma}
+\newtheorem{ax}[thm]{Axiom}
 
@@
   \begin{itemize}
-  \item $\IsType {(-2)} X$ if $X$ is a contractible type, {\em ie} $X$
+  \item $\IsType {(-2)} X$ if $X$ is a contractible type, \ie $X$
     is pointed by $c:X$, and every other point in $X$ is connected to $c$.
@@
 \label{sec:gener-pb}
+\kq{Do we keep this name ?}
 
@@
 
-The Giraud axiom asserts that, if $f:X\to Y$ is a surjection, then the
-colimit of $C(f)$ is $Y$. Viewing homotopy type theory as a
+\begin{ax}[Giraud]
+  If $f:X\to Y$ is a surjection, then the
+colimit of $C(f)$ is $Y$.
+\end{ax}
+
+Viewing homotopy type theory as a
 type-theoristic version of higher topos theory, \cite{lurie} suggests that this
 axiom can be proved, although we admit it in this paper.
+
+Formalization of colimits is based on the formalization of limits in~\cite{lumsdaine}.
 
@@
   classified by $\chi$ is {\em dense} in $E$ if its $\modal$-closure
-  is $E$ seen as a subobject of $E$, {\em ie}
+  is $E$ seen as a subobject of $E$, \ie
   $$\forall e:E,~ \left(\sum_{e':E} e=e'\right) \simeq (\modal (\chi~e)).$$ 
@@
 
-We mimic the construction in~\cite{maclanemoerdijk}
+We mimic the construction in~\cite{maclanemoerdijk}. 
+From any left exact modality on $\HProp$, we extend the new principles
+it gives to every $\Type_n,\,n\geqslant 0$.
+Actually, if $n_0$ is a fixed truncation index, and $\modal$ a left
+exact modality on $\Type_{n_0}$, then we can in the same
+way extend the properties of the modality to any $\Type_n$, for
+$n > n_0$.
 
+Once the first modality is defined, the extension to all $\Type_n$ is
+automatic. It means that the new principles we want to add in the new
+theory must be introduced ine the first modality.
+
+The sheafification process works exactly that way
+in~\cite{maclanemoerdijk}~: from a left exact modality on the internal
+logic (a Lawvere-Tierney topology), we define a new left exact
+modality on the whole topos. We view that as the induction step
+between $\HProp$ and $\Type_0$, and we want to extend it.
+
+\kq{Bof ; maybe rewrite this}
+
+As in most of sheafification process, it will be done in
+two steps~:
+\begin{enumerate}[(i)]
+\item {\em separation~:} In this step, we {\em remove} the lack of
+  sheafness from the type we are considering.
+\item {\em completion~:} In this step, we add what lacks now to the
+  separated type to be a sheaf.
+\end{enumerate}
 \subsection{For h-propositions}
@@
 One can easily show that it defines a left exact modality.
+
+The $\lnot\lnot$-modality allows to work with a propositional law of
+excluded middle.
 
@@
 \begin{proof}
-  As $\mu_T$ is an embedding, we know by Giraud axiom that $\square T$ is the colimit
+  As $\mu_T$ is an surjection, we know by Giraud axiom that $\square T$ is the colimit
   of $C(\mu_T)$. If we can show that $C(\mu_T) = C(\id)$, the the
@@
 suffices to show that for all $a,b:T$, $\modal (a=b) = (\mu_T a =
-\mu_T b)$, which is true by univalence.
+\mu_T b)$. By univalence, we want arrows in both ways, forming an
+equivalence.
+\begin{itemize}
+\item Suppose $p : (\mu_T a = \mu_T b)$. Then projecting $p$ along
+  first components yields $q : \prod_{t:T} \modal(a=t) = \modal (b=t)
+  $.
+  Taking for example $t=b$, we deduce $\modal (a=b) = \modal(a=a)$,
+  and the latter is inhabited by $\eta_{a=a} 1$.
+\item Suppose now $p : \modal(a=b)$. Let $\iota$ be the first
+  projection from $\square T \to (T \to \Type_n^\modal)$. $\iota$ is
+  an embedding, thus it suffices to prove $\iota (\mu_T a) = \iota
+  (\mu_T b)$, \ie $\prod_{t:T}\modal (a=t) = \modal (b=t)$. The latter
+  remains true by univalence.
+\end{itemize}
+The fact that these two form an equivalence is technical, we refer to
+the formalization for an explicit proof.
 \end{proof}
@@
 \bibitem[HTT]{lurie}
-Jacob Lurie
-\newblock{\em Higher Topos Theory}
+Jacob Lurie.
+\newblock{\em Higher Topos Theory}.
 
-\bibitem[lumsdaine]{lumsdaine}
-Lumsdaine
-\newblock{Diagrams and limits}
+\bibitem[AKL]{lumsdaine}
+Peter LeFanu Lumsdaine, Jeremy Avigad and Chris Kapulkin.
+\newblock{\em Homotopy limits in type theory}.
 

### 2015-01-13 paper/sheaf_lics.tex

@@
 
+We have to be careful~: if we want to propagate the properties of the
+first modality $\modal_{-1}$, all the higher modalities $\modal_n$ must be compatible
+with $\modal_{-1}$, \ie we want the property
+\begin{prop}
+  If $P:\HProp$, and $\widehat P$ is $P$ seen as a $\Type_n$, then $\modal_n \widehat P = \modal_{-1} P$, and the
+  following coherence diagram commute 
+  $$\xymatrix{
+    \modal_{-1} P \ar@{=}[r] & \modal_n \widehat P \\
+    P \ar@{=}[r] \ar[u]^{\eta_{-1}} & \widehat P \ar[u]_{\eta_n}
+  }$$
+\end{prop}
+
 The sheafification process works exactly that way
@@
 lemma is central in the proof~:
-\begin{lem}
+\begin{lem}\label{lem:sepiscolim}
   Let $T:\Type_{n+1}$. Then $\square T$ is the colimit of the closed
@@
 
-Now that $\square T$ is a colimit, we can easily define an inverse to
-$\lambda\,(f:\square T \to U),~x \circ \mu_T$ for any separated type
-$U$, using the universal property of the colimit.
+Now, let $Q$ be a separated $\Type_{n+1}$, and $f:T \to Q$. Then the
+following diagram commutes
+$$\xymatrix{ 
+    \cdots \overline{\Delta_3} \tar[r] \ar[rd]_{f\circ \pi_1} & \overline{\Delta_2}
+    \dar[r] \ar[d]^{f\circ \pi_1} &
+    \overline{\Delta_1} = T \ar[ld]^{f\circ \pi_1}\\
+    & Q &
+  }$$
+But we know (lemma~\ref{lem:sepiscolim}) that $\square T$ is the
+colimit of the closed diagonals diagram, thus there is an universal
+arrow $\square T \to Q$.
+
+This proves the following proposition
+\begin{prop}
+  $(\square,\mu)$ defines a reflective subuniverse on $\Type_{n+1}$.
+\end{prop}
 
@@
 (Bx)$, equal when precomposed with $\pi_1$.
+$$\xymatrix{
+  \sum_{e:E} (\chi\, e) \dar[r]^{f\circ\pi_1}_{g\circ \pi_1} \ar[d]_{\mathrm{dense}}& \sum_{x :A} (Bx) \\
+  E \dar[ru]^f_g&
+}$$
+We can restrict the previous diagram to 
+$$\xymatrix{
+  \sum_{e:E} (\chi\, e) \dar[rr]^{\pi_1\circ f\circ\pi_1}_{\pi\circ g\circ \pi_1} \ar[d]_{\mathrm{dense}}&& A \\
+  E \dar[rru]^{\pi_1\circ f}_{\pi_1\circ g}&
+},$$ 
+and as $A$ is separated, $\pi_1\circ f = \pi_2 \circ g$.
+For the second components, let $x:E$. Notice that for all $x:E$,
+$\sum_{y:E} x = y$ has a dense subobject, $\sum_{y:\sum_{e:E} (\chi\,
+  e)} x=y_1$~:
 
-As $A$ is separated, for any $x:E$, the first components of $f x$ and
-$g x$ are equal. For the second components, notice that for all $x:E$,
-$\sum_{y:E} x = y$ has a dense subobject, $\sum_{y:\sum_{e:E} (\chi\,
-  e)} x=y_1$. 
-Using the separation property with these type, one can show the second
-components of $f x$ and $g x$ are equals too, with respect to the
-first path.
+$$\xymatrix{
+  \sum_{y:\sum_{e:E} (\chi\,
+  e)} x=y_1 \dar[rrr]^{\pi_2\circ f\circ\pi_1\circ \pi_1}_{\pi_2\circ g\circ \pi_1\circ \pi_1} \ar[d]_{\mathrm{dense}}&&& B\,x \\
+  \sum_{y:E} x = y \dar[rrru]^{\pi_2\circ f\circ \pi_1}_{\pi_2\circ g\circ \pi_1}&
+}$$
+Using the sepatation property of $B\,x$, one can show that second
+components, transported correctly along the first components equality,
+are equal. The complete proof can be found in the formalisation.
+This proves the following proposition
+\begin{prop}
+  $(\square,\mu)$ defines a modality on $\Type_{n+1}$.
+\end{prop}
+
+As this modality is just a step in the construction, we will not show
+that it is left exact, we will do it only for the sheafification modality.
 

### 2015-01-13 paper/sheaf_lics.tex
good sheafification

@@
 with $\modal_{-1}$, \ie we want the property
-\begin{prop}
+\begin{prop}\label{prop:hprop}
   If $P:\HProp$, and $\widehat P$ is $P$ seen as a $\Type_n$, then $\modal_n \widehat P = \modal_{-1} P$, and the
@@
 and a left exact modality $\modal$ on $\Type_n$, compatible with the
-modality on $\HProp$.
+modality on $\HProp$ in the sense of proposition~\ref{prop:hprop}.
 
@@
 \end{prop}
+It will imply that $\star$ is compatible with $\HProp$ in the sense of proposition~\ref{prop:hprop}.
 \section{Future Works}

### 2015-01-13 paper/sheaf_lics.tex
minor changes on xy definitions

--- 2015-01-13 paper/sheaf_lics.tex
@@
 \usepackage[all]{xy}
-\def\dar[#1]{\ar@<2pt>[#1]\ar@<-2pt>[#1]} %double arrows in xy
-\def\tar[#1]{\ar@<2pt>[#1]\ar@<0pt>[#1]\ar@<-2pt>[#1]} %triple arrows in xy
+\def\dar[#1]#2{\ar@<-#2>[#1]\ar@<#2>[#1]} %double arrows in xy
+\def\tar[#1]#2{\ar@<#2>[#1]\ar@<0pt>[#1]\ar@<-#2>[#1]} %triple arrows in xy
 \DeclareMathOperator{\Type}{Type}
@@
 \DeclareMathOperator{\IsContr}{IsContr}
-\newcommand{\IsType}[1]{\mathrm{Is-}#1\mathrm{-type}}
+
+\def\mymathhyphen{{\hbox{-}}}
+
+\newcommand{\IsType}[1]{\mathrm{Is\mymathhyphen}#1\mathrm{\mymathhyphen
+    type}}
 \newcommand{\modal}{\ensuremath{\ocircle}}
@@
   $$\xymatrix{
-    C(f) := \cdots~ X \times_Y X \times_Y X \tar[r] & X \times_Y X \dar[r] & X
+    C(f) := \cdots~ X \times_Y X \times_Y X \tar[r]{4pt} & X \times_Y X \dar[r]{2pt} & X
   }$$
@@
   diagonal diagram
-  $$\xymatrix{ 
-    \cdots \overline{\Delta_3} \tar[r] & \overline{\Delta_2} \dar[r] &
+  $$\xymatrix @C=4em  { 
+    \cdots \overline{\Delta_3} \tar[r]{4pt} & \overline{\Delta_2} \dar[r]{2pt} &
     \overline{\Delta_1}
@@
 following diagram commutes
-$$\xymatrix{ 
-    \cdots \overline{\Delta_3} \tar[r] \ar[rd]_{f\circ \pi_1} & \overline{\Delta_2}
-    \dar[r] \ar[d]^{f\circ \pi_1} &
+$$\xymatrix @C=4em{ 
+    \cdots \overline{\Delta_3} \tar[r]{4pt} \ar[rd]_{f\circ \pi_1} & \overline{\Delta_2}
+    \dar[r]{2pt} \ar[d]^{f\circ \pi_1} &
     \overline{\Delta_1} = T \ar[ld]^{f\circ \pi_1}\\
@@
 (Bx)$, equal when precomposed with $\pi_1$.
-$$\xymatrix{
-  \sum_{e:E} (\chi\, e) \dar[r]^{f\circ\pi_1}_{g\circ \pi_1} \ar[d]_{\mathrm{dense}}& \sum_{x :A} (Bx) \\
-  E \dar[ru]^f_g&
+$$\xymatrix @R=4em @C=4em{
+  \sum_{e:E} (\chi\, e) \ar@<-2pt>[r]_{g\circ\pi_1} \ar@<2pt>[r]^{f\circ \pi_1} \ar[d]_{\mathrm{dense}}& \sum_{x :A} (Bx) \\
+  E \ar@<-2pt>[ru]_{g} \ar@<2pt>[ru]^{f}&
 }$$
 We can restrict the previous diagram to 
-$$\xymatrix{
-  \sum_{e:E} (\chi\, e) \dar[rr]^{\pi_1\circ f\circ\pi_1}_{\pi\circ g\circ \pi_1} \ar[d]_{\mathrm{dense}}&& A \\
-  E \dar[rru]^{\pi_1\circ f}_{\pi_1\circ g}&
-},$$ 
+$$\xymatrix @R=4em @C=5em{
+  \sum_{e:E} (\chi\, e) \ar@<-2pt>[r]_{\pi_1\circ g\circ\pi_1} \ar@<2pt>[r]^{\pi_1\circ f\circ \pi_1} \ar[d]_{\mathrm{dense}}& \sum_{x :A} (Bx) \\
+  E \ar@<-2pt>[ru]_{\pi_1\circ g} \ar@<2pt>[ru]^{\pi_1\circ f}&
+}$$
 and as $A$ is separated, $\pi_1\circ f = \pi_2 \circ g$.
@@
   \sum_{y:\sum_{e:E} (\chi\,
-  e)} x=y_1 \dar[rrr]^{\pi_2\circ f\circ\pi_1\circ \pi_1}_{\pi_2\circ g\circ \pi_1\circ \pi_1} \ar[d]_{\mathrm{dense}}&&& B\,x \\
-  \sum_{y:E} x = y \dar[rrru]^{\pi_2\circ f\circ \pi_1}_{\pi_2\circ g\circ \pi_1}&
+  e)} x=y_1 \dar[rrr]{2pt}^{\pi_2\circ f\circ\pi_1\circ \pi_1}_{\pi_2\circ g\circ \pi_1\circ \pi_1} \ar[d]_{\mathrm{dense}}&&& B\,x \\
+  \sum_{y:E} x = y \dar[rrru]{2pt}^{\pi_2\circ f\circ \pi_1}_{\pi_2\circ g\circ \pi_1}&
 }$$

### 2015-01-13 paper/sheaf_lics.tex
add .bib
beginning of the introduction

@@
   principles inspired by category theory and homotopy theory, and
-  which corresponds closely to higher toposes.
+  which corresponds closely to higher topoi.
   % 
@@
 
+The sheaf construction, or sheafification~\cite{maclanemoerdijk}, is a
+very powerful tool to construct new topoi from already existing ones,
+allowing logical principles---that can not be proven to be true or false
+in the old topos---to be valided (or invalided) in the new topos.
+%
+The most common use of sheafification is to construct a boolean
+topos, thus validating the classical principle of the law of excluded
+middle (EM), from an intuitionistic topos. 
+%
+An other famous (revisited) application has been developed by Paul
+Cohen~\cite{cohen1966} to prove that the continuum hypothesis is
+independent of the usual axioms of Zermelo-Fr\ae nkel set theory, even
+in presence of the axiom of choice (AC).
+
+Even if similar questions have been considered around the Curry-Howard
+isomorphism, there is no canonical way to extend a programming language
+such as type theory while keeping consistency automatically. 
+%
+For instance, much effort have been done to provide a computational
+content to the law of excluded middle in order to define a constructive
+version of classical logic. This has lead to various calculi, with
+most notably the $\lambda \mu$-calculus of
+Parigot~\cite{parigot1993classical}, but this has not allow to define
+a new version of type theory with classical principles.
+%
+Other works have tried to extend continuation-passing-style (CPS)
+transformation to type theory, but they have been faced with the
+difficulty that the CPS transformation is incompatible with (full) dependent
+sums~\cite{barthe2002cps}, which puts emphasis on the tedious link
+between the axiom of choice and the law of excluded middle in type theory.
+
+Nevertheless the axiom of choice Coquand, Herbelin, Krivine.
+%
+% Some works have tried to provide a realization of the axiom of choice 
+% compatible with classical logic
+
 \begin{itemize}
@@
   Giraud's axiom
-\item
-  previous attempts to extend Gödel's double negation to 
-  Martin-Löf type theory
-\item 
-  compare to the work of Parigot on $\lambda\mu$-calculus (Classical proofs as programs), and Coquand 
-  (A realization of the negative interpretation of the Axiom of
-  Choice)
-\item
-  work of Hugo Herbelin, A Constructive Proof of Dependent Choice,
-  Compatible with Classical Logic
-\item
-  Barthe (CPS Translating Inductive and Coinductive Types)
-  ``No translation is possible along the same lines for small $\Sigma$-types
-  and sum types with dependent case.''
-\item
-
 \end{itemize}
@@
     \sum_{e:E} (\chi~ e) \ar[r]^f \ar[d]_{\pi_1} & F \\
-    E \ar@{-->}[ru]&
+    E \ar@{-->}[ru]_{!}&
   }$$
@@
     \sum_{e:E} (\chi~ e) \ar[r]^f \ar[d]_{\pi_1} & F \\
-    E \ar@{-->}[ru]&
+    E \ar@{-->}[ru]_{\exists !}&
   }$$
@@
 % http://www.michaelshell.org/tex/ieeetran/bibtex/
-%\bibliographystyle{IEEEtran}
+\bibliographystyle{IEEEtran}
+\bibliography{sheaf_lics}
 % argument is your BibTeX string definitions and bibliography database(s)
@@
 % (used to reserve space for the reference number labels box)
-\begin{thebibliography}{4}
+% \begin{thebibliography}{5}
 
-\bibitem[RS13]{sets_in_hott}
-Egbert Rijke and Bas Spitters.
-\newblock Sets in homotopy type theory.
-\newblock 2013.
+% \bibitem[RS13]{sets_in_hott}
+% Egbert Rijke and Bas Spitters.
+% \newblock Sets in homotopy type theory.
+% \newblock 2013.
 
-\bibitem[{Uni}13]{hottbook}
-{Univalent Foundations Project}.
-\newblock {\em Homotopy Type Theory: Univalent Foundations for Mathematics}.
-\newblock 2013.
+% \bibitem[{Uni}13]{hottbook}
+% {Univalent Foundations Project}.
+% \newblock {\em Homotopy Type Theory: Univalent Foundations for Mathematics}.
+% \newblock 2013.
 
-\bibitem[MM92]{maclanemoerdijk}
-Saunders MacLane and Ieke Moerdijk.
-\newblock {\em Sheaves in Geometry and Logic}.
-\newblock Springer-Verlag, 1992.
+% \bibitem[MM92]{maclanemoerdijk}
+% Saunders MacLane and Ieke Moerdijk.
+% \newblock {\em Sheaves in Geometry and Logic}.
+% \newblock Springer-Verlag, 1992.
 
-\bibitem[HTT]{lurie}
-Jacob Lurie.
-\newblock{\em Higher Topos Theory}.
+% \bibitem[HTT]{lurie}
+% Jacob Lurie.
+% \newblock{\em Higher Topos Theory}.
 
-\bibitem[AKL]{lumsdaine}
-Peter LeFanu Lumsdaine, Jeremy Avigad and Chris Kapulkin.
-\newblock{\em Homotopy limits in type theory}.
+% \bibitem[AKL]{lumsdaine}
+% Peter LeFanu Lumsdaine, Jeremy Avigad and Chris Kapulkin.
+% \newblock{\em Homotopy limits in type theory}.
 
-\end{thebibliography}
+% \end{thebibliography}
 

### 2015-01-13 paper/sheaf_lics.tex
sheafification is a refl subu

--- 2015-01-13 paper/sheaf_lics.tex
@@
 As any separated type $T$ embeds in $\left(\Type_n^\modal\right)^T$,
-it suffices to take the closure of $T$ to get a sheaf.
+it suffices to take the closure of $T$ to get a sheaf, and the unit
+$\nu_T:T \to \star T$ is obvious.
+\begin{prop}
+  $(\star,\nu)$ defines a reflective subuniverse.
+\end{prop}
+\begin{proof}
+  Let $T,Q:\Type_{n+1}$ such that $Q$ is a sheaf. Let $f:T\to Q$.
+  Because $Q$ is a sheaf, it is in particular separated ; thus we can
+  extend $f$ to $\square f:\square T\to Q$.
+
+  But as $\star T$ is the closure of $\square T$, $\square T$ is dense
+  into $\star T$, so the sheaf property of $Q$ allows to extend
+  $\square f$ to $\star f:\star T \to Q$.
+
+  As all these steps are universal, the composition is.
+\end{proof}
 

### 2015-01-14 paper/sheaf_lics.tex
work on introduction

@@
 
-The sheaf construction, or sheafification~\cite{maclanemoerdijk}, is a
-very powerful tool to construct new topoi from already existing ones,
-allowing logical principles---that can not be proven to be true or false
-in the old topos---to be valided (or invalided) in the new topos.
+Sheafification~\cite{maclanemoerdijk} is a very powerful geometric
+construction that has been initially stated in topology and has
+quickly been lifted to mathematical logic.
 %
-The most common use of sheafification is to construct a boolean
+In the field of topos theory, it provides a way to construct new topoi
+from already existing ones, allowing logical principles---that can not
+be proven to be true or false in the old topos---to be valid (or
+invalid) in the new topos.
+%
+The most common use of sheafification is the construction a boolean
 topos, thus validating the classical principle of the law of excluded
-middle (EM), from an intuitionistic topos. 
+middle (EM), from an intuitionistic topos using the dense topology. 
 %
-An other famous (revisited) application has been developed by Paul
+An other famous application has been developed by
 Cohen~\cite{cohen1966} to prove that the continuum hypothesis is
 independent of the usual axioms of Zermelo-Fr\ae nkel set theory, even
-in presence of the axiom of choice (AC).
+in presence of the axiom of choice (AC). The initial work of Cohen
+uses forcing but can be rephrased in terms of
+sheafification~\cite{maclanemoerdijk}.
 
 Even if similar questions have been considered around the Curry-Howard
-isomorphism, there is no canonical way to extend a programming language
-such as type theory while keeping consistency automatically. 
+isomorphism, there is no canonical way to extend a programming
+language such as type theory with new logical or computational
+principle while keeping consistency automatically.
 %
 For instance, much effort have been done to provide a computational
-content to the law of excluded middle in order to define a constructive
-version of classical logic. This has lead to various calculi, with
-most notably the $\lambda \mu$-calculus of
-Parigot~\cite{parigot1993classical}, but this has not allow to define
-a new version of type theory with classical principles.
+content to the law of excluded middle in order to define a
+constructive version of classical logic. This has lead to various
+calculi, with most notably the $\lambda \mu$-calculus of
+Parigot~\cite{parigot1993classical}, but this line of work has not
+appeared to be fruitful to define a new version of type theory with
+classical principles.
 %
@@
 
-Nevertheless the axiom of choice Coquand, Herbelin, Krivine.
+Nevertheless the axiom of choice has shown to be realizable by
+computational meaning in a classical setting by techniques turning
+around the notion of (modified) bar induction
+\cite{berardi1998computational}, Krivine's
+realizability~\cite{krivine2003dependent} and even more recently with
+restriction on elimination of dependent sums and lazy
+evaluation~\cite{herbelin2012constructive}.
 %
-% Some works have tried to provide a realization of the axiom of choice 
-% compatible with classical logic
+
+Unfortunately, all those works do not share the generality and
+simplicity of sheafification in topos theory.  
+%
+However, type theory is known to be quite closed to topos theory so
+one could wonder why similar techniques have not been developed in the
+field of type theory. 
+%
+The answer to this question has been given recently by the advent of
+homotopy type theory~\cite{hottbook}, which is an extension of
+Martin-Löf type theory with principles inspired by (higher) category
+theory and homotopy theory, such as higher inductive
+types~\cite{lumsdaine2011higher,lumsdaine2013higher} and
+Voevodsky's univalence principle~\cite{kapulkin2012simplicial}.
+
+This new point of view on type theory has revealed the
+homotopy structure of types where for instance (homotopical)
+propositions are just types with an irrelevant equality and
+(homotopical) sets are types with a propositional equality. 
+%
+When restricted to such propositions and sets, type theory corresponds
+quite closely to topos theory but the mismatch starts when considering
+higher homotopical types. 
+%
+Fortunately, an higher version of topos theory has been developed
+recently by mathematicians, synthesized in the monograph of Lurie on
+higher topos theory~\cite{lurie}. 
+%
+Lurie has developed their all the tools that have been defined in
+topos theory, but in an higher setting. In particular, the theory of
+sheaves has been lifted to higher topos theory.
+%
+As the notion of higher topoi appears to correspond very closely
+to homotopy type theory, this provides a new hope that tackling the problem of
+extending the power of type theory using sheafification is actually possible.
+
+Nevertheless, the adaptation of the sheafification in higher topos
+theory to homotopy type theory is not straightforward because the
+construction in higher topoi is restricted to the initial
+Grothendieck which is still very topological, and not very amenable to
+formalization in type theory. It seems more promising to use a
+generalized sheafification, called Lawvere-Tierney
+sheafification~\cite{maclanemoerdijk}, but this construction has not
+been considered yet in the setting of higher topos theory.
+
+\subsection{Lawvere-Tierney Topology}
+
+\subsection{Overview of the Result}
+
+The main contribution of this article is presents a computer check
 
 \begin{itemize}
-\item 
-  links between HoTT and higher topos theory 
 \item 

### 2015-01-14 paper/sheaf_lics.tex
first draft of introduction

@@
 \usepackage{xspace}
+\usepackage[colorlinks=true]{hyperref}
 
@@
 \newcommand{\modal}{\ensuremath{\ocircle}}
+\newcommand \True {\mathrm{true}}
+\newcommand \False {\mathrm{false}}
+\newcommand \closure[1] {\overline{#1}}
+\newcommand \Char[1] {\chi_{#1}}%{\mathrm{char}(#1)}
+\newcommand \E {\mathcal{E}}
+\newcommand \Hom[1] {\mathrm{Hom}_{#1}}
+\newcommand \Sh[1] {\mathrm{Sh}_{#1}}
 
@@
 
-  Sheafification is a popular tool in (higher) topos theory which
+  Sheafification is a popular tool in topos theory which
   allows to extend the internal logic of a topos with new
   principles. One of its most famous applications is the possibility
-  to transform a topos into a boolean (classical) topos using the
+  to transform a topos into a boolean topos using the
   dense topology, which corresponds in essence to Gödel's double
@@
 
-Unfortunately, all those works do not share the generality and
-simplicity of sheafification in topos theory.  
+Unfortunately, all those works---although they provide very nice and
+sophisticated theoretical results---do not share the generality and
+simplicity of sheafification in topos theory and are not amenable to a
+direct integration into proof assistant based on type theory such as
+Coq or Agda.  
 %
@@
 Voevodsky's univalence principle~\cite{kapulkin2012simplicial}. 
+%
 
 This new point of view on type theory has revealed the
-homotopy structure of types where for instance (homotopical)
-propositions are just types with an irrelevant equality and
-(homotopical) sets are types with a propositional equality. 
+homotopy structure of types where for instance homotopical
+propositions (or $\Type_{-1}$) are just types with an irrelevant equality and
+homotopical sets (or $\Type_{0}$)  are types with a propositional
+equality and so on for $\Type_{n}$.
 %
-When restricted to such propositions and sets, type theory corresponds
-quite closely to topos theory but the mismatch starts when considering
-higher homotopical types. 
+The development of univalence has also shed some light on the
+difficulty to make the AC and EM coexist in type theory. Indeed, it
+has been shown that a naive (non-propositional) version of EM is
+inconsistent with univalence.
+
+When restricted to homotopical propositions and sets, homotopy type
+theory corresponds quite closely to topos theory but the mismatch
+starts when considering higher homotopical types.
 %
-Fortunately, an higher version of topos theory has been developed
+Fortunately, a higher version of topos theory has been developed
 recently by mathematicians, synthesized in the monograph of Lurie on
@@
 %
-Lurie has developed their all the tools that have been defined in
-topos theory, but in an higher setting. In particular, the theory of
+Lurie has developed all the tools that have been defined in
+topos theory, but in a higher setting. In particular, the theory of
 sheaves has been lifted to higher topos theory.
@@
 
+Lawvere-Tierney sheafification in a topos $\E$ is based on an abstract
+point of view on the topology to be considered, being simply an
+endomorphism on the classifying object $\Omega$ of $\E$  
+%
+$$
+j : \Omega \rightarrow \Omega
+$$
+%
+that is required to preserve $\True$ ($j \ \True = \True$), to be
+idempotent ($j \circ j = j$) and compatible with products ($j \circ
+\wedge = \wedge \circ (j, j)$).
+%
+A typical example is given by double negation.
+
+Every Lawvere-Tierney topology $j$ induces a closure operator
+$A \mapsto \closure{A}$ on subojects. If we see a subobject $A$ of $E$
+as a characteristic function $\Char{A}$, the closure $\closure{A}$
+corresponds to the subobject of $E$ whose characteristic function is 
+%
+$$
+\Char{\closure{A}} = j \circ \Char{A}.
+$$
+%
+A suboject $A$ of $E$ is said to
+be dense when $\closure{A} = E$.
+
+The idea is then to define sheaves in $\E$ as objects of $\E$ for
+which it is impossible to make a distinction between objects and their
+dense subobjects. This idea is formalized by saying that for every
+dense subobject $A$ of $E$, the following canonical map is an
+isomorphism
+%
+\begin{equation}\label{equ:sheaf_def}
+\Hom{\E}(E,F) \rightarrow \Hom{\E}(A,F).
+\end{equation}
+%
+On can show that $\Sh{\E}$, the full sub-category of $\E$ given by
+sheaves, is again a topos, with classifying object
+%
+$$
+\Omega_j = \{ P \in \E \ | \ j P  \leftrightarrow P \}.
+$$
+%
+Thus, in case of the double negation, the resulting topos is boolean  
+and admits classical reasoning.
+
+But furthermore, one can defined a left adjoint, the sheafification
+functor, 
+%
+$$
+a_j : \E \rightarrow \Sh{\E}
+$$
+to the inclusion functor, which exhibits $\Sh{\E}$ as a reflective
+subcatefory of $\E$ (which is a particular case of localization). This
+means that logical principles valid in $\E$ are still valid in
+$\Sh{\E}$.
+
 \subsection{Overview of the Result}
 
-The main contribution of this article is presents a computer check
+The main contribution of this article is the development of a
+computer-checked definition of Lawvere-Tierney sheafification in the
+setting of homotopy type theory.
+%
+This provides a way to give a meaning to the propositional law of
+excluded middle, and more generally, combined with the previous work on
+forcing in type theory~\cite{jaber2012extending}, it allows to reuse
+results derived for topos theory, such as independence results.
 
+To extend Lawvere-Tierney sheafification to homotopy type theory, the
+first thing to realize is that the construction can not be done in one
+step anymore. It must be done by induction on the level of homotopy of
+types. More precisely, the first layer of sheafification is defined
+for $\Type_{0}$ , given a topology on $\Type_{-1}$. This part
+corresponds to sheafification for topoi. Then, assuming that the
+sheafification has been constructed up to level $n$, one can defined
+the sheafification for $\Type_{n+1}$.
+%
+This inductive process in the definition of sheafification is similar
+to construction in the Grothendieck setting, where the
+sheafification is obtained by applying the plus-construction $n+1$ to
+$\Type_{n}$~\cite[Chapter~6]{lurie}.
+
+This inductive step requires to formalize the notion of left-exact
+reflective subuniverse, which corresponds to a stratified
+version of left-exact modalities introduced
+in~\cite[Chapter~7]{hottbook}. A Lawvere-Tierney topology can thus be
+seen as a left-exact modality on $\Type_{-1}$ and sheafification as
+a way to extend it as a left-exact modality on $\Type_{n}$ for any $n$.
+
+Compare to the definition of sheafification in topos theory, the main
+change comes from the fact that in homotopy type theory, there is not
+a single object classifier but rather an $n$-object classifier for
+each $n\geq -2$ which classifies functions with $n$-truncated homotopy
+fibers~\cite{sets_in_hott}. This means that when monomorphisms are used in topos
+theoretic definition, the correct generalization may be $n$-truncated homotopy
+fibers or monomorphims (\ie $-1$-truncated homotopy
+fibers) depending on the context.
+
+For instance, the notion of separated objects (which is a technical
+tool in the construction of the sheafification), for which the morphism
+(\ref{equ:sheaf_def}) is just a monomorphism, is generalized in two
+different ways.  
+%
+The subobjects to be considered are given by $n$-truncated fibers,
+while the morphism (\ref{equ:sheaf_def}) is still required to be a
+monomorphism.
+
+In other technical difficulty is the fact that sheafification in topos
+theory uses the fact that epimorphisms are coequalizers of their
+kernel pairs. This properties does not holds in higher topoi, but it
+can be replaced by Giraud's axiom which basically allows to replace 
+ kernel pairs by \v{C}ech nerves.
+
+Finally, one of the main difficulty in our formalization is that many
+equlity which are trival (b yproof irrelevance) in the topos theoretic
+setting becomes relevant and requires a carreful use of univalence
+with appropriate isomorphisms.
+
+\subsection{Plan of the Paper}
 \begin{itemize}
-\item 
-  modalities in HoTT
-\item 
-  Giraud's axiom
+\item Section~\ref{sec:hott}, we define the foundations on which
+  our worked is based.
+\item Section~\ref{sec:sheaves}, we explain the construction of the
+  sheafification functor.
+\item Section~\ref{sec:formalization} presents the details of the
+  formalization in Coq.
+\item Finally, we present future works.
 \end{itemize}
 
-Plan of the paper~:
-\begin{itemize}
-\item In section~\ref{sec:hott}, we define the foundations on which
-  our worked is based.
-\item In section~\ref{sec:sheaves}, we explain the construction of the
-  sheafification functor.
-\item Finally, we discuss related works and present future works.
-\end{itemize}
-\section{Homotopy Type Theory}
+\section{Preliminaries on Homotopy Type Theory}
 \label{sec:hott}
@@
 \label{ssec:hott}
-
-\nt{
-  Il faut au minimum introduire proprement les n-Types, le squash
-  sur les props 
-}
 
@@
 It will imply that $\star$ is compatible with $\HProp$ in the sense of proposition~\ref{prop:hprop}.
-\section{Future Works}
-\label{sec:future-works}
-
-Prove that the Giraud axiom holds in homotopy type theory.
-
-Extend the sheafification to all types, even if non truncated.
 
@@
 
-\section{Conclusion}
-The conclusion goes here.
+\section{Formalization} 
+\label{sec:formalization}
+
+A Coq formalization is available at
+\url{https://github.com/KevinQuirin/sheafification/}.
+
+\nt{describe how many line for each, be clear about what has not been formalized}
+
+\section{Conclusion and Future Works}
+\label{sec:future-works}
+
+Prove that the Giraud axiom holds in homotopy type theory.
+
+Extend the sheafification to all types, even if non truncated.
 

### 2015-01-14 paper/sheaf_lics.tex
redistribute section II

@@
 
+ \IEEEpubid{}
+
 \begin{document}
@@
 $$
-\Omega_j = \{ P \in \E \ | \ j P  \leftrightarrow P \}.
+\Omega_j = \{ P \in \E \ | \ j P  = P \}.
 $$
@@
 \subsection{Plan of the Paper}
-\begin{itemize}
-\item Section~\ref{sec:hott}, we define the foundations on which
-  our worked is based.
-\item Section~\ref{sec:sheaves}, we explain the construction of the
+
+Section~\ref{sec:hott} defines the foundations of homotopy type theory
+and the definition of modalities, on which this paper is based.
+%
+Section~\ref{sec:sheaves} defines the notion of sheaves in homotopy
+type theory.
+%
+Section~\ref{sec:sheafification} depicts the construction of the
 sheafification functor.
-\item Section~\ref{sec:formalization} presents the details of the
+%
+Section~\ref{sec:formalization} presents the details of the
   formalization in Coq.
-\item Finally, we present future works.
-\end{itemize}
+%
+Finally, Section~\ref{sec:future-works} concludes on future works.
 
@@
 
-\subsection{Homotopy Type Theory}
+In this section, we review some basic definitions in homotopy type
+theory that are central in our formalization but not specific to
+sheafification. 
+%
+Section~\ref{ssec:hott} introduces the notion of homotopy types and
+classifying objects as defined in~\cite{sets_in_hott}.
+%
+Section~\ref{sec:colim-homot-type} introduces the notion of colimits
+in homotopy type theory. 
+%
+And Section~\ref{sec:giraud-ax} defines (a particular case of)
+Giraud's axiom on higher topoi that we need in the definition of
+sheafification.
+
+The definition of Section~\ref{ssec:hott} are part of the Coq library
+available at \url{https://github.com/HoTT/HoTT}, while other definitions
+and theorems are specific to our formalization.
+
+ \subsection{Homotopy Types and Classifying Objects}
 \label{ssec:hott}
@@
 
-\subsection{Left exact Modalities}
+\subsection{Colimits in Homotopy Type Theory }
+\label{sec:colim-homot-type}
+
+\nt{Kevin -> complete this subsection}
+
+Formalization of colimits is based on the formalization of limits in~\cite{lumsdaine}.
+
+Definition of graphs, diagrams, is\_colimit.
+
+Maybe the HIT of colimit, at least say that using HIT, all colimits
+exists.
+
+\subsection{Giraud's axiom}
+\label{sec:giraud-ax}
+
+\begin{defi}
+  Let $f:X \to Y$ be a map, and $p:\nat$. The $p$-pullback of $f$,
+  noted $X\times_Y \cdots\times_Y X$ is
+  the limit of the diagram
+  $$\xymatrix{
+    X \ar[rd]^f & \cdots & X \ar[ld]_f \\
+      &    Y   &
+  }$$
+  with $p$ copies of $X$. The $0$-pullback of $f$ is $\Unit$.
+\end{defi}
+
+In homotopy type theory, we have 
+$$X\times_Y \cdots\times_Y X = \sum_{x:X^n} (f x_1 = f x_2) \land
+\cdots \land (f x_{n-1} = f x_n).$$
+\begin{defi}
+  If $f:X \to Y$ is a map, then the {\em \v{C}ech nerve} of $f$ is the
+  diagram
+  $$\xymatrix{
+    C(f) := \cdots~ X \times_Y X \times_Y X \tar[r]{4pt} & X \times_Y X \dar[r]{2pt} & X
+  }$$
+where the arrows are the canonical projections.
+\end{defi}
+
+\begin{ax}[Giraud]
+  If $f:X\to Y$ is a surjection, then the
+colimit of $C(f)$ is $Y$.
+\end{ax}
+
+Viewing homotopy type theory as a
+type-theoristic version of higher topos theory, \cite{lurie} suggests that this
+axiom can be proved, although we admit it in this paper.
+
+\section{Left exact Modalities}
 \label{sec:lexmod}
 
-\nt{
-To be checked:
-  Condition iv of Thm 7.7.5 of HoTT
-}
+\cite{shulman-higher-modalities}
+
+\subsection{Definition}
 
@@
 
-A left exact modality leads to a new theory with new principles. A
-basic example is the {\em open modality of $P$} for a $P:\HProp$, $\modal_P T = P \to T$. 
-When one works in the subuniverse defined by this modality, defining
-an inhabitant of a type $T$ is actually giving an inhabitant of $P \to
-T$. Thus, being in this subuniverse is just adding the axiom $P$ to
-the theory. Sadly, adding an axiom this way does not add any
-computational content to $P$, thus is not completely satisfactory.
+\subsection{Examples of Left Exact Modalities}
+
+A basic example is the {\em open modality of $P$} for a $P:\HProp$,
+$\modal_P T = P \to T$.  When one works in the subuniverse defined by
+this modality, defining an inhabitant of a type $T$ is actually giving
+an inhabitant of $P \to T$. Thus, being in this subuniverse is just
+adding the axiom $P$ to the theory. Sadly, adding an axiom this way
+does not add any computational content to $P$, thus is not completely
+satisfactory.
 
@@
 
-\subsection{Generalized pullbacks}
-\label{sec:gener-pb}
-\kq{Do we keep this name ?}
-
-\begin{defi}
-  Let $f:X \to Y$ be a map, and $p:\nat$. The $p$-pullback of $f$,
-  noted $X\times_Y \cdots\times_Y X$ is
-  the limit of the diagram
-  $$\xymatrix{
-    X \ar[rd]^f & \cdots & X \ar[ld]_f \\
-      &    Y   &
-  }$$
-  with $p$ copies of $X$. The $0$-pullback of $f$ is $\Unit$.
-\end{defi}
-
-In homotopy type theory, we have 
-$$X\times_Y \cdots\times_Y X = \sum_{x:X^n} (f x_1 = f x_2) \land
-\cdots \land (f x_{n-1} = f x_n).$$
-\begin{defi}
-  If $f:X \to Y$ is a map, then the {\em \v{C}ech nerve} of $f$ is the
-  diagram
-  $$\xymatrix{
-    C(f) := \cdots~ X \times_Y X \times_Y X \tar[r]{4pt} & X \times_Y X \dar[r]{2pt} & X
-  }$$
-with canonical projections.
-\end{defi}
-
-\begin{ax}[Giraud]
-  If $f:X\to Y$ is a surjection, then the
-colimit of $C(f)$ is $Y$.
-\end{ax}
-
-Viewing homotopy type theory as a
-type-theoristic version of higher topos theory, \cite{lurie} suggests that this
-axiom can be proved, although we admit it in this paper.
-
-Formalization of colimits is based on the formalization of limits in~\cite{lumsdaine}.
+\subsection{The new type theory arising from a left exact modality}
+
+A left exact modality leads to a new theory with new principles.
 

### 2015-01-14 paper/sheaf_lics.tex
explicit section on modalities

@@
 
-\newcommand{\IsType}[1]{\mathrm{Is\mymathhyphen}#1\mathrm{\mymathhyphen
-    type}}
+\newcommand{\IsType}[1]{\mathop{\mathrm{Is\mymathhyphen}#1\mathrm{\mymathhyphen
+      type}} }
 \newcommand{\modal}{\ensuremath{\ocircle}}
@@
 
-Using higher inductive types, one can define for any type $X$ $\|X\| : \HProp$
-generated by 
-\begin{itemize}
-\item a function $|\cdot|_X : X \to \|X\|$,
-\item paths $x=y$ for all $x,y:\|X\|$.
-\end{itemize}
-$\|X\|$ is called the {\em (propositional) truncation of $X$}. The recursion principle asserts~:
-\begin{lem}
-  For any $A:\Type$ and $B:\HProp$, if $f:A \to B$ then there is an
-  induced $g:\|A\|\to B$ such that $g(|a|)= f(a)$ for any $a:A$.
-\end{lem}
-We refer to~\cite{hottbook} for more details.
-
 Working in $\Type_n$ allows to use the subobjects classification as in~\cite{sets_in_hott}~: for
@@
 $$t_f = \lambda a,~(\mathrm{fib}_f(f(a)),(a,\mathrm{idpath})).$$
+
+Using higher inductive types, one can define for any type $X$ $\|X\| : \HProp$
+generated by 
+\begin{itemize}
+\item a function $|\cdot|_X : X \to \|X\|$,
+\item paths $x=y$ for all $x,y:\|X\|$.
+\end{itemize}
+$\|X\|$ is called the {\em (propositional) truncation of $X$}. The recursion principle asserts~:
+\begin{lem}
+  For any $A:\Type$ and $B:\HProp$, if $f:A \to B$ then there is an
+  induced $g:\|A\|\to B$ such that $g(|a|)= f(a)$ for any $a:A$.
+\end{lem}
+We refer to~\cite{hottbook} for more details.
 
@@
 
-A left exact modality leads to a new theory with new principles.
+A left exact modality leads to a new homotopy type theory with
+potentially new principles.
+
+Definition of Sigma-Types.
+
+Preservation of univalence.
+
+Consistency issues.
 
@@
 
+Talk about issues in universe levels. Lake of cumulativity for
+$\Type_n$ and impossibility to define by induction the sheafification 
+because universe levels must be the same for every $n$.
+
 \section{Conclusion and Future Works}

### 2015-01-14 paper/sheaf_lics.tex
paper
resolve (wird) conflicts
add: graph, diagram, is_colimit, colimit

@@
 \DeclareMathOperator{\IsContr}{IsContr}
+\DeclareMathOperator{\precompose}{\mathrm{precompose}}
 
@@
       type}} }
+
 \newcommand{\modal}{\ensuremath{\ocircle}}
@@
 
-Definition of graphs, diagrams, is\_colimit.
-
-Maybe the HIT of colimit, at least say that using HIT, all colimits
-exists.
+In section~\ref{sec:giraud-ax}, we need to consider colimits of
+diagrams.
+The following definition comes from~\cite{lumsdaine}~:
+\begin{defi}
+  A {\em graph} $G$ is the data of
+  \begin{itemize}
+  \item a type $G_0$ of vertices ;
+  \item for all $i,j:G_0$, a type $G_1(i,j)$ of edges.
+  \end{itemize}
+
+  A {\em diagram} $D$ on a graph $G$ is the data of
+  \begin{itemize}
+  \item for all $i:G_0$, a type $D_0(i)$ ;
+  \item for all $i,j:G_0$ and all $\phi : G_1(i,j)$, a map $D_1(\phi)
+    : D_0(i) \to D_0(j)$
+  \end{itemize}
+\end{defi}
+
+Now, we define what it means for a type to be a colimit of a diagram. 
+\begin{defi}
+Let $G$ be a graph, and $D$ be a diagram on $G$. 
+Let $P:\Type$ together with
+\begin{itemize}
+\item a map $q_i : D_0(i) \to P$ for all
+vertex $i:G_0$, \ie $$q : \prod_{i:G_0} D_0(i) \to P$$
+\item for all vertices $i,j:G_0$ and all edges $\phi:G_1(i,j)$, a path
+  $p_{i,j}^\phi : q_j \circ D_1(\phi) = q_i$, \ie
+  $$p : \prod_{i,j:G_0} \prod_{\phi:G_1(i,j)} q_j \circ D_1(\phi) = q_i.$$
+\end{itemize}
+
+Then $P$ is the {\em colimit} of $D$ if for any other $X:\Type$, the
+map
+$$\lambda f:P \to X,~ \left( \lambda i,~f \circ q_i~;~ \lambda i\, j\,
+  \phi,~ \precompose_f (p(i, j, g))) \right)$$
+is an equivalence.
+\end{defi}
+
+Essentially, being a colimit means making the diagram commutes, and
+being universal for this property.
+
+Note that using higer inductive types, every diagram $D$ on a graph
+$G$ admits a colimit, the type $Q$ whose constructors are
+\begin{itemize}
+\item $q:\prod_{i:G_0} D_0(i) \to Q$
+\item $p:\prod_{i,j:G_0} \prod_{\phi:G_1(i,j)} q_j \circ D_1(\phi) = q_i$
+\end{itemize}
 
@@
 % that's all folks
+
 \end{document}

### 2015-01-15 paper/sheaf_lics.tex
section 2 up to giro's axiom

--- 2015-01-14 paper/sheaf_lics.tex
@@
 \newtheorem{ax}[thm]{Axiom}
+
+\newcommand \defeq {\overset{de\hspace{-0.2ex}f}{=}}
 
@@
 \newcommand \Sh[1] {\mathrm{Sh}_{#1}}
+\newcommand \squash[1] {\| #1 \| }
+\newcommand \fib[2] {\mathrm{fib}_{#1}(#2)}
+\newcommand \colim[1] {\mathrm{colim}(Q)}
 
@@
   % 
-  The same construction has not been considered in Martin-Löf type
+  The same construction has not been developed in Martin-Löf type
   theory because of a mismatch between topos theory and type theory. This
@@
 %
-For instance, much effort have been done to provide a computational
+For instance, much efforts have been done to provide a computational
 content to the law of excluded middle in order to define a
@@
 
-This new point of view on type theory has revealed the
-homotopy structure of types where for instance homotopical
-propositions (or $\Type_{-1}$) are just types with an irrelevant equality and
-homotopical sets (or $\Type_{0}$)  are types with a propositional
-equality and so on for $\Type_{n}$.
+This new point of view on type theory has revealed the homotopy
+structure of types where for instance mere propositions (or
+$\Type_{-1}$) are just types with an
+irrelevant equality and sets (or $\Type_{0}$) are types
+with a propositional equality and so on for $\Type_{n}$.
 %
 The development of univalence has also shed some light on the
-difficulty to make the AC and EM coexist in type theory. Indeed, it
+difficulty to make AC and EM coexist in type theory. Indeed, it
 has been shown that a naive (non-propositional) version of EM is
@@
 
-When restricted to homotopical propositions and sets, homotopy type
+When restricted to mere propositions and sets, homotopy type
 theory corresponds quite closely to topos theory but the mismatch
-starts when considering higher homotopical types.
+starts when considering higher homotopy types.
 %
@@
 %
-As the notion of higher topoi appears to correspond very closely
-to homotopy type theory, this provides a new hope that tackling the problem of
-extending the power of type theory using sheafification is actually possible.
+As the notion of higher topoi appears to correspond very closely to
+homotopy type theory, this provides a new hope that tackling the
+problem of extending the power of homotopy type theory using
+sheafification is actually possible.
 
@@
 theory to homotopy type theory is not straightforward because the
-construction in higher topoi is restricted to the initial
-Grothendieck which is still very topological, and not very amenable to
-formalization in type theory. It seems more promising to use a
-generalized sheafification, called Lawvere-Tierney
+construction in higher topoi is restricted to the initial Grothendieck
+setting which is still very topologically oriented, and not very
+amenable to formalization in type theory. It seems more promising to
+use a generalized notion of sheafification, called Lawvere-Tierney
 sheafification~\cite{maclanemoerdijk}, but this construction has not
-been considered yet in the setting of higher topos theory.
-
-\subsection{Lawvere-Tierney Topology}
+been considered yet in the setting of higher topos theory. This rises
+two issues that this paper addresses: (i) how to lift the notion of
+Lawvere-Tierney sheafification to higher topos theory and (ii) is it
+possible to formalize this new definition in homotopy type theory.
+
+\subsection{Lawvere-Tierney Sheafification in Topos Theory}
 
@@
 %
-On can show that $\Sh{\E}$, the full sub-category of $\E$ given by
+One can show that $\Sh{\E}$, the full sub-category of $\E$ given by
 sheaves, is again a topos, with classifying object
@@
 
-But furthermore, one can defined a left adjoint, the sheafification
-functor, 
+But furthermore, one can defined a left adjoint to the inclusion, the
+sheafification functor
 %
@@
 $$
-to the inclusion functor, which exhibits $\Sh{\E}$ as a reflective
+which exhibits $\Sh{\E}$ as a reflective
 subcatefory of $\E$ (which is a particular case of localization). This
@@
 first thing to realize is that the construction can not be done in one
-step anymore. It must be done by induction on the level of homotopy of
-types. More precisely, the first layer of sheafification is defined
-for $\Type_{0}$ , given a topology on $\Type_{-1}$. This part
-corresponds to sheafification for topoi. Then, assuming that the
-sheafification has been constructed up to level $n$, one can defined
-the sheafification for $\Type_{n+1}$.
+single step anymore. It must rather be performed by induction on the
+level of homotopy types. More precisely, the first layer of
+sheafification is defined for $\Type_{0}$, given a topology on
+$\Type_{-1}$. This part corresponds to sheafification for topoi. Then,
+assuming that the sheafification has been constructed up to level $n$,
+one can defined the sheafification for $\Type_{n+1}$.
 %
@@
 sheafification is obtained by applying the plus-construction $n+1$ to
-$\Type_{n}$~\cite[Chapter~6]{lurie}.
+$n$-types~\cite[Chapter~6]{lurie}.
 
 This inductive step requires to formalize the notion of left-exact
-reflective subuniverse, which corresponds to a stratified
-version of left-exact modalities introduced
-in~\cite[Chapter~7]{hottbook}. A Lawvere-Tierney topology can thus be
-seen as a left-exact modality on $\Type_{-1}$ and sheafification as
-a way to extend it as a left-exact modality on $\Type_{n}$ for any $n$.
+reflective subuniverse, which corresponds to a stratified version of
+left-exact modalities introduced in~\cite[Chapter~7]{hottbook}. A
+Lawvere-Tierney topology can thus be seen as a left-exact modality on
+$\Type_{-1}$ and sheafification as an inductive process that extends
+it as a left-exact modality on $\Type_{n}$ for any $n$.
 
@@
 fibers or monomorphims (\ie $-1$-truncated homotopy
-fibers) depending on the context.
+fibers) depending on the purpose of the definition.
 
 For instance, the notion of separated objects (which is a technical
-tool in the construction of the sheafification), for which the morphism
+notion used in the construction of the sheafification functor), \ie
+objects for which the morphism
 (\ref{equ:sheaf_def}) is just a monomorphism, is generalized in two
@@
 %
-The subobjects to be considered are given by $n$-truncated fibers,
+The subobjects to be considered are given by $n$-truncated homotopy fibers,
 while the morphism (\ref{equ:sheaf_def}) is still required to be a
@@
 can be replaced by Giraud's axiom which basically allows to replace 
- kernel pairs by \v{C}ech nerves.
+ kernel pairs by \v{C}ech nerves in the property.
 
 Finally, one of the main difficulty in our formalization is that many
-equlity which are trival (b yproof irrelevance) in the topos theoretic
+equalities which are trival (by proof irrelevance) in the topos theoretic
 setting becomes relevant and requires a carreful use of univalence
@@
 Section~\ref{sec:hott} defines the foundations of homotopy type theory
-and the definition of modalities, on which this paper is based.
-%
-Section~\ref{sec:sheaves} defines the notion of sheaves in homotopy
+and presents a particular case of GIraud's axiom. 
+%
+Section\ref{sec:lexmod} introduces the notion of left-exact stratified
+modalities, on which this paper is based.
+%
+Section~\ref{sec:sheaves} defines Lawvere-Tierney sheaves in homotopy
 type theory.
 %
-Section~\ref{sec:sheafification} depicts the construction of the
+Section~\ref{sec:sheafification} explains the construction of the
 sheafification functor.
 %
-Section~\ref{sec:formalization} presents the details of the
-  formalization in Coq.
+Section~\ref{sec:formalization} discusses the details of the
+formalization in Coq, in particular issues concerning universe
+polymorphism.
 %
@@
 sheafification.
-
+%
 The definition of Section~\ref{ssec:hott} are part of the Coq library
@@
 
+As a prerequisite, we encourage the reader be familiar with type theory and
+in particular the point of view developed
+in~\cite{hottbook}. Nevertheless, we recall most of the central definitions 
+that we use so that the paper is sufficiently self-contained.
+%
+
  \subsection{Homotopy Types and Classifying Objects}
@@
 
-Our work is mainly based on the stratification by $\Type_n$~:
+One of the most direct application of homotopical notion to type
+theory is the introduction of homotopy types. 
+%
+Using the analogy that points in a space correspond to elements of a
+type and that paths between two points correspond to 
+elements of the corresponding identity type (which defines equality in type theory),  
+%
+an $n$-type is simply a type
+for which equality becomes trivial up to level $n$. 
+%
+Voevodsky has realized that this notion admits a compact inductive definition
+internal to type theory, given by
+% Our work is mainly based on the stratification by $\Type_n$~:
 \begin{defi}
@@
 \end{defi}
-For a type $X$, to be in $\Type_n$ means that path spaces of $X$ are
-trivial after $n+1$ iteration.
-We define some shortcuts
+% For a type $X$, to be in $\Type_n$ means that path spaces of $X$ are
+% trivial after $n+1$ iteration.
+%
+We also define some syntactic sugar for contractible types and mere propositions.
+
 \begin{itemize}
@@
 
-Working in $\Type_n$ allows to use the subobjects classification as in~\cite{sets_in_hott}~: for
-all type $B$, $\Type_n$ classifies subobjects of $B$ with $n$-truncated
-fibers ; we have an equivalence
+Using $\Type_n$ instead of just $\Type$ is the first step to connect
+type theory to higher topos theory. The next step is to exhibit a
+hierarchy of subobject classifiers on $n$-truncated homotopy fibers.
+%
+The homotopy fiber $\fib{f}(b)$ of a function $f$ at element $b$ is
+defined as 
+$$
+\fib{f}{b} \defeq \sum_{a:A} f(a) = b.
+$$ 
+%
+A function $f$ is with $n$-truncated homotopy fibers when $\fib{f}{b}$
+is an $\Type_n$ for any $b$.  
+%
+Following~\cite{sets_in_hott}, it is possible to show that, for any
+homotopy level $n$ and any type $B$, $\Type_n$ classifies subobjects
+of $B$ with $n$-truncated homotopy fibers in the sense that there is
+an equivalence
+%
 $$\left(\sum_{A:\Type} \sum_{f:A \to B} \prod_{b\in B}
-\IsType n~\mathrm{fib}_f(b)\right) \simeq (B \to \Type_n)$$
-and a pullback (with $\Type_n^\bullet$ the universe of pointed
+\IsType n~\fib{f}{b}\right) \simeq (B \to \Type_n)$$
+%
+and a pullback (with $\Type_n^\bullet \defeq \sum_{A:\Type_n} A$ the universe of pointed
 $n$-truncated types)~:
@@
 \xymatrix{
-  A \ar[r]^{t_f} \ar[d]_f & \Type_n^\bullet \ar[d]^{\pi_1}\\
-  B \ar[r]_{\chi_f} & \Type_n
+  A \ar[r]^{\hspace{-1em} t_f} \ar[d]_f & \Type_n^\bullet \ar[d]^{\pi_1}\\
+  B \ar[r]_{\hspace{-1em} \chi_f} & \Type_n
 }
 $$
-forall $f$ with $n$-truncated fibers, with 
-$$t_f = \lambda a,~(\mathrm{fib}_f(f(a)),(a,\mathrm{idpath})).$$
-
-Using higher inductive types, one can define for any type $X$ $\|X\| : \HProp$
-generated by 
+for any $f$ with $n$-truncated homotopy fibers, with 
+$$t_f = \lambda a,~(\fib{f}{f(a)},(a,\mathrm{idpath})).$$
+
+Finally, we need to use mere propositions has basic elements of a logic
+when characterizing properties of types or functions. 
+%
+But as some constructors do not preserve the level of homotopy type
+(\eg sums or $\Sigma$-types), we need to introduce $\squash{X}$, the
+{\em (propositional) truncation} of $X$ (also called bracket type, or
+squash type). This operation on types allows to truncate a type down
+to a mere proposition. Then for instance, the existential
+quantification on mere propositions ``there exists x:A such that
+P(x)'' can be defined as $ \squash{\sum_{x:A} P(x)}.  $
+
+Also proposition truncation could be seen as a primitive type former
+in homotopy type theory, it can be defined using higher inductive
+types. We adopt this point of view as it allows to maintain the slogan
+that homotopy type theory is type theory plus univalence and higher
+inductive types.
+%
+
+Given any type type $X$, its propositional truncation 
+$\squash{X} : \HProp$ is generated by 
 \begin{itemize}
-\item a function $|\cdot|_X : X \to \|X\|$,
-\item paths $x=y$ for all $x,y:\|X\|$.
+\item a function $|\cdot|_X : X \to \squash{X}$,
+\item for any $x,y:\squash{X}$, a path $x=y$.
 \end{itemize}
-$\|X\|$ is called the {\em (propositional) truncation of $X$}. The recursion principle asserts~:
+% 
+The recursion principle $\squash{A}$ of asserts that any mere
+proposition that follows from $A$ already follows from $\squash{A}$.
 \begin{lem}
   For any $A:\Type$ and $B:\HProp$, if $f:A \to B$ then there is an
-  induced $g:\|A\|\to B$ such that $g(|a|)= f(a)$ for any $a:A$.
+  induced $g:\squash{X}\to B$ such that $g(|a|)= f(a)$ for any $a:A$.
 \end{lem}
-We refer to~\cite{hottbook} for more details.
+%
+We refer the reader to~\cite{hottbook} for more details on propositional truncation.
 
@@
 
-\nt{Kevin -> complete this subsection}
-
-Formalization of colimits is based on the formalization of limits in~\cite{lumsdaine}.
-
-Definition of graphs, diagrams, is\_colimit.
-
-Maybe the HIT of colimit, at least say that using HIT, all colimits
-exists.
+In the definition of (a special case of) Giraud's axiom given in
+Section~\ref{sec:giraud-ax}, the colimit of the {\em \v{C}ech nerve}
+plays a central role.
+% 
+Following the notion of (homotopy) limits of diagrams over graphs that
+has been defined in homotopy type theory~\cite{lumsdaine}, we present
+here a definition of colimits of diagrams overs graphs. 
+%
+The main difference between limits and colimits is that limits are
+simply given by $\Sigma$-types, and thus exists already in traditional
+type theory, whereas the situation is more complicated for colimits as
+it requires the use of higher inductive types.
+
+Following~\cite{lumsdaine}, we introduce the notion of graph and
+diagram over a graph.
+%
+\begin{defi}
+  A {\em graph} $G$ is the data of
+  \begin{itemize}
+  \item a type $G_0$ of vertices ;
+  \item for any $i,j:G_0$, a type $G_1(i,j)$ of edges.
+  \end{itemize}
+
+  A {\em diagram} $D$ over a graph $G$ is the data of
+  \begin{itemize}
+  \item for any $i:G_0$, a type $D_0(i)$ ;
+  \item for any $i,j:G_0$ and all $\phi : G_1(i,j)$, a map $D_1(\phi)
+    : D_0(i) \to D_0(j)$
+  \end{itemize}
+\end{defi}
+
+An colimit of a diagram $D$ over a graph $G$ is given by a type $P$
+that defines a cone of $D$, plus the universal property that for any
+type $X$, the canonical map that transforms a function from $f : P
+\rightarrow X$ to a cone of $D$ on $X$ is an isomorphism.
+% 
+\begin{defi}
+Let $G$ be a graph, and $D$ be a diagram on $G$. 
+Let $P:\Type$ together with
+\begin{itemize}
+\item a map $q_i : D_0(i) \to P$ for any
+vertex $i:G_0$, \ie $$q : \prod_{i:G_0} D_0(i) \to P$$
+\item for any vertices $i,j:G_0$ and all edges $\phi:G_1(i,j)$, a path
+  $p_{i,j}^\phi : q_j \circ D_1(\phi) = q_i$, \ie
+  $$p : \prod_{i,j:G_0} \prod_{\phi:G_1(i,j)} q_j \circ D_1(\phi) = q_i.$$
+\end{itemize}
+
+Then $P$ is the {\em colimit} of $D$ if for any other $X:\Type$, the
+map
+$$\lambda f:P \to X,~ \left( \lambda i,~f \circ q_i~;~ \lambda i\, j\,
+  \phi,~ \precompose_f (p(i, j, g))) \right)$$
+is an equivalence.
+
+\nt{say what is $\precompose_f $}
+\end{defi}
+
+% Essentially, being a colimit means making the diagram commutes, and
+% being universal for this property.
+
+Using higer inductive types, every diagram $D$ on a graph $G$ admits a
+colimit in homotopy type thoery. This colimit $\colim D$ is given
+%
+\begin{itemize}
+\item a function $q:\prod_{i:G_0} D_0(i) \to \colim D$
+\item for any $i,j:G_0$ and $\phi:G_1(i,j)$, a path
+  $q_j \circ D_1(\phi) = q_i$.
+\end{itemize}
+%
+Although we have proven in our formalization that $\colim D$ is
+actually a colimit over $D$, we do not detail the proof as the
+existence of colimits is not use in the definition of the
+sheafification process.
 
@@
     is an equivalence.
-  \item For all $A:\Type_p$ and $B:A \to \Type_p$ such that $P(A)$
+  \item for any $A:\Type_p$ and $B:A \to \Type_p$ such that $P(A)$
     and $\prod_{x:A} P(B x)$, then $P\left( \sum_{x:A} B(x)\right)$
-  \item For all $A:\Type_p$ and $x,y:A$, if $\modal A$ is
+  \item for any $A:\Type_p$ and $x,y:A$, if $\modal A$ is
     contractible, then $\modal (x=y)$ is contractible.
@@
 
-Moreover, left exactness implies in particular fibers preservations~:
+Moreover, left exactness implies in particular homotopy fibers preservations~:
 \begin{prop}
@@
 
-We will define sheaves by induction on the homotopical truncation level of types
-~: the base case will be any left exact modality on $\HProp$, and we
+We will define sheaves by induction on the level of homotopy type:
+the base case will be any left exact modality on $\HProp$, and we
 give in this section some definitions useful for the inductive case.
@@
   $$\forall e:E,~ \left(\sum_{e':E} e=e'\right) \simeq (\modal (\chi~e)).$$ 
-  Moreover, for all $x:A$ we need the following coherence diagram to
+  Moreover, for any $x:A$ we need the following coherence diagram to
   commute
@@
 
-It follows from fibers preservation that any $n$-subobject of a type seen as a $n$-subobject
+It follows from homotopy fibers preservation that any $n$-subobject of a type seen as a $n$-subobject
 of its closure is closed.
@@
     arrow $f~:~E\to F$ to $\restriction f A$ the restriction of $f$ to $A$.
-  \item if $A \to E$ with $n$-truncated fibers, $\Phi_E^\chi$ is the
+  \item if $A \to E$ with $n$-truncated homotopy fibers, $\Phi_E^\chi$ is the
     map sending $f~:~E \to f$ to $\restriction f A$.
@@
 \begin{defi}[Separated Type]
-  A type $F$ in $\Type_{n+1}$ is {\em separated} if for all type $E$, and
+  A type $F$ in $\Type_{n+1}$ is {\em separated} if for any type $E$, and
   all dense subobject of $E$ classified by $\chi~:~E \to \Type_n$,
@@
   A type $F$ of $\Type_{n+1}$ is a {\em $(n+1)$-sheaf} if it is
-  separated, and for all type $E$ and all dense subobject of $E$
+  separated, and for any type $E$ and all dense subobject of $E$
   classified by $\chi~:~E \to \Type_{-1}$, $M_E^\chi$ is an
@@
   \item $\Type_n^\modal$ is a sheaf.
-  \item If $A:\Type_{n+1}$ and $B:A \to \Type_{n+1}$ such that for all
+  \item If $A:\Type_{n+1}$ and $B:A \to \Type_{n+1}$ such that for any
     $a:A$, $(B~a)$ is a sheaf, then $\prod_{a:A}B~a$ is a sheaf.
@@
 \end{enumerate}
-\subsection{For h-propositions}
+\subsection{For mere propositions}
 \label{ssec:h-propositions}
@@
 \begin{prop}
-  For all $T:\Type_{n+1}$, $\square T$ is separated.  
+  For any $T:\Type_{n+1}$, $\square T$ is separated.  
 \end{prop}
@@
 By induction on $k$, and the preservation of products by $\modal$, it
-suffices to show that for all $a,b:T$, $\modal (a=b) = (\mu_T a =
+suffices to show that for any $a,b:T$, $\modal (a=b) = (\mu_T a =
 \mu_T b)$. By univalence, we want arrows in both ways, forming an
@@
 and as $A$ is separated, $\pi_1\circ f = \pi_2 \circ g$.
-For the second components, let $x:E$. Notice that for all $x:E$,
+For the second components, let $x:E$. Notice that for any $x:E$,
 $\sum_{y:E} x = y$ has a dense subobject, $\sum_{y:\sum_{e:E} (\chi\,

### 2015-01-15 paper/sheaf_lics.tex
section 2 completed

@@
   in particular to give a meaning to the (propositional) law of
-  excluded middle inside homotopy type theory, without axiom, being
+  excluded middle inside homotopy type theory, being
   compatible with the full type-theoretic axiom of choice.
@@
 theory uses the fact that epimorphisms are coequalizers of their
-kernel pairs. This properties does not holds in higher topoi, but it
-can be replaced by Giraud's axiom which basically allows to replace 
- kernel pairs by \v{C}ech nerves in the property.
+kernel pairs. This properties does not holds in higher topoi, but
+using Giraud-Rezk-Lurie axioms of $\infty$-topoi, the same property
+holds by replacing kernel pairs by \v{C}ech nerves in the statement.
 
@@
 Section~\ref{sec:hott} defines the foundations of homotopy type theory
-and presents a particular case of GIraud's axiom. 
-%
-Section\ref{sec:lexmod} introduces the notion of left-exact stratified
+and presents a particular case of Giraud-Rezk-Lurie axioms. 
+%
+Section~\ref{sec:lexmod} introduces the notion of left-exact stratified
 modalities, on which this paper is based.
@@
 %
-And Section~\ref{sec:giraud-ax} defines (a particular case of)
-Giraud's axiom on higher topoi that we need in the definition of
-sheafification.
+And Section~\ref{sec:giraud-ax} defines a particular case of
+Giraud-Rezk-Lurie axioms on higher topoi that we need in the
+definition of sheafification.
 %
@@
 %
-The homotopy fiber $\fib{f}(b)$ of a function $f$ at element $b$ is
+The homotopy fiber $\fib{f}{b}$ of a function $f$ at element $b$ is
 defined as 
@@
 %
-A function $f$ is with $n$-truncated homotopy fibers when $\fib{f}{b}$
+A function $f$ is with $n$-truncated homotopy fibers (or simply
+$n$-truncated function) when $\fib{f}{b}$
 is an $\Type_n$ for any $b$.  
@@
 %
-$$\left(\sum_{A:\Type} \sum_{f:A \to B} \prod_{b\in B}
-\IsType n~\fib{f}{b}\right) \simeq (B \to \Type_n)$$
-%
-and a pullback (with $\Type_n^\bullet \defeq \sum_{A:\Type_n} A$ the universe of pointed
-$n$-truncated types)~:
+$$\chi : \left(\sum_{A:\Type} \sum_{f:A \to B} \prod_{b\in B}
+\IsType \
+\fib{f}{b}\right) \xrightarrow{\sim} 
+ (B \to \Type_n)$$
+%
+ such that the following diagram is a pullback for any $f$ with
+ $n$-truncated homotopy fibers:
 $$
@@
 $$
-for any $f$ with $n$-truncated homotopy fibers, with 
+with $\Type_n^\bullet \defeq \sum_{A:\Type_n} A$ the universe of pointed
+$n$-truncated types and 
 $$t_f = \lambda a,~(\fib{f}{f(a)},(a,\mathrm{idpath})).$$
@@
 
-In the definition of (a special case of) Giraud's axiom given in
+In the definition of (a special case of) Giraud-Rezk-Lurie axiom given in
 Section~\ref{sec:giraud-ax}, the colimit of the {\em \v{C}ech nerve}
@@
 
-An colimit of a diagram $D$ over a graph $G$ is given by a type $P$
+A colimit of a diagram $D$ over a graph $G$ is given by a type $P$
 that defines a cone of $D$, plus the universal property that for any
-type $X$, the canonical map that transforms a function from $f : P
+type $X$, the canonical map that transforms a function $f : P
 \rightarrow X$ to a cone of $D$ on $X$ is an isomorphism.
@@
 Although we have proven in our formalization that $\colim D$ is
-actually a colimit over $D$, we do not detail the proof as the
+actually a colimit over $D$, we do not detail the proof here as the
 existence of colimits is not use in the definition of the
-sheafification process.
-
-\subsection{Giraud's axiom}
+sheafification process. We only make use of the following special case
+of Giraud-Rezk-Lurie axioms.
+
+\subsection{On Giraud-Rezk-Lurie axioms}
 \label{sec:giraud-ax}
 
+The Giraud-Rezk-Lurie axioms are the $\infty$-version of Giraud's
+axioms that characterize a topos. Namely, there are 4 axioms on a
+$(\infty,1)$-category that have been shown to be equivalent to be
+equivalent to $(\infty,1)$-topos~\cite[Chapter 6]{lurie}. This is not
+the purpose of this paper to discuss those axioms, we only state here
+a particular case of the last axiom which says that every groupoid
+object is effective. 
+%
+This particular case connects an epimorphism (or $-1$-truncated
+function) to the colimit of its \v{C}ech nerve. 
+%
+This is the $\infty$-generalization of the fact that, in a topos,
+every epimorphism is the colimit of its kernel pair, which plays a
+central part in the definition of sheafification.
+
 \begin{defi}
-  Let $f:X \to Y$ be a map, and $p:\nat$. The $p$-pullback of $f$,
-  noted $X\times_Y \cdots\times_Y X$ is
+  Let $f:X \to Y$ be a map, and $p:\nat$. The $p$-pullback of $f$
+  (also called the $p$-fold fiber product of $X$ over $Y$),
+  noted $X\times_Y \cdots\times_Y X$, is
   the limit of the diagram
@@
 
-In homotopy type theory, we have 
-$$X\times_Y \cdots\times_Y X = \sum_{x:X^n} (f x_1 = f x_2) \land
-\cdots \land (f x_{n-1} = f x_n).$$
+As explained in Section~\ref{sec:colim-homot-type}, every limit exists
+in (homotopy) type theory and can be described by a $\Sigma$-type. 
+%
+The particular case of a $p$-pullback is given by
+%
+$$
+X\times_Y \cdots\times_Y X = \sum_{x:X^n} (f x_1 = f x_2) \land
+\cdots \land (f x_{n-1} = f x_n).
+$$
+%
+The \v{C}ech nerve of a map $f$ can then be described as the
+simplicial object that, at degree $p$, is given by the
+$(p+1)$-pullback of $f$. The \v{C}ech nerve can thus also be described
+as a diagram over $\Delta^{op}$ (where $\Delta$ is the simplicial
+category, seen as a graph). For simplicity, we only give an intuitive
+definition here, the complete definition can be found in our
+formalization.
+%
 \begin{defi}
-  If $f:X \to Y$ is a map, then the {\em \v{C}ech nerve} of $f$ is the
-  diagram
+  Let $f$ be a map from $X$ to $Y$. The {\em \v{C}ech nerve} $C(f)$ of $f$
+  is given by the diagram
   $$\xymatrix{
@@
 
-\begin{ax}[Giraud]
-  If $f:X\to Y$ is a surjection, then the
-colimit of $C(f)$ is $Y$.
+\begin{ax}[Giraud-Rezk-Lurie]
+  For any epimorphism $f : X \to Y$, the colimit of its \v{C}ech nerve
+  $C(f)$ is $Y$.
 \end{ax}
 
-Viewing homotopy type theory as a
-type-theoristic version of higher topos theory, \cite{lurie} suggests that this
-axiom can be proved, although we admit it in this paper.
+With the point of view that homotopy type theory is a type-theoristic
+version of higher topos theory, Theorem 6.1.0.6 of~\cite{lurie}
+suggests that this axiom can be proved in homotopy type theory. 
+%
+But proving this axiom seems to be a quite complex task on its own,
+and we have decided to admit it in this paper.
 

### 2015-01-15 paper/sheaf_lics.tex
more on modalities

@@
 \newcommand \colim[1] {\mathrm{colim}(Q)}
+\newcommand \zero {\mathbf{0}}
+\newcommand \one {\mathbf{1}}
 
@@
 setting becomes relevant and requires a carreful use of univalence
-with appropriate isomorphisms.
+with appropriate isomorphisms and a lot of reasoning on equalities
+between paths.
 
@@
 
-\cite{shulman-higher-modalities}
-
-\subsection{Definition}
-
-We use the following definition of truncated left exact modalities~:
+This section is devoted to the definition and analyze of left exact
+modalities as they have been introduced in \cite{hottbook} and
+\cite{shulman-higher-modalities}. Our definition differs slightly from
+those work in that we restrict the definition of the modality to be
+$n$-truncated types for a given truncation index $n$, instead of on
+all types.
+%
+This implies that we cannot reuse directly all results developed
+in~\cite{hottbook} as some care must be taken to check that truncation
+levels are preserved.
+
+Note that similar notion of modal operator have been already studied
+extensively in non-dependent type
+theory~\cite{benton1998computational}. In those works, they connect
+modalities to computational monad in programming
+languages~\cite{moggi-monad}. However, in our setting, modalities
+correspond to idempotent monads, a property is rarely true 
+in functional programming, except for the identity monad.
+
+The notion of left exact modalities plays a central role in this paper
+because Lawvere-Tierney topologies corresponds exactly to left exact
+modalities at $-1$ and sheafification is described as an inductive
+process that produces a left exact modality at level $n+1$ from a
+level exact modality at level $n$.
+
+Section~\ref{sec:definition} presents the definition and basic
+properties of left exact modalities. 
+%
+Section~\ref{sec:new-type-theory} discusses how a left exact modality
+induces a ne homotopy theory, and in which case this theory is
+consistent.
+%
+Finally, Section~\ref{sec:examples-left-exact} introduces some
+example, and in particular the modality for the dense topology.
+
+\subsection{Definition and Basic Properties}
+\label{sec:definition}
+
+% We use the following definition of truncated left exact modalities:
 \begin{defi}
-  Let $p$ be a truncation index. Then a left exact modality is the
-  data of
+  \label{sec:defin-basic-prop-1}
+  Let $n\geq -1$ be a truncation index. A left exact modality at level
+  $n$ is the data of
   \begin{enumerate}[(i)]
   \item A predicate $P:\Type_p \to \HProp$
-  \item For every $p$-truncated type $A$, a $p$-truncated type
+  \item For every $n$-truncated type $A$, a $n$-truncated type
     $\modal A$ such that $P(\modal A)$
-  \item For every $p$-truncated type $A$, a map $\eta_A:A \to
+  \item For every $n$-truncated type $A$, a map $\eta_A:A \to
     \modal A$
@@
     \setcounter{enumi}{3}
-  \item For every $p$-truncated types $A$ and $B$, if $P(B)$ then
+  \item For every $n$-truncated types $A$ and $B$, if $P(B)$ then
     $$\left\{
@@
 
-The type of all $p$-types such that $P$ will be noted $\Type^\modal_p$
-
-Since basic operations (dependent products, products, sigma types) are
-stable for truncation levels, all theorem in~\cite{hottbook}, chapter
-7.7 remains true.
-In particular, if $X:\Type_p$ verifies $P(X)$, then for any $x,y:X$, $P(x=y)$.
-
+The type of all modal $n$-types, \ie types for which $P$ holds, is
+defined as %$\Type^\modal_p$, so 
+$$
+\Type^\modal_p \ \defeq \sum_{T : \Type_n} P (T).
+$$
+%
+Since basic type formers (dependent products, products, sigma types)
+preserves truncation levels, we can leverage all theorems on modalites
+presented in~\cite[Chapter 7.7]{hottbook}.
+
+We now summarize the main properties that are used in our
+formalization.
+
+\begin{prop}
+
+stability properties
+\vspace{1em}
+
+\nt{state precisely some properties here, at least the 4 below}
+
+\vspace{1em}
+
+\begin{itemize}
+
+\item
+$\modal \one =\one $
+\item
+Stability under dependent product (7.7.2 of \cite{hottbook})
+\item
+ In particular, if $X:\Type_p$ verifies
+$P(X)$, then for any $x,y:X$, $P(x=y)$.
+\item
 Another useful property is that we can extend property {\it (iii)} to
 dependent product.
-
-Moreover, left exactness implies in particular homotopy fibers preservations~:
+\end{itemize}
+\end{prop}
+
+As the definition of sheafification is based on subobject classifiers
+and thus on homotopy fibers, we need to look at how left exact
+modalities preserves homotopy fibers.
 \begin{prop}
+\label{sec:defin-basic-prop}
 For any $n$-truncated types $X$ and $Y$,
@@
 }$$
-where $\pi_1 \circ \gamma = \eta_X$, and $\pi_2 \circ \gamma$ is the
-usual modalisation of paths.
+where $\pi_1 \circ \gamma = \eta_X \circ \pi_1$, and $\pi_2 \circ
+\gamma$ is given by the modalisation of paths.
 \end{prop}
-
-As we want, a left exact modality preserves $\HProp$~:
+\begin{proof}
+\nt{ça a du sens de mettre une preuve, non ? c'est non trivial et
+  nouveau par rapport aux travaux sur les modalités.}
+\end{proof}
+
+In the same way, left exact modalities preserve homotopy types.
 \begin{prop}
-  If $P:\HProp$, then $\modal \widehat P : \HProp$, where $\widehat P$
-  is $P$ seen as a $p$-type.
+  Let $k \leq n$.
+  If $P:\Type_k$, then $\modal \widehat P : \Type_k$, where $\widehat P$
+  is $P$ seen as a $n$-type.
 \end{prop}
+\begin{proof}
+  An $n$-truncated type $P$ can equivalently be described as a type for
+  which the unique map to $\one$ is with $n$-truncated fibers. Thus, the
+  property is a direct corollary of
+  Proposition~\ref{sec:defin-basic-prop} and the fact that $\modal \one =
+  \one$.
+\end{proof}
+
+\subsection{The new type theory arising from a left exact modality}
+\label{sec:new-type-theory}
+
+As noticed in \cite{hottbook}, left exact modalities corresponds 
+to sub-$(\infty,1)$-topoi.
+
+A left exact modality leads to a new homotopy type theory with
+potentially new principles.
+
+Definition of Sigma-Types.
+
+Preservation of univalence.
+
+\nt{state the property on IsEquiv to show preservation of univalence}
+
+Every left exact modality induces a new type theory, but with more
+properties, that can make it inconsistent. We now provide a simple
+characterization of left exact modalities that give rise to a
+consistent theory.
+
+\begin{prop}
+  A left exact modality $\modal$ induces a consistent type theory if
+  and only if $\modal \zero$ can not be inhabited in the initial type
+  theory. In that case, the modality is said to be consistent.
+\end{prop}
+\begin{proof}
+  By condition (iv) of Definition~\ref{sec:defin-basic-prop-1},
+  $\modal \zero$ is an initial object of $\Type^\modal_p$, and thus
+  corresponds to false for modal mere proposition.
+  % 
+  As $\modal \one = \one$, $\Type^\modal_p$ is consistent when
+  $\modal \zero \neq \one$, that is when there is no proof of
+  $\modal \zero$.
+\end{proof}
 
 \subsection{Examples of Left Exact Modalities}
+\label{sec:examples-left-exact}
 
@@
 extend it to every $\Type_n$ without collapsing every type.
-
-\subsection{The new type theory arising from a left exact modality}
-
-A left exact modality leads to a new homotopy type theory with
-potentially new principles.
-
-Definition of Sigma-Types.
-
-Preservation of univalence.
-
-Consistency issues.
 
@@
 \url{https://github.com/KevinQuirin/sheafification/}.
+%
+It is based on the homotopy type theory library available at
+\url{https://github.com/HoTT/HoTT}
 

### 2015-01-15 paper/sheaf_lics.tex
first complete pass on section 3

@@
 As noticed in \cite{hottbook}, left exact modalities corresponds 
-to sub-$(\infty,1)$-topoi.
-
-A left exact modality leads to a new homotopy type theory with
-potentially new principles.
-
-Definition of Sigma-Types.
-
-Preservation of univalence.
-
-\nt{state the property on IsEquiv to show preservation of univalence}
-
-Every left exact modality induces a new type theory, but with more
+to sub-$(\infty,1)$-topoi. Indeed, modal types defines a new type
+theory as modality preserves all type former, such as dependent
+products, dependent sums or inductive types (\nt{double check that}).
+
+Furthermore, when $A$ and $B$ are modal and $f : A \to B$, the mere
+proposition of $f$ being an equivalence is also modal. This means that
+univalence also holds in $\Type^\modal_p$, so it induces a good notion
+of homotopy type theory.
+
+Every left exact modality induces a new homotopy type theory, but with more
 properties, that can make it inconsistent. We now provide a simple
@@
 
-A basic example is the {\em open modality of $P$} for a $P:\HProp$,
-$\modal_P T = P \to T$.  When one works in the subuniverse defined by
-this modality, defining an inhabitant of a type $T$ is actually giving
-an inhabitant of $P \to T$. Thus, being in this subuniverse is just
-adding the axiom $P$ to the theory. Sadly, adding an axiom this way
-does not add any computational content to $P$, thus is not completely
-satisfactory.
-
-Another example is the $\lnot\lnot$-modality, $\modal_{\lnot\lnot} T = (T
-\to \bot) \to \bot$. Although we might think it enables us to use
-classical reasonings in the corresponding reflective subuniverse, the
-main problem is that every type in the new universe is an $\HProp$,
-which is very restrictive. 
-We will use this modality only in the $\HProp$ layer, and try to
-extend it to every $\Type_n$ without collapsing every type.
+A first example of left exact modality for every level $p$ is the {\em
+  open modality of $P$} for a $P:\HProp$ defined as
+$\modal_P T = P \to T$.
+%
+When one works in
+the subuniverse defined by this modality, defining an inhabitant of a
+type $T$ amounts to give an inhabitant of $P \to T$. 
+%
+Thus, being in this subuniverse corresponds to adding the axiom $P$ to the
+theory. Of course, adding an axiom this way does not add any computational
+content to $P$, and the consistency of the modallity corresponds to
+the fact that $P \to \zero$ (or $\lnot P$) is not inhabited.
+
+Another example of left exact modality for every level $p$ is the
+double negation modality,
+$\modal_{\lnot\lnot} T = (T \to \zero) \to \zero$.
+%
+This modality enables to use
+classical reasonings in the corresponding reflective subuniverse, but
+every type in the new universe is collapsed to an $\HProp$, ending up
+with porpositional classical logic.
+
+In this paper, we use the double negation modality at level $-1$ only,
+and then extend it to higher homotopy types using sheafification. This
+way, we can add classical reasoning in the type theory without
+collapsing the theory to mere propositions.
 

### 2015-01-15 paper/sheaf_lics.tex
in the middle of section 4

@@
 \subsection{Lawvere-Tierney Sheafification in Topos Theory}
+\label{sec:lawv-tiern-sheaf}
 
@@
 %
-This particular case connects an epimorphism (or $-1$-truncated
+This particular case connects an epimorphism (or a $-1$-connected
 function) to the colimit of its \v{C}ech nerve.
@@
 \nt{ça a du sens de mettre une preuve, non ? c'est non trivial et
-  nouveau par rapport aux travaux sur les modalités.}
+  nouveau par rapport aux travaux sur les modalités. Ou alors c'est
+  la preuve de Shulman qui tu n'as pas faite. En tout cas, il faut clarifier.}
 \end{proof}
@@
 
-We will define sheaves by induction on the level of homotopy type:
-the base case will be any left exact modality on $\HProp$, and we
-give in this section some definitions useful for the inductive case.
+As explained in the introduction, a Lawvere-Tierney topology for an
+topos $\E$, as defined in Section~\ref{sec:lawv-tiern-sheaf},
+corresponds to a left exact modality at level $-1$.
+%
+And sheafification corresponds to lifting this modality to a left
+exact modality at level $0$.
+%
+In homotopy type theory, it is possible to use similar ideas to lift
+any left exact modality at level $n$ to a left exact modality at level
+$n+1$ using sheafification.
+
+In this section, we suppose given a truncation index $n\geqslant -1$,
+and a left exact modality $\modal$ on $\Type_n$ and we introduce the
+definition of sheaves for $(n+1)$-types and basic properties of
+sheaves that will be useful in the next section.
+
+% We will define sheaves by induction on the level of homotopy type:
+% the base case will be any left exact modality on $\HProp$, and we
+% give in this section some definitions useful for the inductive case.
 
@@
 
-In this section, we suppose given a truncation index $n\geqslant -1$,
-and a left exact modality $\modal$ on $\Type_n$.
+To define sheaves, we first have to leverage the notion of dense
+subobjects for the modality $\modal$. The closure operator is defined
+using the $n$-subobject classifier $\Type_n$ introduced in
+Section~\ref{ssec:hott}.
 
 \begin{defi}
-  Let $E$ be a type. The {\em closure} of a subobject of $E$ classified by $\chi$
-  is the subobject of $E$ classified by $\modal \circ \chi$.
-
-  The subobject of $E$ classified by $\chi$ is said {\em closed in
-    $E$} if its closure is itself : $\chi = \modal \circ \chi$.
+  Let $E$ be a type. The {\em closure} of a subobject of $E$ with
+  n-truncated homotopy fibers (or $n$-subobject of $E$, for short),
+  classified by $\chi : E \to \Type_n$, is the subobject of $E$
+  classified by $\modal \circ \chi$.
+
+  An $n$-subobject of $E$ classified by $\chi$ is said to be {\em
+    closed in $E$} if it is equal to its closure, \ie
+  $\chi = \modal \circ \chi$.
 \end{defi}
 
+The definition of a dense $n$-subobject requires more work as the
+characteristic function $\chi_E$ of $E$ seen as an $n$-subobject of
+itself is not the constant function equal to $\one$ as it is the case
+for topoi, but rather the map 
+%
+$$\chi_E  = \lambda e . \ \sum_{e':E} e=e'.$$
+%
+This means that we need to take more care of the homotopy structure of
+this charactestic function, and in particular, the defintiion of a
+dense $n$-subobject comes with an additional coherence condition
+(which is trivial when $n=-1$).
+
 \begin{defi}
-  Let $E$ be a type, and $\chi:E \to \Type_n$. The subobject $A$ of $E$
-  classified by $\chi$ is {\em dense} in $E$ if its $\modal$-closure
-  is $E$ seen as a subobject of $E$, \ie
-  $$\forall e:E,~ \left(\sum_{e':E} e=e'\right) \simeq (\modal (\chi~e)).$$ 
-  Moreover, for any $x:A$ we need the following coherence diagram to
-  commute
+  Let $E$ be a type, and $\chi:E \to \Type_n$. The subobject $\iota :
+  A \to E$ of $E$
+  classified by $\chi$ is {\em dense} in $E$ when its $\modal$-closure
+  is equvialent to $\chi_E$, \ie
+  %
+  $$\forall e:E,~ \left(\sum_{e':E} e=e'\right) \simeq (\modal
+  (\chi~e)).$$
+  % 
+  and moreover, for any $x:A$, the following diagram 
+  commutes, 
   $$\xymatrix{
-    \sum_{e':A} x = e' \ar@{=}[r] \ar[d]_\iota & (\chi~x) \ar[d]^{\eta_{(\chi~x)}}\\
+    \sum_{e':A} x = e' \ar@{=}[r] \ar[d]_{\iota'} & (\chi~x) \ar[d]^{\eta_{(\chi~x)}}\\
     \sum_{e':E} x = e' \ar@{=}[r] & \modal (\chi~x)
   }$$
-  where $\iota: x \mapsto ({x_1}_1 ; x_2)$.
+  where $\iota': x \mapsto (\iota (\pi_1 x) ; \pi_2 x)$.
 \end{defi}
 
-It follows from homotopy fibers preservation that any $n$-subobject of a type seen as a $n$-subobject
-of its closure is closed.
-
-\begin{defi}[Restriction]
+It follows from homotopy fibers preservation that any $n$-subobject of
+a type seen as a $n$-subobject of its closure is closed.
+
+As recall in Section~\ref{sec:lawv-tiern-sheaf}, in topos theory,
+sheaves are elements that can not distinguish between objects and
+their dense sub-objects. Here, we need to distinguish between
+dense $(-1)$-subobjects, that will be used in the definition of
+sheaves, and dense $n$-subobjects, that will be used in the definition
+of separated types.
+
+\begin{defi}
   For any $A, E:\Type$ and $F:\Type_{n+1}$, we define~:
   \begin{itemize}
-  \item if $A \hookrightarrow E$, $M_E^\chi$ is the map sending an
+  \item if $m : A \hookrightarrow E$ is a monomorphism, we define 
+    %
+    $$
+    M_E^\chi : (E \to F) \to (A \to F)
+    $$
+    %
+    is the map sending an
     arrow $f~:~E\to F$ to $\restriction f A$ the restriction of $f$ to $A$.

### 2015-01-15 paper/sheaf_lics.tex
begin formalization

--- 2015-01-15 paper/sheaf_lics.tex
@@
 \DeclareMathOperator{\IsContr}{IsContr}
-\DeclareMathOperator{\precompose}{\mathrm{precompose}}
+\DeclareMathOperator{\postcompose}{\mathrm{postcompose}}
 
@@
 $$\lambda f:P \to X,~ \left( \lambda i,~f \circ q_i~;~ \lambda i\, j\,
-  \phi,~ \precompose_f (p(i, j, g))) \right)$$
+  \phi,~ \postcompose_f (p(i, j, g))) \right),$$
+where $\postcompose_f : \phi = \psi \to \phi \circ f = \psi \circ f$,
 is an equivalence.
-
-\nt{say what is $\precompose_f $}
 \end{defi}
@@
 
+Formalisation of subobject classifiers has been done in
+\texttt{sub\_object\_classifier.v}, 284 lines.
+
+Colimits, $p$-pullbacks and \v{C}ech nerves has been formalized in
+\texttt{colimits.v, cech\_nerve.v}, 552 lines.
+
+Reflective subuniverses and modalities are formalized in
+\texttt{reflective\_subuniverse.v, modalities.v}, 1053 lines.
+
+Definition of th reflective subuniverse on $\HProp$ is given in
+\texttt{sheaf\_base\_case.v}, 186 lines.
+
+Section~\ref{sec:sheaves} is formalized in
+\texttt{sheaf\_def\_and\_thm.v}, 1029 lines.
+
+Section~\ref{sec:sheafification} is formalized in
+\texttt{sheaf\_induction.v}, 2340 lines.
+
+At total, with other files containing technical lemmas, the project
+contains 6322 lines.
+
 \nt{describe how many line for each, be clear about what has not been formalized}

### 2015-01-15 paper/sheaf_lics.tex
section 4 completed, with remarks

--- 2015-01-15 paper/sheaf_lics.tex
@@
 sheaves are elements that can not distinguish between objects and
-their dense sub-objects. Here, we need to distinguish between
+their dense sub-objects.
+
+\begin{defi}
+  For any type $A$ and $E$, $\iota : A -> E$ and $F:\Type_{n+1}$, we define
+    $$
+    \Phi_E^\iota : (E \to F) \to (A \to F) 
+    $$
+    %
+    as the map sending an
+    arrow $f:E\to F$ to its restriction the restriction of $f \circ m$.
+  % \item if $A \to E$ with $n$-truncated homotopy fibers, $\Phi_E^\chi$ is the
+  %   map sending $f~:~E \to f$ to $\restriction f A$.
+  % \end{itemize}
+\end{defi}
+
+Here, we need to distinguish between
 dense $(-1)$-subobjects, that will be used in the definition of
@@
 
-\begin{defi}
-  For any $A, E:\Type$ and $F:\Type_{n+1}$, we define~:
-  \begin{itemize}
-  \item if $m : A \hookrightarrow E$ is a monomorphism, we define 
-    %
-    $$
-    M_E^\chi : (E \to F) \to (A \to F)
-    $$
-    %
-    is the map sending an
-    arrow $f~:~E\to F$ to $\restriction f A$ the restriction of $f$ to $A$.
-  \item if $A \to E$ with $n$-truncated homotopy fibers, $\Phi_E^\chi$ is the
-    map sending $f~:~E \to f$ to $\restriction f A$.
-  \end{itemize}
-\end{defi}
-
 \begin{defi}[Separated Type]
   A type $F$ in $\Type_{n+1}$ is {\em separated} if for any type $E$, and
-  all dense subobject of $E$ classified by $\chi~:~E \to \Type_n$,
-  $\Phi_E^\chi$ is a monomorphism. In other words, the dotted arrow,
+  all dense $n$-subobject of $\iota : A \to E$,
+  $\Phi_E^\iota$ is a monomorphism. In other words, the dotted arrow,
   if exists, is unique.
@@
   $$\xymatrix{
-    \sum_{e:E} (\chi~ e) \ar[r]^f \ar[d]_{\pi_1} & F \\
+    A \ar[r]^f \ar[d]_{\iota} & F \\
     E \ar@{-->}[ru]_{!}&
@@
   A type $F$ of $\Type_{n+1}$ is a {\em $(n+1)$-sheaf} if it is
-  separated, and for any type $E$ and all dense subobject of $E$
-  classified by $\chi~:~E \to \Type_{-1}$, $M_E^\chi$ is an
+  separated, and for any type $E$ and all dense $(-1)$-subobject of
+  $\iota : A \to E$, $\Phi_E^\iota$ is an
   equivalence. In other words, the dotted arrow exists and is unique.
@@
   $$\xymatrix{
-    \sum_{e:E} (\chi~ e) \ar[r]^f \ar[d]_{\pi_1} & F \\
+    A \ar[r]^f \ar[d]_{\iota} & F \\
     E \ar@{-->}[ru]_{\exists !}&
@@
 Note that these definitions are almost the same as the ones
-in~\cite{maclanemoerdijk}. The main difference is that {\em separated}
-is defined for $n$-subobjects,
-while {\em sheaf} only for $-1$-subobject.
+in~\cite{maclanemoerdijk}. The main difference is that {separated}
+is defined for $n$-subobjects, while {sheaf} only for
+$(-1)$-subobjects
+
+\nt{Il faut introduire c'est propriété, $\Type_n^\modal$ is a sheaf
+  est très importante en particulier. Il faut qu'on comprenne d'où ça
+  vient et que tu fasses un preuve, même un peu idalisée}
 

### 2015-01-15 paper/sheaf_lics.tex
more detailed comment on section 4

@@
   est très importante en particulier. Il faut qu'on comprenne d'où ça
-  vient et que tu fasses un preuve, même un peu idalisée}
-
-From these definitions, one can show that
-\begin{prop}\label{prop:sheaf_prop}
-  \begin{itemize}
-  \item $\Type_n^\modal$ is a sheaf.
-  \item If $A:\Type_{n+1}$ and $B:A \to \Type_{n+1}$ such that for any
+  vient et que tu fasses une preuve, même un peu idalisée. Je dirais
+  que tu peux mettre des lemmes genres nj\_inter en disant que telle
+  flèche est un iso, and co.
+  En tout cas, cette partie peut et doit être bien plus longue.
+}
+
+\begin{prop}\label{prop:sheaf-forall}
+  If $A:\Type_{n+1}$ and $B:A \to \Type_{n+1}$ such that for any
   $a:A$, $(B~a)$ is a sheaf, then $\prod_{a:A}B~a$ is a sheaf.
-  \end{itemize}
+\end{prop}
+
+\begin{prop}\label{prop:sheaf-is-sheaf}
+  $\Type_n^\modal$ is a sheaf.
 \end{prop}
@@
 As $\square T$ embeds in $\left( \Type_n^\modal \right)^T$, we only
-have to show that the latter is separated. But it is the case with
-both parts proposition~\ref{prop:sheaf_prop}.
+have to show that the latter is separated. But it is the case because
+$\Type_n^\modal$ is a sheaf (by Proposition~\ref{prop:sheaf-is-sheaf})
+and a function type is a sheaf as soon
+as its codomain is a sheaf (by Proposition~\ref{prop:sheaf-forall}).
 \end{proof}

### 2015-01-15 paper/sheaf_lics.tex
remarque on the definition of sheaves in HTT

@@
 is defined for $n$-subobjects, while {sheaf} only for
-$(-1)$-subobjects
+$(-1)$-subobjects.
+%
+For higher topos theory, in the Grotehndieck setting, sheaves seems
+to correspond to the second part of our definition only (assuming that
+Lawvere-Tierney topologies can still be shown to subsume Grothendieck
+toplogies for $(\infty,1)$-topoi, which is beyond the scope of this paper).
+%
 
@@
 
+Check if Lawvere-Tierney subsumes Grothendieck in higher topos theory.
+
 % conference papers do not normally have an appendix

### 2015-01-15 paper/sheaf_lics.tex
starting to dive into section 5

--- 2015-01-15 paper/sheaf_lics.tex
@@
 extensively in non-dependent type
-theory~\cite{benton1998computational}. In those works, they connect
+theory~(see \eg \cite{benton1998computational}). In those works, they connect
 modalities to computational monad in programming
 languages~\cite{moggi-monad}. However, in our setting, modalities
-correspond to idempotent monads, a property is rarely true 
-in functional programming, except for the identity monad.
+correspond to idempotent monads, a property that is rarely true 
+in functional programming, except for the identity monad of course.
 
@@
 
-We mimic the construction in~\cite{maclanemoerdijk}. 
-From any left exact modality on $\HProp$, we extend the new principles
-it gives to every $\Type_n,\,n\geqslant 0$.
-Actually, if $n_0$ is a fixed truncation index, and $\modal$ a left
-exact modality on $\Type_{n_0}$, then we can in the same
-way extend the properties of the modality to any $\Type_n$, for
-$n > n_0$.
-
-Once the first modality is defined, the extension to all $\Type_n$ is
-automatic. It means that the new principles we want to add in the new
-theory must be introduced ine the first modality.
-
-We have to be careful~: if we want to propagate the properties of the
-first modality $\modal_{-1}$, all the higher modalities $\modal_n$ must be compatible
-with $\modal_{-1}$, \ie we want the property
+% We mimic the construction in~\cite{maclanemoerdijk}.
+
+From any left exact modality $\modal_{-1}$ on $\HProp$, we define by
+induction the truncation index $n$, a
+left exact modality $\modal_n$ at level $n$ that extends $\modal_{-1}$ in the
+sense that the following property holds.
+%
+% We have to be careful~: if we want to propagate the properties of the
+% first modality $\modal_{-1}$, all the higher modalities $\modal_n$ must be compatible
+% with $\modal_{-1}$, \ie we want the property
 \begin{prop}\label{prop:hprop}
-  If $P:\HProp$, and $\widehat P$ is $P$ seen as a $\Type_n$, then $\modal_n \widehat P = \modal_{-1} P$, and the
-  following coherence diagram commute 
+  For any mere proposition $P$ (with $\widehat P$ is $P$ seen as a
+  $\Type_n$),  $\modal_n \widehat P = \modal_{-1} P$ and the
+  following coherence diagram commutes
   $$\xymatrix{
-    \modal_{-1} P \ar@{=}[r] & \modal_n \widehat P \\
-    P \ar@{=}[r] \ar[u]^{\eta_{-1}} & \widehat P \ar[u]_{\eta_n}
+    P \ar@{=}[r] \ar[d]_{\eta_{-1}} & \widehat P \ar[d]^{\eta_n} \\
+    \modal_{-1} P \ar@{=}[r] & \modal_n \widehat P 
   }$$
@@
 
-The sheafification process works exactly that way
-in~\cite{maclanemoerdijk}~: from a left exact modality on the internal
-logic (a Lawvere-Tierney topology), we define a new left exact
-modality on the whole topos. We view that as the induction step
-between $\HProp$ and $\Type_0$, and we want to extend it.
-
-\kq{Bof ; maybe rewrite this}
-
-As in most of sheafification process, it will be done in
-two steps~:
+Note that once the first modality is defined, the extension to all
+$\Type_n$ is automatic. This means that the new logical principles
+that can be added by the sheaf construction must be added in the
+first modality already. The sheafification is just a way to propagate
+those new principles to other types than just mere proposition.
+
+%   extend the new principles it gives to
+% every $\Type_n,\,n\geqslant 0$.  Actually, if $n_0$ is a fixed
+% truncation index, and $\modal$ a left exact modality on $\Type_{n_0}$,
+% then we can in the same way extend the properties of the modality to
+% any $\Type_n$, for $n > n_0$.
+
+The sheafification in topos theory described in
+Section~\ref{sec:lawv-tiern-sheaf} works in the same way. 
+%
+From a left exact modality on the internal logic (a Lawvere-Tierney
+topology), a new left exact modality is defined on the whole topos.
+%
+This construction may be seen as the first induction step between
+$\HProp$ and $\Type_0$.
+
+\nt{je ne comprends pas ce passage, c'est quoi sheafness ? }
+For a given induction step, the sheafification is defined in two
+steps:
 \begin{enumerate}[(i)]
@@
 \end{enumerate}
-\subsection{For mere propositions}
-\label{ssec:h-propositions}
-
-For the case $n=-1$, one can take any left exact modality on
-$\HProp$. Here we took the $\lnot\lnot$ modality~:
-$$\forall P:\HProp,~\modal_j P = \lnot\lnot P.$$
-One can easily show that it defines a left exact modality.
-
-The $\lnot\lnot$-modality allows to work with a propositional law of
-excluded middle.
+
+% \subsection{For mere propositions}
+% \label{ssec:h-propositions}
+
+% For the case $n=-1$, one can take any left exact modality on
+% $\HProp$. Here we took the $\lnot\lnot$ modality~:
+% $$\forall P:\HProp,~\modal_j P = \lnot\lnot P.$$
+% One can easily show that it defines a left exact modality.
+
+% The $\lnot\lnot$-modality allows to work with a propositional law of
+% excluded middle.
+
+In the rest of the section, we suppose given a truncation index
+$n\geqslant -1$, and a left exact modality $\modal_n$ on $\Type_n$,
+compatible with a left exact modality $\modal_{-1}$ on $\HProp$ in the
+sense of proposition~\ref{prop:hprop}.
 
@@
 \label{ssec:from-type-separated}
-In this section, we suppose given a truncation index $n\geqslant -1$,
-and a left exact modality $\modal$ on $\Type_n$, compatible with the
-modality on $\HProp$ in the sense of proposition~\ref{prop:hprop}.
-
-Let $T$ be in $\Type_{n+1}$. We define $\square T$ as the image of
+
+Let $T : \Type_{n+1}$. We define $\square T$ as the image of
 $\modal^T \circ \{\cdot\}_T$, as in
@@
 (t':T),~t=t'$.
+\nt{parler du système de factorization et HoTT book}
 In type theory words, 
@@
 
-$$\xymatrix{
+$$\xymatrix@C=8em@R=4em{
   \sum_{y:\sum_{e:E} (\chi\,
-  e)} x=y_1 \dar[rrr]{2pt}^{\pi_2\circ f\circ\pi_1\circ \pi_1}_{\pi_2\circ g\circ \pi_1\circ \pi_1} \ar[d]_{\mathrm{dense}}&&& B\,x \\
-  \sum_{y:E} x = y \dar[rrru]{2pt}^{\pi_2\circ f\circ \pi_1}_{\pi_2\circ g\circ \pi_1}&
+  e)} x=y_1 \ar@<2pt>[r]^{\qquad \pi_2\circ f\circ\pi_1\circ \pi_1}
+\ar@<-2pt>[r]_{\qquad \pi_2\circ g\circ \pi_1\circ \pi_1}
+\ar[d]_{\mathrm{dense}}& B\,x \\
+  \sum_{y:E} x = y \ar@<2pt>[ur]^{\pi_2\circ f\circ \pi_1} \ar@<-2pt>[ur]_{\pi_2\circ g\circ \pi_1}&
 }$$
@@
 
-Talk about issues in universe levels. Lake of cumulativity for
+Talk about issues in universe levels \cite{sozeau2014universe}. Lake of cumulativity for
 $\Type_n$ and impossibility to define by induction the sheafification 

### 2015-01-16 paper/sheaf_lics.tex
test of username
added a section on the eli-mono fact. system

@@
 %
+Section~\ref{sec:epi-mono-fact} presents the usual epi-mono
+factorization system in the light of homotopy type theory.
+%
 Section~\ref{sec:colim-homot-type} introduces the notion of colimits
@@
 We refer the reader to~\cite{hottbook} for more details on propositional truncation.
+
+\subsection{The epi-mono Factorization System}
+\label{sec:epi-mono-fact}
+
+\nt{define the image here, with (-1)-truncated and (-1)-connected
+  factorization system}
 
@@
 
-Let $T : \Type_{n+1}$. We define $\square T$ as the image of
+Let $T : \Type_{n+1}$. We define $\square T$ as the image (see
+Section~\ref{sec:epi-mono-fact}) of
 $\modal^T \circ \{\cdot\}_T$, as in
@@
 (t':T),~t=t'$.
-\nt{parler du système de factorization et HoTT book}
+\nt{parler du système de factorization}
 In type theory words, 

### 2015-01-16 paper/sheaf_lics.tex
reorganize a bit section 5

@@
 \item
-Stability under dependent product (7.7.2 of \cite{hottbook})
+Stability under product and dependent product (7.7.2 of \cite{hottbook})
 \item
@@
 
-\nt{Il faut introduire c'est propriété, $\Type_n^\modal$ is a sheaf
+\nt{Il faut mieux introduire ces propriétés, $\Type_n^\modal$ is a sheaf
   est très importante en particulier. Il faut qu'on comprenne d'où ça
@@
 \begin{proof}
-  As $\mu_T$ is an surjection, we know by Giraud axiom that $\square T$ is the colimit
+  As $\mu_T$ is an surjection, using the particular case of the last
+  Giraud-Rezk-Lurie axiom introduced in Section~\ref{sec:giraud-ax},
+  we know that $\square T$ is the colimit
   of $C(\mu_T)$. If we can show that $C(\mu_T) = C(\id)$, the the
@@
 
+\nt{put the definition of $\modal_{n+1}$ upfront}
+
 If $T$ is already a separated type, the following lemma alllows us
@@
 
-\subsection{Summary}
-\label{ssec:summary}
-
 If $T:\Type_{n+1}$, the sheafification of $T$ is
@@
 It will imply that $\star$ is compatible with $\HProp$ in the sense of proposition~\ref{prop:hprop}.
+
+\subsection{Summary}
+\label{ssec:summary}
+
+\nt{Recap what has been done, and give the final result about EM with
+  the dense topology as a corollary.}
 
@@
 
-Prove that the Giraud axiom holds in homotopy type theory.
+Prove that the Giraud-Rezk-Lurie axiom holds in homotopy type theory,
+or at least the particular case we need.
 

### 2015-01-16 paper/sheaf_lics.tex
add some theorems about modalities

@@
 
+Now, let a map $f:A\to B$ be called {\em (-1)-truncated} if
+every fibers of $f$ is an $\HProp$, and {\em
+  (-1)-connected} if every fibers $f$ is merely
+inhabited, \ie such that $\|\fib f y\|$ for all $y$.
+
+\begin{defi}
+  Let $f:A\to B$ be a function. The {\em image} of $f$ is
+  $$\im(f) = \sum_{y:B} \|\fib f y\|.$$
+\end{defi}
+
+It is then fairly obvious that the canonical function $\hat f : A \to
+\im(f)$ is (-1)-connected, thus the following proposition arise
+(\cite{hottbook}, Lemma 7.6.4)~:
+\begin{prop}
+  An map $f:A\to B$ factors as a (-1)-connected function followed by a
+  (-1)-truncated function.
+\end{prop}
+
 \nt{define the image here, with (-1)-truncated and (-1)-connected
@@
   $n$ is the data of
+  \kq{Do you prefer $n$ over $p$ ?}
   \begin{enumerate}[(i)]
@@
 \begin{prop}
-
-stability properties
 \vspace{1em}
@@
 
-\item
-$\modal \one =\one $
-\item
-Stability under product and dependent product (7.7.2 of \cite{hottbook})
-\item
- In particular, if $X:\Type_p$ verifies
-$P(X)$, then for any $x,y:X$, $P(x=y)$.
-\item
-Another useful property is that we can extend property {\it (iii)} to
-dependent product.
+\item $\one$ seen as a $\Type_p$ is modal, ie $\modal \one =\one$.
+\item (\cite{hottbook}, Lemma 7.7.2)
+  Let $A:\Type$ and $B:A \to \Type_p^\modal$. Then $P\left(\prod_{a:A} (B\,
+  a)\right)$. 
+\item (\cite{hottbook}, Corollary 7.7.2) For any $A,B:\Type_p$,
+  $\modal (A\times B) = (\modal A) \times (\modal B)$.
+\item Let $A:\Type_p\modal$. Then for any $x,y:X$, $P(x=y)$.
+\item (\cite{hottbook}, Theorem 7.7.4) 
+  {\it (iv} can be extended to dependent product. Precisely~:
+
+  Let $A:\Type_p$ and $B:\modal A \to \Type_p^\modal$. For any map
+  $g:\prod_{a:A} B(\eta_A(a))$ there exists a map $f:\prod_{z:\modal
+    A} (B\, z)$ such that for all $a:A$, $f(\eta_A(a)) = g(a)$.
 \end{itemize}
@@
   nouveau par rapport aux travaux sur les modalités. Ou alors c'est
-  la preuve de Shulman qui tu n'as pas faite. En tout cas, il faut clarifier.}
+  la preuve de Shulman qui tu n'as pas faite. En tout cas, il faut
+  clarifier.}
+
+The central result used here is the following
+\begin{lem}
+  Let $X:\Type_p$, $Y:\Type_p\modal$ and $f:X\to Y$. If for all $y:Y$,
+  $\modal (\fib f y)$ is contractible, then $\modal X = Y$.
+\end{lem}
+
+Here, $\sum_{x:\modal X} (\modal f x = \eta_Y y)$ is modal, thus we
+can define $$f:\modal \left(\sum_{x:X}  (f x = y)\right) \to
+\sum_{x:\modal X} (\modal f x = \eta_Y y).$$
+We need to check that the image of its fibers by $\modal$ are
+contractible.
+\kq{Finish that}
 \end{proof}

### 2015-01-16 paper/sheaf_lics.tex
fibers preservation

@@
 
-\nt{state precisely some properties here, at least the 4 below}
-
 \vspace{1em}
@@
 \item (\cite{hottbook}, Theorem 7.7.4) 
-  {\it (iv} can be extended to dependent product. Precisely~:
+  {\it (iv)} can be extended to dependent product. Precisely~:
 
@@
 Here, $\sum_{x:\modal X} (\modal f x = \eta_Y y)$ is modal, thus we
-can define $$f:\modal \left(\sum_{x:X}  (f x = y)\right) \to
+can define $$\phi:\modal \left(\sum_{x:X}  (f x = y)\right) \to
 \sum_{x:\modal X} (\modal f x = \eta_Y y).$$
-We need to check that the image of its fibers by $\modal$ are
+We just need to check that the image of its fibers by $\modal$ are
 contractible.
+
+From definition of left exactness, one can deduce the following~:
+\begin{lem}
+  Let $A,B:\Type_p$. Let $f:A\to B$. If $\modal A$ and $\modal B$ are
+  contractible, then so is $\fib f b$ for any $b:B$.
+\end{lem}
+We can rewrite $\fib \phi {x;p}$ in terms of the fiber of a fonction
+$\fib {\eta_X} a \to \fib {\eta_Y} b$.
+
+But we have the lemma~:
+\begin{lem}
+  Let $A:\Type_p$. Then for any $y:\modal A$, $\modal(\fib {\eta_A}
+  y)$ is contractible.
+\end{lem}
+
+Finally, $\modal(\fib \phi {x;p})$ is contractible, and the result is proved.
 \kq{Finish that}
@@
 \end{prop}
+\begin{proof}
+  Firstly, we prove that $\Type_n^\modal$ is separated. Let $E:\Type$
+  and $\chi:E \to \Type_n$, dense in $E$. Let $\phi_1,\phi_2:E \to
+  \Type_n^\modal$, such that $\phi_1 \circ \pi_1 = \phi_2 \circ
+  \pi_1$. Finally, let $x:E$. We show $\phi_1 x = \phi_2 x$ using
+  univalence.
+
+  As $\chi$ is dense, we have a term $m_x : \modal(\chi\, x)$.
+  But as $\phi_2 x$ is modal, thus we can obtain a term $h_x : \chi\,
+  x$. 
+  As $\phi_1$ and $\phi_2$ are equal on $\sum_{e:E}\chi\, e$, we
+  have an arrow $\phi_1 x \to \phi_2 x$.
+  The same method leads to an arrow $\phi_2 x \to \phi_1 x$, and we
+  prove they are each other inverse.
+
+  Now, we prove that $\Type_n^\modal$ is a sheaf. Let $E:\Type$ and
+  $\chi:E \to \HProp$, dense in $E$. Let $f:\sum_{e:E} \chi\, e \to
+  \Type_n^\modal$. We want to extend $f$ to $E \to \Type_n^\modal$.
+
+  We define $g$ as 
+  $$g(e) = \modal \left( \fib \phi e \right), $$
+  where
+  $$\phi : \left(\sum_{b:\left(\sum_{e:E} (\chi\, e)\right)} (f\,
+    b)\right) \to E$$
+  defined by $\phi(x) = (x_1)_1$.
+  Using the following properties on fibers, we can prove that it
+  defines an equivalence.
+  \begin{lem}
+    Let $A,B,C:\Type_n$, $f:A\to B$ and $g:B\to C$.
+    Then
+    \begin{itemize}
+    \item If $g$ is an embedding, then 
+      $$\forall b:B,~\fib f b = \fib {g\circ f} {g\, b}.$$
+    \item If all fibers of $f$ and $g$ are $n$-truncated, then
+      $$\forall c:C,~\modal(\fib {g \circ f} c) =
+      \modal \left(  
+        \sum_{w:\fib g c} \modal (\fib f {w_1})
+      \right).$$
+    \end{itemize}
+  \end{lem}
+
+\end{proof}
 

### 2015-01-16 paper/sheaf_lics.tex
pass on section 2

@@
 
-\subsection{The epi-mono Factorization System}
+\subsection{The Epi-Mono Factorization System}
 \label{sec:epi-mono-fact}
 
-Now, let a map $f:A\to B$ be called {\em (-1)-truncated} if
-every fibers of $f$ is an $\HProp$, and {\em
-  (-1)-connected} if every fibers $f$ is merely
-inhabited, \ie such that $\|\fib f y\|$ for all $y$.
+We now recall the presence of a factorization system in homotopy type
+theory, constituted by monomorphism and epimorphism.
+
+A map $f:A\to B$ is a \emph{monomorphism} when it is (-1)-truncated
+(see definition in Section~\ref{ssec:hott}), and an \emph{epimorphism}
+when it is (-1)-connected, that is when every fiber of $f$ is merely
+inhabited (\ie $\|\fib f y\|$ holds for all $y$).
 
 \begin{defi}
-  Let $f:A\to B$ be a function. The {\em image} of $f$ is
-  $$\im(f) = \sum_{y:B} \|\fib f y\|.$$
+  Let $f:A\to B$ be a function. The {\em image} of $f$ is defined as 
+  $$\im(f) \defeq \sum_{y:B} \|\fib f y\|.$$
 \end{defi}
@@
 It is then fairly obvious that the canonical function $\hat f : A \to
-\im(f)$ is (-1)-connected, thus the following proposition arise
+\im(f)$ is (-1)-connected, thus the following proposition holds
 (\cite{hottbook}, Lemma 7.6.4)~:
 \begin{prop}
-  An map $f:A\to B$ factors as a (-1)-connected function followed by a
-  (-1)-truncated function.
+  A map $f:A\to B$ factors through $\im(f)$ as a (-1)-connected
+  function followed by a (-1)-truncated function.
 \end{prop}
 
-\nt{define the image here, with (-1)-truncated and (-1)-connected
-  factorization system}
+Note that there are other factorization systems constituted by
+$n$-truncated morphism and $n$-connected morphisms, for every
+truncated level $n$, but they do not appear to be useful in the
+definition of sheafification.
 
@@
 map
-$$\lambda f:P \to X,~ \left( \lambda i,~f \circ q_i~;~ \lambda i\, j\,
-  \phi,~ \postcompose_f (p(i, j, g))) \right),$$
+$$\lambda f:P \to X, \left( \lambda i,~f \circ q_i~;~ \lambda i\, j\,
+  \phi,~ \postcompose_f (p(i, j, g)) \right)$$
 where $\postcompose_f : \phi = \psi \to \phi \circ f = \psi \circ f$,
@@
   $n$ is the data of
-  \kq{Do you prefer $n$ over $p$ ?}
   \begin{enumerate}[(i)]
-  \item A predicate $P:\Type_p \to \HProp$
+  \item A predicate $P:\Type_n \to \HProp$
   \item For every $n$-truncated type $A$, a $n$-truncated type
@@
     is an equivalence.
-  \item for any $A:\Type_p$ and $B:A \to \Type_p$ such that $P(A)$
+  \item for any $A:\Type_n$ and $B:A \to \Type_n$ such that $P(A)$
     and $\prod_{x:A} P(B x)$, then $P\left( \sum_{x:A} B(x)\right)$
-  \item for any $A:\Type_p$ and $x,y:A$, if $\modal A$ is
+  \item for any $A:\Type_n$ and $x,y:A$, if $\modal A$ is
     contractible, then $\modal (x=y)$ is contractible.
@@
 $$
-\Type^\modal_p \ \defeq \sum_{T : \Type_n} P (T).
+\Type^\modal_n \ \defeq \sum_{T : \Type_n} P (T).
 $$
@@
 presented in~\cite[Chapter 7.7]{hottbook}.
-
+%
 We now summarize the main properties that are used in our
-formalization.
+formalization. We do not recall the proofs as they are not specific to
+our presentation (also the setting using truncated types provides
+minor changes), but they can be found in our Coq form
 
 \begin{prop}
-\vspace{1em}
-
-\vspace{1em}
 
@@
 \item $\one$ seen as a $\Type_p$ is modal, ie $\modal \one =\one$.
-\item (\cite{hottbook}, Lemma 7.7.2)
-  Let $A:\Type$ and $B:A \to \Type_p^\modal$. Then $P\left(\prod_{a:A} (B\,
-  a)\right)$. 
-\item (\cite{hottbook}, Corollary 7.7.2) For any $A,B:\Type_p$,
-  $\modal (A\times B) = (\modal A) \times (\modal B)$.
-\item Let $A:\Type_p\modal$. Then for any $x,y:X$, $P(x=y)$.
-\item (\cite{hottbook}, Theorem 7.7.4) 
-  {\it (iv)} can be extended to dependent product. Precisely~:
-
-  Let $A:\Type_p$ and $B:\modal A \to \Type_p^\modal$. For any map
-  $g:\prod_{a:A} B(\eta_A(a))$ there exists a map $f:\prod_{z:\modal
-    A} (B\, z)$ such that for all $a:A$, $f(\eta_A(a)) = g(a)$.
+\item 
+  Let $A:\Type$ and $B:A \to \Type_n^\modal$. Then $P\left(\prod_{a:A} (B\,
+  a)\right)$ (from \cite[Lemma 7.7.2]{hottbook}). 
+\item For any $A,B:\Type_p$,
+  $\modal (A\times B) = (\modal A) \times (\modal B)$ \\
+(from \cite[Corollary 7.7.2]{hottbook}).
+\item Let $A:\Type_n \modal$. Then for any $x,y:X$, $P(x=y)$.
+\item 
+  {\it (iv)} can be extended to dependent product, that is for any $A:\Type_p$, $B:\modal A \to
+  \Type_p^\modal$ and
+  $g:\prod_{a:A} B(\eta_A(a))$, there exists a map $f:\prod_{z:\modal
+    A} (B\, z)$ such that for all $a:A$, $f(\eta_A(a)) = g(a)$ \\
+ (from \cite[Theorem
+ 7.7.4]{hottbook}).
 \end{itemize}
@@
 \begin{proof}
-\nt{ça a du sens de mettre une preuve, non ? c'est non trivial et
-  nouveau par rapport aux travaux sur les modalités. Ou alors c'est
-  la preuve de Shulman qui tu n'as pas faite. En tout cas, il faut
-  clarifier.}
+% \nt{ça a du sens de mettre une preuve, non ? c'est non trivial et
+%   nouveau par rapport aux travaux sur les modalités. Ou alors c'est
+%   la preuve de Shulman qui tu n'as pas faite. En tout cas, il faut
+%   clarifier.}
 
@@
 \end{lem}
-
+%
 Here, $\sum_{x:\modal X} (\modal f x = \eta_Y y)$ is modal, thus we
@@
 contractible.
-
+%
 From definition of left exactness, one can deduce the following~:

### 2015-01-16 paper/sheaf_lics.tex
fixes in section 2 and add some biblio

@@
 direct integration into proof assistant based on type theory such as
-Coq or Agda.  
+Coq~\cite{coq:refman:8.4} or Agda~\cite{norell2007towards}.  
 %
@@
 use a generalized notion of sheafification, called Lawvere-Tierney
-sheafification~\cite{maclanemoerdijk}, but this construction has not
+sheafification~\cite{tierney1972sheaf,maclanemoerdijk}, but this construction has not
 been considered yet in the setting of higher topos theory. This rises
@@
 our presentation (also the setting using truncated types provides
-minor changes), but they can be found in our Coq form
+minor changes), but they can be found in our Coq formalization.
 
@@
 
+The following property holds for any left-exact modality $\modal$ at
+level $n$.
+
 \begin{itemize}
 
-\item $\one$ seen as a $\Type_p$ is modal, ie $\modal \one =\one$.
+\item $\one$ seen as an $n$-type is modal, ie $\modal \one =\one$.
 \item 
-  Let $A:\Type$ and $B:A \to \Type_n^\modal$. Then $P\left(\prod_{a:A} (B\,
-  a)\right)$ (from \cite[Lemma 7.7.2]{hottbook}). 
-\item For any $A,B:\Type_p$,
-  $\modal (A\times B) = (\modal A) \times (\modal B)$ \\
-(from \cite[Corollary 7.7.2]{hottbook}).
-\item Let $A:\Type_n \modal$. Then for any $x,y:X$, $P(x=y)$.
+  For any  $A:\Type$ and $B:A \to \Type_n^\modal$,  $P\left(\prod_{a:A} (B\,
+  a)\right)$. \\
+(from \cite[Lemma 7.7.2]{hottbook})
+\item For any $A,B:\Type_n$,
+  $\modal (A\times B) = (\modal A) \times (\modal B)$. \\
+(from \cite[Corollary 7.7.2]{hottbook})
+\item For any $A:\Type_n^\modal$ and $x,y:X$, $P(x=y)$.
 \item 
-  {\it (iv)} can be extended to dependent product, that is for any $A:\Type_p$, $B:\modal A \to
-  \Type_p^\modal$ and
+  {\it (iv)} can be extended to dependent product, that is for any $A:\Type_n$, $B:\modal A \to
+  \Type_n^\modal$ and
   $g:\prod_{a:A} B(\eta_A(a))$, there exists a map $f:\prod_{z:\modal
-    A} (B\, z)$ such that for all $a:A$, $f(\eta_A(a)) = g(a)$ \\
+    A} (B\, z)$ such that for all $a:A$, $f(\eta_A(a)) = g(a)$. \\
  (from \cite[Theorem
- 7.7.4]{hottbook}).
+ 7.7.4]{hottbook})
 \end{itemize}
@@
 As the definition of sheafification is based on subobject classifiers
-and thus on homotopy fibers, we need to look at how left exact
-modalities preserves homotopy fibers.
+and thus on homotopy fibers, we need to take a closer look at the way left
+exact modalities preserve homotopy fibers.
+%
 \begin{prop}

### 2015-01-16 paper/sheaf_lics.tex
continue Formalization

--- 2015-01-16 paper/sheaf_lics.tex
@@
 contains 6322 lines.
-
-\nt{describe how many line for each, be clear about what has not been formalized}
+%\nt{describe how many line for each, be clear about what has not been formalized}
+
+In our formalization, some things are admitted, either because they
+are just an adaptation of a proposition already formalized
+in~\cite{hottbook}, either because they are true, but very technical~:
+\begin{itemize}
+\item in \texttt{modalities.v}, an equality is admitted in the proof
+  of proposition~\ref{sec:defin-basic-prop}.
+\item in \texttt{nat\_lemmas.v}, the proof that $\IsHProp (n\leqslant
+  m)$ is admitted.
+\item in \texttt{sheaf\_induction.v}, in the proof that
+  $C(\mu_T)=(C\id)$ (see lemma~\ref{lem:sepiscolim}), we admit the proof
+  of equality of projections.
+\item in \texttt{sheaf\_induction.v}, we admit some very technical
+  proofs in proofs that $\square$ and $\star$ are modalities.
+\item in \texttt{sheaf\_induction.v}, cumulativity is an axiom,
+  because of universes issues.
+\end{itemize}
+
+Some other lemmas are proved, then admitted. We do it so that Coq
+doesn't constrain too much universes, without having to explicitly
+annotate universes everywhere in the formalization.
+
+The main issue in this formalization is the gestion of universes in
+Coq~\cite{sozeau2014universe}. If Coq can handle cumulativity on all
+$\Type$, it is not the case for the sigma type $\Type_n$. This issue
+can be solved by adding an axiom of cumulativity for $\Type_n$, but as
+it would not have any computational content, it would really
+complicates the proofs.
+
+Moreover, even the definition of successive sheafification types is
+not allowed~:
+
+\[ \left\{
+    \begin{array}{l}
+      \modal_{-1} = \lambda T:\HProp,~ \lnot\lnot T \\
+      \displaystyle{\modal_{n+1} = \lambda T:\Type_{n+1}},\\
+      \quad \displaystyle{\sum_{u:T \to \Type_n^\modal} \modal_{-1} 
+      \left\|
+      \sum_{a:T} u = (\lambda t,~\modal_n (a=t))
+      \right\|}
+    \end{array}
+  \right.,\]
+because Coq asks for every $\modal_n$ to have the same universes levels.
 

### 2015-01-16 paper/sheaf_lics.tex
in HoTT, mono -> embedding, epi -> surjection

@@
 theoretic definition, the correct generalization may be $n$-truncated homotopy
-fibers or monomorphims (\ie $-1$-truncated homotopy
+fibers or embeddings (\ie $-1$-truncated homotopy
 fibers) depending on the purpose of the definition.
@@
 objects for which the morphism
-(\ref{equ:sheaf_def}) is just a monomorphism, is generalized in two
+(\ref{equ:sheaf_def}) is just an embedding, is generalized in two
 different ways.  
@@
 The subobjects to be considered are given by $n$-truncated homotopy fibers,
-while the morphism (\ref{equ:sheaf_def}) is still required to be a
-monomorphism.
+while the morphism (\ref{equ:sheaf_def}) is still required to be an
+embedding.
 
@@
 $$\chi : \left(\sum_{A:\Type} \sum_{f:A \to B} \prod_{b\in B}
-\IsType \
+\IsType n\
 \fib{f}{b}\right) \xrightarrow{\sim} 
@@
 
-Finally, we need to use mere propositions has basic elements of a logic
+Finally, we need to use mere propositions as basic elements of a logic
 when characterizing properties of types or functions. 
@@
 
-\subsection{The Epi-Mono Factorization System}
+\subsection{The Orthogonal Factorization System}
 \label{sec:epi-mono-fact}
@@
 We now recall the presence of a factorization system in homotopy type
-theory, constituted by monomorphism and epimorphism.
-
-A map $f:A\to B$ is a \emph{monomorphism} when it is (-1)-truncated
-(see definition in Section~\ref{ssec:hott}), and an \emph{epimorphism}
+theory, constituted by embeddings and surjections.
+
+A map $f:A\to B$ is an \emph{embedding} when it is (-1)-truncated
+(see definition in Section~\ref{ssec:hott}), and a \emph{surjection}
 when it is (-1)-connected, that is when every fiber of $f$ is merely
@@
 %
-This particular case connects an epimorphism (or a $-1$-connected
+This particular case connects a surjection (or a $-1$-connected
 function) to the colimit of its \v{C}ech nerve.
@@
 \begin{ax}[Giraud-Rezk-Lurie]
-  For any epimorphism $f : X \to Y$, the colimit of its \v{C}ech nerve
+  For any surjection $f : X \to Y$, the colimit of its \v{C}ech nerve
   $C(f)$ is $Y$.
@@
   all dense $n$-subobject of $\iota : A \to E$,
-  $\Phi_E^\iota$ is a monomorphism. In other words, the dotted arrow,
+  $\Phi_E^\iota$ is an embedding. In other words, the dotted arrow,
   if exists, is unique.
@@
   \right.,\]
-because Coq asks for every $\modal_n$ to have the same universes levels.
+because Coq asks for every $\modal_n$ to have the same universes
+levels. Thus, the induction presented in this paper cannot be
+formalized at the moment. We only present the induction step.
 

### 2015-01-16 paper/sheaf_lics.tex
pass on section 5.1

--- 2015-01-16 paper/sheaf_lics.tex
@@
 
-\newcommand{\IsType}[1]{\mathop{\mathrm{Is\mymathhyphen}#1\mathrm{\mymathhyphen
-      type}} }
+\newcommand{\IsType}[1]
+{\mathop{\mathrm{Is\mymathhyphen}#1\mathrm{\mymathhyphen type}} }
 
@@
 \newcommand \squash[1] {\| #1 \| }
+\newcommand \separated {\mathop{\square} }
 \newcommand \fib[2] {\mathrm{fib}_{#1}(#2)}
@@
 
-The central result used here is the following
+The proof is based on the following central result.
 \begin{lem}
@@
 %
-Here, $\sum_{x:\modal X} (\modal f x = \eta_Y y)$ is modal, thus we
+As $\sum_{x:\modal X} (\modal f x = \eta_Y y)$ is modal, we
 can define $$\phi:\modal \left(\sum_{x:X}  (f x = y)\right) \to
@@
 \kq{Finish that}
+
+\nt{Indeed, I can't follow the proof for the moment}
 \end{proof}
@@
 \begin{defi}
-  For any type $A$ and $E$, $\iota : A -> E$ and $F:\Type_{n+1}$, we define
+  For any type $A$ and $E$, $\iota : A \to E$ and $F:\Type_{n+1}$, we define
     $$
@@
 
-\nt{Il faut mieux introduire ces propriétés, $\Type_n^\modal$ is a sheaf
-  est très importante en particulier. Il faut qu'on comprenne d'où ça
-  vient et que tu fasses une preuve, même un peu idalisée. Je dirais
-  que tu peux mettre des lemmes genres nj\_inter en disant que telle
-  flèche est un iso, and co.
-  En tout cas, cette partie peut et doit être bien plus longue.
-}
-
-\begin{prop}\label{prop:sheaf-forall}
-  If $A:\Type_{n+1}$ and $B:A \to \Type_{n+1}$ such that for any
-  $a:A$, $(B~a)$ is a sheaf, then $\prod_{a:A}B~a$ is a sheaf.
-\end{prop}
+A fundamental property in the construction of the sheafification is
+the fact that the type of modal types at level $n$ defines a sheaf. 
+%
+This property initially holds in the topos theoretic construction, but
+it requires much work here, as we have to deal with equalities. In
+particular, this property uses in particular the specific definition
+of separated types and sheaves, which shows that we had actually no
+choice in their statement.
 
@@
 \begin{proof}
-  Firstly, we prove that $\Type_n^\modal$ is separated. Let $E:\Type$
+  First, we prove that $\Type_n^\modal$ is separated. Let $E:\Type$
   and $\chi:E \to \Type_n$, dense in $E$. Let $\phi_1,\phi_2:E \to
   \Type_n^\modal$, such that $\phi_1 \circ \pi_1 = \phi_2 \circ
-  \pi_1$. Finally, let $x:E$. We show $\phi_1 x = \phi_2 x$ using
+  \pi_1$ and let $x:E$. We show $\phi_1 x = \phi_2 x$ using
   univalence.
@@
   As $\chi$ is dense, we have a term $m_x : \modal(\chi\, x)$.
-  But as $\phi_2 x$ is modal, thus we can obtain a term $h_x : \chi\,
+  But as $\phi_2 x$ is modal, we can obtain a term $h_x : \chi\,
   x$. 
@@
 
+An other fundamental property on sheaves is that the type of (dependent)
+functions is a sheaf as soon as its codomain is a sheaf.
+
+\begin{prop}\label{prop:sheaf-forall}
+  If $A:\Type_{n+1}$ and $B:A \to \Type_{n+1}$ such that for any
+  $a:A$, $(B~a)$ is a sheaf, then $\prod_{a:A}B~a$ is a sheaf.
+\end{prop}
+\begin{proof}
+  \nt{add a short proof here}
+\end{proof}
+
 \section{Construction of sheafification}
@@
 From any left exact modality $\modal_{-1}$ on $\HProp$, we define by
-induction the truncation index $n$, a
+induction on the truncation index $n$, a
 left exact modality $\modal_n$ at level $n$ that extends $\modal_{-1}$ in the
@@
 
-Note that once the first modality is defined, the extension to all
-$\Type_n$ is automatic. This means that the new logical principles
-that can be added by the sheaf construction must be added in the
-first modality already. The sheafification is just a way to propagate
-those new principles to other types than just mere proposition.
+Note that once the first modality $\modal_{-1}$ is defined, its
+extension to all $\Type_n$ is automatic. This means that the new
+logical principles that can be added by the sheaf construction must be
+added in the first modality already. Sheafification is just a way
+to propagate those new principles to types than are not mere
+propositions.
 
@@
 
-Let $T : \Type_{n+1}$. We define $\square T$ as the image (see
+Let $T : \Type_{n+1}$. We define $\separated T$ as the image (see
 Section~\ref{sec:epi-mono-fact}) of
@@
   T \ar[r]^{\{\cdot\}_T} \ar[d]_{\mu_T} & \Type_n^T \ar[d]^{\modal^T} \\
-  \square T \ar[r]& \left( \Type_n^\modal \right)^T
+  \separated T \ar[r]& \left( \Type_n^\modal \right)^T
 }, $$
@@
 (t':T),~t=t'$.
-\nt{parler du système de factorization}
-In type theory words, 
-\begin{align*}
-\square T &= \im (\lambda~t:T,~\lambda~ t',~ \modal (t = t')) \\
-          &= \sum_{u:T \to \Type_n^\modal} \left\| \sum_{a:X} 
-            (\lambda t,~\modal (a=t)) = u\right\|
-\end{align*}
-
-There again, the separation step has the same definition as
-in~\cite{maclanemoerdijk}, using $\Type_n^\modal$ instead of the
-$j$-subobject classifier.
-
+%
+$\separated T$ can be given explicitly by
+%
+$$
+\begin{array}{rcl}
+\separated T &\defeq & \im (\lambda~t:T,~\lambda~ t',~ \modal (t = t')) \\
+          & \defeq & \sum_{u:T \to \Type_n^\modal} \left\| \sum_{a:X} 
+            (\lambda t,~\modal (a=t)) = u\right\|.
+\end{array}
+$$
+%
+This corresponds to the free separated object used in the topos
+theoretic construction, but using $\Type_n^\modal$ instead of the
+$j$-subobject classifier $\Omega_j$.
+%
 \begin{prop}
-  For any $T:\Type_{n+1}$, $\square T$ is separated.  
+  For any $T:\Type_{n+1}$, $\separated T$ is separated.  
 \end{prop}
@@
 \begin{proof}
-We will use the following lemma~:
+We use the following lemma:
 \begin{lem}
@@
 \end{lem}
-As $\square T$ embeds in $\left( \Type_n^\modal \right)^T$, we only
+As $\separated T$ embeds in $\left( \Type_n^\modal \right)^T$, we only
 have to show that the latter is separated. But it is the case because
@@
 
-We now need to show that $\square$ is universal. The following
-lemma is central in the proof~:
+To show that $\separated$ is a modality, we need t show that $\separated T$
+is universal among separated type below $T$. This comes from the
+following central lemma which connects $\separated T$ to the colimit of
+the \v{C}ech nerve of $\mu_T$.
+
 \begin{lem}\label{lem:sepiscolim}
-  Let $T:\Type_{n+1}$. Then $\square T$ is the colimit of the closed
+  Let $T:\Type_{n+1}$. Then $\separated T$ is the colimit of the closed
   diagonal diagram
@@
 
-This lemma is an adaptation of the sheafification process
-in~\cite{maclanemoerdijk}, where they consider only the kernel pair of
-$\mu_T$ instead of the \v{C}ech nerve.
+Again, this lemma is an adaptation of the sheafification process in
+topos theory, where the kernel pair of $\mu_T$
+is replaced by its \v{C}ech nerve.
 
 \begin{proof}
-  As $\mu_T$ is an surjection, using the particular case of the last
+  As $\mu_T$ is an surjection, using the particular case of 
   Giraud-Rezk-Lurie axiom introduced in Section~\ref{sec:giraud-ax},
-  we know that $\square T$ is the colimit
-  of $C(\mu_T)$. If we can show that $C(\mu_T) = C(\id)$, the the
-  result will follow.
-
-  Let $k:\nat$, let's show that 
+  we know that $\separated T$ is the colimit of $C(\mu_T)$. 
+%
+  Thus the result follows from the fact that $C(\mu_T) = C(\id)$,
+  which we now show.
+
+  Let $k:\nat$, we will show that 
   \begin{align*}
@@
 \item Suppose now $p : \modal(a=b)$. Let $\iota$ be the first
-  projection from $\square T \to (T \to \Type_n^\modal)$. $\iota$ is
+  projection from $\separated T \to (T \to \Type_n^\modal)$. $\iota$ is
   an embedding, thus it suffices to prove $\iota (\mu_T a) = \iota
@@
   }$$
-But we know (lemma~\ref{lem:sepiscolim}) that $\square T$ is the
+But we know (lemma~\ref{lem:sepiscolim}) that $\separated T$ is the
 colimit of the closed diagonals diagram, thus there is an universal
-arrow $\square T \to Q$.
-
-This proves the following proposition
+arrow $\separated T \to Q$.
+%
+This is enough to state the following proposition.
 \begin{prop}
-  $(\square,\mu)$ defines a reflective subuniverse on $\Type_{n+1}$.
+  $(\separated,\mu)$ defines a reflective subuniverse on $\Type_{n+1}$.
 \end{prop}
 
-We now have to prove point {\it (v)} in the definition of
-modality. Let $A:\Type_{n+1}$ be a sheaf and $B:A \to \Type_{n+1}$ be
+To show that $\separated$ is a modality, it remains to show point {\it (v)}. Let $A:\Type_{n+1}$ be a sheaf and $B:A \to \Type_{n+1}$ be
 a sheaf family. We want to show that $\sum_{x:A} (Bx)$ is
@@
 \begin{prop}
-  $(\square,\mu)$ defines a modality on $\Type_{n+1}$.
+  $(\separated,\mu)$ defines a modality on $\Type_{n+1}$.
 \end{prop}
 
-As this modality is just a step in the construction, we will not show
-that it is left exact, we will do it only for the sheafification modality.
+As this modality is just a step in the construction, we do not show
+that it is left exact, we will only do it for the sheafification modality.
 
@@
   Because $Q$ is a sheaf, it is in particular separated ; thus we can
-  extend $f$ to $\square f:\square T\to Q$.
-
-  But as $\star T$ is the closure of $\square T$, $\square T$ is dense
+  extend $f$ to $\separated f:\separated T\to Q$.
+
+  But as $\star T$ is the closure of $\separated T$, $\separated T$ is dense
   into $\star T$, so the sheaf property of $Q$ allows to extend
-  $\square f$ to $\star f:\star T \to Q$.
+  $\separated f$ to $\star f:\star T \to Q$.
 
@@
 contains 6322 lines.
-%\nt{describe how many line for each, be clear about what has not been formalized}
-
-In our formalization, some things are admitted, either because they
-are just an adaptation of a proposition already formalized
-in~\cite{hottbook}, either because they are true, but very technical~:
-\begin{itemize}
-\item in \texttt{modalities.v}, an equality is admitted in the proof
-  of proposition~\ref{sec:defin-basic-prop}.
-\item in \texttt{nat\_lemmas.v}, the proof that $\IsHProp (n\leqslant
-  m)$ is admitted.
-\item in \texttt{sheaf\_induction.v}, in the proof that
-  $C(\mu_T)=(C\id)$ (see lemma~\ref{lem:sepiscolim}), we admit the proof
-  of equality of projections.
-\item in \texttt{sheaf\_induction.v}, we admit some very technical
-  proofs in proofs that $\square$ and $\star$ are modalities.
-\item in \texttt{sheaf\_induction.v}, cumulativity is an axiom,
-  because of universes issues.
-\end{itemize}
-
-Some other lemmas are proved, then admitted. We do it so that Coq
-doesn't constrain too much universes, without having to explicitly
-annotate universes everywhere in the formalization.
-
-The main issue in this formalization is the gestion of universes in
-Coq~\cite{sozeau2014universe}. If Coq can handle cumulativity on all
-$\Type$, it is not the case for the sigma type $\Type_n$. This issue
-can be solved by adding an axiom of cumulativity for $\Type_n$, but as
-it would not have any computational content, it would really
-complicates the proofs.
-
-Moreover, even the definition of successive sheafification types is
-not allowed~:
-
-\[ \left\{
-    \begin{array}{l}
-      \modal_{-1} = \lambda T:\HProp,~ \lnot\lnot T \\
-      \displaystyle{\modal_{n+1} = \lambda T:\Type_{n+1}},\\
-      \quad \displaystyle{\sum_{u:T \to \Type_n^\modal} \modal_{-1} 
-      \left\|
-      \sum_{a:T} u = (\lambda t,~\modal_n (a=t))
-      \right\|}
-    \end{array}
-  \right.,\]
-because Coq asks for every $\modal_n$ to have the same universes levels.
+
+\nt{describe how many line for each, be clear about what has not been formalized}
 

### 2015-01-16 paper/sheaf_lics.tex

@@
 contains 6322 lines.
-
-\nt{describe how many line for each, be clear about what has not been formalized}
+%\nt{describe how many line for each, be clear about what has not been formalized}
+
+In our formalization, some things are admitted, either because they
+are just an adaptation of a proposition already formalized
+in~\cite{hottbook}, either because they are true, but very technical~:
+\begin{itemize}
+\item in \texttt{modalities.v}, an equality is admitted in the proof
+  of proposition~\ref{sec:defin-basic-prop}.
+\item in \texttt{nat\_lemmas.v}, the proof that $\IsHProp (n\leqslant
+  m)$ is admitted.
+\item in \texttt{sheaf\_induction.v}, in the proof that
+  $C(\mu_T)=(C\id)$ (see lemma~\ref{lem:sepiscolim}), we admit the proof
+  of equality of projections.
+\item in \texttt{sheaf\_induction.v}, we admit some very technical
+  proofs in proofs that $\square$ and $\star$ are modalities.
+\item in \texttt{sheaf\_induction.v}, cumulativity is an axiom,
+  because of universes issues.
+\end{itemize}
+
+Some other lemmas are proved, then admitted. We do it so that Coq
+doesn't constrain too much universes, without having to explicitly
+annotate universes everywhere in the formalization.
+
+The main issue in this formalization is the gestion of universes in
+Coq~\cite{sozeau2014universe}. If Coq can handle cumulativity on all
+$\Type$, it is not the case for the sigma type $\Type_n$. This issue
+can be solved by adding an axiom of cumulativity for $\Type_n$, but as
+it would not have any computational content, it would really
+complicates the proofs.
+
+Moreover, even the definition of successive sheafification types is
+not allowed~:
+
+\[ \left\{
+    \begin{array}{l}
+      \modal_{-1} = \lambda T:\HProp,~ \lnot\lnot T \\
+      \displaystyle{\modal_{n+1} = \lambda T:\Type_{n+1}},\\
+      \quad \displaystyle{\sum_{u:T \to \Type_n^\modal} \modal_{-1} 
+      \left\|
+      \sum_{a:T} u = (\lambda t,~\modal_n (a=t))
+      \right\|}
+    \end{array}
+  \right.,\]
+because Coq asks for every $\modal_n$ to have the same universes
+levels. Thus, the induction presented in this paper cannot be
+formalized at the moment. We only present the induction step.
 

### 2015-01-16 paper/sheaf_lics.tex
formalization update lines

--- 2015-01-16 paper/sheaf_lics.tex
@@
 \url{https://github.com/HoTT/HoTT}
-
-Formalisation of subobject classifiers has been done in
+\begin{itemize}
+\item Formalisation of subobject classifiers has been done in
 \texttt{sub\_object\_classifier.v}, 284 lines.
-
+\item 
 Colimits, $p$-pullbacks and \v{C}ech nerves has been formalized in
 \texttt{colimits.v, cech\_nerve.v}, 552 lines.
-
+\item 
 Reflective subuniverses and modalities are formalized in
 \texttt{reflective\_subuniverse.v, modalities.v}, 1053 lines.
-
+\item 
 Definition of th reflective subuniverse on $\HProp$ is given in
 \texttt{sheaf\_base\_case.v}, 186 lines.
-
+\item 
 Section~\ref{sec:sheaves} is formalized in
 \texttt{sheaf\_def\_and\_thm.v}, 1029 lines.
-
+\item 
 Section~\ref{sec:sheafification} is formalized in
 \texttt{sheaf\_induction.v}, 2340 lines.
-
+\end{itemize}
 At total, with other files containing technical lemmas, the project

### 2015-01-16 paper/sheaf_lics.tex
pass on section 6

@@
 \begin{defi}
-  $\IsType n$ is defined by induction on $n\geqslant -2$~:
+  $\IsType n$ is defined by induction on $n\geqslant -2$:
   \begin{itemize}
@@
 \im(f)$ is (-1)-connected, thus the following proposition holds
-(\cite{hottbook}, Lemma 7.6.4)~:
+(\cite{hottbook}, Lemma 7.6.4):
 \begin{prop}
@@
 and any map $f:X \to Y$, the modalisation of fiber of $f$ above any element $y:Y$
-is the fiber of $\modal f$ above $\eta_Y y$~:
+is the fiber of $\modal f$ above $\eta_Y y$:
 $$\modal \left(\sum_{x:X}  (f x = y)\right) = \sum_{x:\modal X} (\modal f x = \eta_Y y).$$
@@
 %
-From definition of left exactness, one can deduce the following~:
+From definition of left exactness, one can deduce the following:
 \begin{lem}
@@
 
-But we have the lemma~:
+But we have the lemma:
 \begin{lem}
@@
   % \item if $A \to E$ with $n$-truncated homotopy fibers, $\Phi_E^\chi$ is the
-  %   map sending $f~:~E \to f$ to $\restriction f A$.
+  %   map sending $f:~E \to f$ to $\restriction f A$.
   % \end{itemize}
@@
 %
-% We have to be careful~: if we want to propagate the properties of the
+% We have to be careful: if we want to propagate the properties of the
 % first modality $\modal_{-1}$, all the higher modalities $\modal_n$ must be compatible
@@
 \begin{enumerate}[(i)]
-\item {\em separation~:} In this step, we {\em remove} the lack of
+\item {\em separation:} In this step, we {\em remove} the lack of
   sheafness from the type we are considering.
-\item {\em completion~:} In this step, we add what lacks now to the
+\item {\em completion:} In this step, we add what lacks now to the
   separated type to be a sheaf.
@@
 % For the case $n=-1$, one can take any left exact modality on
-% $\HProp$. Here we took the $\lnot\lnot$ modality~:
+% $\HProp$. Here we took the $\lnot\lnot$ modality:
 % $$\forall P:\HProp,~\modal_j P = \lnot\lnot P.$$
@@
 $\sum_{y:E} x = y$ has a dense subobject, $\sum_{y:\sum_{e:E} (\chi\,
-  e)} x=y_1$~:
+  e)} x=y_1$:
 
@@
 If $T$ is already a separated type, the following lemma alllows us
-to build a sheaf~:
+to build a sheaf:
 \begin{lem}
@@
 
-The last thing to prove is the cumulativity of our sheafifications~:
+The last thing to prove is the cumulativity of our sheafifications:
 \begin{prop}
@@
 A Coq formalization is available at
-\url{https://github.com/KevinQuirin/sheafification/}.
+\url{https://github.com/KevinQuirin/sheafification}.
 %
 It is based on the homotopy type theory library available at
-\url{https://github.com/HoTT/HoTT}
+\url{https://github.com/HoTT/HoTT}.
+
+We provide a more detailed insight of where properties have been
+formalized, and what is the size of the formalisation for each part:
 \begin{itemize} 
-\item Formalisation of subobject classifiers has been done in
-\texttt{sub\_object\_classifier.v}, 284 lines.
-\item 
-Colimits, $p$-pullbacks and \v{C}ech nerves has been formalized in
-\texttt{colimits.v, cech\_nerve.v}, 552 lines.
+\item  Homotopy fibers and subobject classifiers are formalized in
+\texttt{sub\_object\_classifier.v} (284 lines).
+%
+\item Colimits, $p$-pullbacks and \v{C}ech nerves are formalized in
+\texttt{colimits.v, cech\_nerve.v} (552 lines).
 \item
 Reflective subuniverses and modalities are formalized in
-\texttt{reflective\_subuniverse.v, modalities.v}, 1053 lines.
+\texttt{reflective\_subuniverse.v, modalities.v} (1053 lines).
 \item 
-Definition of th reflective subuniverse on $\HProp$ is given in
-\texttt{sheaf\_base\_case.v}, 186 lines.
+%
+  The definition of the dense topology as a left exact modality on
+  $\HProp$ is given in \texttt{sheaf\_base\_case.v} (186 lines).
 \item
 Section~\ref{sec:sheaves} is formalized in
-\texttt{sheaf\_def\_and\_thm.v}, 1029 lines.
+\texttt{sheaf\_def\_and\_thm.v} (1029 lines).
 \item
 Section~\ref{sec:sheafification} is formalized in
-\texttt{sheaf\_induction.v}, 2340 lines.
+\texttt{sheaf\_induction.v} (2340 lines).
 \end{itemize}
-At total, with other files containing technical lemmas, the project
-contains 6322 lines.
-%\nt{describe how many line for each, be clear about what has not been formalized}
-
-In our formalization, some things are admitted, either because they
-are just an adaptation of a proposition already formalized
-in~\cite{hottbook}, either because they are true, but very technical~:
-\begin{itemize}
-\item in \texttt{modalities.v}, an equality is admitted in the proof
-  of proposition~\ref{sec:defin-basic-prop}.
-\item in \texttt{nat\_lemmas.v}, the proof that $\IsHProp (n\leqslant
-  m)$ is admitted.
-\item in \texttt{sheaf\_induction.v}, in the proof that
-  $C(\mu_T)=(C\id)$ (see lemma~\ref{lem:sepiscolim}), we admit the proof
-  of equality of projections.
-\item in \texttt{sheaf\_induction.v}, we admit some very technical
-  proofs in proofs that $\square$ and $\star$ are modalities.
-\item in \texttt{sheaf\_induction.v}, cumulativity is an axiom,
-  because of universes issues.
-\end{itemize}
-
-Some other lemmas are proved, then admitted. We do it so that Coq
-doesn't constrain too much universes, without having to explicitly
-annotate universes everywhere in the formalization.
-
-The main issue in this formalization is the gestion of universes in
-Coq~\cite{sozeau2014universe}. If Coq can handle cumulativity on all
-$\Type$, it is not the case for the sigma type $\Type_n$. This issue
-can be solved by adding an axiom of cumulativity for $\Type_n$, but as
-it would not have any computational content, it would really
-complicates the proofs.
-
-Moreover, even the definition of successive sheafification types is
-not allowed~:
-
+
+Overall, with other files containing technical lemmas, the project
+contains 6322 lines, and it could be reduced a bit by improving the
+way Coq tries to rewrite and apply lemmas automatically. 
+%
+This constitute a significant amount of work but the part dedicated to
+sheaves and sheafification is only 4000, which seems quite reasonable
+and is encouraging, because it suggests that homotopy type theory
+provides a convenient tool to formalize some part of the theory of
+higher topoi.
+
+% In our formalization, some things are admitted, either because they
+% are just an adaptation of a proposition already formalized
+% in~\cite{hottbook}, either because they are true, but very technical:
+% \begin{itemize}
+% \item in \texttt{modalities.v}, an equality is admitted in the proof
+%   of proposition~\ref{sec:defin-basic-prop}.
+% \item in \texttt{nat\_lemmas.v}, the proof that $\IsHProp (n\leqslant
+%   m)$ is admitted.
+% \item in \texttt{sheaf\_induction.v}, in the proof that
+%   $C(\mu_T)=(C\id)$ (see lemma~\ref{lem:sepiscolim}), we admit the proof
+%   of equality of projections.
+% \item in \texttt{sheaf\_induction.v}, we admit some very technical
+%   proofs in proofs that $\square$ and $\star$ are modalities.
+% \item in \texttt{sheaf\_induction.v}, cumulativity is an axiom,
+%   because of universes issues.
+% \end{itemize}
+
+In our formalization of sheafification defined
+Section~\ref{sec:sheafification} (done in 
+\texttt{sheaf\_induction.v}), five very technical facts required to
+show that $\square$ and $\modal_{n+1}$ are modalities have been
+admitted. Nevertheless, we are confident that they can be proven and
+have been admitted only by lack of time.
+
+More importantly, there are few axioms in the formalization that are
+due to an unsatisfactory management of polymorphic universe level in
+Coq~\cite{sozeau2014universe}.
+%
+\nt{Say a word about polymorphic universe level, consistency check,
+  and the fact that polymorphic universe level are nevertheless
+  mandatory for our formalization}
+
+%
+If Coq handles cumulativity on all $\Type$ natively, it is not
+the case for the $\Sigma$-type $\Type_n$, which require propositional
+resizing. 
+%
+This issue could be solved by adding an axiom of cumulativity
+for $\Type_n$ with an explicit management of universes. 
+%
+But as it would not have any computational content, it
+would really complicates the proofs as the axiom would appear
+everywhere cumulativity is needed and it would need explicitly
+annotation for universe levels everywhere in the formalization.
+
+We have chosen a more simple but less satisfactory solution. 
+%
+When the absence of cumulativity rises an universe issue for a lemma,
+we have proven the lemma and then admit the exact same lemma, just to
+let Coq infer less constrained universe levels.
+
+An other issue with universe polymorphism lies in their management for
+recursive definitions. Indeed, the definition of successive sheafification types is
+not allowed:
+%
 \[ \left\{
@@
     \end{array}
-  \right.,\]
+  \right.\]
+%
 because Coq asks for every $\modal_n$ to have the same universes
-levels. Thus, the induction presented in this paper cannot be
-formalized at the moment. We only present the induction step.
-
-Talk about issues in universe levels \cite{sozeau2014universe}. Lake of cumulativity for
-$\Type_n$ and impossibility to define by induction the sheafification 
-because universe levels must be the same for every $n$.
+levels in the recursive definition. 
+%
+Thus, the induction step presented in this paper has been formalized,
+but the complete induction can not be formalized at the moment. 
+%
+This restriction in our formalization can not be solved without
+generalizing universe polymorphism for recursive definition.
+%
+% In particular, compoatcumulativity is an axiom, because of
+% universes issues.
 

### 2015-01-16 paper/sheaf_lics.tex
add proof of depprod, compatibility, left exactness + typos

@@
 %
-As $\sum_{x:\modal X} (\modal f x = \eta_Y y)$ is modal, we
-can define $$\phi:\modal \left(\sum_{x:X}  (f x = y)\right) \to
-\sum_{x:\modal X} (\modal f x = \eta_Y y).$$
-We just need to check that the image of its fibers by $\modal$ are
+It is straighforward to define a map
+$$\phi:\sum_{x:X}  (f x = y)\to
+\sum_{x:\modal X} (\modal f x = \eta_Y y),$$
+using $\eta$ functions.
+We just need to check that every $\modal(\fib \phi {x;p})$ is
 contractible.
+Technical transformations allow to prove
+$$\fib\phi{x;p} = \fib s {y;p^{-1}}$$
+for some $a:\modal X$, $b:\modal Y$ and 
+$$s:\fib{\eta_X}a \to \fib{\eta_Y}b.$$
 %
@@
 \end{lem}
-We can rewrite $\fib \phi {x;p}$ in terms of the fiber of a fonction
-$\fib {\eta_X} a \to \fib {\eta_Y} b$.
-
+Thus, we just need to prove that $\modal(\fib {\eta_X} a)$ and
+$\modal(\fib {\eta_Y} b)$ are contractible.
 But we have the lemma:
@@
 
-Finally, $\modal(\fib \phi {x;p})$ is contractible, and the result is proved.
-\kq{Finish that}
-
-\nt{Indeed, I can't follow the proof for the moment}
+Finally, $\modal(\fib s{y;p^{-1}})$ is contractible, so $\modal(\fib \phi {x;p})$ also, and the result is proved.
+%\kq{Finish that}
+
+%\nt{Indeed, I can't follow the proof for the moment}
 \end{proof}
@@
 every type in the new universe is collapsed to an $\HProp$, ending up
-with porpositional classical logic.
+with propositional classical logic.
 
@@
 \begin{proof}
-  \nt{add a short proof here}
+  \begin{itemize}
+  \item {\em Separation:} Let $E:\Type$, $\chi:E \to \Type_n$ dense
+    and $\phi_1,\phi_2:E\to \prod_{a:A} (B\, a)$ equal on
+    $\sum_{e:E}(\chi\, e)$ \ie $\phi_1\circ \pi_1 = \phi_2\circ
+    \pi_1$.
+    Then for any $a:A$, $(\lambda x:E,~\phi_1\, x\,
+    a)$
+    and $(\lambda x:E,~\phi_2\, x\, a)$
+    coincide on $\sum_{e:E}(\chi\, e)$, and as $B\, a$ is separated,
+    they coincide also on all $E$.
+  \item {\em Sheaf:} Let $E:\Type$, $\chi:E\to \HProp$ dense and
+    $f:\sum_{e:E} (\chi\, e)\to \prod_{a:A} (B\,a)$. Let $a:A$ ; the
+    map $(\lambda x,~f\,x \,a)$ is valued in the sheaf $B\, a$, so it
+    can be extended to all $E$, allowing $f$ to be extended to all $E$.
+  \end{itemize}
+%  \nt{add a short proof here}
 \end{proof}
@@
 \nt{je ne comprends pas ce passage, c'est quoi sheafness ? }
+
+\kq{c'est une vision plutôt grothendieck en fait. La separation ``enlève''
+  ce qui empêche à $T$ d'être un faisceau, et la sheafification
+  rajoute ce qui manque pour être un faisceau. Mais c'était plutôt mal
+  dit, effectivement}
+
 For a given induction step, the sheafification is defined in two
@@
 \begin{enumerate}[(i)]
-\item {\em separation:} In this step, we {\em remove} the lack of
-  sheafness from the type we are considering.
+\item {\em separation:} In this step, we construct a {\em free}
+  separated object.
 \item {\em completion:} In this step, we add what lacks now to the
@@
 Section~\ref{sec:epi-mono-fact}) of
-$\modal^T \circ \{\cdot\}_T$, as in
+$\modal_n^T \circ \{\cdot\}_T$, as in
 $$\xymatrix{
-  T \ar[r]^{\{\cdot\}_T} \ar[d]_{\mu_T} & \Type_n^T \ar[d]^{\modal^T} \\
+  T \ar[r]^{\{\cdot\}_T} \ar[d]_{\mu_T} & \Type_n^T \ar[d]^{\modal_n^T} \\
   \separated T \ar[r]& \left( \Type_n^\modal \right)^T
@@
 \begin{array}{rcl}
-\separated T &\defeq & \im (\lambda~t:T,~\lambda~ t',~ \modal (t = t')) \\
-          & \defeq & \sum_{u:T \to \Type_n^\modal} \left\| \sum_{a:X} 
-            (\lambda t,~\modal (a=t)) = u\right\|.
+\separated T &\defeq & \im (\lambda~t:T,~\lambda~ t',~ \modal_n (t = t')) \\
+          & \defeq & \sum_{u:T \to \Type_n^\modal} \left\| \sum_{a:T} 
+            (\lambda t,~\modal_n (a=t)) = u\right\|.
 \end{array}
@@
 
-To show that $\separated$ is a modality, we need t show that $\separated T$
+To show that $\separated$ is a modality, we need to show that $\separated T$
 is universal among separated type below $T$. This comes from the
@@
       \land \mu_T t_{k-1} \mu_T t_k)\\
-    &= \sum_{t:T^k} \modal (t_1 = t_2 \land \cdots
+    &= \sum_{t:T^k} \modal_n (t_1 = t_2 \land \cdots
       \land t_{k-1} = t_k)
   \end{align*}
-By induction on $k$, and the preservation of products by $\modal$, it
-suffices to show that for any $a,b:T$, $\modal (a=b) = (\mu_T a =
+By induction on $k$, and the preservation of products by $\modal_n$, it
+suffices to show that for any $a,b:T$, $\modal_n (a=b) = (\mu_T a =
 \mu_T b)$. By univalence, we want arrows in both ways, forming an
@@
 \item Suppose $p : (\mu_T a = \mu_T b)$. Then projecting $p$ along
-  first components yields $q : \prod_{t:T} \modal(a=t) = \modal (b=t)
+  first components yields $q : \prod_{t:T} \modal_n(a=t) = \modal_n (b=t)
   $.
-  Taking for example $t=b$, we deduce $\modal (a=b) = \modal(a=a)$,
+  Taking for example $t=b$, we deduce $\modal_n (a=b) = \modal_n(a=a)$,
   and the latter is inhabited by $\eta_{a=a} 1$.
-\item Suppose now $p : \modal(a=b)$. Let $\iota$ be the first
+\item Suppose now $p : \modal_n(a=b)$. Let $\iota$ be the first
   projection from $\separated T \to (T \to \Type_n^\modal)$. $\iota$ is
   an embedding, thus it suffices to prove $\iota (\mu_T a) = \iota
-  (\mu_T b)$, \ie $\prod_{t:T}\modal (a=t) = \modal (b=t)$. The latter
+  (\mu_T b)$, \ie $\prod_{t:T}\modal_n (a=t) = \modal_n (b=t)$. The latter
   remains true by univalence.
@@
 This is enough to state the following proposition.
-\begin{prop}
+\begin{prop}\label{prop:sep-subu}
   $(\separated,\mu)$ defines a reflective subuniverse on $\Type_{n+1}$.
@@
 This proves the following proposition
-\begin{prop}
+\begin{prop}\label{prop:sep-mod}
   $(\separated,\mu)$ defines a modality on $\Type_{n+1}$.
@@
 
-\nt{put the definition of $\modal_{n+1}$ upfront}
-
-If $T$ is already a separated type, the following lemma alllows us
-to build a sheaf:
+%\nt{put the definition of $\modal_{n+1}$ upfront}
+If $T:\Type_{n+1}$, let  $\modal_{n+1}T$ be the cloture of $\square T$
+seen as a subobject of $T \to \Type_n^\modal$. Explicitly
+$$\modal_{n+1} T = \sum_{u:T \to \Type_n^\modal} \modal_{-1}\left\| \sum_{a:T} 
+            (\lambda t,~\modal_n (a=t)) = u\right\|$$
+
+We begin by proving that $\modal_{n+1}$ is compatible with $\modal_n$
+on lower homotopy types.
+\begin{prop}
+  If $T:\Type_n$, then $\modal_{n+1}T = \modal_n T$.
+\end{prop}
+\begin{proof}
+  We prove it by induction on $n$~:
+  \begin{itemize}
+  \item For $n=-1$: Let $T:\HProp$. Then
+    \begin{align*}
+      \modal_{0} T &= \sum_{u:T \to \Type_n^\modal} \modal_{-1}\left\| \sum_{a:T} 
+      (\lambda t,~\modal_{-1} (a=t)) = u\right\| \\
+      &= \sum_{u:T \to \Type_n^\modal} \modal_{-1}\left( \sum_{a:T} 
+      (\lambda t,~\modal_{-1} (a=t)) = u\right)
+    \end{align*}
+    because the type inside the truncation is already in $\HProp$.
+    Now, let define $\phi : \modal_{-1} T \to \modal_0T$ by
+    $$\phi t = (\lambda t',\, \modal_{-1}(t=t') ; \eta
+    (t;\mathrm{idpath}))$$
+    and $\psi : \modal_0T\to \modal_{-1} T$ by obtaining the
+    witness $a:T$, and letting
+    $$\psi (u;x) = \eta_T a.$$
+    These two maps form an equivalence.
+  \item Suppose now that $\modal_{n+1}$ is compatible with all $\modal_k$ on
+    lower homotopy types. Let $\modal_{n+2}$ be as above, and let
+    $T:\Type_{n+1}$. Then
+    \begin{align*}
+      \modal_{n+2} T &= \sum_{u:T \to \Type_{n+1}^\modal} \modal_{-1}\left\| \sum_{a:T} 
+            (\lambda t,~\modal_{n+1} (a=t)) = u\right\| \\
+                     &= \sum_{u:T \to \Type_{n+1}^\modal} \modal_{-1}\left\| \sum_{a:T} 
+            (\lambda t,~\modal_{n} (a=t)) = u\right\|
+    \end{align*}
+    because $\modal_{n+1}$ is compatible with $\modal_{n}$, and
+    $(a=t):\Type_n$.
+
+    It lefts to prove that $u:T\to\Type_n^\modal$, \ie that for all
+    $t:T$, $\IsType n (u\, t)$.
+    But for any truncation index $p$, the predicate $\IsType p:\HProp$
+    is a a sheaf, so we can get rid of $\modal_{-1}$, and pick up what
+    is inside the truncation. Thus, we know that for all $t:T$, $u\, t
+    = \modal_n(a=t) : \Type_n$.
+  \end{itemize}
+\end{proof}
+
+This also proves that $\modal_{n+1}$ is compatible with $\modal_{-1}$
+in the sense of proposition~\ref{prop:hprop}.
+
+To prove that $\modal_{n+1} T$ is a sheaf for any $T:\Type_{n+1}$, we
+use the following:
 \begin{lem}
-  Let $T:\Type_{n+1}$ be separated, and $U$ be a sheaf. If $T$ embeds
-  in $U$, and is closed in $U$, then $T$ is a sheaf.
+  Let $X:\Type_{n+1}$ be separated, and $U$ be a sheaf. If $X$ embeds
+  in $U$, and is closed in $U$, then $X$ is a sheaf.
 \end{lem}
 
-As any separated type $T$ embeds in $\left(\Type_n^\modal\right)^T$,
-it suffices to take the closure of $T$ to get a sheaf, and the unit
-$\nu_T:T \to \star T$ is obvious.
+As $T\to \Type_n^\modal$ is a sheaf, and any object is closed into its
+cloture, $\modal_{n+1}T$ is indeed a sheaf.
+
 \begin{prop}
-  $(\star,\nu)$ defines a reflective subuniverse.
+  $(\modal_{n+1},\nu)$ defines a reflective subuniverse.
 \end{prop}
@@
 
-  But as $\star T$ is the closure of $\separated T$, $\separated T$ is dense
-  into $\star T$, so the sheaf property of $Q$ allows to extend
-  $\separated f$ to $\star f:\star T \to Q$.
+  But as $\modal_{n+1} T$ is the closure of $\separated T$, $\separated T$ is dense
+  into $\modal_{n+1} T$, so the sheaf property of $Q$ allows to extend
+  $\separated f$ to $\modal_{n+1} f:\modal_{n+1} T \to Q$.
 
@@
 
-If $T:\Type_{n+1}$, the sheafification of $T$ is
-$$\star T = \sum_{u:T \to \Type_n^\modal} \lnot\lnot\left\| \sum_{a:X} 
-            (\lambda t,~\modal (a=t)) = u\right\|$$
-
-We already know that $\star$ defines a modality. It remains to show
-that this modality is left exact.
-
-The last thing to prove is the cumulativity of our sheafifications:
+Using the same idea of proof as in proposition~\ref{prop:sep-mod}, we have
 \begin{prop}
-  If $T:\Type_n$, then $\modal T = \star \widehat T$ (where $\widehat
-  T$ is $T$ seen as a $\Type_{n+1}$.
+  $(\modal_{n+1},\nu)$ defines a reflective subuniverse.
 \end{prop}
-It will imply that $\star$ is compatible with $\HProp$ in the sense of proposition~\ref{prop:hprop}.
+
+The last step is the left-exactness of $\modal_{n+1}$. Let
+$T:\Type_{n+1}$, and $x,y:T$. Suppose that $\Contr(\modal_{n+1} T)$.
+Thanks to compatibility between $\modal_{n+1}$ and $\modal_n$ for
+$n-\Type$, it suffices to show that $\modal_n(x=y)$ is contractible.
+
+The same proof as in proposition~\ref{prop:sep-subu} leads to
+\begin{prop}
+  For all $a,b:T$, $\modal_n(a=b) = (\nu_T a = \nu_T b)$.
+\end{prop}
+As $\modal_{n+1} T$ is contractible, any path space of $\modal T$ is
+contractible, in particular $(\nu_T a=\nu_T b)$.
 

### 2015-01-18 paper/sheaf_lics.tex
improve code a little bit / Explain admits from equality of diagrams
Summary+ admits not solved

@@
 
-\begin{prop}
+\begin{prop}\label{prop:consistent}
   A left exact modality $\modal$ induces a consistent type theory if
@@
 \begin{prop}
-  $(\modal_{n+1},\nu)$ defines a reflective subuniverse.
+  $(\modal_{n+1},\nu)$ defines a modality.
 \end{prop}
@@
 \label{ssec:summary}
+
+Starting from any left-exact modality $\modal_{-1}$ on $\HProp$, we
+defined in this paper a new left-exact modality $\modal$ on
+$\sum_{T:\Type} \sum_{n} \IsType n T$, which is exactly $\modal_{-1}$
+when restricted to $\HProp$.
+
+If $\modal_{-1}$ is the double negation modality on $\HProp$, it is obvious
+that $\modal_{-1}\zero=\zero$, hence the homotopy type theory induced by
+$\modal$ is consistent by proposition~\ref{prop:consistent}.
+
+In topos theory, the topos of Lawvere-Tierney sheaves for the double
+negation topology is a boolean topos. In homotopy type theory, this
+result can be expressed as:
+
+\begin{prop}
+  If $\modal_{-1}=\modal_{\lnot\lnot}$, and $\modal$ is the modality
+  as defined above, then the propositional excluded middle is true in
+  the induced type theory.
+\end{prop}
 
@@
 Section~\ref{sec:sheafification} (done in 
-\texttt{sheaf\_induction.v}), five very technical facts required to
+\texttt{sheaf\_induction.v}), three very technical facts required to
 show that $\square$ and $\modal_{n+1}$ are modalities have been
@@
 have been admitted only by lack of time.
+In the proof that diagrams are equal in lemma~\ref{lem:sepiscolim},
+the equality of projections is admitted, because if this proof is
+obvious on paper, it becomes really technical when formalized.
+As a consequence, the fact that $P$ is a colimit of the diagonal
+closed diagram is admitted too, but again, this proof should be really
+short as it is on paper.
 

### 2015-01-18 paper/sheaf_lics.tex
update admits in Formalization
add summary/conclusion/acknowlegment

@@
 
-% An example of a floating figure using the graphicx package.
-% Note that \label must occur AFTER (or within) \caption.
-% For figures, \caption should occur after the \includegraphics.
-% Note that IEEEtran v1.7 and later has special internal code that
-% is designed to preserve the operation of \label within \caption
-% even when the captionsoff option is in effect. However, because
-% of issues like this, it may be the safest practice to put all your
-% \label just after \caption rather than within \caption{}.
-%
-% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
-% option should be used if it is desired that the figures are to be
-% displayed while in draft mode.
-%
-%\begin{figure}[!t]
-%\centering
-%\includegraphics[width=2.5in]{myfigure}
-% where an .eps filename suffix will be assumed under latex, 
-% and a .pdf suffix will be assumed for pdflatex; or what has been declared
-% via \DeclareGraphicsExtensions.
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure}
-
-% Note that IEEE typically puts floats only at the top, even when this
-% results in a large percentage of a column being occupied by floats.
-
-% An example of a double column floating figure using two subfigures.
-% (The subfig.sty package must be loaded for this to work.)
-% The subfigure \label commands are set within each subfloat command,
-% and the \label for the overall figure must come after \caption.
-% \hfil is used as a separator to get equal spacing.
-% Watch out that the combined width of all the subfigures on a 
-% line do not exceed the text width or a line break will occur.
-%
-%\begin{figure*}[!t]
-%\centering
-%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
-%\label{fig_first_case}}
-%\hfil
-%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
-%\label{fig_second_case}}
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure*}
-%
-% Note that often IEEE papers with subfigures do not employ subfigure
-% captions (using the optional argument to \subfloat[]), but instead will
-% reference/describe all of them (a), (b), etc., within the main caption.
-% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
-% labels, the optional argument to \subfloat must be present. If a
-% subcaption is not desired, just leave its contents blank,
-% e.g., \subfloat[].
-
-% An example of a floating table. Note that, for IEEE style tables, the
-% \caption command should come BEFORE the table and, given that table
-% captions serve much like titles, are usually capitalized except for words
-% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
-% and up, which are usually not capitalized unless they are the first or
-% last word of the caption. Table text will default to \footnotesize as
-% IEEE normally uses this smaller font for tables.
-% The \label must come after \caption as always.
-%
-%\begin{table}[!t]
-%% increase table row spacing, adjust to taste
-%\renewcommand{\arraystretch}{1.3}
-% if using array.sty, it might be a good idea to tweak the value of
-% \extrarowheight as needed to properly center the text within the cells
-%\caption{An Example of a Table}
-%\label{table_example}
-%\centering
-%% Some packages, such as MDW tools, offer better commands for making tables
-%% than the plain LaTeX2e tabular which is used here.
-%\begin{tabular}{|c||c|}
-%\hline
-%One & Two\\
-%\hline
-%Three & Four\\
-%\hline
-%\end{tabular}
-%\end{table}
-
-% Note that the IEEE does not put floats in the very first column
-% - or typically anywhere on the first page for that matter. Also,
-% in-text middle ("here") positioning is typically not used, but it
-% is allowed and encouraged for Computer Society conferences (but
-% not Computer Society journals). Most IEEE journals/conferences use
-% top floats exclusively. 
-% Note that, LaTeX2e, unlike IEEE journals/conferences, places
-% footnotes above bottom floats. This can be corrected via the
-% \fnbelowfloat command of the stfloats package.
-
 \section{Formalization} 
@@
 provides a convenient tool to formalize some part of the theory of
-higher topoi.
+higher topoi. As a comparison, this \TeX\xspace file contains
+approximatively 1800 lines.
 
@@
 Section~\ref{sec:sheafification} (done in 
-\texttt{sheaf\_induction.v}), three very technical facts required to
-show that $\square$ and $\modal_{n+1}$ are modalities have been
+\texttt{sheaf\_induction.v}), two very technical facts required to
+show that $\square$ is a modality have been
 admitted. Nevertheless, we are confident that they can be proven and
@@
 
-Prove that the Giraud-Rezk-Lurie axiom holds in homotopy type theory,
-or at least the particular case we need.
-
-Extend the sheafification to all types, even if non truncated.
-
-Check if Lawvere-Tierney subsumes Grothendieck in higher topos theory.
+In this paper, we defined a way to extend type theory with new
+principles, using the already known method in topos theory,
+sheafification. Using this method, one an add some principles, like
+excluded middle, to type theory, but with computational content, while
+doing it with axiom break the extraction of proofs.
+
+Beyond the result, this work allows to illustrate some homotopy type theoritical
+techniques, as well as use of proof assistants.
+
+\kq{J'ai un peu copié sur la conclusion de Licata ``Eilenberg MacLane
+  spaces''. Si ça pose problème, je changerai.}
+
+In future works, we would like to improve this construction in two
+ways:
+\begin{itemize}
+\item At the moment, the sheafification process only works with
+  truncated types. But some classical types in homotopy type theory
+  are not truncated, like the circle $\mathbb S^1$. 
+  Thus, we plan to extend the sheafification to all types, even if non
+  truncated
+\item In this work, the Giraud-Rezk-Lurie axiom is admitted. The work
+  done in~\cite{lurie} suggests, if we see homotopy type theory as
+  higher topos theory, that this axiome could be provable. A
+  computer-checked proof of this fact would improve our developpement.
+\end{itemize}
+Moreover, in topos theory, we know that Lawvere-Tierney subsumes
+Grothendieck (\cite{maclanemoerdijk}, section~V.4). Higher Lawvere-Tierney sheaves are
+presented here, and higher Grothendieck sheaves in~\cite{lurie} ; we want to
+check if this remains true in higher topos theory.
 
@@
 
-The authors would like to thank...
+We would like to thank all maintainers of the Coq/HoTT library, who
+allowed us to base our formalization on something. We would also like
+to thank Bas Spitters for the helpful discussions on the subject.
 
@@
 
-% \bibitem[RS13]{sets_in_hott}
-% Egbert Rijke and Bas Spitters.
-% \newblock Sets in homotopy type theory.
-% \newblock 2013.
-
-% \bibitem[{Uni}13]{hottbook}
-% {Univalent Foundations Project}.
-% \newblock {\em Homotopy Type Theory: Univalent Foundations for Mathematics}.
-% \newblock 2013.
-
-% \bibitem[MM92]{maclanemoerdijk}
-% Saunders MacLane and Ieke Moerdijk.
-% \newblock {\em Sheaves in Geometry and Logic}.
-% \newblock Springer-Verlag, 1992.
-
-% \bibitem[HTT]{lurie}
-% Jacob Lurie.
-% \newblock{\em Higher Topos Theory}.
-
-% \bibitem[AKL]{lumsdaine}
-% Peter LeFanu Lumsdaine, Jeremy Avigad and Chris Kapulkin.
-% \newblock{\em Homotopy limits in type theory}.
-
-% \end{thebibliography}
-
 % that's all folks

### 2015-01-19 paper/sheaf_lics.tex
minor changes

@@
   \sum_{x:X} (f x = y) \ar[r]^\eta \ar[d]_\gamma & \modal \left(\sum_{x:X}  (f x = y)\right) \ar@{=}[dl]\\
-  \sum_{x:\modal X} (\modal f x = \eta_Y y) &
-}$$
-where $\pi_1 \circ \gamma = \eta_X \circ \pi_1$, and $\pi_2 \circ
-\gamma$ is given by the modalisation of paths.
+  \sum_{x:\modal X} (\modal f x = \eta_Y y) & }$$
+where $\pi_1 \circ \gamma = \eta_X \circ \pi_1$, and
+$\pi_2 \circ \gamma$ is given by the modalisation of
+paths.\footnote{This proposition comes from on old version of the
+  formalization of left exact modalities in the HoTT library, but is
+  not present anymore in the library.}
 \end{prop}
@@
 %
-This modality enables to use
-classical reasonings in the corresponding reflective subuniverse, but
-every type in the new universe is collapsed to an $\HProp$, ending up
-with propositional classical logic.
+This modality is consistent as $\modal_{\lnot\lnot} \zero = \zero$ can
+not be inhabited.
+%
+Using the double negation modality enables to use classical reasonings
+in the corresponding reflective subuniverse, but every type in the new
+universe is collapsed to an $\HProp$, ending up with propositional
+classical logic.
 
@@
 it requires much work here, as we have to deal with equalities. In
-particular, this property uses in particular the specific definition
+particular, this property uses the specific definition
 of separated types and sheaves, which shows that we had actually no
@@
     and $\psi : \modal_0T\to \modal_{-1} T$ by obtaining the
-    witness $a:T$, and letting
+    witness $a:T$ (which is possible because we are trying to inhabit
+    a modal proposition), and letting
     $$\psi (u;x) = \eta_T a.$$
-    These two maps form an equivalence.
+    These two maps form an equivalence (the section and retraction are
+    trivial because the equivalence is between mere propositions).
   \item Suppose now that $\modal_{n+1}$ is compatible with all $\modal_k$ on
@@
     \begin{align*}
-      \modal_{n+2} T &= \sum_{u:T \to \Type_{n+1}^\modal} \modal_{-1}\left\| \sum_{a:T} 
+      \modal_{n+2} T &= \hspace{-1em} \sum_{u:T \to \Type_{n+1}^\modal}
+                       \hspace{-1em} \modal_{-1}\left\| \sum_{a:T} 
             (\lambda t,~\modal_{n+1} (a=t)) = u\right\| \\
-                     &= \sum_{u:T \to \Type_{n+1}^\modal} \modal_{-1}\left\| \sum_{a:T} 
+                     &= \hspace{-1em} \sum_{u:T \to
+                       \Type_{n+1}^\modal} 
+\hspace{-1em} \modal_{-1}\left\| \sum_{a:T} 
             (\lambda t,~\modal_{n} (a=t)) = u\right\|
@@
 
-    It lefts to prove that $u:T\to\Type_n^\modal$, \ie that for all
-    $t:T$, $\IsType n (u\, t)$.
-    But for any truncation index $p$, the predicate $\IsType p:\HProp$
-    is a a sheaf, so we can get rid of $\modal_{-1}$, and pick up what
-    is inside the truncation. Thus, we know that for all $t:T$, $u\, t
-    = \modal_n(a=t) : \Type_n$.
+    It remains to prove that for every $(u,x)$ inhabited the
+    $\Sigma$-type above, $u$ is in $T\to\Type_n^\modal$, \ie that for
+    every $t:T$, $\IsType n (u\, t)$.  But for any truncation index $p$,
+    the predicate $\IsType p:\HProp$ is a a sheaf, so we can get rid
+    of $\modal_{-1}$ and of the truncation, which tells us that for
+    every 
+    $t:T$, $u\, t = \modal_n(a=t) : \Type_n$.
   \end{itemize}
@@
 Thanks to compatibility between $\modal_{n+1}$ and $\modal_n$ for
-$n-\Type$, it suffices to show that $\modal_n(x=y)$ is contractible.
+$\Type_n$, it suffices to show that $\modal_n(x=y)$ is contractible.
 
@@
 provides a convenient tool to formalize some part of the theory of
-higher topoi. As a comparison, this \TeX\xspace file contains
+higher topoi. As a matter of comparison, this \TeX\xspace file contains
 approximatively 1800 lines.
@@
 show that $\square$ is a modality have been
-admitted. Nevertheless, we are confident that they can be proven and
+admitted. 
+\nt{say that it comes from a lack of lemmas on paths between paths in
+  the HoTT library}
+%
+Nevertheless, we are confident that they can be proven and
 have been admitted only by lack of time.
@@
 %
-When the absence of cumulativity rises an universe issue for a lemma,
+When the absence of cumulativity rises a universe issue for a lemma,
 we have proven the lemma and then admit the exact same lemma, just to
@@
 levels in the recursive definition. 
+%
+\nt{explain why it is an issue.}
 %

### 2015-01-19 paper/sheaf_lics.tex
typos

--- 2015-01-18 paper/sheaf_lics.tex
@@
 Every Lawvere-Tierney topology $j$ induces a closure operator
-$A \mapsto \closure{A}$ on subojects. If we see a subobject $A$ of $E$
+$A \mapsto \closure{A}$ on subobjects. If we see a subobject $A$ of $E$
 as a characteristic function $\Char{A}$, the closure $\closure{A}$
@@
 %
-A suboject $A$ of $E$ is said to
+A subobject $A$ of $E$ is said to
 be dense when $\closure{A} = E$.
@@
 which exhibits $\Sh{\E}$ as a reflective
-subcatefory of $\E$ (which is a particular case of localization). This
+subcategory of $\E$ (which is a particular case of localization). This
 means that logical principles valid in $\E$ are still valid in
@@
 to construction in the Grothendieck setting, where the
-sheafification is obtained by applying the plus-construction $n+1$ to
+sheafification is obtained by applying the plus-construction $n+1$
+times to
 $n$-types~\cite[Chapter~6]{lurie}.
@@
 
-Given any type type $X$, its propositional truncation 
+Given any type $X$, its propositional truncation 
 $\squash{X} : \HProp$ is generated by 
@@
   For any $A:\Type$ and $B:\HProp$, if $f:A \to B$ then there is an
-  induced $g:\squash{X}\to B$ such that $g(|a|)= f(a)$ for any $a:A$.
+  induced $g:\squash{A}\to B$ such that $g(|a|)= f(a)$ for any $a:A$.
 \end{lem}
@@
 actually a colimit over $D$, we do not detail the proof here as the
-existence of colimits is not use in the definition of the
+existence of colimits is not used in the definition of the
 sheafification process. We only make use of the following special case
@@
 %
-This particular case connects a surjection (or a $-1$-connected
+This particular case connects a surjection (or a $(-1)$-connected
 function) to the colimit of its \v{C}ech nerve.
@@
 Section~\ref{sec:new-type-theory} discusses how a left exact modality
-induces a ne homotopy theory, and in which case this theory is
+induces a new homotopy theory, and in which case this theory is
 consistent.
@@
 is the fiber of $\modal f$ above $\eta_Y y$:
-$$\modal \left(\sum_{x:X}  (f x = y)\right) = \sum_{x:\modal X} (\modal f x = \eta_Y y).$$
+$$\modal \left(\sum_{x:X}  (f x = y)\right) = \sum_{x:\modal X}
+(\modal f x = \eta_Y y),$$
+where $\modal f:\modal X \to \modal Y$ is defined via the equivalence {\it (iv)} and $\eta_Y$.
 
@@
 This means that we need to take more care of the homotopy structure of
-this charactestic function, and in particular, the defintiion of a
+this charactestic function, and in particular, the definition of a
 dense $n$-subobject comes with an additional coherence condition
@@
   $.
-  Taking for example $t=b$, we deduce $\modal_n (a=b) = \modal_n(a=a)$,
-  and the latter is inhabited by $\eta_{a=a} 1$.
+  Taking for example $t=b$, we deduce $\modal_n (a=b) = \modal_n(b=b)$,
+  and the latter is inhabited by $\eta_{b=b} 1$.
 \item Suppose now $p : \modal_n(a=b)$. Let $\iota$ be the first
@@
 components, transported correctly along the first components equality,
-are equal. The complete proof can be found in the formalisation.
+are equal. The complete proof can be found in the formalization.
 This proves the following proposition
@@
 seen as a subobject of $T \to \Type_n^\modal$. Explicitly
-$$\modal_{n+1} T = \sum_{u:T \to \Type_n^\modal} \modal_{-1}\left\| \sum_{a:T} 
+$$\modal_{n+1} T \defeq \sum_{u:T \to \Type_n^\modal} \modal_{-1}\left\| \sum_{a:T} 
             (\lambda t,~\modal_n (a=t)) = u\right\|$$
@@
     But for any truncation index $p$, the predicate $\IsType p:\HProp$
-    is a a sheaf, so we can get rid of $\modal_{-1}$, and pick up what
+    is a sheaf, so we can get rid of $\modal_{-1}$, and pick up what
     is inside the truncation. Thus, we know that for all $t:T$, $u\, t
@@
   done in~\cite{lurie} suggests, if we see homotopy type theory as
-  higher topos theory, that this axiome could be provable. A
+  higher topos theory, that this axiom could be provable. A
   computer-checked proof of this fact would improve our developpement.

### 2015-01-19 paper/sheaf_lics.tex

@@
 
-    It lefts to prove that $u:T\to\Type_n^\modal$, \ie that for all
-    $t:T$, $\IsType n (u\, t)$.
-    But for any truncation index $p$, the predicate $\IsType p:\HProp$
-    is a sheaf, so we can get rid of $\modal_{-1}$, and pick up what
-    is inside the truncation. Thus, we know that for all $t:T$, $u\, t
-    = \modal_n(a=t) : \Type_n$.
+    It remains to prove that for every $(u,x)$ inhabiting the
+    $\Sigma$-type above, $u$ is in $T\to\Type_n^\modal$, \ie that for
+    every $t:T$, $\IsType n (u\, t)$.  But for any truncation index $p$,
+    the predicate $\IsType p:\HProp$ is a a sheaf, so we can get rid
+    of $\modal_{-1}$ and of the truncation, which tells us that for
+    every 
+    $t:T$, $u\, t = \modal_n(a=t) : \Type_n$.
   \end{itemize}

### 2015-01-19 paper/sheaf_lics.tex
add coqwc
pass on section 6

@@
 \newcommand \squash[1] {\| #1 \| }
-\newcommand \separated {\mathop{\square} }
+\newcommand \separated {\mathop{\square_{n+1}} }
 \newcommand \fib[2] {\mathrm{fib}_{#1}(#2)}
@@
 %
+The presentation of this paper uses an informal presentation of
+homotopy type theory, working entirely inside of Univalent
+Foundations. 
+ %
 Section~\ref{sec:formalization} discusses the details of the
-formalization in Coq, in particular issues concerning universe
-polymorphism.
+formalization in Coq using formal type theory, in particular issues
+concerning universe polymorphism.
 %
@@
 theory as modality preserves all type former, such as dependent
-products, dependent sums or inductive types (\nt{double check that}).
+products or dependent sums.
+% or inductive types (\nt{double check that}).
 
@@
 
-As explained in the introduction, a Lawvere-Tierney topology for an
+As explained in the introduction, a Lawvere-Tierney topology for a
 topos $\E$, as defined in Section~\ref{sec:lawv-tiern-sheaf},
-corresponds to a left exact modality at level $-1$.
-%
-And sheafification corresponds to lifting this modality to a left
+corresponds to a left exact modality at level $-1$
+%
+and sheafification corresponds to lifting this modality to a left
 exact modality at level $0$.
@@
 In homotopy type theory, it is possible to use similar ideas to lift
-any left exact modality at level $n$ to a left exact modality at level
-$n+1$ using sheafification.
+any left exact modality at level $-1$ to a left exact modality at level
+$n$ using sheafification, by induction on $n$.
 
@@
 characteristic function $\chi_E$ of $E$ seen as an $n$-subobject of
-itself is not the constant function equal to $\one$ as it is the case
+itself is not the constant function equal to $\one$, as it is the case
 for topoi, but rather the map 
@@
 
-A fundamental property in the construction of the sheafification is
-the fact that the type of modal types at level $n$ defines a sheaf. 
+A fundamental property in the construction of the sheafification
+functor is the fact that the type of modal types at level $n$ defines
+a sheaf.
 %
@@
 
-\section{Construction of sheafification}
+\section{Sheafification of Left Exact Modalities}
 \label{sec:sheafification}
@@
 
-\nt{je ne comprends pas ce passage, c'est quoi sheafness ? }
-
-\kq{c'est une vision plutôt grothendieck en fait. La separation ``enlève''
-  ce qui empêche à $T$ d'être un faisceau, et la sheafification
-  rajoute ce qui manque pour être un faisceau. Mais c'était plutôt mal
-  dit, effectivement}
+% \kq{c'est une vision plutôt grothendieck en fait. La separation ``enlève''
+%   ce qui empêche à $T$ d'être un faisceau, et la sheafification
+%   rajoute ce qui manque pour être un faisceau. Mais c'était plutôt mal
+%   dit, effectivement}
 
@@
 steps:
+%
 \begin{enumerate}[(i)]
-\item {\em separation:} In this step, we construct a {\em free}
-  separated object.
-\item {\em completion:} In this step, we add what lacks now to the
-  separated type to be a sheaf.
+\item {\em separation:} From any $T$ in $\Type_{n+1}$, we construct
+  its {\em free} separated object $\separated T$ by ``removing'' what
+  prevents $T$ from being a sheaf.
+\item {\em completion:} We add what is missing for the free 
+  separated type to be a sheaf by using closure in a sheaf.
 \end{enumerate}
@@
 %\nt{put the definition of $\modal_{n+1}$ upfront}
-If $T:\Type_{n+1}$, let  $\modal_{n+1}T$ be the cloture of $\square T$
-seen as a subobject of $T \to \Type_n^\modal$. Explicitly
-$$\modal_{n+1} T \defeq \sum_{u:T \to \Type_n^\modal} \modal_{-1}\left\| \sum_{a:T} 
-            (\lambda t,~\modal_n (a=t)) = u\right\|$$
-
-We begin by proving that $\modal_{n+1}$ is compatible with $\modal_n$
-on lower homotopy types.
+For any $T$ in $\Type_{n+1}$, 
+$\modal_{n+1}T$ is defined as the closure of $\separated T$,
+seen as a subobject of $T \to \Type_n^\modal$. 
+%
+Explicitly
+$$
+\modal_{n+1} T \ \defeq \sum_{u:T \to \Type_n^\modal} \modal_{-1}\left\| \sum_{a:T} 
+            (\lambda t,~\modal_n (a=t)) = u\right\|.
+$$
+
+To prove that $\modal_{n+1} T$ is a sheaf for any $T:\Type_{n+1}$, we
+use the following lemma.
+\begin{lem}
+  Let $X:\Type_{n+1}$ and $U$ be a sheaf. If $X$ embeds
+  in $U$, and is closed in $U$, then $X$ is a sheaf.
+\end{lem}
+
+As $T\to \Type_n^\modal$ is a sheaf, and $\modal_{n+1}T$ is closed in
+$T\to \Type_n^\modal$, $\modal_{n+1}T$ is a sheaf. We now prove that
+it forms a reflective subuniverse.
+
 \begin{prop}
-  If $T:\Type_n$, then $\modal_{n+1}T = \modal_n T$.
+  $(\modal_{n+1},\nu)$ defines a reflective subuniverse.
 \end{prop}
 \begin{proof}
-  We prove it by induction on $n$~:
+  Let $T,Q:\Type_{n+1}$ such that $Q$ is a sheaf. Let $f:T\to Q$.
+  Because $Q$ is a sheaf, it is in particular separated;
+  % 
+  thus we can extend $f$ to $\separated f:\separated T\to Q$.
+
+  But as $\modal_{n+1} T$ is the closure of $\separated T$, $\separated T$ is dense
+  into $\modal_{n+1} T$, so the sheaf property of $Q$ allows to extend
+  $\separated f$ to $\modal_{n+1} f:\modal_{n+1} T \to Q$.
+
+  As all these steps are universal, the composition is.
+\end{proof}
+
+Using the same technique as in proposition~\ref{prop:sep-mod}, we have
+\begin{prop}
+  $(\modal_{n+1},\nu)$ defines a modality.
+\end{prop}
+
+It remains to show that $\modal_{n+1}$ is left exact and is compatible
+with $\modal_{-1}$. To do that, we need to extend the notion of
+compatibility and that actually every modality $\modal_{n+1}$ is
+compatible with $\modal_n$ on lower homotopy types.
+\begin{prop}
+  If $T:\Type_n$, then $\modal_{n+1} \widehat T = \modal_n T$.
+\end{prop}
+\begin{proof}
+  We prove it by induction on $n$:
   \begin{itemize}
@@
 
-This also proves that $\modal_{n+1}$ is compatible with $\modal_{-1}$
-in the sense of proposition~\ref{prop:hprop}.
-
-To prove that $\modal_{n+1} T$ is a sheaf for any $T:\Type_{n+1}$, we
-use the following:
-\begin{lem}
-  Let $X:\Type_{n+1}$ be separated, and $U$ be a sheaf. If $X$ embeds
-  in $U$, and is closed in $U$, then $X$ is a sheaf.
-\end{lem}
-
-As $T\to \Type_n^\modal$ is a sheaf, and any object is closed into its
-cloture, $\modal_{n+1}T$ is indeed a sheaf.
-
-\begin{prop}
-  $(\modal_{n+1},\nu)$ defines a reflective subuniverse.
-\end{prop}
-\begin{proof}
-  Let $T,Q:\Type_{n+1}$ such that $Q$ is a sheaf. Let $f:T\to Q$.
-  Because $Q$ is a sheaf, it is in particular separated ; thus we can
-  extend $f$ to $\separated f:\separated T\to Q$.
-
-  But as $\modal_{n+1} T$ is the closure of $\separated T$, $\separated T$ is dense
-  into $\modal_{n+1} T$, so the sheaf property of $Q$ allows to extend
-  $\separated f$ to $\modal_{n+1} f:\modal_{n+1} T \to Q$.
-
-  As all these steps are universal, the composition is.
-\end{proof}
-
-Using the same idea of proof as in proposition~\ref{prop:sep-mod}, we have
-\begin{prop}
-  $(\modal_{n+1},\nu)$ defines a modality.
-\end{prop}
+This proves in particular that $\modal_{n+1}$ is compatible with
+$\modal_{-1}$ in the sense of proposition~\ref{prop:hprop}.
 
 The last step is the left-exactness of $\modal_{n+1}$. Let
-$T:\Type_{n+1}$, and $x,y:T$. Suppose that $\Contr(\modal_{n+1} T)$.
+$T$ be in $\Type_{n+1}$ such that $\Contr(\modal_{n+1} T)$.
 Thanks to compatibility between $\modal_{n+1}$ and $\modal_n$ for
-$\Type_n$, it suffices to show that $\modal_n(x=y)$ is contractible.
-
-The same proof as in proposition~\ref{prop:sep-subu} leads to
+$\Type_n$, left exactness means that for any $x,y : T$, $\modal_n(x=y)$ is contractible.
+
+Using a proof by univalence as we have done for proving $\modal_n (a=b) = (\mu_T a =
+\mu_T b)$ in Proposition~\ref{prop:sep-subu}, we can show that:
 \begin{prop}
@@
 \end{prop}
-As $\modal_{n+1} T$ is contractible, any path space of $\modal T$ is
-contractible, in particular $(\nu_T a=\nu_T b)$.
+
+As $\modal_{n+1} T$ is contractible, any path space of $\modal_{n+1} T$ is
+contractible, in particular $(\nu_T a=\nu_T b)$, which proves left
+exactness.
 
@@
 Starting from any left-exact modality $\modal_{-1}$ on $\HProp$, we
-defined in this paper a new left-exact modality $\modal$ on
-$\sum_{T:\Type} \sum_{n} \IsType n T$, which is exactly $\modal_{-1}$
-when restricted to $\HProp$.
-
-If $\modal_{-1}$ is the double negation modality on $\HProp$, it is obvious
-that $\modal_{-1}\zero=\zero$, hence the homotopy type theory induced by
-$\modal$ is consistent by proposition~\ref{prop:consistent}.
+have defined for any truncation level $n$, a new left-exact modality
+$\modal_n$ on $\Type_n$, which is exactly $\modal_{-1}$ when
+restricted to $\HProp$.
+
+When $\modal_{-1}$ is consistent (in the sense of
+proposition~\ref{prop:consistent}), 
+$\modal_{n}\zero=\modal_{-1}\zero$ is also not inhabited, hence the homotopy type theory induced by
+$\modal_n$ is consistent. 
+%
+In particular, the modality induced by the double negation modality on
+$\HProp$ is consistent.
 
@@
 \begin{prop}
-  If $\modal_{-1}=\modal_{\lnot\lnot}$, and $\modal$ is the modality
-  as defined above, then the propositional excluded middle is true in
-  the induced type theory.
+  $(\modal_{\lnot\lnot})_n$, the modality obtained by sheafification
+  of the double negation modality,
+  %
+  induces a type theory where the propositional excluded middle holds.
 \end{prop}
 
-\nt{Recap what has been done, and give the final result about EM with
-  the dense topology as a corollary.}
+% \nt{Recap what has been done, and give the final result about EM with
+%   the dense topology as a corollary.}
 
@@
 way Coq tries to rewrite and apply lemmas automatically. 
+The \texttt{coqwc} tool counts 2000 lines of specifications
+(definitions, lemmas, theorems, propositions) and 3700 lines of proof overall.
 %
@@
 \texttt{sheaf\_induction.v}), two very technical facts required to
-show that $\square$ is a modality have been
+show that $\separated$ is a modality have been
 admitted. 
@@
 generalizing universe polymorphism for recursive definition.
-%
-% In particular, compoatcumulativity is an axiom, because of
-% universes issues.
 

### 2015-01-19 paper/sheaf_lics.tex
pass on section 6

@@
 \begin{prop}
-  If $T:\Type_n$, then $\modal_{n+1} \widehat T = \modal_n T$.
+  If $T:\Type_n$, then $\modal_{n+1} \widehat T = \modal_n T$, where $\widehat T$ is $T$ seen as a
+  $\Type_{n+1}$.
 \end{prop}
@@
     \begin{align*}
-      \modal_{0} T &= \sum_{u:T \to \Type_n^\modal} \modal_{-1}\left\| \sum_{a:T} 
+      \modal_{0} \widehat T &= \sum_{u:T \to \Type_n^\modal} \modal_{-1}\left\| \sum_{a:T} 
       (\lambda t,~\modal_{-1} (a=t)) = u\right\| \\
@@
     \begin{align*}
-      \modal_{n+2} T &= \hspace{-1em} \sum_{u:T \to \Type_{n+1}^\modal}
+      \modal_{n+2} \widehat  T &= \hspace{-1em} \sum_{u:T \to \Type_{n+1}^\modal}
                        \hspace{-1em} \modal_{-1}\left\| \sum_{a:T} 
@@
     because $\modal_{n+1}$ is compatible with $\modal_{n}$, and
-    $(a=t):\Type_n$.
+    $(a=t)$ is in $\Type_n$.
 
@@
 %
-It is based on the homotopy type theory library available at
-\url{https://github.com/HoTT/HoTT}.
-
-We provide a more detailed insight of where properties have been
-formalized, and what is the size of the formalisation for each part:
+It is based on the Coq/HoTT library available at
+\url{https://github.com/HoTT/HoTT} and uses the version of Coq
+designed for that library.
+%
+
+\subsection{Content of the formalization}
+
+We provide a more detailed insight of the structure of our formalization:
 \begin{itemize} 
 \item  Homotopy fibers and subobject classifiers are formalized in
-\texttt{sub\_object\_classifier.v} (284 lines).
+\texttt{sub\_object\_classifier.v}.% (284 lines).
 %
 \item Colimits, $p$-pullbacks and \v{C}ech nerves are formalized in
-\texttt{colimits.v, cech\_nerve.v} (552 lines).
+\texttt{colimits.v, cech\_nerve.v}.% (552 lines).
 \item
 Reflective subuniverses and modalities are formalized in
-\texttt{reflective\_subuniverse.v, modalities.v} (1053 lines).
+\texttt{reflective\_subuniverse.v, modalities.v}. % (1053 lines).
 \item 
@@
   The definition of the dense topology as a left exact modality on
-  $\HProp$ is given in \texttt{sheaf\_base\_case.v} (186 lines).
+  $\HProp$ is given in \texttt{sheaf\_base\_case.v}. % (186 lines).
 \item
 Section~\ref{sec:sheaves} is formalized in
-\texttt{sheaf\_def\_and\_thm.v} (1029 lines).
+\texttt{sheaf\_def\_and\_thm.v}. % (1029 lines).
 \item
 Section~\ref{sec:sheafification} is formalized in
-\texttt{sheaf\_induction.v} (2340 lines).
+\texttt{sheaf\_induction.v}. % (2340 lines).
 \end{itemize}
@@
 The \texttt{coqwc} tool counts 2000 lines of specifications
-(definitions, lemmas, theorems, propositions) and 3700 lines of proof overall.
-%
-This constitute a significant amount of work but the part dedicated to
-sheaves and sheafification is only 4000, which seems quite reasonable
-and is encouraging, because it suggests that homotopy type theory
-provides a convenient tool to formalize some part of the theory of
-higher topoi. As a matter of comparison, this \TeX\xspace file contains
-approximatively 1800 lines.
+(definitions, lemmas, theorems, propositions) and 3700 lines of proof
+script.
+%
+This constitutes a significant amount of work but the part dedicated
+to sheaves and sheafification is only 2200 lines of proof script,
+which seems quite reasonable and encouraging, because it suggests
+that homotopy type theory provides a convenient tool to formalize some
+part of the theory of higher topoi. As a matter of comparison, this
+\TeX\xspace file contains approximatively 1800 lines.
 
@@
 
+\subsection{Limitations of the formalization}
+
 In our formalization of sheafification defined
@@
 admitted. 
-\nt{say that it comes from a lack of lemmas on paths between paths in
-  the HoTT library}
-%
-Nevertheless, we are confident that they can be proven and
-have been admitted only by lack of time.
+%
+It is due to a lack of important lemmas on paths between paths in
+the Coq/HoTT library.
+%
+Nevertheless, we are confident that with enrichment of the Coq/HoTT
+library, they can be proven and have been admitted only by lack of
+time.
+%
 In the proof that diagrams are equal in lemma~\ref{lem:sepiscolim},
@@
 obvious on paper, it becomes really technical when formalized.
+%
+This suggest that we should try to fine a better way to encode
+projections, with a better computational content.
+%
 As a consequence, the fact that $P$ is a colimit of the diagonal
@@
 due to an unsatisfactory management of polymorphic universe level in
-Coq~\cite{sozeau2014universe}.
-%
-\nt{Say a word about polymorphic universe level, consistency check,
-  and the fact that polymorphic universe level are nevertheless
-  mandatory for our formalization}
+Coq.
+%
+Universes are used in type theory to ensure consistency by checking
+that definitions are well-stratified according to a certain hierarchy.
+%
+Universe polymorphism~\cite{sozeau2014universe} supports generic
+definitions over universes, reusable at different levels.
+%
+Although the presence of universe polymorphism is mandatory for our
+formalization, its implementation is still to rigid to allow a
+complete formalization of our work.
 
@@
 %
-But as it would not have any computational content, it
-would really complicates the proofs as the axiom would appear
+But as it would not have any computational content, such a solution
+would really complicate the proofs as the axiom would appear
 everywhere cumulativity is needed and it would need explicitly
@@
 
-We have chosen a more simple but less satisfactory solution. 
-%
-When the absence of cumulativity rises a universe issue for a lemma,
-we have proven the lemma and then admit the exact same lemma, just to
-let Coq infer less constrained universe levels.
+We have chosen a more simple but maybe less satisfactory solution. 
+%
+When the absence of cumulativity on $\Type_n$ rises a universe issue
+for a lemma, we have proven the lemma and then admit the exact same
+lemma, just letting Coq infer less constrained universe levels.
 
 An other issue with universe polymorphism lies in their management for
-recursive definitions. Indeed, the definition of successive sheafification types is
-not allowed:
-%
-\[ \left\{
-    \begin{array}{l}
-      \modal_{-1} = \lambda T:\HProp,~ \lnot\lnot T \\
-      \displaystyle{\modal_{n+1} = \lambda T:\Type_{n+1}},\\
-      \quad \displaystyle{\sum_{u:T \to \Type_n^\modal} \modal_{-1} 
+recursive definitions. Indeed, the following recursive definition of
+sheafification:
+%
+\[ \begin{array}{l}
+   \modal : \forall \ (n : nat), \ \Type_n \to \Type_n 
+   \\
+    \modal_{-1\phantom{n}} \defeq \lambda T:\HProp,~ \lnot\lnot T \\
+      \displaystyle{\modal_{n+1} \defeq  \lambda T:\Type_{n+1}},\\
+      \hspace{5em} \displaystyle{\sum_{u:T \to \Type_n^\modal} \modal_{-1} 
       \left\|
-      \sum_{a:T} u = (\lambda t,~\modal_n (a=t))
+      \sum_{a:T} u= (\lambda t,~\modal_n (a=t))
       \right\|}
     \end{array}
-  \right.\]
-%
-because Coq asks for every $\modal_n$ to have the same universes
-levels in the recursive definition. 
-%
-\nt{explain why it is an issue.}
+\]
+%
+is not allowed. 
+%
+This is because Coq forces the universe of the first $\Type_n$
+occurring to be the same for every $n$, whereas the universe of the
+$\Type_{n+1}$ occurring in $\modal_{n+1}$ should be at least one level
+higher as the one of $\Type_n$ occurring in $\modal_{n+1}$ because of
+the use of $\Sigma$-type over $T \to \Type_n^\modal$.
+%
+% \nt{explain why it is an issue.}
 %
 Thus, the induction step presented in this paper has been formalized,
-but the complete induction can not be formalized at the moment. 
+but the complete recursive sheafification can not be defined for the
+moment.
 %
@@
 generalizing universe polymorphism for recursive definition.
+%
 

### 2015-01-19 paper/sheaf_lics.tex

@@
 
-In this paper, we defined a way to extend type theory with new
-principles, using the already known method in topos theory,
-sheafification. Using this method, one an add some principles, like
-excluded middle, to type theory, but with computational content, while
-doing it with axiom break the extraction of proofs.
-
-Beyond the result, this work allows to illustrate some homotopy type theoritical
-techniques, as well as use of proof assistants.
-
-\kq{J'ai un peu copié sur la conclusion de Licata ``Eilenberg MacLane
-  spaces''. Si ça pose problème, je changerai.}
+In this paper, we have defined a way to extend homotopy type theory
+with new principles, leveraging Lawvere-Tierney sheafification to
+higher topos theory. 
+%
+Using sheafification, it is possible to add new logical principles,
+like the law of excluded middle, to homotopy type theory and this
+without axiom.
+%
+Beyond the result, our work is part of a more general line of works
+which aim to illustrate that homotopy type theory is a very promising
+theory for formalizing mathematics inside a proof assistant.
 
@@
 \begin{itemize}
-\item At the moment, the sheafification process only works with
-  truncated types. But some classical types in homotopy type theory
-  are not truncated, like the circle $\mathbb S^1$. 
-  Thus, we plan to extend the sheafification to all types, even if non
-  truncated
-\item In this work, the Giraud-Rezk-Lurie axiom is admitted. The work
-  done in~\cite{lurie} suggests, if we see homotopy type theory as
-  higher topos theory, that this axiom could be provable. A
-  computer-checked proof of this fact would improve our developpement.
+\item 
+  At the moment, sheafification only works with truncated
+  types. Also this is fine from a programming point of view, 
+  many important types in homotopy theory are not
+  truncated, like the circle $\mathbb S^1$.  Thus, we plan to extend
+  the sheafification to all types, even if non truncated by using a
+  limit construction.
+\item 
+  In this work, a particular case of a Giraud-Rezk-Lurie axiom has
+  been admitted. The work done in~\cite[Chapter 6]{lurie} suggests,
+  considering that homotopy type theory is a language for higher topos
+  theory, that this axiom is provable. We plan to investigate if the
+  higher topos theoretic proof can be adapted to our setting. 
+  %
+  % A computer-checked proof of this fact would improve our
+  % developpement.
 \end{itemize}
-Moreover, in topos theory, we know that Lawvere-Tierney subsumes
-Grothendieck (\cite{maclanemoerdijk}, section~V.4). Higher Lawvere-Tierney sheaves are
-presented here, and higher Grothendieck sheaves in~\cite{lurie} ; we want to
-check if this remains true in higher topos theory.
+Moreover, in topos theory, Lawvere-Tierney subsumes Grothendieck
+(\cite{maclanemoerdijk}, section~V.4) in the sense that any
+Grothendieck topology gives rise to a Lawvere-Tierney topology with
+the same notion of sheaves. Higher Lawvere-Tierney sheaves are
+presented here, and higher Grothendieck sheaves have been defined
+in~\cite{lurie}. It should be interesting to check if the subsumption
+remains true in higher topos theory.
 
@@
 We would like to thank all maintainers of the Coq/HoTT library, who
-allowed us to base our formalization on something. We would also like
-to thank Bas Spitters for the helpful discussions on the subject.
+provide a very solid base for our formalization. We also would like
+to thank Bas Spitters for helpful discussions on the subject.
 

### 2015-01-19 paper/sheaf_lics.tex
minor correction

@@
 annotation for universe levels everywhere in the formalization.
-
+%
 We have chosen a more simple but maybe less satisfactory solution. 
@@
 
-An other issue with universe polymorphism lies in their management for
+An other issue with universe polymorphism lies in the management of
 recursive definitions. Indeed, the following recursive definition of
-sheafification:
+sheafification
 %
@@
 This is because Coq forces the universe of the first $\Type_n$
-occurring to be the same for every $n$, whereas the universe of the
-$\Type_{n+1}$ occurring in $\modal_{n+1}$ should be at least one level
-higher as the one of $\Type_n$ occurring in $\modal_{n+1}$ because of
-the use of $\Sigma$-type over $T \to \Type_n^\modal$.
+occurring in the definition to be the same for every $n$, whereas the
+universe of the first $\Type_{n+1}$ occurring in $\modal_{n+1}$ should be at
+least one level higher as the one of $\Type_n$ occurring in
+$\modal_{n}$ because of the use of $\Sigma$-type over
+$T \to \Type_n^\modal$.
 %

### 2015-01-19 paper/sheaf_lics.tex
typo

--- 2015-01-19 paper/sheaf_lics.tex
@@
 Moreover, in topos theory, Lawvere-Tierney subsumes Grothendieck
-(\cite{maclanemoerdijk}, section~V.4) in the sense that any
+(\cite[Section~V.4]{maclanemoerdijk}) in the sense that any
 Grothendieck topology gives rise to a Lawvere-Tierney topology with

### 2015-01-19 paper/sheaf_lics.tex
orthogonal system

--- 2015-01-19 paper/sheaf_lics.tex
@@
 
-\subsection{The Orthogonal Factorization System}
+\subsection{The (n-Connected,n-Truncated) Factorization System}
 \label{sec:epi-mono-fact}
@@
 
-It is then fairly obvious that the canonical function $\hat f : A \to
-\im(f)$ is (-1)-connected, thus the following proposition holds
-(\cite{hottbook}, Lemma 7.6.4):
+The canonical function $\hat f : A \to \im(f)$ is (-1)-connected,
+being the left component of an orthogonal factorization system which
+satisfies~\cite[Lemma 7.6.4]{hottbook}:
 \begin{prop}
-  A map $f:A\to B$ factors through $\im(f)$ as a (-1)-connected
-  function followed by a (-1)-truncated function.
+  A map $f:A\to B$ factors uniquely (up-to homotopy) through
+  $\im(f)$ as a (-1)-connected function followed by a (-1)-truncated
+  function.
 \end{prop}
 
-Note that there are other factorization systems constituted by
-$n$-truncated morphism and $n$-connected morphisms, for every
-truncated level $n$, but they do not appear to be useful in the
-definition of sheafification.
+Note that this generalizes to orthogonal factorization systems
+constituted by $n$-truncated morphism and $n$-connected morphisms, for
+every truncated level $n$. We will only use the factorization system
+at level  $-1$ in the definition of sheafification.
 

### 2015-01-19 paper/sheaf_lics.tex
minor changes

@@
 
-As a prerequisite, we encourage the reader be familiar with type theory and
+As a prerequisite, we encourage the reader to be familiar with type theory and
 in particular the point of view developed
@@
 %
+We note $\prod_{x:T} U$ for the dependent product, $\sum_{x:T} U$ for
+the dependent sum, and $\pi_1, \pi_2$ for the first and second
+projection of a dependent pair. We use informal mathematical language
+instead of type theory whenever it is possible to ease the
+reading, without making our statement imprecise. In particular,
+(higher) inductive types are defined using itemization to avoid an
+overhead of notation.
 
@@
 in homotopy type theory, it can be defined using higher inductive
-types. We adopt this point of view as it allows to maintain the slogan
-that homotopy type theory is type theory plus univalence and higher
-inductive types.
+types~\cite{lumsdaine2011higher,lumsdaine2013higher}. We adopt this
+point of view as it allows to maintain the slogan that homotopy type
+theory is type theory plus univalence and higher inductive types.
 %
@@
 
-In our formalization of sheafification defined
+In our formalization of sheafification defined in
 Section~\ref{sec:sheafification} (done in 
@@
 %
-It is due to a lack of important lemmas on paths between paths in
+It is due to a lack of technical lemmas on paths between paths in
 the Coq/HoTT library.
 %
-Nevertheless, we are confident that with enrichment of the Coq/HoTT
+Nevertheless, we are confident that with an enrichment of the Coq/HoTT
 library, they can be proven and have been admitted only by lack of
@@
 %
-Using sheafification, it is possible to add new logical principles,
-like the law of excluded middle, to homotopy type theory and this
+Using sheafification, it is possible to add new logical principles to
+homotopy type theory, like the law of excluded middle, and this
 without axiom.
 %
-Beyond the result, our work is part of a more general line of works
+Beyond this result, our work is part of a more general line of works
 which aim to illustrate that homotopy type theory is a very promising
@@
 
-In future works, we would like to improve this construction in two
+As a future work, we would like to improve this construction in two
 ways:
@@
   At the moment, sheafification only works with truncated
-  types. Also this is fine from a programming point of view, 
-  many important types in homotopy theory are not
-  truncated, like the circle $\mathbb S^1$.  Thus, we plan to extend
-  the sheafification to all types, even if non truncated by using a
+  types. This is fine from a programming point of view, 
+  but many important types in homotopy theory are not
+  truncated, \eg the circle $\mathbb S^1$.  Thus, we plan to extend
+  the sheafification to all types by using a
   limit construction.
@@
   considering that homotopy type theory is a language for higher topos
-  theory, that this axiom is provable. We plan to investigate if the
+  theory, that this axiom is provable. We plan to investigate whether the
   higher topos theoretic proof can be adapted to our setting.
@@
 \end{itemize}
-Moreover, in topos theory, Lawvere-Tierney subsumes Grothendieck
-(\cite[Section~V.4]{maclanemoerdijk}) in the sense that any
+Moreover, in topos theory, Lawvere-Tierney subsumes Grothendieck~\cite[Section~V.4]{maclanemoerdijk} in the sense that any
 Grothendieck topology gives rise to a Lawvere-Tierney topology with

### 2015-01-19 paper/sheaf_lics.tex
pass on intro

@@
   principles inspired by category theory and homotopy theory, and
-  which corresponds closely to higher topoi.
+  which corresponds closely to higher toposes.
   % 
@@
 %
-In the field of topos theory, it provides a way to construct new topoi
+In the field of topos theory, it provides a way to construct new toposes
 from already existing ones, allowing logical principles---that can not
@@
 simplicity of sheafification in topos theory and are not amenable to a
-direct integration into proof assistant based on type theory such as
+direct integration into a proof assistant based on type theory such as
 Coq~\cite{coq:refman:8.4} or Agda~\cite{norell2007towards}.  
@@
 %
-Lurie has developed all the tools that have been defined in
+Lurie has presented all the tools that have been defined in
 topos theory, but in a higher setting. In particular, the theory of
@@
 %
-As the notion of higher topoi appears to correspond very closely to
+As the notion of higher toposes appears to correspond very closely to
 homotopy type theory, this provides a new hope that tackling the
@@
 theory to homotopy type theory is not straightforward because the
-construction in higher topoi is restricted to the initial Grothendieck
+construction in higher toposes is restricted to the initial Grothendieck
 setting which is still very topologically oriented, and not very
@@
 Lawvere-Tierney sheafification in a topos $\E$ is based on an abstract
-point of view on the topology to be considered, being simply an
+point of view on the topology to be considered, being simply defined by an
 endomorphism on the classifying object $\Omega$ of $\E$  
@@
 To extend Lawvere-Tierney sheafification to homotopy type theory, the
-first thing to realize is that the construction can not be done in one
+first thing to understand is that the construction can not be done in one
 single step anymore. It must rather be performed by induction on the
@@
 sheafification is defined for $\Type_{0}$, given a topology on
-$\Type_{-1}$. This part corresponds to sheafification for topoi. Then,
+$\Type_{-1}$. This part corresponds to sheafification for toposes. Then,
 assuming that the sheafification has been constructed up to level $n$,
@@
 each $n\geq -2$ which classifies functions with $n$-truncated homotopy
-fibers~\cite{sets_in_hott}. This means that when monomorphisms are used in topos
-theoretic definition, the correct generalization may be $n$-truncated homotopy
+fibers~\cite{sets_in_hott}. This means that when monomorphisms are used in topos-theoretic definition, the correct generalization may be $n$-truncated homotopy
 fibers or embeddings (\ie $-1$-truncated homotopy
@@
 
-In other technical difficulty is the fact that sheafification in topos
-theory uses the fact that epimorphisms are coequalizers of their
-kernel pairs. This properties does not holds in higher topoi, but
-using Giraud-Rezk-Lurie axioms of $\infty$-topoi, the same property
+Another technical difficulty is the fact that sheafification in topos
+theory uses the property that epimorphisms are coequalizers of their
+kernel pairs. This properties does not hold anymore in higher toposes, but
+using Giraud-Rezk-Lurie axioms of $\infty$-toposes, the same property
 holds by replacing kernel pairs by \v{C}ech nerves in the statement.
@@
 Finally, one of the main difficulty in our formalization is that many
-equalities which are trival (by proof irrelevance) in the topos theoretic
+equalities which are trival (by proof irrelevance) in the topos-theoretic
 setting becomes relevant and requires a carreful use of univalence
@@
 
-Section~\ref{sec:hott} defines the foundations of homotopy type theory
-and presents a particular case of Giraud-Rezk-Lurie axioms. 
+Section~\ref{sec:hott} presents the foundations of homotopy type theory
+and defines a particular case of Giraud-Rezk-Lurie axioms. 
 %
@@
 %
-The presentation of this paper uses an informal presentation of
+The definitions of this paper uses an informal presentation of
 homotopy type theory, working entirely inside of Univalent
@@
 And Section~\ref{sec:giraud-ax} defines a particular case of
-Giraud-Rezk-Lurie axioms on higher topoi that we need in the
+Giraud-Rezk-Lurie axioms on higher toposes that we need in the
 definition of sheafification.
 %
-The definition of Section~\ref{ssec:hott} are part of the Coq library
+The definition of Section~\ref{ssec:hott} and \ref{sec:epi-mono-fact}
+are part of the Coq library
 available at \url{https://github.com/HoTT/HoTT}, while other definitions
@@
 %
-We note $\prod_{x:T} U$ for the dependent product, $\sum_{x:T} U$ for
-the dependent sum, and $\pi_1, \pi_2$ for the first and second
+We note $\prod_{x:T} U$ for dependent products, $\sum_{x:T} U$ for
+dependent sums, and $\pi_1, \pi_2$ for the first and second
 projection of a dependent pair. We use informal mathematical language
-instead of type theory whenever it is possible to ease the
-reading, without making our statement imprecise. In particular,
+instead of type theory whenever it is possible, to ease the
+reading without making our statement imprecise. In particular,
 (higher) inductive types are defined using itemization to avoid an
@@
 As noticed in \cite{hottbook}, left exact modalities corresponds 
-to sub-$(\infty,1)$-topoi. Indeed, modal types defines a new type
+to sub-$(\infty,1)$-toposes. Indeed, modal types defines a new type
 theory as modality preserves all type former, such as dependent
@@
 itself is not the constant function equal to $\one$, as it is the case
-for topoi, but rather the map 
+for toposes, but rather the map 
 %
@@
 Lawvere-Tierney topologies can still be shown to subsume Grothendieck
-toplogies for $(\infty,1)$-topoi, which is beyond the scope of this paper).
+toplogies for $(\infty,1)$-toposes, which is beyond the scope of this paper).
 %
@@
 %
-This property initially holds in the topos theoretic construction, but
+This property initially holds in the topos-theoretic construction, but
 it requires much work here, as we have to deal with equalities. In
@@
 %
-This corresponds to the free separated object used in the topos
-theoretic construction, but using $\Type_n^\modal$ instead of the
+This corresponds to the free separated object used in the topos-theoretic construction, but using $\Type_n^\modal$ instead of the
 $j$-subobject classifier $\Omega_j$.
@@
 that homotopy type theory provides a convenient tool to formalize some
-part of the theory of higher topoi. As a matter of comparison, this
+part of the theory of higher toposes. As a matter of comparison, this
 \TeX\xspace file contains approximatively 1800 lines.
@@
   theory, that this axiom is provable. We plan to investigate whether the
-  higher topos theoretic proof can be adapted to our setting.
+  higher-topos-theoretic proof can be adapted to our setting.
   %

### 2015-01-20 paper/sheaf_lics.tex
new pass up to section 4

@@
 from already existing ones, allowing logical principles---that can not
-be proven to be true or false in the old topos---to be valid (or
+be proved to be true or false in the old topos---to be valid (or
 invalid) in the new topos.
@@
 $n$-truncated function) when $\fib{f}{b}$
-is an $\Type_n$ for any $b$.  
+is in $\Type_n$ for any $b$.  
 %
@@
 %
- such that the following diagram is a pullback for any $f$ with
- $n$-truncated homotopy fibers:
+ such that the diagram
 $$
@@
 $$
-with $\Type_n^\bullet \defeq \sum_{A:\Type_n} A$ the universe of pointed
+is a pullback for any $f$ with
+ $n$-truncated homotopy fibers where $\Type_n^\bullet \defeq
+ \sum_{A:\Type_n} A$ is the universe of pointed
 $n$-truncated types and 
@@
 But as some constructors do not preserve the level of homotopy type
-(\eg sums or $\Sigma$-types), we need to introduce $\squash{X}$, the
+(\eg sums or $\Sigma$-types), so we need to introduce $\squash{X}$, the
 {\em (propositional) truncation} of $X$ (also called bracket type, or
@@
 
-Given any type $X$, its propositional truncation 
-$\squash{X} : \HProp$ is generated by 
+Given any type $A$, its propositional truncation 
+$\squash{A} : \HProp$ is generated by 
 \begin{itemize}
-\item a function $|\cdot|_X : X \to \squash{X}$,
-\item for any $x,y:\squash{X}$, a path $x=y$.
+\item a function $|\cdot|_A : A \to \squash{A}$,
+\item for any $x,y:\squash{A}$, a path $x=y$.
 \end{itemize}
 % 
-The recursion principle $\squash{A}$ of asserts that any mere
+The recursion principle of $\squash{A}$  asserts that any mere
 proposition that follows from $A$ already follows from $\squash{A}$.
@@
 Using higer inductive types, every diagram $D$ on a graph $G$ admits a
-colimit in homotopy type thoery. This colimit $\colim D$ is given
+colimit in homotopy type thoery. This colimit $\colim D$ is given by
 %
@@
 %
-Although we have proven in our formalization that $\colim D$ is
+Although we have proved in our formalization that $\colim D$ is
 actually a colimit over $D$, we do not detail the proof here as the
@@
 axioms that characterize a topos. Namely, there are 4 axioms on a
-$(\infty,1)$-category that have been shown to be equivalent to be
-equivalent to $(\infty,1)$-topos~\cite[Chapter 6]{lurie}. This is not
-the purpose of this paper to discuss those axioms, we only state here
-a particular case of the last axiom which says that every groupoid
-object is effective. 
+$(\infty,1)$-category that have been shown to be equivalent to
+$(\infty,1)$-topos~\cite[Chapter 6]{lurie}. This is not the purpose of
+this paper to discuss those axioms, we only state here a particular
+case of the last axiom which says that every groupoid object is
+effective.
 %
@@
 
-With the point of view that homotopy type theory is a type-theoristic
+With the point of view that homotopy type theory is a type-theoretic
 version of higher topos theory, Theorem 6.1.0.6 of~\cite{lurie}
@@
 \cite{shulman-higher-modalities}. Our definition differs slightly from
-those work in that we restrict the definition of the modality to be
+those work in that we restrict the definition of the modality to be on
 $n$-truncated types for a given truncation index $n$, instead of on
@@
 because Lawvere-Tierney topologies corresponds exactly to left exact
-modalities at $-1$ and sheafification is described as an inductive
+modalities at level $-1$ and sheafification is described as an inductive
 process that produces a left exact modality at level $n+1$ from a
@@
 $$
-\Type^\modal_n \ \defeq \sum_{T : \Type_n} P (T).
+\Type_n^\modal \ \defeq \sum_{T : \Type_n} P (T).
 $$
@@
 properties, that can make it inconsistent. We now provide a simple
-characterization of left exact modalities that give rise to a
+characterization of left exact modalities that gives rise to a
 consistent theory.
@@
 A first example of left exact modality for every level $p$ is the {\em
-  open modality of $P$} for a $P:\HProp$ defined as
+  open modality of mere proposition $P$} defined as
 $\modal_P T = P \to T$.
@@
 %
-Thus, being in this subuniverse corresponds to adding the axiom $P$ to the
+Thus, working in this subuniverse amounts to add the axiom $P$ to the
 theory. Of course, adding an axiom this way does not add any computational
@@
 
-Another example of left exact modality for every level $p$ is the
+Another example of a left exact modality for every level $p$ is the
 double negation modality,
@@
 In this section, we suppose given a truncation index $n\geqslant -1$,
-and a left exact modality $\modal$ on $\Type_n$ and we introduce the
+and a left exact modality $\modal_n$ on $\Type_n$ and we introduce the
 definition of sheaves for $(n+1)$-types and basic properties of
@@
 To define sheaves, we first have to leverage the notion of dense
-subobjects for the modality $\modal$. The closure operator is defined
+subobjects for the modality $\modal_n$. The closure operator is defined
 using the $n$-subobject classifier $\Type_n$ introduced in
@@
   classified by $\chi : E \to \Type_n$, is the subobject of $E$
-  classified by $\modal \circ \chi$.
+  classified by $\modal_n \circ \chi$.
 
@@
     closed in $E$} if it is equal to its closure, \ie
-  $\chi = \modal \circ \chi$.
+  $\chi = \modal_n \circ \chi$.
 \end{defi}
@@
   A \to E$ of $E$
-  classified by $\chi$ is {\em dense} in $E$ when its $\modal$-closure
+  classified by $\chi$ is {\em dense} in $E$ when its $\modal_n$-closure
   is equvialent to $\chi_E$, \ie
   %
-  $$\forall e:E,~ \left(\sum_{e':E} e=e'\right) \simeq (\modal
-  (\chi~e)).$$
+  $$\forall e:E,~ \left(\sum_{e':E} e=e'\right) \simeq \modal_n
+  (\chi~e).$$
   % 
@@
     \sum_{e':A} x = e' \ar@{=}[r] \ar[d]_{\iota'} & (\chi~x) \ar[d]^{\eta_{(\chi~x)}}\\
-    \sum_{e':E} x = e' \ar@{=}[r] & \modal (\chi~x)
+    \sum_{e':E} x = e' \ar@{=}[r] & \modal_n (\chi~x)
   }$$
@@
 
-  As $\chi$ is dense, we have a term $m_x : \modal(\chi\, x)$.
+  As $\chi$ is dense, we have a term $m_x : \modal_n(\chi\, x)$.
   But as $\phi_2 x$ is modal, we can obtain a term $h_x : \chi\,
@@
   We define $g$ as 
-  $$g(e) = \modal \left( \fib \phi e \right), $$
+  $$g(e) = \modal_n \left( \fib \phi e \right), $$
   where
@@
     \item If all fibers of $f$ and $g$ are $n$-truncated, then
-      $$\forall c:C,~\modal(\fib {g \circ f} c) =
-      \modal \left(  
-        \sum_{w:\fib g c} \modal (\fib f {w_1})
+      $$\forall c:C,~\modal_n(\fib {g \circ f} c) =
+      \modal_n \left(  
+        \sum_{w:\fib g c} \modal_n (\fib f {w_1})
       \right).$$
@@
 Nevertheless, we are confident that with an enrichment of the Coq/HoTT
-library, they can be proven and have been admitted only by lack of
+library, they can be proved and have been admitted only by lack of
 time.
@@
 When the absence of cumulativity on $\Type_n$ rises a universe issue
-for a lemma, we have proven the lemma and then admit the exact same
+for a lemma, we have proved the lemma and then admit the exact same
 lemma, just letting Coq infer less constrained universe levels.

### 2015-01-20 paper/sheaf_lics.tex
complete pass

@@
 %
-We note $\prod_{x:T} U$ for dependent products, $\sum_{x:T} U$ for
-dependent sums, and $\pi_1, \pi_2$ for the first and second
-projection of a dependent pair. We use informal mathematical language
-instead of type theory whenever it is possible, to ease the
-reading without making our statement imprecise. In particular,
-(higher) inductive types are defined using itemization to avoid an
-overhead of notation.
+Given a type $T$ and a type family $U : T \to \Type$, we note
+$\prod_{x:T} U x$ for the dependent product, $\sum_{x:T} U x$ for the dependent
+sum, and $\pi_1, \pi_2$ for the first and second projection of a
+dependent pair (noted $(a;b)$). We use informal mathematical language
+instead of type theory whenever it is possible, to ease the reading
+without making our statement imprecise. In particular, (higher)
+inductive types are defined using itemization to avoid an overhead of
+notation.
 
@@
 
-As explained in the introduction, a Lawvere-Tierney topology for a
-topos $\E$, as defined in Section~\ref{sec:lawv-tiern-sheaf},
-corresponds to a left exact modality at level $-1$
+As explained in in Section~\ref{sec:lawv-tiern-sheaf}, a
+Lawvere-Tierney topology for a topos $\E$ corresponds to a left exact
+modality at level $-1$
 %
@@
   $$\xymatrix{
-    \sum_{e':A} x = e' \ar@{=}[r] \ar[d]_{\iota'} & (\chi~x) \ar[d]^{\eta_{(\chi~x)}}\\
-    \sum_{e':E} x = e' \ar@{=}[r] & \modal_n (\chi~x)
+    \sum_{e':A} x = e' \ar@{=}[r] \ar[d]_{\iota'} & (\chi \ x)
+    \ar[d]^{\eta_{(\chi \ x)}}\\
+    \sum_{e':E} x = e' \ar@{=}[r] & \modal_n (\chi \ x)
   }$$
@@
     as the map sending an
-    arrow $f:E\to F$ to its restriction the restriction of $f \circ m$.
+    arrow $f:E\to F$ to its restriction $f \circ m$.
   % \item if $A \to E$ with $n$-truncated homotopy fibers, $\Phi_E^\chi$ is the
@@
 it requires much work here, as we have to deal with equalities. In
-particular, this property uses the specific definition
-of separated types and sheaves, which shows that we had actually no
-choice in their statement.
+particular, this property uses the specific definitions
+of separated types and sheaves, which has constituted a guide in the
+choice of their precise statements.
 
 \begin{prop}\label{prop:sheaf-is-sheaf}
-  $\Type_n^\modal$ is a sheaf.
+  $\Type_n^\modal$ is a sheaf.\footnote{Note that we omit the $n$
+on $\modal_n$ in $\Type_n^\modal$ to ease the reading.}
 \end{prop}
@@
 added in the first modality already. Sheafification is just a way
-to propagate those new principles to types than are not mere
+to propagate those new principles to types that are not mere
 propositions.
@@
 
-  Let $k:\nat$, we will show that 
+  Let $k:\nat$, we have to show that 
   \begin{align*}
@@
 
-To show that $\separated$ is a modality, it remains to show point {\it (v)}. Let $A:\Type_{n+1}$ be a sheaf and $B:A \to \Type_{n+1}$ be
-a sheaf family. We want to show that $\sum_{x:A} (Bx)$ is
-separated. Let $E$ be a type, and $\sum_{e:E} (\chi\,e)$ a dense
-subobject of E.
+To show that $\separated$ is a modality, it remains to show point
+{\it (v)} of the definition.
+%
+Let $A:\Type_{n+1}$ be a sheaf and $B:A \to \Type_{n+1}$ be a sheaf
+family. We want to show that $\sum_{x:A} (Bx)$ is separated. Let $E$
+be a type, and $\sum_{e:E} (\chi\,e)$ a dense subobject of E.
 
@@
 
-As this modality is just a step in the construction, we do not show
-that it is left exact, we will only do it for the sheafification modality.
+As this modality is just a step in the construction, we do not need to
+show that it is left exact, we will only do it for the sheafification
+modality.
 
@@
 %
-Explicitly
+$\modal_{n+1}T$ can be given explicitly by
 $$
@@
 with $\modal_{-1}$. To do that, we need to extend the notion of
-compatibility and that actually every modality $\modal_{n+1}$ is
+compatibility and show that actually every modality $\modal_{n+1}$ is
 compatible with $\modal_n$ on lower homotopy types.
@@
     every $t:T$, $\IsType n (u\, t)$.  But for any truncation index $p$,
-    the predicate $\IsType p:\HProp$ is a a sheaf, so we can get rid
+    the predicate $\IsType p:\HProp$ is a sheaf, so we can get rid
     of $\modal_{-1}$ and of the truncation, which tells us that for
@@
 
-The last step is the left-exactness of $\modal_{n+1}$. Let
-$T$ be in $\Type_{n+1}$ such that $\Contr(\modal_{n+1} T)$.
-Thanks to compatibility between $\modal_{n+1}$ and $\modal_n$ for
-$\Type_n$, left exactness means that for any $x,y : T$, $\modal_n(x=y)$ is contractible.
+The last step is the left-exactness of $\modal_{n+1}$. Let $T$ be in
+$\Type_{n+1}$ such that $\Contr(\modal_{n+1} T)$.  Thanks to the just
+shown compatibility between $\modal_{n+1}$ and $\modal_n$ for
+$\Type_n$, left exactness means that for any $x,y : T$,
+$\modal_n(x=y)$ is contractible.
 
@@
 have defined for any truncation level $n$, a new left-exact modality
-$\modal_n$ on $\Type_n$, which is exactly $\modal_{-1}$ when
+$\modal_n$ on $\Type_n$, which corresponds to $\modal_{-1}$ when
 restricted to $\HProp$.
@@
   %
-  induces a type theory where the propositional excluded middle holds.
+  induces a type theory where the propositional excluded middle law holds.
 \end{prop}
+
+Combined with forcing in type theory~\cite{jaber2012extending}, it
+should be possible to lift the proof of independence of the continuum
+hypothesis to a classical setting, which is where the continuum hypothesis is
+really meaningful.  However, we haven't work out the details and left
+this for future work.
 
@@
 \url{https://github.com/HoTT/HoTT} and uses the version of Coq
-designed for that library.
-%
+tailored for that library.
+%
+After reviewing the content and some statistics about the
+formalization in Section~\ref{sec:cont-form}, we present the
+limitations of our formalization in Section~\ref{sec:limit-form}, in
+particular the issues relative to universe polymorphism.
 
 \subsection{Content of the formalization}
+\label{sec:cont-form}
 
@@
 \subsection{Limitations of the formalization}
+\label{sec:limit-form}
 
@@
 %
+This is not surprising because those lemmas are defined in a
+call-by-need approach and there is no notion of a complete set of
+lemmas on paths for the moment. 
+%
+This is the subject of ongoing work, in particular those trying to
+define a syntax based on the cubical set model~\cite{bezem2013model}
+of Coquand et al.  
+%
 Nevertheless, we are confident that with an enrichment of the Coq/HoTT
@@
 closed diagram is admitted too, but again, this proof should be really
-short as it is on paper.
+as short as it is on paper.
 
 More importantly, there are few axioms in the formalization that are
-due to an unsatisfactory management of polymorphic universe level in
+due to an unsatisfactory management of polymorphic universe levels in
 Coq.
@@
 formalization, its implementation is still to rigid to allow a
-complete formalization of our work.
-
-%
-If Coq handles cumulativity on all $\Type$ natively, it is not
+complete formalization of our work for the following reasons.
+
+%
+If Coq handles cumulativity on $\Type$ natively, it is not
 the case for the $\Sigma$-type $\Type_n$, which require propositional
@@
 When the absence of cumulativity on $\Type_n$ rises a universe issue
-for a lemma, we have proved the lemma and then admit the exact same
-lemma, just letting Coq infer less constrained universe levels.
+for a lemma (it occurs twice in the formalization), we have proved the
+lemma and then admit the exact same lemma, just letting Coq infer less
+constrained universe levels.
 
@@
 %
-This restriction in our formalization can not be solved without
-generalizing universe polymorphism for recursive definition.
-%
+This restriction in our formalization may be solved by
+generalizing universe polymorphism for recursive definition
+%
+or with the use of general resizing rules.
 
@@
 Using sheafification, it is possible to add new logical principles to
-homotopy type theory, like the law of excluded middle, and this
+homotopy type theory, such as the law of excluded middle, and this
 without axiom.
 %
-Beyond this result, our work is part of a more general line of works
+Beyond this result, our work is part of a more general line of work
 which aim to illustrate that homotopy type theory is a very promising
@@
 We would like to thank all maintainers of the Coq/HoTT library, who
-provide a very solid base for our formalization. We also would like
+have provided a very solid base for our formalization. We also would like
 to thank Bas Spitters for helpful discussions on the subject.

### 2015-01-20 paper/sheaf_lics.tex
added Resk completion

@@
 types~\cite{lumsdaine2011higher,lumsdaine2013higher} and
-Voevodsky's univalence principle~\cite{kapulkin2012simplicial}. 
-%
-
+Voevodsky's univalence principle~\cite{kapulkin2012simplicial}, which
+says that for any types $T$ and $U$, the canonical map 
+$$
+(T = U) \to (T \simeq U)
+$$ 
+%
+which transports equalities to equivalences is an equivalence.
+%
 This new point of view on type theory has revealed the homotopy
@@
 paths.\footnote{This proposition comes from on old version of the
-  formalization of left exact modalities in the HoTT library, but is
+  formalization of left exact modalities in the Coq/HoTT library, but is
   not present anymore in the library.}
@@
 $\modal_{n}$ because of the use of $\Sigma$-type over
-$T \to \Type_n^\modal$.
+$T \to \Type_n^\modal$ and equality on the return type of $\modal_n$.
 %
@@
 moment.
+
+Note that the same increasing in the universe levels occurs in the
+Resk completion for categories~\cite{resk}. In the definition of the
+completion, they use the Yoneda embedding and representable functors,
+which is similar to our use of characteristic functions.
+
 %
 This restriction in our formalization may be solved by
-generalizing universe polymorphism for recursive definition
-%
-or with the use of general resizing rules.
+generalizing the management of universe polymorphism for recursive definition
+%
+or by the use of general ``resizing axiom'' which is still under
+discussion in the community.
 

### 2015-01-20 paper/sheaf_lics.tex
types + cocone + isSheaf (IsTrunc p)

@@
 
-But furthermore, one can defined a left adjoint to the inclusion, the
+Furthermore, one can define a left adjoint to the inclusion, the
 sheafification functor
@@
 assuming that the sheafification has been constructed up to level $n$,
-one can defined the sheafification for $\Type_{n+1}$.
+one can define the sheafification for $\Type_{n+1}$.
 %
@@
 reflective subuniverse, which corresponds to a stratified version of
-left-exact modalities introduced in~\cite[Chapter~7]{hottbook}. A
+left-exact modality introduced in~\cite[Chapter~7]{hottbook}. A
 Lawvere-Tierney topology can thus be seen as a left-exact modality on
@@
 fibers~\cite{sets_in_hott}. This means that when monomorphisms are used in topos-theoretic definition, the correct generalization may be $n$-truncated homotopy
-fibers or embeddings (\ie $-1$-truncated homotopy
+fibers or embeddings (\ie $(-1)$-truncated homotopy
 fibers) depending on the purpose of the definition.
@@
 %
-The definition of Section~\ref{ssec:hott} and \ref{sec:epi-mono-fact}
+The definitions of Section~\ref{ssec:hott} and \ref{sec:epi-mono-fact}
 are part of the Coq library
@@
 \begin{itemize}
-\item $\IsContr := \IsType {(-2)}$ and $\Contr := \Type_{-2}$
-\item $\IsHProp := \IsType {(-1)}$ and $\HProp := \Type_{-1}$
+\item $\IsContr \defeq \IsType {(-2)}$ and $\Contr \defeq \Type_{-2}$
+\item $\IsHProp \defeq \IsType {(-1)}$ and $\HProp \defeq \Type_{-1}$
 \end{itemize}
@@
 
-A map $f:A\to B$ is an \emph{embedding} when it is (-1)-truncated
+A map $f:A\to B$ is an \emph{embedding} when it is $(-1)$-truncated
 (see definition in Section~\ref{ssec:hott}), and a \emph{surjection}
-when it is (-1)-connected, that is when every fiber of $f$ is merely
+when it is $(-1)$-connected, that is when every fiber of $f$ is merely
 inhabited (\ie $\|\fib f y\|$ holds for all $y$).
@@
 Note that this generalizes to orthogonal factorization systems
-constituted by $n$-truncated morphism and $n$-connected morphisms, for
+constituted by $n$-truncated morphisms and $n$-connected morphisms, for
 every truncated level $n$. We will only use the factorization system
@@
 The main difference between limits and colimits is that limits are
-simply given by $\Sigma$-types, and thus exists already in traditional
+simply given by $\Sigma$-types, and thus exist already in traditional
 type theory, whereas the situation is more complicated for colimits as
@@
 A colimit of a diagram $D$ over a graph $G$ is given by a type $P$
-that defines a cone of $D$, plus the universal property that for any
+that defines a cocone of $D$, plus the universal property that for any
 type $X$, the canonical map that transforms a function $f : P
-\rightarrow X$ to a cone of $D$ on $X$ is an isomorphism.
+\rightarrow X$ to a cocone of $D$ on $X$ is an isomorphism.
 % 
@@
 
-Using higer inductive types, every diagram $D$ on a graph $G$ admits a
-colimit in homotopy type thoery. This colimit $\colim D$ is given by
+Using higher inductive types, every diagram $D$ on a graph $G$ admits a
+colimit in homotopy type theory. This colimit $\colim D$ is given by
 %
@@
   }$$
-  with $p$ copies of $X$. The $0$-pullback of $f$ is $\Unit$.
+  with $p$ copies of $X$. The $0$-pullback of $f$ is $\one$.
 \end{defi}
@@
 The notion of left exact modalities plays a central role in this paper
-because Lawvere-Tierney topologies corresponds exactly to left exact
+because Lawvere-Tierney topologies correspond exactly to left exact
 modalities at level $-1$ and sheafification is described as an inductive
@@
 Finally, Section~\ref{sec:examples-left-exact} introduces some
-example, and in particular the modality for the dense topology.
+examples, and in particular the modality for the dense topology.
 
@@
 
-As noticed in \cite{hottbook}, left exact modalities corresponds 
+As noticed in \cite{hottbook}, left exact modalities correspond
 to sub-$(\infty,1)$-toposes. Indeed, modal types defines a new type
@@
 theory. Of course, adding an axiom this way does not add any computational
-content to $P$, and the consistency of the modallity corresponds to
+content to $P$, and the consistency of the modality corresponds to
 the fact that $P \to \zero$ (or $\lnot P$) is not inhabited.
@@
 This means that we need to take more care of the homotopy structure of
-this charactestic function, and in particular, the definition of a
+this characteristic function, and in particular, the definition of a
 dense $n$-subobject comes with an additional coherence condition
@@
   classified by $\chi$ is {\em dense} in $E$ when its $\modal_n$-closure
-  is equvialent to $\chi_E$, \ie
+  is equivalent to $\chi_E$, \ie
   %
@@
     $\Sigma$-type above, $u$ is in $T\to\Type_n^\modal$, \ie that for
-    every $t:T$, $\IsType n (u\, t)$.  But for any truncation index $p$,
-    the predicate $\IsType p:\HProp$ is a sheaf, so we can get rid
+    every $t:T$, $\IsType n (u\, t)$.  But for any truncation index
+    $p$,
+    the type $\IsType p X:\HProp$ is a sheaf as soon as $X$ is, so we can get rid
     of $\modal_{-1}$ and of the truncation, which tells us that for
@@
 Although the presence of universe polymorphism is mandatory for our
-formalization, its implementation is still to rigid to allow a
+formalization, its implementation is still too rigid to allow a
 complete formalization of our work for the following reasons.

### 2015-01-20 paper/sheaf_lics.tex
some more typos

@@
 theory uses the property that epimorphisms are coequalizers of their
-kernel pairs. This properties does not hold anymore in higher toposes, but
+kernel pairs. This property does not hold anymore in higher toposes, but
 using Giraud-Rezk-Lurie axioms of $\infty$-toposes, the same property
@@
 equalities which are trival (by proof irrelevance) in the topos-theoretic
-setting becomes relevant and requires a carreful use of univalence
+setting become relevant and require a careful use of univalence
 with appropriate isomorphisms and a lot of reasoning on equalities
@@
 Since basic type formers (dependent products, products, sigma types)
-preserves truncation levels, we can leverage all theorems on modalites
+preserve truncation levels, we can leverage all theorems on modalities
 presented in~\cite[Chapter 7.7]{hottbook}.
@@
 would really complicate the proofs as the axiom would appear
-everywhere cumulativity is needed and it would need explicitly
-annotation for universe levels everywhere in the formalization.
+everywhere cumulativity is needed and it would need explicit
+annotations for universe levels everywhere in the formalization.
 %

### 2015-01-20 paper/sheaf_lics.tex
again more typos

@@
 \begin{lem}
-  Let $X:\Type_p$, $Y:\Type_p\modal$ and $f:X\to Y$. If for all $y:Y$,
+  Let $X:\Type_p$, $Y:\Type_p^\modal$ and $f:X\to Y$. If for all $y:Y$,
   $\modal (\fib f y)$ is contractible, then $\modal X = Y$.
@@
 sheaves are elements that can not distinguish between objects and
-their dense sub-objects.
+their dense subobjects.
 

### 2015-10-12 paper/sheaf_lics.tex
finish Simon's proof

@@
 $(\infty,1)$-category that have been shown to be equivalent to
-$(\infty,1)$-topos~\cite[Chapter 6]{lurie}. This is not the purpose of
-this paper to discuss those axioms, we only state here a particular
-case of the last axiom which says that every groupoid object is
-effective.
+$(\infty,1)$-topos~\cite[Chapter 6]{lurie}. In~\cite{boulier}, a
+type-theoritic version of these axioms are proposed.
 %
@@
 \begin{defi}
-  Let $f:X \to Y$ be a map, and $p:\nat$. The $p$-pullback of $f$
-  (also called the $p$-fold fiber product of $X$ over $Y$),
-  noted $X\times_Y \cdots\times_Y X$, is
-  the limit of the diagram
+  Let $f:X \to Y$ be a map. $T_f$ is the higher inductive type given
+  by
+  \begin{itemize}
+  \item $t:~X \to T_f$
+  \item $\alpha:~\forall a\,b:X,~f a = f b \to t a = t b$
+  \item $\alpha_1:~\forall a:X,~\alpha\, a\, a\, 1 = 1$
+  \end{itemize}
+  We view $T_f$ as the coequalizer of
   $$\xymatrix{
-    X \ar[rd]^f & \cdots & X \ar[ld]_f \\
-      &    Y   &
+    \sum_{a,b:X} f a = f b \dar[r]{4pt}^-{\pi_1}_-{\pi_2} & X
   }$$
-  with $p$ copies of $X$. The $0$-pullback of $f$ is $\one$.
+  preserving the identity.
+
+  We call $\tilde f$ the map $T_f \to Y$ given by induction.
 \end{defi}
-
-As explained in Section~\ref{sec:colim-homot-type}, every limit exists
-in (homotopy) type theory and can be described by a $\Sigma$-type. 
-%
-The particular case of a $p$-pullback is given by
-%
-$$
-X\times_Y \cdots\times_Y X = \sum_{x:X^n} (f x_1 = f x_2) \land
-\cdots \land (f x_{n-1} = f x_n).
-$$
 %
@@
 simplicial object that, at degree $p$, is given by the
-$(p+1)$-pullback of $f$. The \v{C}ech nerve can thus also be described
-as a diagram over $\Delta^{op}$ (where $\Delta$ is the simplicial
-category, seen as a graph). For simplicity, we only give an intuitive
-definition here, the complete definition can be found in our
-formalization.
+$p$th iteration of $T$. 
 %
@@
   $$\xymatrix{
-    C(f) := \cdots~ X \times_Y X \times_Y X \tar[r]{4pt} & X \times_Y X \dar[r]{2pt} & X
+    C(f) := X \ar[r]^t & T_f \ar[r]^t & T_{\tilde f} \ar[r] &\cdots
+% \cdots~ X \times_Y X \times_Y X \tar[r]{4pt} & X \times_Y X \dar[r]{2pt} & X
   }$$
-where the arrows are the canonical projections.
 \end{defi}
 
-\begin{ax}[Giraud-Rezk-Lurie]
+\begin{thm}[Giraud-Rezk-Lurie]
   For any surjection $f : X \to Y$, the colimit of its \v{C}ech nerve
   $C(f)$ is $Y$.
-\end{ax}
+\end{thm}
 
 With the point of view that homotopy type theory is a type-theoretic
-version of higher topos theory, Theorem 6.1.0.6 of~\cite{lurie}
-suggests that this axiom can be proved in homotopy type theory. 
-%
-But proving this axiom seems to be a quite complex task on its own,
-and we have decided to admit it in this paper.
+version of higher topos theory, this seems to be the analogous of Theorem 6.1.0.6 of~\cite{lurie}.
+%
+The proof of this theorem can be found in~\cite{boulier}.
 
@@
 
+\begin{defi}
+  Let $X:\Type$. Let $\mathring T_X$ be the higher inductive type
+  generated by
+  \begin{itemize}
+  \item $\mathring t:~\|X\|_{n+1} \to \mathring T_X$
+  \item $\mathring \alpha:~\forall a\, b:\|X\|_{n+1},~\modal (a=b) \to
+    \mathring t a = \mathring t b$
+  \item $\mathring \alpha_1:~\forall a:\|X\|_{n+1},~
+    \mathring \alpha\, a \, a\, (\eta_{a=a} 1) = 1$
+  \end{itemize}
+
+  We view $\mathring T$ as the coequalizer of
+  $$\xymatrix{\sum_{a,b:\|X\|_{n+1}} \modal (a=b) \dar[r]{4pt}^-{\pi_1}_-{\pi_2}
+  & \|X\|_{n+1}}$$
+  preserving $\eta_{a=a} 1$.
+
+  We consider the diagram $\mathring T$ :
+  $$\xymatrix{\|X\|_{n+1} \ar[r] & \|\mathring T_{X}\|_{n+1} \ar[r] & \|\mathring
+  T_{\mathring T_X} \|_{n+1} \ar[r] & \cdots} $$
+\end{defi}
+
 \begin{lem}\label{lem:sepiscolim}
-  Let $T:\Type_{n+1}$. Then $\separated T$ is the colimit of the closed
-  diagonal diagram
-  $$\xymatrix @C=4em  { 
-    \cdots \overline{\Delta_3} \tar[r]{4pt} & \overline{\Delta_2} \dar[r]{2pt} &
-    \overline{\Delta_1}
-  }$$
-where $\Delta_k$ is the $k$-pullback of $\id : T \to T$.
+  Let $T:\Type_{n+1}$. Then $\separated T$ is the $(n+1)$-colimit of the
+  diagram $\mathring T$.
 \end{lem}
@@
 
-\begin{proof}
+We will need the following lemma:
+
+\begin{lem}
+  Let $A,S:\Type_{n+1}$, $S$ separated, and $f:A \to S$. Then if 
+  $$\forall a,b:A,~f a = f b \simeq \modal (a=b),$$ then 
+  $$\forall a,b:\|T_f\|_{n+1},~|\tilde f|_{n+1} a = |\tilde f|_{n+1} b \simeq \modal (a=b).$$
+\end{lem}
+
+\begin{proof}[Sketch of proof]
+  By induction on truncation, we need to show that 
+  $$\forall a,b:T_f,~\tilde f |a|_{n+1} = \tilde f |b|_{n+1} \simeq
+  \modal (|a|_{n+1}=|b|_{n+1}).$$
+
+  We use the encode-decode method to characterize $\tilde f |a|_{n+1}
+  = x$, and the result follows. We refer to the formalization for details.
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:sepiscolim}]
   As $\mu_T$ is an surjection, using the particular case of 
   Giraud-Rezk-Lurie axiom introduced in Section~\ref{sec:giraud-ax},
-  we know that $\separated T$ is the colimit of $C(\mu_T)$. 
-%
-  Thus the result follows from the fact that $C(\mu_T) = C(\id)$,
+  we know that $\separated T$ is the $(n+1)$-colimit of $\|C(\mu_T)\|_{n+1}$. 
+%
+  Thus the result follows from the fact that $\|C(\mu_T)\|_{n+1} =
+  \mathring T$,
   which we now show.
 
-  Let $k:\nat$, we have to show that 
+  By induction, we need to show that for any $A,B:\Type_{n+1}$ such
+  that $\varphi:~A \simeq B$, with an arrow $f:A \to \separated T$, 
+  $$T_f = \mathring T_B.$$
+  For this, we use 
   \begin{align*}
-    &\sum_{t:T^k} (\mu_T t_1 = \mu_T t_2 \land \cdots
-      \land \mu_T t_{k-1} \mu_T t_k)\\
-    &= \sum_{t:T^k} \modal_n (t_1 = t_2 \land \cdots
-      \land t_{k-1} = t_k)
+    T_f &= \mathrm{Coeq}_1 \left( 
+          \xymatrix{
+          \sum_{a,b:A} f a = f b \dar[r]{4pt}^-{\pi_1}_-{\pi_2} & A
+  }
+                                                                  \right)
+    \\
+    \mathring T_B &= \mathrm{Coeq}_1 \left( 
+                    \xymatrix{\sum_{a,b:B} \modal (a=b) \dar[r]{4pt}^-{\pi_1}_-{\pi_2}
+                                                                & B
+                                                                  \right)
   \end{align*}
-By induction on $k$, and the preservation of products by $\modal_n$, it
-suffices to show that for any $a,b:T$, $\modal_n (a=b) = (\mu_T a =
-\mu_T b)$. By univalence, we want arrows in both ways, forming an
-equivalence.
-\begin{itemize}
-\item Suppose $p : (\mu_T a = \mu_T b)$. Then projecting $p$ along
-  first components yields $q : \prod_{t:T} \modal_n(a=t) = \modal_n (b=t)
-  $.
-  Taking for example $t=b$, we deduce $\modal_n (a=b) = \modal_n(b=b)$,
-  and the latter is inhabited by $\eta_{b=b} 1$.
-\item Suppose now $p : \modal_n(a=b)$. Let $\iota$ be the first
-  projection from $\separated T \to (T \to \Type_n^\modal)$. $\iota$ is
-  an embedding, thus it suffices to prove $\iota (\mu_T a) = \iota
-  (\mu_T b)$, \ie $\prod_{t:T}\modal_n (a=t) = \modal_n (b=t)$. The latter
-  remains true by univalence.
-\end{itemize}
-The fact that these two form an equivalence is technical, we refer to
-the formalization for an explicit proof.
+
+%   Let $k:\nat$, we have to show that 
+%   \begin{align*}
+%     &\sum_{t:T^k} (\mu_T t_1 = \mu_T t_2 \land \cdots
+%       \land \mu_T t_{k-1} \mu_T t_k)\\
+%     &= \sum_{t:T^k} \modal_n (t_1 = t_2 \land \cdots
+%       \land t_{k-1} = t_k)
+%   \end{align*}
+% By induction on $k$, and the preservation of products by $\modal_n$, it
+% suffices to show that for any $a,b:T$, $\modal_n (a=b) = (\mu_T a =
+% \mu_T b)$. By univalence, we want arrows in both ways, forming an
+% equivalence.
+% \begin{itemize}
+% \item Suppose $p : (\mu_T a = \mu_T b)$. Then projecting $p$ along
+%   first components yields $q : \prod_{t:T} \modal_n(a=t) = \modal_n (b=t)
+%   $.
+%   Taking for example $t=b$, we deduce $\modal_n (a=b) = \modal_n(b=b)$,
+%   and the latter is inhabited by $\eta_{b=b} 1$.
+% \item Suppose now $p : \modal_n(a=b)$. Let $\iota$ be the first
+%   projection from $\separated T \to (T \to \Type_n^\modal)$. $\iota$ is
+%   an embedding, thus it suffices to prove $\iota (\mu_T a) = \iota
+%   (\mu_T b)$, \ie $\prod_{t:T}\modal_n (a=t) = \modal_n (b=t)$. The latter
+%   remains true by univalence.
+% \end{itemize}
+% The fact that these two form an equivalence is technical, we refer to
+% the formalization for an explicit proof.
 \end{proof}

### 2015-11-02 paper/sheaf_lics.tex
Finish

@@
 \end{defi}
+
+Colimits behave well with truncations, in the following sense:
+
+\begin{lem}
+  Let $D$ be a diagram, $m$ a truncation index, and
+  $P:\Type_{m}$. Let's call $P$ a $m$-colimit of $D$ if
+  for any other $X:\Type_m$, the
+  map
+  $$\lambda f:P \to X, \left( \lambda i,~f \circ q_i~;~ \lambda i\, j\,
+    \phi,~ \postcompose_f (p(i, j, g)) \right)$$
+  is an equivalence.
+
+  Then, if $\|D\|_m$ is the same diagram as $D$, but where every type
+  is $m$-truncated, $P$ is a $m$-colimit of $\|D\|_m$.
+\end{lem}
 
@@
   Let $A,S:\Type_{n+1}$, $S$ separated, and $f:A \to S$. Then if 
-  $$\forall a,b:A,~f a = f b \simeq \modal (a=b),$$ then 
+  \begin{equation}
+    \label{eq:Omono}\tag{$\star$}
+    \forall a,b:A,~f a = f b \simeq \modal (a=b),
+  \end{equation}
+  then
   $$\forall a,b:\|T_f\|_{n+1},~|\tilde f|_{n+1} a = |\tilde f|_{n+1} b \simeq \modal (a=b).$$
@@
 
-  By induction, we need to show that for any $A,B:\Type_{n+1}$ such
-  that $\varphi:~A \simeq B$, with an arrow $f:A \to \separated T$, 
-  $$T_f = \mathring T_B.$$
+  First, we show that $\mu_T$ satisfies condition~(\ref{eq:Omono}),
+  \ie{} $\forall a,b:T$, $\modal_n (a=b) = (\mu_T a =
+  \mu_T b)$. By univalence, we want arrows in both ways, forming an
+  equivalence.
+  \begin{itemize}
+  \item Suppose $p : (\mu_T a = \mu_T b)$. Then projecting $p$ along
+    first components yields $q : \prod_{t:T} \modal_n(a=t) = \modal_n (b=t)
+    $.
+    Taking for example $t=b$, we deduce $\modal_n (a=b) = \modal_n(b=b)$,
+    and the latter is inhabited by $\eta_{b=b} 1$.
+  \item Suppose now $p : \modal_n(a=b)$. Let $\iota$ be the first
+    projection from $\separated T \to (T \to \Type_n^\modal)$. $\iota$ is
+    an embedding, thus it suffices to prove $\iota (\mu_T a) = \iota
+    (\mu_T b)$, \ie $\prod_{t:T}\modal_n (a=t) = \modal_n (b=t)$. The latter
+    remains true by univalence.
+  \end{itemize}
+  The fact that these two form an equivalence is technical, we refer to
+  the formalization for an explicit proof.
+
+  Then, by induction, we need to show that for any $A,B:\Type_{n+1}$ such
+  that $\varphi:~A \simeq B$, with an arrow $f:A \to \separated T$
+  satisfying~(\ref{eq:Omono}), 
+  $$T_f \simeq \mathring T_B.$$
   For this, we use 
@@
                     \xymatrix{\sum_{a,b:B} \modal (a=b) \dar[r]{4pt}^-{\pi_1}_-{\pi_2}
-                                                                & B
+                                                                & B}
                                                                   \right)
@@
 
-%   Let $k:\nat$, we have to show that 
-%   \begin{align*}
-%     &\sum_{t:T^k} (\mu_T t_1 = \mu_T t_2 \land \cdots
-%       \land \mu_T t_{k-1} \mu_T t_k)\\
-%     &= \sum_{t:T^k} \modal_n (t_1 = t_2 \land \cdots
-%       \land t_{k-1} = t_k)
-%   \end{align*}
-% By induction on $k$, and the preservation of products by $\modal_n$, it
-% suffices to show that for any $a,b:T$, $\modal_n (a=b) = (\mu_T a =
-% \mu_T b)$. By univalence, we want arrows in both ways, forming an
-% equivalence.
-% \begin{itemize}
-% \item Suppose $p : (\mu_T a = \mu_T b)$. Then projecting $p$ along
-%   first components yields $q : \prod_{t:T} \modal_n(a=t) = \modal_n (b=t)
-%   $.
-%   Taking for example $t=b$, we deduce $\modal_n (a=b) = \modal_n(b=b)$,
-%   and the latter is inhabited by $\eta_{b=b} 1$.
-% \item Suppose now $p : \modal_n(a=b)$. Let $\iota$ be the first
-%   projection from $\separated T \to (T \to \Type_n^\modal)$. $\iota$ is
-%   an embedding, thus it suffices to prove $\iota (\mu_T a) = \iota
-%   (\mu_T b)$, \ie $\prod_{t:T}\modal_n (a=t) = \modal_n (b=t)$. The latter
-%   remains true by univalence.
-% \end{itemize}
-% The fact that these two form an equivalence is technical, we refer to
-% the formalization for an explicit proof.
+  By~(\ref{eq:Omono}), $\sum_{a,b:A} f a = f b \simeq \sum_{a,b:A} \modal
+  (a=b)$, and by $\varphi$, $\sum_{a,b:A} f a = f b \simeq
+  \sum_{a,b:B} \modal ( a = b)$.
+
+Thus, $T_f \simeq \mathring T$.
 \end{proof}
 
-Now, let $Q$ be a separated $\Type_{n+1}$, and $f:T \to Q$. Then the
+Now, let $Q$ be any separated $\Type_{n+1}$, and $f:X \to Q$. Then the
 following diagram commutes
-$$\xymatrix @C=4em{ 
-    \cdots \overline{\Delta_3} \tar[r]{4pt} \ar[rd]_{f\circ \pi_1} & \overline{\Delta_2}
-    \dar[r]{2pt} \ar[d]^{f\circ \pi_1} &
-    \overline{\Delta_1} = T \ar[ld]^{f\circ \pi_1}\\
+
+$$\xymatrix{
+\|X\|_{n+1} \ar[r] \ar[rd] & \|\mathring T_{X}\|_{n+1} \ar[r] \ar[d] & \|\mathring
+  T_{\mathring T_X} \|_{n+1} \ar[ld]\cdots \\
   & Q &
-  }$$
+} $$
+
+% $$\xymatrix @C=4em{ 
+%     \cdots \overline{\Delta_3} \tar[r]{4pt} \ar[rd]_{f\circ \pi_1} & \overline{\Delta_2}
+%     \dar[r]{2pt} \ar[d]^{f\circ \pi_1} &
+%     \overline{\Delta_1} = T \ar[ld]^{f\circ \pi_1}\\
+%     & Q &
+%   }$$
 But we know (lemma~\ref{lem:sepiscolim}) that $\separated T$ is the
-colimit of the closed diagonals diagram, thus there is an universal
+$(n+1)$-colimit of the diagram $\mathring T$, thus there is an universal
 arrow $\separated T \to Q$.
@@
 Using a proof by univalence as we have done for proving $\modal_n (a=b) = (\mu_T a =
-\mu_T b)$ in Proposition~\ref{prop:sep-subu}, we can show that:
+\mu_T b)$ in Proposition~\ref{prop:sepiscolim}, we can show that:
 \begin{prop}
@@
 Overall, with other files containing technical lemmas, the project
-contains 6322 lines, and it could be reduced a bit by improving the
+contains 8081 lines, and it could be reduced a bit by improving the
 way Coq tries to rewrite and apply lemmas automatically. 
-The \texttt{coqwc} tool counts 2000 lines of specifications
-(definitions, lemmas, theorems, propositions) and 3700 lines of proof
+The \texttt{coqwc} tool counts 1750 lines of specifications
+(definitions, lemmas, theorems, propositions) and 6330 lines of proof
 script.
@@
 
-In our formalization of sheafification defined in
-Section~\ref{sec:sheafification} (done in 
-\texttt{sheaf\_induction.v}), two very technical facts required to
-show that $\separated$ is a modality have been
-admitted. 
-%
-It is due to a lack of technical lemmas on paths between paths in
-the Coq/HoTT library.
-%
-This is not surprising because those lemmas are defined in a
-call-by-need approach and there is no notion of a complete set of
-lemmas on paths for the moment. 
-%
-This is the subject of ongoing work, in particular those trying to
-define a syntax based on the cubical set model~\cite{bezem2013model}
-of Coquand et al.  
-%
-Nevertheless, we are confident that with an enrichment of the Coq/HoTT
-library, they can be proved and have been admitted only by lack of
-time.
-%
-In the proof that diagrams are equal in lemma~\ref{lem:sepiscolim},
-the equality of projections is admitted, because if this proof is
-obvious on paper, it becomes really technical when formalized.
-%
-This suggest that we should try to fine a better way to encode
-projections, with a better computational content.
-%
-As a consequence, the fact that $P$ is a colimit of the diagonal
-closed diagram is admitted too, but again, this proof should be really
-as short as it is on paper.
-
-More importantly, there are few axioms in the formalization that are
-due to an unsatisfactory management of polymorphic universe levels in
-Coq.
-%
+In the formalization, we had to use the type-in-type option, to handle
+the universe issues we faced.
+
 Universes are used in type theory to ensure consistency by checking
@@
 annotations for universe levels everywhere in the formalization.
-%
-We have chosen a more simple but maybe less satisfactory solution. 
 %
@@
   but many important types in homotopy theory are not
-  truncated, \eg the circle $\mathbb S^1$.  Thus, we plan to extend
+  truncated, \eg the circle $\mathbb S^2$.  Thus, we plan to extend
   the sheafification to all types by using a
@@
 \item 
-  In this work, a particular case of a Giraud-Rezk-Lurie axiom has
-  been admitted. The work done in~\cite[Chapter 6]{lurie} suggests,
-  considering that homotopy type theory is a language for higher topos
-  theory, that this axiom is provable. We plan to investigate whether the
-  higher-topos-theoretic proof can be adapted to our setting.
-  %
-  % A computer-checked proof of this fact would improve our
-  % developpement.
+  We would like to have much more examples of left-exact modalities on
+  $\HProp$, in order to extend type theory with different principles
+  than excluded middle.
 \end{itemize}
@@
 We would like to thank all maintainers of the Coq/HoTT library, who
-have provided a very solid base for our formalization. We also would like
+have provided a very solid base for our formalization. 
+We also would like
 to thank Bas Spitters for helpful discussions on the subject.

### 2015-11-27 paper/sheaf_lics.tex
Add missing files

@@
 \newcommand \one {\mathbf{1}}
+\newcommand \two {\mathbf{2}}
 
@@
 Using a proof by univalence as we have done for proving $\modal_n (a=b) = (\mu_T a =
-\mu_T b)$ in Proposition~\ref{prop:sepiscolim}, we can show that:
+\mu_T b)$ in Proposition~\ref{lem:sepiscolim}, we can show that:
 \begin{prop}
@@
 
+\section{Extension to $\Type$}
+\label{sec:extension-type}
+
+Now, we have defined a (countably) infinite family of modalities
+$\Type_i \to \Type_i$. Now:
+
+\begin{lem}
+  Let $\modal_i:\Type_i \to \Type_i$ be a modality. Then $\modal_i
+  \circ \|\cdot\|_i : \Type \to \Type$ is a modality in the sense
+  of~\cite[Section 7.7]{hottbook}
+\end{lem}
+
+Taking $i$ as a non-fixed parameter, that would allow us to work in an
+universe where the forced principle ({\em e.g.} mere excluded middle) is
+true for any explicit truncated type. Indeed, $i$ can be chosen
+dynamically along a proof, and thus be increased as much as needed.
+
+\subsection{New objects}
+\label{sec:new-objects}
+
+Working in $\Type^\modal$, one can consider new objects. For example,
+let $c:nat \to \nat$ implementing the Collatz function. Then, for any
+$n:\nat$, the type $C_n \defeq \sum_{k:\nat}\left( c^k(n) = 1 \land k\text{ is the
+  smallest such $\nat$}\right)$ is an $\HProp$.
+Thus, for any $n$, one have $C_n + \lnot C_n$, and one can {\em
+  constructively} build a map $\varphi:\nat \to \two$ such that
+$\varphi(n) = \True$ if and only if $n$ satisfies the Collatz
+conjecture.
+Actually, one can defined such a map for any family of $\HProp$
+indexed by $\nat$, for example with the family $C_n \defeq \text{«\,The $n$-th Turing Machine halts\,»}$.
+
 \section{Formalization} 
@@
 %
-\item Colimits, $p$-pullbacks and \v{C}ech nerves are formalized in
-\texttt{colimits.v, cech\_nerve.v}.% (552 lines).
+\item Colimits and \v{C}ech nerves are formalized in
+\texttt{Limit, T.v, OT.v OT\_Tf.v, T\_telescope.v, Tf\_Omono\_sep.v}.% (552 lines).
 \item
@@
 Note that the same increasing in the universe levels occurs in the
-Resk completion for categories~\cite{resk}. In the definition of the
+Rezk completion for categories~\cite{rezk}. In the definition of the
 completion, they use the Yoneda embedding and representable functors,

### 2015-11-27 paper/sheaf_lics.tex
[typo] An other => another

@@
 %
-An other famous application has been developed by
+Another famous application has been developed by
 Cohen~\cite{cohen1966} to prove that the continuum hypothesis is

### 2015-11-27 paper/sheaf_lics.tex
[grammar] Allow to => allow one to

@@
 types~\cite{lumsdaine2011higher,lumsdaine2013higher}. We adopt this
-point of view as it allows to maintain the slogan that homotopy type
+point of view as it allows one to maintain the slogan that homotopy type
 theory is type theory plus univalence and higher inductive types.
@@
 theory~(see \eg \cite{benton1998computational}). In those works, they connect
-modalities to computational monad in programming
+modalities to computational monads in programming
 languages~\cite{moggi-monad}. However, in our setting, modalities
@@
 contractible.
-Technical transformations allow to prove
+Technical transformations allow one to prove
 $$\fib\phi{x;p} = \fib s {y;p^{-1}}$$

### 2015-11-27 paper/sheaf_lics.tex
[wording] improve wording in concessive statement

@@
 
-Also proposition truncation could be seen as a primitive type former
-in homotopy type theory, it can be defined using higher inductive
+Although proposition truncation could be seen as a primitive type former
+in homotopy type theory, it can also be defined using higher inductive
 types~\cite{lumsdaine2011higher,lumsdaine2013higher}. We adopt this

### 2015-11-30 paper/sheaf_lics.tex
[grammar] monad => monads
include some reviewers remarks

@@
 \newtheorem{ax}[thm]{Axiom}
+\newtheorem{cond}[thm]{Condition}
 
@@
 \DeclareMathOperator{\IsContr}{IsContr}
-\DeclareMathOperator{\postcompose}{\mathrm{postcompose}}
+\DeclareMathOperator{\precompose}{\mathrm{precompose}}
 
@@
 %
-The most common use of sheafification is the construction a boolean
+One use of sheafification is the construction a boolean
 topos, thus validating the classical principle of the law of excluded
@@
 $\Type_{-1}$. This part corresponds to sheafification for toposes. Then,
-assuming that the sheafification has been constructed up to level $n$,
+assuming that the sheafification has been constructed above level $n$,
 one can define the sheafification for $\Type_{n+1}$.
@@
 times to
-$n$-types~\cite[Chapter~6]{lurie}.
+$n$-types~\cite[Chapter~6]{lurie}. Unfortunately, some types does not
+have a finite homotopy level, hence this construction cannot reach
+every type.
 
@@
 inductive types are defined using itemization to avoid an overhead of
-notation.
+notation. In the rest of the paper, $\Type$ must be seen in an
+universe-polymorphic way.
 
@@
 $$\lambda f:P \to X, \left( \lambda i,~f \circ q_i~;~ \lambda i\, j\,
-  \phi,~ \postcompose_f (p(i, j, g)) \right)$$
-where $\postcompose_f : \phi = \psi \to \phi \circ f = \psi \circ f$,
+  \phi,~ \precompose_f (p(i, j, g)) \right)$$
+where $\precompose_f : \phi = \psi \to \phi \circ f = \psi \circ f$,
 is an equivalence.
@@
   $$\lambda f:P \to X, \left( \lambda i,~f \circ q_i~;~ \lambda i\, j\,
-    \phi,~ \postcompose_f (p(i, j, g)) \right)$$
+    \phi,~ \precompose_f (p(i, j, g)) \right)$$
   is an equivalence.
@@
   %
-  $$\forall e:E,~ \left(\sum_{e':E} e=e'\right) \simeq \modal_n
+  $$\prod_{e:E}~ \left(\sum_{e':E} e=e'\right) \simeq \modal_n
   (\chi~e).$$
@@
     as the map sending an
-    arrow $f:E\to F$ to its restriction $f \circ m$.
+    arrow $f:E\to F$ to its restriction $f \circ \iota$.
   % \item if $A \to E$ with $n$-truncated homotopy fibers, $\Phi_E^\chi$ is the
@@
 
-An other fundamental property on sheaves is that the type of (dependent)
+Another fundamental property on sheaves is that the type of (dependent)
 functions is a sheaf as soon as its codomain is a sheaf.
@@
 % with $\modal_{-1}$, \ie we want the property
-\begin{prop}\label{prop:hprop}
+\begin{cond}\label{cond:hprop}
   For any mere proposition $P$ (with $\widehat P$ is $P$ seen as a
@@
   }$$
-\end{prop}
+\end{cond}
 
@@
 compatible with a left exact modality $\modal_{-1}$ on $\HProp$ in the
-sense of proposition~\ref{prop:hprop}.
+sense of condition~\ref{cond:hprop}.
 
@@
 This proves in particular that $\modal_{n+1}$ is compatible with
-$\modal_{-1}$ in the sense of proposition~\ref{prop:hprop}.
+$\modal_{-1}$ in the sense of condition~\ref{cond:hprop}.
 
@@
 
-An other issue with universe polymorphism lies in the management of
+Another issue with universe polymorphism lies in the management of
 recursive definitions. Indeed, the following recursive definition of

### 2015-12-03 paper/sheaf_lics.tex
some corrections about Giraud

@@
 Section~\ref{sec:hott} presents the foundations of homotopy type theory
-and defines a particular case of Giraud-Rezk-Lurie axioms. 
+and present a analogous of~\cite[Corollary 6.2.3.5]{lurie}. 
 %
@@
 %
-And Section~\ref{sec:giraud-ax} defines a particular case of
+And Section~\ref{sec:giraud-ax} defines an analogue of
 Giraud-Rezk-Lurie axioms on higher toposes that we need in the
@@
 $(\infty,1)$-category that have been shown to be equivalent to
-$(\infty,1)$-topos~\cite[Chapter 6]{lurie}. In~\cite{boulier}, a
-type-theoritic version of these axioms are proposed.
-%
-This particular case connects a surjection (or a $(-1)$-connected
-function) to the colimit of its \v{C}ech nerve.
-%
-This is the $\infty$-generalization of the fact that, in a topos,
-every epimorphism is the colimit of its kernel pair, which plays a
-central part in the definition of sheafification.
+$(\infty,1)$-topos~\cite[Chapter 6]{lurie}.
+%
+The consequence we want to use here is the fact that a surjection
+(\ie{} $(-1)$-connected function) is the colimit of its \v{C}ech
+nerve.
+%
+In~\cite{boulier}, the authors propose an analogue of this property~:
+they give, for any map $f$, a diagram $D_f$ whose colimit is $\im(f)$.
 
@@
 \end{defi}
-%
-The \v{C}ech nerve of a map $f$ can then be described as the
-simplicial object that, at degree $p$, is given by the
-$p$th iteration of $T$. 
-%
+
+Then, the desired diagram is the mapping telescope of the iterations
+of $T$.
 \begin{defi}
-  Let $f$ be a map from $X$ to $Y$. The {\em \v{C}ech nerve} $C(f)$ of $f$
+  Let $f$ be a map from $X$ to $Y$. Then $C(f)$
   is given by the diagram
@@
     C(f) := X \ar[r]^t & T_f \ar[r]^t & T_{\tilde f} \ar[r] &\cdots
-% \cdots~ X \times_Y X \times_Y X \tar[r]{4pt} & X \times_Y X \dar[r]{2pt} & X
   }$$
@@
 
-\begin{thm}[Giraud-Rezk-Lurie]
-  For any surjection $f : X \to Y$, the colimit of its \v{C}ech nerve
-  $C(f)$ is $Y$.
+Let's recall the main theorem:
+\begin{thm}[Colimit of $C(f)$~\cite{boulier}]\label{cech}
+  For any surjection $f : X \to Y$, the colimit of $C(f)$ is $Y$.
 \end{thm}
 
-With the point of view that homotopy type theory is a type-theoretic
-version of higher topos theory, this seems to be the analogous of Theorem 6.1.0.6 of~\cite{lurie}.
-%
-The proof of this theorem can be found in~\cite{boulier}.
+% \subsection{On Giraud-Rezk-Lurie axioms}
+% \label{sec:giraud-ax}
+
+% The Giraud-Rezk-Lurie axioms are the $\infty$-version of Giraud's
+% axioms that characterize a topos. Namely, there are 4 axioms on a
+% $(\infty,1)$-category that have been shown to be equivalent to
+% $(\infty,1)$-topos~\cite[Chapter 6]{lurie}. In~\cite{boulier}, a
+% type-theoritic version of these axioms are proposed.
+% %
+% This particular case connects a surjection (or a $(-1)$-connected
+% function) to the colimit of its \v{C}ech nerve.
+% %
+% This is the $\infty$-generalization of the fact that, in a topos,
+% every epimorphism is the colimit of its kernel pair, which plays a
+% central part in the definition of sheafification.
+
+% \begin{defi}
+%   Let $f:X \to Y$ be a map. $T_f$ is the higher inductive type given
+%   by
+%   \begin{itemize}
+%   \item $t:~X \to T_f$
+%   \item $\alpha:~\forall a\,b:X,~f a = f b \to t a = t b$
+%   \item $\alpha_1:~\forall a:X,~\alpha\, a\, a\, 1 = 1$
+%   \end{itemize}
+%   We view $T_f$ as the coequalizer of
+%   $$\xymatrix{
+%     \sum_{a,b:X} f a = f b \dar[r]{4pt}^-{\pi_1}_-{\pi_2} & X
+%   }$$
+%   preserving the identity.
+
+%   We call $\tilde f$ the map $T_f \to Y$ given by induction.
+% \end{defi}
+% %
+% The \v{C}ech nerve of a map $f$ can then be described as the
+% simplicial object that, at degree $p$, is given by the
+% $p$th iteration of $T$. 
+% %
+% \begin{defi}
+%   Let $f$ be a map from $X$ to $Y$. The {\em \v{C}ech nerve} $C(f)$ of $f$
+%   is given by the diagram
+%   $$\xymatrix{
+%     C(f) := X \ar[r]^t & T_f \ar[r]^t & T_{\tilde f} \ar[r] &\cdots
+% % \cdots~ X \times_Y X \times_Y X \tar[r]{4pt} & X \times_Y X \dar[r]{2pt} & X
+%   }$$
+% \end{defi}
+
+% \begin{thm}[Giraud-Rezk-Lurie]
+%   For any surjection $f : X \to Y$, the colimit of its \v{C}ech nerve
+%   $C(f)$ is $Y$.
+% \end{thm}
+
+% With the point of view that homotopy type theory is a type-theoretic
+% version of higher topos theory, this seems to be the analogous of Theorem 6.1.0.6 of~\cite{lurie}.
+% %
+% The proof of this theorem can be found in~\cite{boulier}.
 
@@
 \begin{proof}[Proof of lemma~\ref{lem:sepiscolim}]
-  As $\mu_T$ is an surjection, using the particular case of 
-  Giraud-Rezk-Lurie axiom introduced in Section~\ref{sec:giraud-ax},
+  As $\mu_T$ is an surjection, using the theorem~\ref{cech} introduced in Section~\ref{sec:giraud-ax},
   we know that $\separated T$ is the $(n+1)$-colimit of $\|C(\mu_T)\|_{n+1}$. 

### 2015-12-03 paper/sheaf_lics.tex
some corrections

@@
 
-Unfortunately, all those works---although they provide very nice and
-sophisticated theoretical results---do not share the generality and
-simplicity of sheafification in topos theory and are not amenable to a
+All those works---although they provide very nice and
+sophisticated theoretical results---are not as general as the process
+of sheafification in topos theory and are not amenable to a
 direct integration into a proof assistant based on type theory such as
@@
 
+In what follows, we work in the context of homotopy type theory, seen
+as Martin-Löf type thepry, with univalence axiom (and thus functional
+extensionality) and higher inductive types.
+
 \subsection{Plan of the Paper}
@@
 The definitions of Section~\ref{ssec:hott} and \ref{sec:epi-mono-fact}
-are part of the Coq library
-available at \url{https://github.com/HoTT/HoTT}, while other definitions
+are part, or direct applications of the Coq library~\cite{hottlib}, while other definitions
 and theorems are specific to our formalization.
@@
 $\pi_2 \circ \gamma$ is given by the modalisation of
-paths.\footnote{This proposition comes from on old version of the
-  formalization of left exact modalities in the Coq/HoTT library, but is
-  not present anymore in the library.}
+paths.
 \end{prop}
@@
 the subuniverse defined by this modality, defining an inhabitant of a
-type $T$ amounts to give an inhabitant of $P \to T$. 
-%
-Thus, working in this subuniverse amounts to add the axiom $P$ to the
-theory. Of course, adding an axiom this way does not add any computational
-content to $P$, and the consistency of the modality corresponds to
-the fact that $P \to \zero$ (or $\lnot P$) is not inhabited.
+type $T$ amounts to give an inhabitant of $P \to T$ (but where $P$ is
+still a type in the original theory).
+% %
+% Thus, working in this subuniverse amounts to add the axiom $P$ to the
+% theory. Of course, adding an axiom this way does not add any computational
+% content to $P$, and the consistency of the modality corresponds to
+% the fact that $P \to \zero$ (or $\lnot P$) is not inhabited.
 
@@
   % 
-  and moreover, for any $x:A$, the following diagram 
-  commutes, 
-  $$\xymatrix{
-    \sum_{e':A} x = e' \ar@{=}[r] \ar[d]_{\iota'} & (\chi \ x)
-    \ar[d]^{\eta_{(\chi \ x)}}\\
-    \sum_{e':E} x = e' \ar@{=}[r] & \modal_n (\chi \ x)
-  }$$
-  where $\iota': x \mapsto (\iota (\pi_1 x) ; \pi_2 x)$.
+  % and moreover, for any $x:A$, the following diagram 
+  % commutes, 
+  % $$\xymatrix{
+  %   \sum_{e':A} x = e' \ar@{=}[r] \ar[d]_{\iota'} & (\chi \ x)
+  %   \ar[d]^{\eta_{(\chi \ x)}}\\
+  %   \sum_{e':E} x = e' \ar@{=}[r] & \modal_n (\chi \ x)
+  % }$$
+  % where $\iota': x \mapsto (\iota (\pi_1 x) ; \pi_2 x)$.
 \end{defi}
@@
   all dense $n$-subobject of $\iota : A \to E$,
-  $\Phi_E^\iota$ is an embedding. In other words, the dotted arrow,
-  if exists, is unique.
-
-  $$\xymatrix{
-    A \ar[r]^f \ar[d]_{\iota} & F \\
-    E \ar@{-->}[ru]_{!}&
-  }$$
+  $\Phi_E^\iota$ is an embedding.%  In other words, the dotted arrow,
+  % if exists, is unique.
+
+  % $$\xymatrix{
+  %   A \ar[r]^f \ar[d]_{\iota} & F \\
+  %   E \ar@{-->}[ru]_{!}&
+  % }$$
 \end{defi}
+
+With topos theory point of view, it means that given a map $A\to F$,
+if there is an extension $\tilde f:E\to F$, then it is unique.
+Actually, there is a simpler characterisation:
+
+\begin{prop}
+  A type $F$ is $\Type_{n+1}$ is separated if, and only if all its
+  path types are $n$-modal, ie
+  $$\mathrm{separated}(F) = \prod_{x,y:F}~\modal_n(x=y) = (x=y).$$
+\end{prop}
 
@@
   $\iota : A \to E$, $\Phi_E^\iota$ is an
-  equivalence. In other words, the dotted arrow exists and is unique.
-
-  $$\xymatrix{
-    A \ar[r]^f \ar[d]_{\iota} & F \\
-    E \ar@{-->}[ru]_{\exists !}&
-  }$$
+  equivalence. % In other words, the dotted arrow exists and is unique.
+
+  % $$\xymatrix{
+  %   A \ar[r]^f \ar[d]_{\iota} & F \\
+  %   E \ar@{-->}[ru]_{\exists !}&
+  % }$$
 \end{defi}
+
+In topos-theoretic words, it means that given a map $A\to F$, one can
+extend it uniquely in $\tilde f:E \to F$.
 
@@
 \begin{cond}\label{cond:hprop}
-  For any mere proposition $P$ (with $\widehat P$ is $P$ seen as a
+  For any mere proposition $P$ (where $\widehat P$ is $P$ seen as a
   $\Type_n$),  $\modal_n \widehat P = \modal_{-1} P$ and the
@@
   $$\xymatrix{
-    P \ar@{=}[r] \ar[d]_{\eta_{-1}} & \widehat P \ar[d]^{\eta_n} \\
-    \modal_{-1} P \ar@{=}[r] & \modal_n \widehat P 
+    P \ar@{->}^{\sim}[r] \ar[d]_{\eta_{-1}} & \widehat P \ar[d]^{\eta_n} \\
+    \modal_{-1} P \ar@{->}^{\sim}[r] & \modal_n \widehat P 
   }$$

### 2015-12-03 paper/sheaf_lics.tex
some corrections

@@
 \begin{defi}
-  Let $E$ be a type, and $\chi:E \to \Type_n$. The subobject $\iota :
+  Let $E$ be a type, and $\chi:E \to \Type_n$. The $n$-subobject $\iota :
   A \to E$ of $E$
@@
 family. We want to show that $\sum_{x:A} (Bx)$ is separated. Let $E$
-be a type, and $\sum_{e:E} (\chi\,e)$ a dense subobject of E.
+be a type, and $\sum_{e:E} (\chi\,e)$ a dense $n$-subobject of E.
 
@@
 For the second components, let $x:E$. Notice that for any $x:E$,
-$\sum_{y:E} x = y$ has a dense subobject, $\sum_{y:\sum_{e:E} (\chi\,
+$\sum_{y:E} x = y$ has a dense $n$-subobject, $\sum_{y:\sum_{e:E} (\chi\,
   e)} x=y_1$:

### 2015-12-11 paper/paper_acm.tex
add ACM version

--- 2015-12-03 paper/sheaf_lics.tex
@@
-\documentclass[conference]{IEEEtran}
+%-----------------------------------------------------------------------------
+%
+%               Template for sigplanconf LaTeX Class
+%
+% Name:         sigplanconf-template.tex
+%
+% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
+%               file for SIGPLAN conference proceedings.
+%
+% Guide:        Refer to "Author's Guide to the ACM SIGPLAN Class,"
+%               sigplanconf-guide.pdf
+%
+% Author:       Paul C. Anagnostopoulos
+%               Windfall Software
+%               978 371-2316
+%               paul@windfall.com
+%
+% Created:      15 February 2005
+%
+%-----------------------------------------------------------------------------
+
+\documentclass[preprint,10pt,numbers]{sigplanconf}
+
+% The following \documentclass options may be useful:
+
+% preprint      Remove this option only once the paper is in final form.
+% 10pt          To set in 10-point type instead of 9-point.
+% 11pt          To set in 11-point type instead of 9-point.
+% numbers       To obtain numeric citation style instead of author/year.
+
+\usepackage{amsmath}
 
@@
 %
-\usepackage[cmex10]{amsmath}
+%\usepackage[cmex10]{amsmath}
 \usepackage{amssymb,amsthm,amsfonts}
@@
 \newtheorem{cond}[thm]{Condition}
+\newtheorem{exm}[thm]{Example}
 
@@
 
- \IEEEpubid{}
+\newcommand{\cL}{{\cal L}}
 
 \begin{document}
-%
-% paper title
-% Titles are generally capitalized except for words such as a, an, and, as,
-% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
-% not capitalized unless they are the first or last word of the title.
-% Linebreaks \\ can be used within to get better formatting as desired.
-% Do not put math or special symbols in the title.
+
+\special{papersize=8.5in,11in}
+\setlength{\pdfpageheight}{\paperheight}
+\setlength{\pdfpagewidth}{\paperwidth}
+
+\conferenceinfo{LICS '16}{July 5--8, 2016, New York City, USA}
+\copyrightyear{2016}
+\copyrightdata{978-1-nnnn-nnnn-n/yy/mm}
+\copyrightdoi{nnnnnnn.nnnnnnn}
+
+% Uncomment the publication rights you want to use.
+%\publicationrights{transferred}
+%\publicationrights{licensed}     % this is the default
+%\publicationrights{author-pays}
+
+\titlebanner{banner above paper title}        % These are ignored unless
+\preprintfooter{short description of paper}   % 'preprint' option specified.
+
 \title{Lawvere-Tierney Sheafification\\ in Homotopy Type Theory}
-
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
-\author{\IEEEauthorblockN{Kevin Quirin and Nicolas Tabareau}
-\IEEEauthorblockA{Inria \\ \'Ecole des Mines de Nantes \\ Nantes, France}}
-
-% conference papers do not typically use \thanks and this command
-% is locked out in conference mode. If really needed, such as for
-% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
-% after \documentclass
-
-% make the title area
+% \subtitle{Subtitle Text, if any}
+
+\authorinfo{Kevin Quirin}
+           {\'Ecole des Mines de Nantes}
+           {kevin.quirin@inria.fr}
+\authorinfo{Nicolas Tabareau}
+           {Inria}
+           {nicolas.tabareau@inria.fr}
+
 \maketitle
 
-% As a general rule, do not put math, special symbols or citations
-% in the abstract
 \begin{abstract}
-
   Sheafification is a popular tool in topos theory which
@@
   compatible with the full type-theoretic axiom of choice.
-
 \end{abstract}
 
-% no keywords
+\category{CR-number}{subcategory}{third-level}
+
+% general terms are not compulsory anymore,
+% you may leave them out
+\terms
+term1, term2
+
+\keywords
+keyword1, keyword2
 
@@
 \end{defi}
-% For a type $X$, to be in $\Type_n$ means that path spaces of $X$ are
+% For a type $X$, to be in $\Type_n$ means that paths spaces of $X$ are
 % trivial after $n+1$ iteration.
@@
 %
-$$\chi : \left(\sum_{A:\Type} \sum_{f:A \to B} \prod_{b\in B}
+\begin{align*}
+  \chi &: \left(\sum_{A:\Type} \sum_{f:A \to B} \prod_{b\in B}
 \IsType n\
-\fib{f}{b}\right) \xrightarrow{\sim} 
- (B \to \Type_n)$$
+\fib{f}{b}\right) \\ &\qquad\xrightarrow{\sim} 
+ (B \to \Type_n)
+\end{align*}
 %
@@
 map
-$$\lambda f:P \to X, \left( \lambda i,~f \circ q_i~;~ \lambda i\, j\,
-  \phi,~ \precompose_f (p(i, j, g)) \right)$$
+$$\lambda f:P \to X, \left( \lambda i,~f \circ q_i\, ;\, \lambda i\, j\,
+  \phi,\ \precompose_f (p_{i, j}^\phi)) \right)$$
 where $\precompose_f : \phi = \psi \to \phi \circ f = \psi \circ f$,
@@
   map
-  $$\lambda f:P \to X, \left( \lambda i,~f \circ q_i~;~ \lambda i\, j\,
-    \phi,~ \precompose_f (p(i, j, g)) \right)$$
+  $$\lambda f:P \to X, \left( \lambda i,~f \circ q_i\, ;\, \lambda i\, j\,
+  \phi,\ \precompose_f (p_{i, j}^\phi)) \right)$$
   is an equivalence.
@@
 \begin{defi}
-  Let $f:X \to Y$ be a map. $T_f$ is the higher inductive type given
+  Let $f:X \to Y$ be a map. The kernel pair $T_f$ of $f$ is the higher inductive type given
   by
@@
 \begin{defi}
-  Let $f$ be a map from $X$ to $Y$. Then $C(f)$
+  Let $f$ be a map from $X$ to $Y$. Then the iterated kernel pair of
+  $f$ $C(f)$
   is given by the diagram
@@
 using $\eta$ functions.
-We just need to check that every $\modal(\fib \phi {x;p})$ is
+We just need to check that every $\modal$-fiber $\modal(\fib \phi {x;p})$ is
 contractible.
@@
 Here, we need to distinguish between
-dense $(-1)$-subobjects, that will be used in the definition of
+dense $(-1)$-sub\-objects, that will be used in the definition of
 sheaves, and dense $n$-subobjects, that will be used in the definition
@@
     \item If $g$ is an embedding, then 
-      $$\forall b:B,~\fib f b = \fib {g\circ f} {g\, b}.$$
+      $$\prod_{b:B}\fib f b = \fib {g\circ f} {g\, b}.$$
     \item If all fibers of $f$ and $g$ are $n$-truncated, then
-      $$\forall c:C,~\modal_n(\fib {g \circ f} c) =
+      $$\prod_{c:C}\modal_n(\fib {g \circ f} c) =
       \modal_n \left(  
@@
 \separated T &\defeq & \im (\lambda~t:T,~\lambda~ t',~ \modal_n (t = t')) \\
-          & \defeq & \sum_{u:T \to \Type_n^\modal} \left\| \sum_{a:T} 
+          & \defeq & \sum\limits_{u:T \to \Type_n^\modal} \left\| \sum_{a:T} 
             (\lambda t,~\modal_n (a=t)) = u\right\|.
@@
 
-  We use the encode-decode method to characterize $\tilde f |a|_{n+1}
+  We use the encode-decode~\cite[Section 8.9]{hottbook} method to characterize $\tilde f |a|_{n+1}
   = x$, and the result follows. We refer to the formalization for details.
@@
 
-\section{Extension to $\Type$}
+\section{Extension to Type}
 \label{sec:extension-type}
 
-Now, we have defined a (countably) infinite family of modalities
-$\Type_i \to \Type_i$. Now:
-
-\begin{lem}
-  Let $\modal_i:\Type_i \to \Type_i$ be a modality. Then $\modal_i
+In the previous section, we defined a (countably) infinite family of
+modalities $\Type_i \to \Type_i$. One can extend them to whole
+$\Type$ by composing with truncation:
+
+\begin{lem}\label{lem:type}
+  Let $\modal_i:\Type_i \to \Type_i$ be a modality. Then $\modal
+  \defeq \modal_i
   \circ \|\cdot\|_i : \Type \to \Type$ is a modality in the sense
@@
 
+\begin{exm}
+  If $\modal_{-1}$ is the double negation modality on $\HProp$ and
+  $i=-1$, for any type $X$:
+  \begin{align*}
+  \modal X &\defeq \lnot\lnot \|X\|  \\
+           &= \lnot\lnot X
+  \end{align*}
+  by univalence for h-propositions.
+
+ Thus, $\modal$ is exactly the double negation modality on $\Type$
+  described in~\ref{sec:examples-left-exact}.
+\end{exm}
+
+Chosing $i\geqslant 0$ is a refinement of this double negation
+modality on $\Type$: it will collapse every type to a $\Type_i$,
+instead of an $\HProp$.
+
 Taking $i$ as a non-fixed parameter, that would allow us to work in an
-universe where the forced principle ({\em e.g.} mere excluded middle) is
+universe where the new principle ({\em e.g.} mere excluded middle) is
 true for any explicit truncated type. Indeed, $i$ can be chosen
@@
 
-\subsection{New objects}
-\label{sec:new-objects}
-
-Working in $\Type^\modal$, one can consider new objects. For example,
-let $c:nat \to \nat$ implementing the Collatz function. Then, for any
-$n:\nat$, the type $C_n \defeq \sum_{k:\nat}\left( c^k(n) = 1 \land k\text{ is the
-  smallest such $\nat$}\right)$ is an $\HProp$.
-Thus, for any $n$, one have $C_n + \lnot C_n$, and one can {\em
-  constructively} build a map $\varphi:\nat \to \two$ such that
-$\varphi(n) = \True$ if and only if $n$ satisfies the Collatz
-conjecture.
-Actually, one can defined such a map for any family of $\HProp$
-indexed by $\nat$, for example with the family $C_n \defeq \text{«\,The $n$-th Turing Machine halts\,»}$.
+Unfortunately, one lose canonicity when working in $\Type^\modal$;
+indeed, for example, the term 
+$$\left( \lambda P:\HProp,~\left\{\begin{array}{l} 0 \text{ if $P$ is
+                                    false} \\ 1 \text{ if $P$ is
+                                    true} \end{array}\right. \right)
+                            P~:~\nat$$
+will not reduce to a numeral if the chosen property $P$ is not
+decidable. However, we keep canonicity for ``old'' terms, i.e. the
+ones constructed without using excluded middle.
+
+% \subsection{New objects}
+% \label{sec:new-objects}
+
+% Working in $\Type^\modal$, one can consider new objects. For example,
+% let $c:nat \to \nat$ implementing the Collatz function. Then, for any
+% $n:\nat$, the type $$C_n \defeq \sum_{k:\nat}\left( c^k(n) = 1 \land k\text{ is the
+%   smallest such $\nat$}\right)$$ is an $\HProp$.
+% Thus, for any $n$, one have $C_n + \lnot C_n$, and one can {\em
+%   constructively} build a map $\varphi:\nat \to \two$ such that
+% $\varphi(n) = \True$ if and only if $n$ satisfies the Collatz
+% conjecture.
+% Actually, one can defined such a map for any family of $\HProp$
+% indexed by $\nat$, for example with the family $C_n \defeq \text{«\,The $n$-th Turing Machine halts\,»}$.
 
@@
 
-A Coq formalization is available at
-\url{https://github.com/KevinQuirin/sheafification}.
-%
-It is based on the Coq/HoTT library available at
-\url{https://github.com/HoTT/HoTT} and uses the version of Coq
-tailored for that library.
+A Coq formalization based on the Coq/HoTT library~\cite{hottlib} is available at
+\url{github.com/KevinQuirin/sheafification}.
 %
@@
 \begin{itemize} 
-\item  Homotopy fibers and subobject classifiers are formalized in
-\texttt{sub\_object\_classifier.v}.% (284 lines).
-%
-\item Colimits and \v{C}ech nerves are formalized in
-\texttt{Limit, T.v, OT.v OT\_Tf.v, T\_telescope.v, Tf\_Omono\_sep.v}.% (552 lines).
+\item Colimits and iterated kernel pairs are formalized in
+\texttt{Limit}, \texttt{T.v}, \texttt{OT.v}v \texttt{OT\_Tf.v}, \texttt{T\_telescope.v},\\ \texttt{Tf\_Omono\_sep.v}.% (552 lines).
 \item
-Reflective subuniverses and modalities are formalized in
-\texttt{reflective\_subuniverse.v, modalities.v}. % (1053 lines).
+Reflective subuniverses and modalities are formalized in\\
+\texttt{reflective\_subuniverse.v}, \texttt{modalities.v}. % (1053 lines).
 \item 
@@
 
-In the formalization, we had to use the type-in-type option, to handle
+In the formalization, we had to use the \texttt{type-in-type} option, to handle
 the universe issues we faced.
@@
 %
-When the absence of cumulativity on $\Type_n$ rises a universe issue
-for a lemma (it occurs twice in the formalization), we have proved the
-lemma and then admit the exact same lemma, just letting Coq infer less
-constrained universe levels.
+Note that we also could have resolved the issue by giving explicitely
+the universe levels we wanted, like in the Coq/HoTT library\cite{hottlib}.
 
@@
 \item 
-  At the moment, sheafification only works with truncated
-  types. This is fine from a programming point of view, 
-  but many important types in homotopy theory are not
-  truncated, \eg the circle $\mathbb S^2$.  Thus, we plan to extend
-  the sheafification to all types by using a
-  limit construction.
+  The extension to whole $\Type$ in lemma~\ref{lem:type} is not
+  totally satisfactory, as every type is collapsed to a truncated
+  one. But some types in homotopy type theorey are not
+  truncated\cite[Example 8.8.6]{hottbook}.
+
+  % At the moment, sheafification only works with truncated
+  % types. This is fine from a programming point of view, 
+  % but many important types in homotopy theory are not
+  % truncated, \eg the circle $\mathbb S^2$.  Thus, we plan to extend
+  % the sheafification to all types by using a
+  % limit construction.
 \item 
@@
 \end{itemize}
-Moreover, in topos theory, Lawvere-Tierney subsumes Grothendieck~\cite[Section~V.4]{maclanemoerdijk} in the sense that any
-Grothendieck topology gives rise to a Lawvere-Tierney topology with
+Moreover, in topos theory, Lawvere-Tierney subsumes Gro\-thendieck~\cite[Section~V.4]{maclanemoerdijk} in the sense that any
+Grothen\-dieck topology gives rise to a Lawvere-Tierney topology with
 the same notion of sheaves. Higher Lawvere-Tierney sheaves are
@@
 
-% conference papers do not normally have an appendix
-
-% use section* for acknowledgment
 \section*{Acknowledgment}
@@
 
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-% references section
-
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-\bibliographystyle{IEEEtran}
+\bibliographystyle{abbrvnat}
 \bibliography{sheaf_lics}
-% argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{IEEEabrv,../bib/paper}
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
-% \begin{thebibliography}{5}
-
-% that's all folks
 

### 2015-12-11 paper/paper_acm_long.tex
add long version

--- 2015-12-11 paper/paper_acm.tex
@@
 has been defined in homotopy type theory~\cite{lumsdaine}, we present
-here a definition of colimits of diagrams overs graphs. 
+here a definition of colimits of diagrams overs graphs, presented in~\cite{sets_in_hott}. 
 %

### 2015-12-11 paper/paper_acm.tex
reduce space

@@
 
+\newenvironment{mymath}%
+{\vspace{-0.7em}\begin{equation*}}%
+  {\vspace{-0.7em}\end{equation*}}
+
 \begin{document}
@@
 says that for any types $T$ and $U$, the canonical map 
-$$
+\begin{mymath}
 (T = U) \to (T \simeq U)
-$$ 
+\end{mymath}
 %
@@
 %
-$$
+\begin{mymath}
 j : \Omega \rightarrow \Omega
-$$
+\end{mymath}
 %
@@
 %
-$$
+\begin{mymath}
 \Char{\closure{A}} = j \circ \Char{A}.
-$$
+\end{mymath}
 %
@@
 %
-$$
+\begin{mymath}
 \Omega_j = \{ P \in \E \ | \ j P  = P \}.
-$$
+\end{mymath}
 %
@@
 %
-$$
+\begin{mymath}
 a_j : \E \rightarrow \Sh{\E}
-$$
+\end{mymath}
 which exhibits $\Sh{\E}$ as a reflective
@@
 defined as 
-$$
+\begin{mymath}
 \fib{f}{b} \defeq \sum_{a:A} f(a) = b.
-$$ 
+\end{mymath} 
 %
@@
  such that the diagram
-$$
+\begin{mymath}
 \xymatrix{
@@
 }
-$$
+\end{mymath}
 is a pullback for any $f$ with
@@
 $n$-truncated types and 
-$$t_f = \lambda a,~(\fib{f}{f(a)},(a,\mathrm{idpath})).$$
+\begin{mymath}t_f = \lambda a,~(\fib{f}{f(a)},(a,\mathrm{idpath})).\end{mymath}
 
@@
   Let $f:A\to B$ be a function. The {\em image} of $f$ is defined as 
-  $$\im(f) \defeq \sum_{y:B} \|\fib f y\|.$$
+  \begin{mymath}\im(f) \defeq \sum_{y:B} \|\fib f y\|.\end{mymath}
 \end{defi}
@@
 \item a map $q_i : D_0(i) \to P$ for any
-vertex $i:G_0$, \ie $$q : \prod_{i:G_0} D_0(i) \to P$$
+vertex $i:G_0$, \ie \begin{mymath}q : \prod_{i:G_0} D_0(i) \to P\end{mymath}
 \item for any vertices $i,j:G_0$ and all edges $\phi:G_1(i,j)$, a path
   $p_{i,j}^\phi : q_j \circ D_1(\phi) = q_i$, \ie
-  $$p : \prod_{i,j:G_0} \prod_{\phi:G_1(i,j)} q_j \circ D_1(\phi) = q_i.$$
+  \begin{mymath}p : \prod_{i,j:G_0} \prod_{\phi:G_1(i,j)} q_j \circ D_1(\phi) = q_i.\end{mymath}
 \end{itemize}
@@
 map
-$$\lambda f:P \to X, \left( \lambda i,~f \circ q_i\, ;\, \lambda i\, j\,
-  \phi,\ \precompose_f (p_{i, j}^\phi)) \right)$$
+\begin{mymath}\lambda f:P \to X, \left( \lambda i,~f \circ q_i\, ;\, \lambda i\, j\,
+  \phi,\ \precompose_f (p_{i, j}^\phi)) \right)\end{mymath}
 where $\precompose_f : \phi = \psi \to \phi \circ f = \psi \circ f$,
@@
   map
-  $$\lambda f:P \to X, \left( \lambda i,~f \circ q_i\, ;\, \lambda i\, j\,
-  \phi,\ \precompose_f (p_{i, j}^\phi)) \right)$$
+  \begin{mymath}\lambda f:P \to X, \left( \lambda i,~f \circ q_i\, ;\, \lambda i\, j\,
+  \phi,\ \precompose_f (p_{i, j}^\phi)) \right)\end{mymath}
   is an equivalence.
@@
   We view $T_f$ as the coequalizer of
-  $$\xymatrix{
+  \begin{mymath}\xymatrix{
     \sum_{a,b:X} f a = f b \dar[r]{4pt}^-{\pi_1}_-{\pi_2} & X
-  }$$
+  }\end{mymath}
   preserving the identity.
@@
   is given by the diagram
-  $$\xymatrix{
+  \begin{mymath}\xymatrix{
     C(f) := X \ar[r]^t & T_f \ar[r]^t & T_{\tilde f} \ar[r] &\cdots
-  }$$
+  }\end{mymath}
 \end{defi}
@@
 %   We view $T_f$ as the coequalizer of
-%   $$\xymatrix{
+%   \begin{mymath}\xymatrix{
 %     \sum_{a,b:X} f a = f b \dar[r]{4pt}^-{\pi_1}_-{\pi_2} & X
-%   }$$
+%   }\end{mymath}
 %   preserving the identity.
@@
 %   is given by the diagram
-%   $$\xymatrix{
+%   \begin{mymath}\xymatrix{
 %     C(f) := X \ar[r]^t & T_f \ar[r]^t & T_{\tilde f} \ar[r] &\cdots
 % % \cdots~ X \times_Y X \times_Y X \tar[r]{4pt} & X \times_Y X \dar[r]{2pt} & X
-%   }$$
+%   }\end{mymath}
 % \end{defi}
@@
   \item For every $n$-truncated types $A$ and $B$, if $P(B)$ then
-    $$\left\{
+    \begin{mymath}\left\{
       \begin{array}{rcl}
@@
         f & \mapsto & f \circ \eta_A
-      \end{array} \right.$$
+      \end{array} \right.\end{mymath}
     is an equivalence.
@@
 defined as %$\Type^\modal_p$, so 
-$$
+\begin{mymath}
 \Type_n^\modal \ \defeq \sum_{T : \Type_n} P (T).
-$$
+\end{mymath}
 %
@@
 is the fiber of $\modal f$ above $\eta_Y y$:
-$$\modal \left(\sum_{x:X}  (f x = y)\right) = \sum_{x:\modal X}
-(\modal f x = \eta_Y y),$$
+\begin{mymath}\modal \left(\sum_{x:X}  (f x = y)\right) = \sum_{x:\modal X}
+(\modal f x = \eta_Y y),\end{mymath}
 where $\modal f:\modal X \to \modal Y$ is defined via the equivalence {\it (iv)} and $\eta_Y$.
@@
 Moreover the following diagram commute
-$$\xymatrix{
+\begin{mymath}\xymatrix{
   \sum_{x:X} (f x = y) \ar[r]^\eta \ar[d]_\gamma & \modal \left(\sum_{x:X}  (f x = y)\right) \ar@{=}[dl]\\
-  \sum_{x:\modal X} (\modal f x = \eta_Y y) & }$$
+  \sum_{x:\modal X} (\modal f x = \eta_Y y) & }\end{mymath}
 where $\pi_1 \circ \gamma = \eta_X \circ \pi_1$, and
@@
 It is straighforward to define a map
-$$\phi:\sum_{x:X}  (f x = y)\to
-\sum_{x:\modal X} (\modal f x = \eta_Y y),$$
+\begin{mymath}\phi:\sum_{x:X}  (f x = y)\to
+\sum_{x:\modal X} (\modal f x = \eta_Y y),\end{mymath}
 using $\eta$ functions.
@@
 Technical transformations allow one to prove
-$$\fib\phi{x;p} = \fib s {y;p^{-1}}$$
+\begin{mymath}\fib\phi{x;p} = \fib s {y;p^{-1}}\end{mymath}
 for some $a:\modal X$, $b:\modal Y$ and 
-$$s:\fib{\eta_X}a \to \fib{\eta_Y}b.$$
+\begin{mymath}s:\fib{\eta_X}a \to \fib{\eta_Y}b.\end{mymath}
 %
@@
 %
-$$\chi_E  = \lambda e . \ \sum_{e':E} e=e'.$$
+\begin{mymath}\chi_E  = \lambda e . \ \sum_{e':E} e=e'.\end{mymath}
 %
@@
   %
-  $$\prod_{e:E}~ \left(\sum_{e':E} e=e'\right) \simeq \modal_n
-  (\chi~e).$$
+  \begin{mymath}\prod_{e:E}~ \left(\sum_{e':E} e=e'\right) \simeq \modal_n
+  (\chi~e).\end{mymath}
   % 
@@
   % commutes, 
-  % $$\xymatrix{
+  % \begin{mymath}\xymatrix{
   %   \sum_{e':A} x = e' \ar@{=}[r] \ar[d]_{\iota'} & (\chi \ x)
@@
   %   \sum_{e':E} x = e' \ar@{=}[r] & \modal_n (\chi \ x)
-  % }$$
+  % }\end{mymath}
   % where $\iota': x \mapsto (\iota (\pi_1 x) ; \pi_2 x)$.
@@
   For any type $A$ and $E$, $\iota : A \to E$ and $F:\Type_{n+1}$, we define
-    $$
+    \begin{mymath}
     \Phi_E^\iota : (E \to F) \to (A \to F) 
-    $$
+    \end{mymath}
     %
@@
 
-  % $$\xymatrix{
+  % \begin{mymath}\xymatrix{
   %   A \ar[r]^f \ar[d]_{\iota} & F \\
   %   E \ar@{-->}[ru]_{!}&
-  % }$$
+  % }\end{mymath}
 \end{defi}
@@
   path types are $n$-modal, ie
-  $$\mathrm{separated}(F) = \prod_{x,y:F}~\modal_n(x=y) = (x=y).$$
+  \begin{mymath}\mathrm{separated}(F) = \prod_{x,y:F}~\modal_n(x=y) = (x=y).\end{mymath}
 \end{prop}
@@
 
-  % $$\xymatrix{
+  % \begin{mymath}\xymatrix{
   %   A \ar[r]^f \ar[d]_{\iota} & F \\
   %   E \ar@{-->}[ru]_{\exists !}&
-  % }$$
+  % }\end{mymath}
 \end{defi}
@@
   We define $g$ as 
-  $$g(e) = \modal_n \left( \fib \phi e \right), $$
+  \begin{mymath}g(e) = \modal_n \left( \fib \phi e \right), \end{mymath}
   where
-  $$\phi : \left(\sum_{b:\left(\sum_{e:E} (\chi\, e)\right)} (f\,
-    b)\right) \to E$$
+  \begin{mymath}\phi : \left(\sum_{b:\left(\sum_{e:E} (\chi\, e)\right)} (f\,
+    b)\right) \to E\end{mymath}
   defined by $\phi(x) = (x_1)_1$.
@@
     \item If $g$ is an embedding, then 
-      $$\prod_{b:B}\fib f b = \fib {g\circ f} {g\, b}.$$
+      \begin{mymath}\prod_{b:B}\fib f b = \fib {g\circ f} {g\, b}.\end{mymath}
     \item If all fibers of $f$ and $g$ are $n$-truncated, then
-      $$\prod_{c:C}\modal_n(\fib {g \circ f} c) =
+      \begin{mymath}\prod_{c:C}\modal_n(\fib {g \circ f} c) =
       \modal_n \left(  
         \sum_{w:\fib g c} \modal_n (\fib f {w_1})
-      \right).$$
+      \right).\end{mymath}
     \end{itemize}
@@
   following coherence diagram commutes
-  $$\xymatrix{
+  \begin{mymath}\xymatrix{
     P \ar@{->}^{\sim}[r] \ar[d]_{\eta_{-1}} & \widehat P \ar[d]^{\eta_n} \\
     \modal_{-1} P \ar@{->}^{\sim}[r] & \modal_n \widehat P 
-  }$$
+  }\end{mymath}
 \end{cond}
@@
 % $\HProp$. Here we took the $\lnot\lnot$ modality:
-% $$\forall P:\HProp,~\modal_j P = \lnot\lnot P.$$
+% \begin{mymath}\forall P:\HProp,~\modal_j P = \lnot\lnot P.\end{mymath}
 % One can easily show that it defines a left exact modality.
@@
 $\modal_n^T \circ \{\cdot\}_T$, as in
-$$\xymatrix{
+\begin{mymath}\xymatrix{
   T \ar[r]^{\{\cdot\}_T} \ar[d]_{\mu_T} & \Type_n^T \ar[d]^{\modal_n^T} \\
   \separated T \ar[r]& \left( \Type_n^\modal \right)^T
-}, $$
+}, \end{mymath}
 where $\{\cdot\}_T$ is the singleton map $\lambda (t:T),~\lambda
@@
 %
-$$
+\begin{mymath}
 \begin{array}{rcl}
@@
 \end{array}
-$$
+\end{mymath}
 %
@@
   We view $\mathring T$ as the coequalizer of
-  $$\xymatrix{\sum_{a,b:\|X\|_{n+1}} \modal (a=b) \dar[r]{4pt}^-{\pi_1}_-{\pi_2}
-  & \|X\|_{n+1}}$$
+  \begin{mymath}\xymatrix{\sum_{a,b:\|X\|_{n+1}} \modal (a=b) \dar[r]{4pt}^-{\pi_1}_-{\pi_2}
+  & \|X\|_{n+1}}\end{mymath}
   preserving $\eta_{a=a} 1$.
@@
   We consider the diagram $\mathring T$ :
-  $$\xymatrix{\|X\|_{n+1} \ar[r] & \|\mathring T_{X}\|_{n+1} \ar[r] & \|\mathring
-  T_{\mathring T_X} \|_{n+1} \ar[r] & \cdots} $$
+  \begin{mymath}\xymatrix{\|X\|_{n+1} \ar[r] & \|\mathring T_{X}\|_{n+1} \ar[r] & \|\mathring
+  T_{\mathring T_X} \|_{n+1} \ar[r] & \cdots} \end{mymath}
 \end{defi}
@@
   then
-  $$\forall a,b:\|T_f\|_{n+1},~|\tilde f|_{n+1} a = |\tilde f|_{n+1} b \simeq \modal (a=b).$$
+  \begin{mymath}\forall a,b:\|T_f\|_{n+1},~|\tilde f|_{n+1} a = |\tilde f|_{n+1} b \simeq \modal (a=b).\end{mymath}
 \end{lem}
@@
   By induction on truncation, we need to show that 
-  $$\forall a,b:T_f,~\tilde f |a|_{n+1} = \tilde f |b|_{n+1} \simeq
-  \modal (|a|_{n+1}=|b|_{n+1}).$$
+  \begin{mymath}\forall a,b:T_f,~\tilde f |a|_{n+1} = \tilde f |b|_{n+1} \simeq
+  \modal (|a|_{n+1}=|b|_{n+1}).\end{mymath}
 
@@
   satisfying~(\ref{eq:Omono}), 
-  $$T_f \simeq \mathring T_B.$$
+  \begin{mymath}T_f \simeq \mathring T_B.\end{mymath}
   For this, we use 
@@
 
-$$\xymatrix{
+\begin{mymath}\xymatrix{
 \|X\|_{n+1} \ar[r] \ar[rd] & \|\mathring T_{X}\|_{n+1} \ar[r] \ar[d] & \|\mathring
@@
   & Q &
-} $$
-
-% $$\xymatrix @C=4em{ 
+} \end{mymath}
+
+% \begin{mymath}\xymatrix @C=4em{ 
 %     \cdots \overline{\Delta_3} \tar[r]{4pt} \ar[rd]_{f\circ \pi_1} & \overline{\Delta_2}
@@
 %     & Q &
-%   }$$
+%   }\end{mymath}
 But we know (lemma~\ref{lem:sepiscolim}) that $\separated T$ is the
@@
 (Bx)$, equal when precomposed with $\pi_1$.
-$$\xymatrix @R=4em @C=4em{
+\begin{mymath}\xymatrix @R=4em @C=4em{
   \sum_{e:E} (\chi\, e) \ar@<-2pt>[r]_{g\circ\pi_1} \ar@<2pt>[r]^{f\circ \pi_1} \ar[d]_{\mathrm{dense}}& \sum_{x :A} (Bx) \\
   E \ar@<-2pt>[ru]_{g} \ar@<2pt>[ru]^{f}&
-}$$
+}\end{mymath}
 We can restrict the previous diagram to 
-$$\xymatrix @R=4em @C=5em{
+\begin{mymath}\xymatrix @R=4em @C=5em{
   \sum_{e:E} (\chi\, e) \ar@<-2pt>[r]_{\pi_1\circ g\circ\pi_1} \ar@<2pt>[r]^{\pi_1\circ f\circ \pi_1} \ar[d]_{\mathrm{dense}}& \sum_{x :A} (Bx) \\
   E \ar@<-2pt>[ru]_{\pi_1\circ g} \ar@<2pt>[ru]^{\pi_1\circ f}&
-}$$
+}\end{mymath}
 and as $A$ is separated, $\pi_1\circ f = \pi_2 \circ g$.
@@
 
-$$\xymatrix@C=8em@R=4em{
+\begin{mymath}\xymatrix@C=8em@R=4em{
   \sum_{y:\sum_{e:E} (\chi\,
@@
   \sum_{y:E} x = y \ar@<2pt>[ur]^{\pi_2\circ f\circ \pi_1} \ar@<-2pt>[ur]_{\pi_2\circ g\circ \pi_1}&
-}$$
+}\end{mymath}
 Using the sepatation property of $B\,x$, one can show that second
@@
 $\modal_{n+1}T$ can be given explicitly by
-$$
+\begin{mymath}
 \modal_{n+1} T \ \defeq \sum_{u:T \to \Type_n^\modal} \modal_{-1}\left\| \sum_{a:T} 
             (\lambda t,~\modal_n (a=t)) = u\right\|.
-$$
+\end{mymath}
 
@@
     Now, let define $\phi : \modal_{-1} T \to \modal_0T$ by
-    $$\phi t = (\lambda t',\, \modal_{-1}(t=t') ; \eta
-    (t;\mathrm{idpath}))$$
+    \begin{mymath}\phi t = (\lambda t',\, \modal_{-1}(t=t') ; \eta
+    (t;\mathrm{idpath}))\end{mymath}
     and $\psi : \modal_0T\to \modal_{-1} T$ by obtaining the
@@
     a modal proposition), and letting
-    $$\psi (u;x) = \eta_T a.$$
+    \begin{mymath}\psi (u;x) = \eta_T a.\end{mymath}
     These two maps form an equivalence (the section and retraction are
@@
 indeed, for example, the term 
-$$\left( \lambda P:\HProp,~\left\{\begin{array}{l} 0 \text{ if $P$ is
+\begin{mymath}\left( \lambda P:\HProp,~\left\{\begin{array}{l} 0 \text{ if $P$ is
                                     false} \\ 1 \text{ if $P$ is
                                     true} \end{array}\right. \right)
-                            P~:~\nat$$
+                            P~:~\nat\end{mymath}
 will not reduce to a numeral if the chosen property $P$ is not
@@
 % let $c:nat \to \nat$ implementing the Collatz function. Then, for any
-% $n:\nat$, the type $$C_n \defeq \sum_{k:\nat}\left( c^k(n) = 1 \land k\text{ is the
-%   smallest such $\nat$}\right)$$ is an $\HProp$.
+% $n:\nat$, the type \begin{mymath}C_n \defeq \sum_{k:\nat}\left( c^k(n) = 1 \land k\text{ is the
+%   smallest such $\nat$}\right)\end{mymath} is an $\HProp$.
 % Thus, for any $n$, one have $C_n + \lnot C_n$, and one can {\em

### 2015-12-11 paper/paper_acm.tex
reduce space

@@
 
-\newenvironment{mymath}%
-{\vspace{-0.7em}\begin{equation*}}%
-  {\vspace{-0.7em}\end{equation*}}
+\newenvironment{mymath}[1][-0.7em]{%
+  \newcommand\mymathaux{\vspace{#1}}%
+  \vspace{#1}%
+  \begin{equation*}%
+  }{ %
+    \mymathaux%
+    % \vspace{#1}
+  \end{equation*}}
 
@@
 %
+\vspace{-0.7em}
 \begin{equation}\label{equ:sheaf_def}
 \Hom{\E}(E,F) \rightarrow \Hom{\E}(A,F).
+\vspace{-0.7em}
 \end{equation}
@@
 sheafification. 
-%
-Section~\ref{ssec:hott} introduces the notion of homotopy types and
-classifying objects as defined in~\cite{sets_in_hott}.
-%
-Section~\ref{sec:epi-mono-fact} presents the usual epi-mono
-factorization system in the light of homotopy type theory.
-%
-Section~\ref{sec:colim-homot-type} introduces the notion of colimits
-in homotopy type theory. 
-%
-And Section~\ref{sec:giraud-ax} defines an analogue of
-Giraud-Rezk-Lurie axioms on higher toposes that we need in the
-definition of sheafification.
-%
+% %
+% Section~\ref{ssec:hott} introduces the notion of homotopy types and
+% classifying objects as defined in~\cite{sets_in_hott}.
+% %
+% Section~\ref{sec:epi-mono-fact} presents the usual epi-mono
+% factorization system in the light of homotopy type theory.
+% %
+% Section~\ref{sec:colim-homot-type} introduces the notion of colimits
+% in homotopy type theory. 
+% %
+% And Section~\ref{sec:giraud-ax} defines an analogue of
+% Giraud-Rezk-Lurie axioms on higher toposes that we need in the
+% definition of sheafification.
+% %
 The definitions of Section~\ref{ssec:hott} and \ref{sec:epi-mono-fact}
@@
 
-Finally, we need to use mere propositions as basic elements of a logic
-when characterizing properties of types or functions. 
-%
-But as some constructors do not preserve the level of homotopy type
-(\eg sums or $\Sigma$-types), so we need to introduce $\squash{X}$, the
-{\em (propositional) truncation} of $X$ (also called bracket type, or
-squash type). This operation on types allows to truncate a type down
-to a mere proposition. Then for instance, the existential
-quantification on mere propositions ``there exists x:A such that
-P(x)'' can be defined as $ \squash{\sum_{x:A} P(x)}.  $
-
-Although proposition truncation could be seen as a primitive type former
-in homotopy type theory, it can also be defined using higher inductive
-types~\cite{lumsdaine2011higher,lumsdaine2013higher}. We adopt this
-point of view as it allows one to maintain the slogan that homotopy type
-theory is type theory plus univalence and higher inductive types.
-%
-
-Given any type $A$, its propositional truncation 
-$\squash{A} : \HProp$ is generated by 
+One has a way, from any type $T$, to construct a type
+$\|T\|_n:\Type_n$ generated by
 \begin{itemize}
-\item a function $|\cdot|_A : A \to \squash{A}$,
-\item for any $x,y:\squash{A}$, a path $x=y$.
+\item a function $|\cdot|_n : T \to \|A\|_n$,
+\item a proof of $\IsType n \|T\|_n$,
 \end{itemize}
-% 
-The recursion principle of $\squash{A}$  asserts that any mere
-proposition that follows from $A$ already follows from $\squash{A}$.
+ satisfying the following universal property:
+% Finally, we need to use mere propositions as basic elements of a logic
+% when characterizing properties of types or functions. 
+% %
+% But as some constructors do not preserve the level of homotopy type
+% (\eg sums or $\Sigma$-types), so we need to introduce $\squash{X}$, the
+% {\em (propositional) truncation} of $X$ (also called bracket type, or
+% squash type). This operation on types allows to truncate a type down
+% to a mere proposition. Then for instance, the existential
+% quantification on mere propositions ``there exists x:A such that
+% P(x)'' can be defined as $ \squash{\sum_{x:A} P(x)}.  $
+
+% Although proposition truncation could be seen as a primitive type former
+% in homotopy type theory, it can also be defined using higher inductive
+% types~\cite{lumsdaine2011higher,lumsdaine2013higher}. We adopt this
+% point of view as it allows one to maintain the slogan that homotopy type
+% theory is type theory plus univalence and higher inductive types.
+% %
+
+% Given any type $A$, its propositional truncation 
+% $\squash{A} : \HProp$ is generated by 
+% \begin{itemize}
+% \item a function $|\cdot|_A : A \to \squash{A}$,
+% \item for any $x,y:\squash{A}$, a path $x=y$.
+% \end{itemize}
+% % 
+% The recursion principle of $\squash{A}$  asserts that any mere
+% proposition that follows from $A$ already follows from $\squash{A}$.
 \begin{lem}
-  For any $A:\Type$ and $B:\HProp$, if $f:A \to B$ then there is an
-  induced $g:\squash{A}\to B$ such that $g(|a|)= f(a)$ for any $a:A$.
+  For any $A:\Type$ and $B:\Type_n$, if $f:A \to B$ then there is an
+  induced $g:\|A\|_n\to B$ such that $g(|a|_n)= f(a)$ for any $a:A$.
 \end{lem}
 %
-We refer the reader to~\cite{hottbook} for more details on propositional truncation.
+We refer the reader to~\cite[7.3]{hottbook} for more details on propositional truncation.
 
@@
 % 
-Following the notion of (homotopy) limits of diagrams over graphs that
-has been defined in homotopy type theory~\cite{lumsdaine}, we present
-here a definition of colimits of diagrams overs graphs, presented in~\cite{sets_in_hott}. 
-%
-The main difference between limits and colimits is that limits are
-simply given by $\Sigma$-types, and thus exist already in traditional
-type theory, whereas the situation is more complicated for colimits as
-it requires the use of higher inductive types.
-
-Following~\cite{lumsdaine}, we introduce the notion of graph and
-diagram over a graph.
-%
-\begin{defi}
-  A {\em graph} $G$ is the data of
-  \begin{itemize}
-  \item a type $G_0$ of vertices ;
-  \item for any $i,j:G_0$, a type $G_1(i,j)$ of edges.
-  \end{itemize}
-
-  A {\em diagram} $D$ over a graph $G$ is the data of
-  \begin{itemize}
-  \item for any $i:G_0$, a type $D_0(i)$ ;
-  \item for any $i,j:G_0$ and all $\phi : G_1(i,j)$, a map $D_1(\phi)
-    : D_0(i) \to D_0(j)$
-  \end{itemize}
-\end{defi}
+Following the definition of graphs and diagrams defined
+in~\cite{lumsdaine}, we recall the definition of colimits of
+diagrams overs graphs presented in~\cite{sets_in_hott}. 
+%
+% The main difference between limits and colimits is that limits are
+% simply given by $\Sigma$-types, and thus exist already in traditional
+% type theory, whereas the situation is more complicated for colimits as
+% it requires the use of higher inductive types.
+
+% Following~\cite{lumsdaine}, we introduce the notion of graph and
+% diagram over a graph.
+% %
+% \begin{defi}
+%   A {\em graph} $G$ is the data of
+%   \begin{itemize}
+%   \item a type $G_0$ of vertices ;
+%   \item for any $i,j:G_0$, a type $G_1(i,j)$ of edges.
+%   \end{itemize}
+
+%   A {\em diagram} $D$ over a graph $G$ is the data of
+%   \begin{itemize}
+%   \item for any $i:G_0$, a type $D_0(i)$ ;
+%   \item for any $i,j:G_0$ and all $\phi : G_1(i,j)$, a map $D_1(\phi)
+%     : D_0(i) \to D_0(j)$
+%   \end{itemize}
+% \end{defi}
 
@@
 \end{defi}
-
-Colimits behave well with truncations, in the following sense:
-
-\begin{lem}
-  Let $D$ be a diagram, $m$ a truncation index, and
-  $P:\Type_{m}$. Let's call $P$ a $m$-colimit of $D$ if
-  for any other $X:\Type_m$, the
-  map
-  \begin{mymath}\lambda f:P \to X, \left( \lambda i,~f \circ q_i\, ;\, \lambda i\, j\,
-  \phi,\ \precompose_f (p_{i, j}^\phi)) \right)\end{mymath}
-  is an equivalence.
-
-  Then, if $\|D\|_m$ is the same diagram as $D$, but where every type
-  is $m$-truncated, $P$ is a $m$-colimit of $\|D\|_m$.
-\end{lem}
+Using higher inductive types, every diagram $D$ on a graph $G$ admits a
+colimit in homotopy type theory.
+
+% Colimits behave well with truncations, in the following sense:
+
+% \begin{lem}
+%   Let $D$ be a diagram, $m$ a truncation index, and
+%   $P:\Type_{m}$. Let's call $P$ a $m$-colimit of $D$ if
+%   for any other $X:\Type_m$, the
+%   map
+%   \begin{mymath}\lambda f:P \to X, \left( \lambda i,~f \circ q_i\, ;\, \lambda i\, j\,
+%   \phi,\ \precompose_f (p_{i, j}^\phi)) \right)\end{mymath}
+%   is an equivalence.
+
+%   Then, if $\|D\|_m$ is the same diagram as $D$, but where every type
+%   is $m$-truncated, $P$ is a $m$-colimit of $\|D\|_m$.
+% \end{lem}
 
@@
 
-Using higher inductive types, every diagram $D$ on a graph $G$ admits a
-colimit in homotopy type theory. This colimit $\colim D$ is given by
+This colimit $\colim D$ is given by
 %
@@
   Let $A,S:\Type_{n+1}$, $S$ separated, and $f:A \to S$. Then if 
+  \vspace{-0.7em}
   \begin{equation}
@@
     \forall a,b:A,~f a = f b \simeq \modal (a=b),
+    \vspace{-0.7em}
   \end{equation}

### 2015-12-11 paper/paper_acm.tex
reduce space v2

@@
 In what follows, we work in the context of homotopy type theory, seen
-as Martin-Löf type thepry, with univalence axiom (and thus functional
+as Martin-Löf type theory, with univalence axiom (and thus functional
 extensionality) and higher inductive types.
@@
 %
-\begin{align*}
-  \chi &: \left(\sum_{A:\Type} \sum_{f:A \to B} \prod_{b\in B}
+\begin{mymath}[0em]
+  \chi : \sum_{A:\Type} \sum_{f:A \to B} \prod_{b\in B}
 \IsType n\
-\fib{f}{b}\right) \\ &\qquad\xrightarrow{\sim} 
+\fib{f}{b} \xrightarrow{\sim} 
  (B \to \Type_n)
-\end{align*}
+\end{mymath}
 %
@@
 % 
-\begin{defi}
+\begin{defi}\label{def:colimit}
 Let $G$ be a graph, and $D$ be a diagram on $G$. 
@@
 \item a map $q_i : D_0(i) \to P$ for any
-vertex $i:G_0$, \ie \begin{mymath}q : \prod_{i:G_0} D_0(i) \to P\end{mymath}
+vertex $i:G_0$, \ie \begin{mymath}[0em]q : \prod_{i:G_0} D_0(i) \to P\end{mymath}
 \item for any vertices $i,j:G_0$ and all edges $\phi:G_1(i,j)$, a path
   $p_{i,j}^\phi : q_j \circ D_1(\phi) = q_i$, \ie
-  \begin{mymath}p : \prod_{i,j:G_0} \prod_{\phi:G_1(i,j)} q_j \circ D_1(\phi) = q_i.\end{mymath}
+  \begin{mymath}[0em]p : \prod_{i,j:G_0} \prod_{\phi:G_1(i,j)} q_j \circ D_1(\phi) = q_i.\end{mymath}
 \end{itemize}
@@
 
-% Colimits behave well with truncations, in the following sense:
-
-% \begin{lem}
-%   Let $D$ be a diagram, $m$ a truncation index, and
-%   $P:\Type_{m}$. Let's call $P$ a $m$-colimit of $D$ if
-%   for any other $X:\Type_m$, the
-%   map
-%   \begin{mymath}\lambda f:P \to X, \left( \lambda i,~f \circ q_i\, ;\, \lambda i\, j\,
-%   \phi,\ \precompose_f (p_{i, j}^\phi)) \right)\end{mymath}
-%   is an equivalence.
-
-%   Then, if $\|D\|_m$ is the same diagram as $D$, but where every type
-%   is $m$-truncated, $P$ is a $m$-colimit of $\|D\|_m$.
-% \end{lem}
+In~\ref{ssec:from-type-separated}, we will need to know how colimits
+behave with respect to truncations. An answer is given by the following:
+
+\begin{lem}
+  Let $D$ be a diagram, $m$ a truncation index, and
+  $P:\Type_{m}$. 
+  Then, if $\|D\|_m$ is the same diagram as $D$, but where every type
+  is $m$-truncated, $P$ is a $m$-colimit\footnote{$P$ is a $m$-colimit
+    if $P$ satisfies the same property as in~\ref{def:colimit} when
+    we replace $\Type$ by $\Type_m$} of $\|D\|_m$.
+\end{lem}
 
@@
 
-This colimit $\colim D$ is given by
-%
-\begin{itemize}
-\item a function $q:\prod_{i:G_0} D_0(i) \to \colim D$
-\item for any $i,j:G_0$ and $\phi:G_1(i,j)$, a path
-  $q_j \circ D_1(\phi) = q_i$.
-\end{itemize}
-%
-Although we have proved in our formalization that $\colim D$ is
-actually a colimit over $D$, we do not detail the proof here as the
-existence of colimits is not used in the definition of the
-sheafification process. We only make use of the following special case
-of Giraud-Rezk-Lurie axioms.
+% This colimit $\colim D$ is given by
+% %
+% \begin{itemize}
+% \item a function $q:\prod_{i:G_0} D_0(i) \to \colim D$
+% \item for any $i,j:G_0$ and $\phi:G_1(i,j)$, a path
+%   $q_j \circ D_1(\phi) = q_i$.
+% \end{itemize}
+% %
+% Although we have proved in our formalization that $\colim D$ is
+% actually a colimit over $D$, we do not detail the proof here as the
+% existence of colimits is not used in the definition of the
+% sheafification process. We only make use of the following special case
+% of Giraud-Rezk-Lurie axioms.
 
@@
 
-Section~\ref{sec:definition} presents the definition and basic
-properties of left exact modalities. 
-%
-Section~\ref{sec:new-type-theory} discusses how a left exact modality
-induces a new homotopy theory, and in which case this theory is
-consistent.
-%
-Finally, Section~\ref{sec:examples-left-exact} introduces some
-examples, and in particular the modality for the dense topology.
+% Section~\ref{sec:definition} presents the definition and basic
+% properties of left exact modalities. 
+% %
+% Section~\ref{sec:new-type-theory} discusses how a left exact modality
+% induces a new homotopy theory, and in which case this theory is
+% consistent.
+% %
+% Finally, Section~\ref{sec:examples-left-exact} introduces some
+% examples, and in particular the modality for the dense topology.
 
@@
 and thus on homotopy fibers, we need to take a closer look at the way left
-exact modalities preserve homotopy fibers.
+exact modalities preserve homotopy fibers. The following is a part of
+the library~\cite{hottlib}.
 %
@@
 is the fiber of $\modal f$ above $\eta_Y y$:
-\begin{mymath}\modal \left(\sum_{x:X}  (f x = y)\right) = \sum_{x:\modal X}
+\begin{mymath}[-0.3em]\modal \left(\sum_{x:X}  (f x = y)\right) = \sum_{x:\modal X}
 (\modal f x = \eta_Y y),\end{mymath}
@@
 Moreover the following diagram commute
-\begin{mymath}\xymatrix{
+\begin{mymath}[-0.3em]\xymatrix{
   \sum_{x:X} (f x = y) \ar[r]^\eta \ar[d]_\gamma & \modal \left(\sum_{x:X}  (f x = y)\right) \ar@{=}[dl]\\
@@
 \end{prop}
-\begin{proof}
-% \nt{ça a du sens de mettre une preuve, non ? c'est non trivial et
-%   nouveau par rapport aux travaux sur les modalités. Ou alors c'est
-%   la preuve de Shulman qui tu n'as pas faite. En tout cas, il faut
-%   clarifier.}
-
-The proof is based on the following central result.
-\begin{lem}
-  Let $X:\Type_p$, $Y:\Type_p^\modal$ and $f:X\to Y$. If for all $y:Y$,
-  $\modal (\fib f y)$ is contractible, then $\modal X = Y$.
-\end{lem}
-%
-It is straighforward to define a map
-\begin{mymath}\phi:\sum_{x:X}  (f x = y)\to
-\sum_{x:\modal X} (\modal f x = \eta_Y y),\end{mymath}
-using $\eta$ functions.
-We just need to check that every $\modal$-fiber $\modal(\fib \phi {x;p})$ is
-contractible.
-Technical transformations allow one to prove
-\begin{mymath}\fib\phi{x;p} = \fib s {y;p^{-1}}\end{mymath}
-for some $a:\modal X$, $b:\modal Y$ and 
-\begin{mymath}s:\fib{\eta_X}a \to \fib{\eta_Y}b.\end{mymath}
-%
-From definition of left exactness, one can deduce the following:
-\begin{lem}
-  Let $A,B:\Type_p$. Let $f:A\to B$. If $\modal A$ and $\modal B$ are
-  contractible, then so is $\fib f b$ for any $b:B$.
-\end{lem}
-Thus, we just need to prove that $\modal(\fib {\eta_X} a)$ and
-$\modal(\fib {\eta_Y} b)$ are contractible.
-But we have the lemma:
-\begin{lem}
-  Let $A:\Type_p$. Then for any $y:\modal A$, $\modal(\fib {\eta_A}
-  y)$ is contractible.
-\end{lem}
-
-Finally, $\modal(\fib s{y;p^{-1}})$ is contractible, so $\modal(\fib \phi {x;p})$ also, and the result is proved.
-%\kq{Finish that}
-
-%\nt{Indeed, I can't follow the proof for the moment}
-\end{proof}
-
+% \begin{proof}
+% % \nt{ça a du sens de mettre une preuve, non ? c'est non trivial et
+% %   nouveau par rapport aux travaux sur les modalités. Ou alors c'est
+% %   la preuve de Shulman qui tu n'as pas faite. En tout cas, il faut
+% %   clarifier.}
+
+% The proof is based on the following central result.
+% \begin{lem}
+%   Let $X:\Type_p$, $Y:\Type_p^\modal$ and $f:X\to Y$. If for all $y:Y$,
+%   $\modal (\fib f y)$ is contractible, then $\modal X = Y$.
+% \end{lem}
+% %
+% It is straighforward to define a map
+% \begin{mymath}\phi:\sum_{x:X}  (f x = y)\to
+% \sum_{x:\modal X} (\modal f x = \eta_Y y),\end{mymath}
+% using $\eta$ functions.
+% We just need to check that every $\modal$-fiber $\modal(\fib \phi {x;p})$ is
+% contractible.
+% Technical transformations allow one to prove
+% \begin{mymath}\fib\phi{x;p} = \fib s {y;p^{-1}}\end{mymath}
+% for some $a:\modal X$, $b:\modal Y$ and 
+% \begin{mymath}s:\fib{\eta_X}a \to \fib{\eta_Y}b.\end{mymath}
+% %
+% From definition of left exactness, one can deduce the following:
+% \begin{lem}
+%   Let $A,B:\Type_p$. Let $f:A\to B$. If $\modal A$ and $\modal B$ are
+%   contractible, then so is $\fib f b$ for any $b:B$.
+% \end{lem}
+% Thus, we just need to prove that $\modal(\fib {\eta_X} a)$ and
+% $\modal(\fib {\eta_Y} b)$ are contractible.
+% But we have the lemma:
+% \begin{lem}
+%   Let $A:\Type_p$. Then for any $y:\modal A$, $\modal(\fib {\eta_A}
+%   y)$ is contractible.
+% \end{lem}
+
+% Finally, $\modal(\fib s{y;p^{-1}})$ is contractible, so $\modal(\fib \phi {x;p})$ also, and the result is proved.
+% %\kq{Finish that}
+
+% %\nt{Indeed, I can't follow the proof for the moment}
+% \end{proof}
+
+\kq{Is the following usefull?}
 In the same way, left exact modalities preserve homotopy types.
@@
   \item For $n=-1$: Let $T:\HProp$. Then
-    \begin{align*}
-      \modal_{0} \widehat T &= \sum_{u:T \to \Type_n^\modal} \modal_{-1}\left\| \sum_{a:T} 
-      (\lambda t,~\modal_{-1} (a=t)) = u\right\| \\
-      &= \sum_{u:T \to \Type_n^\modal} \modal_{-1}\left( \sum_{a:T} 
+    \begin{mymath}
+      \modal_{0} \widehat T = \sum_{u:T \to \Type_n^\modal} \modal_{-1}\left( \sum_{a:T} 
       (\lambda t,~\modal_{-1} (a=t)) = u\right)
-    \end{align*}
+    \end{mymath}
     because the type inside the truncation is already in $\HProp$.
@@
     lower homotopy types. Let $\modal_{n+2}$ be as above, and let
-    $T:\Type_{n+1}$. Then
-    \begin{align*}
-      \modal_{n+2} \widehat  T &= \hspace{-1em} \sum_{u:T \to \Type_{n+1}^\modal}
-                       \hspace{-1em} \modal_{-1}\left\| \sum_{a:T} 
-            (\lambda t,~\modal_{n+1} (a=t)) = u\right\| \\
-                     &= \hspace{-1em} \sum_{u:T \to
+    $T:\Type_{n+1}$. Then, as $\modal_{n+1}$ is compatible with $\modal_{n}$, and
+    $(a=t)$ is in $\Type_n$,
+    \begin{mymath}
+      \modal_{n+2} \widehat = \hspace{-1em} \sum_{u:T \to
         \Type_{n+1}^\modal} 
       \hspace{-1em} \modal_{-1}\left\| \sum_{a:T} 
-            (\lambda t,~\modal_{n} (a=t)) = u\right\|
-    \end{align*}
-    because $\modal_{n+1}$ is compatible with $\modal_{n}$, and
-    $(a=t)$ is in $\Type_n$.
+        (\lambda t,~\modal_{n} (a=t)) = u\right\|.
+    \end{mymath}
 
@@
 
-\begin{exm}
 If $\modal_{-1}$ is the double negation modality on $\HProp$ and
-  $i=-1$, for any type $X$:
-  \begin{align*}
-  \modal X &\defeq \lnot\lnot \|X\|  \\
-           &= \lnot\lnot X
-  \end{align*}
-  by univalence for h-propositions.
-
- Thus, $\modal$ is exactly the double negation modality on $\Type$
+$i=-1$, $\modal$ is exactly the double negation modality on $\Type$
 described in~\ref{sec:examples-left-exact}.
-\end{exm}
 
@@
 Unfortunately, one lose canonicity when working in $\Type^\modal$;
-indeed, for example, the term 
-\begin{mymath}\left( \lambda P:\HProp,~\left\{\begin{array}{l} 0 \text{ if $P$ is
+indeed, for example, the function 
+\begin{mymath}\lambda P:\HProp,~\left\{\begin{array}{l} 0 \text{ if $P$ is
                                     false} \\ 1 \text{ if $P$ is
-                                    true} \end{array}\right. \right)
-                            P~:~\nat\end{mymath}
-will not reduce to a numeral if the chosen property $P$ is not
-decidable. However, we keep canonicity for ``old'' terms, i.e. the
+                                    true} \end{array}\right.~:~\HProp
+                              \to \nat\end{mymath}
+will not reduce to a numeral when applied to a undecidable
+property. However, we keep canonicity for ``old'' terms, i.e. the
 ones constructed without using excluded middle.

### 2015-12-11 paper/paper_acm.tex
reduce space v3

@@
 \usepackage{wasysym}
+\usepackage{microtype}
+
 \usepackage[all]{xy}
@@
   %
-  \begin{mymath}\prod_{e:E}~ \left(\sum_{e':E} e=e'\right) \simeq \modal_n
+  \begin{mymath}[-0.3em]\prod_{e:E}~ \left(\sum_{e':E} e=e'\right) \simeq \modal_n
   (\chi~e).\end{mymath}
@@
   For any type $A$ and $E$, $\iota : A \to E$ and $F:\Type_{n+1}$, we define
-    \begin{mymath}
-    \Phi_E^\iota : (E \to F) \to (A \to F) 
-    \end{mymath}
-    %
-    as the map sending an
-    arrow $f:E\to F$ to its restriction $f \circ \iota$.
+    $\Phi_E^\iota \defeq \cdot \circ \iota : (E \to F) \to (A \to F) $
+    as the map sending an  arrow $f:E\to F$ to its restriction $f \circ \iota$.
   % \item if $A \to E$ with $n$-truncated homotopy fibers, $\Phi_E^\chi$ is the
@@
 
-  We define $g$ as 
-  \begin{mymath}g(e) = \modal_n \left( \fib \phi e \right), \end{mymath}
+  We define $g$ as $g(e) = \modal_n \left( \fib \phi e \right)$,
   where
-  \begin{mymath}\phi : \left(\sum_{b:\left(\sum_{e:E} (\chi\, e)\right)} (f\,
+  \begin{mymath}[0em]\phi : \left(\sum_{b:\left(\sum_{e:E} (\chi\, e)\right)} (f\,
     b)\right) \to E\end{mymath}
   defined by $\phi(x) = (x_1)_1$.
+\kq{Is it necessary to state the lemma ?}
   Using the following properties on fibers, we can prove that it
@@
 }\end{mymath}
-We can restrict the previous diagram to 
-\begin{mymath}\xymatrix @R=4em @C=5em{
-  \sum_{e:E} (\chi\, e) \ar@<-2pt>[r]_{\pi_1\circ g\circ\pi_1} \ar@<2pt>[r]^{\pi_1\circ f\circ \pi_1} \ar[d]_{\mathrm{dense}}& \sum_{x :A} (Bx) \\
-  E \ar@<-2pt>[ru]_{\pi_1\circ g} \ar@<2pt>[ru]^{\pi_1\circ f}&
-}\end{mymath}
-and as $A$ is separated, $\pi_1\circ f = \pi_2 \circ g$.
+% We can restrict the previous diagram to 
+% \begin{mymath}\xymatrix @R=4em @C=5em{
+%   \sum_{e:E} (\chi\, e) \ar@<-2pt>[r]_{\pi_1\circ g\circ\pi_1} \ar@<2pt>[r]^{\pi_1\circ f\circ \pi_1} \ar[d]_{\mathrm{dense}}& \sum_{x :A} (Bx) \\
+%   E \ar@<-2pt>[ru]_{\pi_1\circ g} \ar@<2pt>[ru]^{\pi_1\circ f}&
+% }\end{mymath}
+As $A$ is separated, $\pi_1\circ f = \pi_2 \circ g$.
 For the second components, let $x:E$. Notice that for any $x:E$,
@@
     \begin{mymath}
-      \modal_{n+2} \widehat = \hspace{-1em} \sum_{u:T \to
+      \modal_{n+2} \widehat T= \hspace{-1em} \sum_{u:T \to
         \Type_{n+1}^\modal} 
@@
 
-We provide a more detailed insight of the structure of our formalization:
-\begin{itemize} 
-\item Colimits and iterated kernel pairs are formalized in
-\texttt{Limit}, \texttt{T.v}, \texttt{OT.v}v \texttt{OT\_Tf.v}, \texttt{T\_telescope.v},\\ \texttt{Tf\_Omono\_sep.v}.% (552 lines).
-\item
-Reflective subuniverses and modalities are formalized in\\
-\texttt{reflective\_subuniverse.v}, \texttt{modalities.v}. % (1053 lines).
-\item 
-%
-  The definition of the dense topology as a left exact modality on
-  $\HProp$ is given in \texttt{sheaf\_base\_case.v}. % (186 lines).
-\item
-Section~\ref{sec:sheaves} is formalized in
-\texttt{sheaf\_def\_and\_thm.v}. % (1029 lines).
-\item
-Section~\ref{sec:sheafification} is formalized in
-\texttt{sheaf\_induction.v}. % (2340 lines).
-\end{itemize}
-
-Overall, with other files containing technical lemmas, the project
+% We provide a more detailed insight of the structure of our formalization:
+% \begin{itemize} 
+% \item Colimits and iterated kernel pairs are formalized in
+% \texttt{Limit}, \texttt{T.v}, \texttt{OT.v}v \texttt{OT\_Tf.v}, \texttt{T\_telescope.v},\\ \texttt{Tf\_Omono\_sep.v}.% (552 lines).
+% \item
+% Reflective subuniverses and modalities are formalized in\\
+% \texttt{reflective\_subuniverse.v}, \texttt{modalities.v}. % (1053 lines).
+% \item 
+% %
+%   The definition of the dense topology as a left exact modality on
+%   $\HProp$ is given in \texttt{sheaf\_base\_case.v}. % (186 lines).
+% \item
+% Section~\ref{sec:sheaves} is formalized in
+% \texttt{sheaf\_def\_and\_thm.v}. % (1029 lines).
+% \item
+% Section~\ref{sec:sheafification} is formalized in
+% \texttt{sheaf\_induction.v}. % (2340 lines).
+% \end{itemize}
+
+Overall, % with other files containing technical lemmas,
+the project
 contains 8081 lines, and it could be reduced a bit by improving the

### 2015-12-11 paper/paper_acm.tex
reduce space v4

@@
 \usepackage{microtype}
+% \usepackage[subtle]{savetrees}
 
@@
 
+\let\oldthebibliography=\thebibliography
+  \let\endoldthebibliography=\endthebibliography
+  \renewenvironment{thebibliography}[1]{%
+    \begin{oldthebibliography}{#1}%
+      \setlength{\parskip}{0ex}%
+      \setlength{\itemsep}{0ex}%
+  }%
+  {%
+    \end{oldthebibliography}%
+  }
+
 \begin{document}
@@
 sums~\cite{barthe2002cps}, which puts emphasis on the tedious link
-between the axiom of choice and the law of excluded middle in type theory.
+between the axiom of choice and the law of excluded middle in type theory\looseness=-1.
 
@@
 \begin{mymath}t_f = \lambda a,~(\fib{f}{f(a)},(a,\mathrm{idpath})).\end{mymath}
-
 One has a way, from any type $T$, to construct a type
@@
 \end{defi}
-
 The canonical function $\hat f : A \to \im(f)$ is (-1)-connected,
@@
   \modal (|a|_{n+1}=|b|_{n+1}).\end{mymath}
-
   We use the encode-decode~\cite[Section 8.9]{hottbook} method to characterize $\tilde f |a|_{n+1}
@@
 } \end{mymath} 
-
 % \begin{mymath}\xymatrix @C=4em{ 
@@
     \end{mymath}
-
     It remains to prove that for every $(u,x)$ inhabiting the

### 2015-12-14 paper/paper_acm.tex
reduce dense

@@
 
+\kq{Maybe delete this subsection?}
 We now recall the presence of a factorization system in homotopy type
@@
   $\chi = \modal_n \circ \chi$.
+
+  An $n$-subobject of $E$ classified by $\chi$ is said to be {\em
+    dense in $E$} if its closure is $E$, \ie
+  $\modal_n \circ \chi = \lambda e. 1$ 
 \end{defi}
 
-The definition of a dense $n$-subobject requires more work as the
-characteristic function $\chi_E$ of $E$ seen as an $n$-subobject of
-itself is not the constant function equal to $\one$, as it is the case
-for toposes, but rather the map 
-%
-\begin{mymath}\chi_E  = \lambda e . \ \sum_{e':E} e=e'.\end{mymath}
-%
-This means that we need to take more care of the homotopy structure of
-this characteristic function, and in particular, the definition of a
-dense $n$-subobject comes with an additional coherence condition
-(which is trivial when $n=-1$).
-
-\begin{defi}
-  Let $E$ be a type, and $\chi:E \to \Type_n$. The $n$-subobject $\iota :
-  A \to E$ of $E$
-  classified by $\chi$ is {\em dense} in $E$ when its $\modal_n$-closure
-  is equivalent to $\chi_E$, \ie
-  %
-  \begin{mymath}[-0.3em]\prod_{e:E}~ \left(\sum_{e':E} e=e'\right) \simeq \modal_n
-  (\chi~e).\end{mymath}
-  % 
-  % and moreover, for any $x:A$, the following diagram 
-  % commutes, 
-  % \begin{mymath}\xymatrix{
-  %   \sum_{e':A} x = e' \ar@{=}[r] \ar[d]_{\iota'} & (\chi \ x)
-  %   \ar[d]^{\eta_{(\chi \ x)}}\\
-  %   \sum_{e':E} x = e' \ar@{=}[r] & \modal_n (\chi \ x)
-  % }\end{mymath}
-  % where $\iota': x \mapsto (\iota (\pi_1 x) ; \pi_2 x)$.
-\end{defi}
+% The definition of a dense $n$-subobject requires more work as the
+% characteristic function $\chi_E$ of $E$ seen as an $n$-subobject of
+% itself is not the constant function equal to $\one$, as it is the case
+% for toposes, but rather the map 
+% %
+% \begin{mymath}\chi_E  = \lambda e . \ \sum_{e':E} e=e'.\end{mymath}
+% %
+% This means that we need to take more care of the homotopy structure of
+% this characteristic function, and in particular, the definition of a
+% dense $n$-subobject comes with an additional coherence condition
+% (which is trivial when $n=-1$).
+
+% \begin{defi}
+%   Let $E$ be a type, and $\chi:E \to \Type_n$. The $n$-subobject $\iota :
+%   A \to E$ of $E$
+%   classified by $\chi$ is {\em dense} in $E$ when its $\modal_n$-closure
+%   is equivalent to $\chi_E$, \ie
+%   %
+%   \begin{mymath}[-0.3em]\prod_{e:E}~ \left(\sum_{e':E} e=e'\right) \simeq \modal_n
+%   (\chi~e).\end{mymath}
+%   % 
+%   % and moreover, for any $x:A$, the following diagram 
+%   % commutes, 
+%   % \begin{mymath}\xymatrix{
+%   %   \sum_{e':A} x = e' \ar@{=}[r] \ar[d]_{\iota'} & (\chi \ x)
+%   %   \ar[d]^{\eta_{(\chi \ x)}}\\
+%   %   \sum_{e':E} x = e' \ar@{=}[r] & \modal_n (\chi \ x)
+%   % }\end{mymath}
+%   % where $\iota': x \mapsto (\iota (\pi_1 x) ; \pi_2 x)$.
+% \end{defi}
 
@@
   For any type $A$ and $E$, $\iota : A \to E$ and $F:\Type_{n+1}$, we define
-    $\Phi_E^\iota \defeq \cdot \circ \iota : (E \to F) \to (A \to F) $
+    $\Phi_E^\iota \defeq - \circ \iota : (E \to F) \to (A \to F) $
     as the map sending an  arrow $f:E\to F$ to its restriction $f \circ \iota$.

### 2015-12-15 paper/paper_acm.tex
reduce

@@
 \usepackage[colorlinks=true]{hyperref}
+
+% \usepackage{mathptmx}
+% \usepackage[scaled=.90]{helvet}
+% \usepackage{courier}
 
@@
 %
-One use of sheafification is the construction a boolean
+One use of sheafification is the construction of a boolean
 topos, thus validating the classical principle of the law of excluded
@@
 equalities which are trival (by proof irrelevance) in the topos-theoretic
-setting become relevant and require a careful use of univalence
-with appropriate isomorphisms and a lot of reasoning on equalities
+setting become relevant and require a lot of reasoning on equalities
 between paths.
@@
 % %
-The definitions of Section~\ref{ssec:hott} and \ref{sec:epi-mono-fact}
-are part, or direct applications of the Coq library~\cite{hottlib}, while other definitions
+The definitions of Section~\ref{ssec:hott} 
+are part, or direct applications of~\cite{hottbook}, while other definitions
 and theorems are specific to our formalization.
@@
 \end{itemize}
-
-Using $\Type_n$ instead of just $\Type$ is the first step to connect
-type theory to higher topos theory. The next step is to exhibit a
-hierarchy of subobject classifiers on $n$-truncated homotopy fibers.
+One has a way, from any type $T$, to construct a type
+$\|T\|_n:\Type_n$ as the HIT generated by
+\begin{itemize}
+\item a function $|\cdot|_n : T \to \|A\|_n$,
+\item a proof of $\IsType n \|T\|_n$,
+\end{itemize}
+ satisfying the following universal property:
+% Finally, we need to use mere propositions as basic elements of a logic
+% when characterizing properties of types or functions. 
+% %
+% But as some constructors do not preserve the level of homotopy type
+% (\eg sums or $\Sigma$-types), so we need to introduce $\squash{X}$, the
+% {\em (propositional) truncation} of $X$ (also called bracket type, or
+% squash type). This operation on types allows to truncate a type down
+% to a mere proposition. Then for instance, the existential
+% quantification on mere propositions ``there exists x:A such that
+% P(x)'' can be defined as $ \squash{\sum_{x:A} P(x)}.  $
+
+% Although proposition truncation could be seen as a primitive type former
+% in homotopy type theory, it can also be defined using higher inductive
+% types~\cite{lumsdaine2011higher,lumsdaine2013higher}. We adopt this
+% point of view as it allows one to maintain the slogan that homotopy type
+% theory is type theory plus univalence and higher inductive types.
+% %
+
+% Given any type $A$, its propositional truncation 
+% $\squash{A} : \HProp$ is generated by 
+% \begin{itemize}
+% \item a function $|\cdot|_A : A \to \squash{A}$,
+% \item for any $x,y:\squash{A}$, a path $x=y$.
+% \end{itemize}
+% % 
+% The recursion principle of $\squash{A}$  asserts that any mere
+% proposition that follows from $A$ already follows from $\squash{A}$.
+\begin{lem}
+  For any $A:\Type$ and $B:\Type_n$, if $f:A \to B$ then there is an
+  induced $g:\|A\|_n\to B$ such that $g(|a|_n)= f(a)$ for any $a:A$.
+\end{lem}
+%
+We refer the reader to~\cite[7.3]{hottbook} for more details on 
+truncations.
+
+% Using $\Type_n$ instead of just $\Type$ is the first step to connect
+% type theory to higher topos theory. The next step is to exhibit a
+% hierarchy of subobject classifiers on $n$-truncated homotopy fibers.
 %
@@
 defined as 
-\begin{mymath}
+\begin{mymath}[-0.3em]
 \fib{f}{b} \defeq \sum_{a:A} f(a) = b.
@@
 %
+Again, we define some sugar. A function $f$ is 
+\begin{itemize}
+\item an {\em embedding} if $f$ is $(-1)$-truncated
+\item a {\em surjection} is every fiber of $f$ is merely inhabited
+  (\ie $\|\fib f y\|$ holds for all $y$).
+\end{itemize}
+Then one can show~\cite[Lemma 7.6.4]{hottbook} that any map $f$
+factors uniquely through $\im(f) \defeq \sum_{y:B} \|\fib f y\|$ as a
+surjection followed by an embedding.
+
 Following~\cite{sets_in_hott}, it is possible to show that, for any
@@
 $n$-truncated types and 
-\begin{mymath}t_f = \lambda a,~(\fib{f}{f(a)},(a,\mathrm{idpath})).\end{mymath}
-One has a way, from any type $T$, to construct a type
-$\|T\|_n:\Type_n$ generated by
-\begin{itemize}
-\item a function $|\cdot|_n : T \to \|A\|_n$,
-\item a proof of $\IsType n \|T\|_n$,
-\end{itemize}
- satisfying the following universal property:
-% Finally, we need to use mere propositions as basic elements of a logic
-% when characterizing properties of types or functions. 
-% %
-% But as some constructors do not preserve the level of homotopy type
-% (\eg sums or $\Sigma$-types), so we need to introduce $\squash{X}$, the
-% {\em (propositional) truncation} of $X$ (also called bracket type, or
-% squash type). This operation on types allows to truncate a type down
-% to a mere proposition. Then for instance, the existential
-% quantification on mere propositions ``there exists x:A such that
-% P(x)'' can be defined as $ \squash{\sum_{x:A} P(x)}.  $
-
-% Although proposition truncation could be seen as a primitive type former
-% in homotopy type theory, it can also be defined using higher inductive
-% types~\cite{lumsdaine2011higher,lumsdaine2013higher}. We adopt this
-% point of view as it allows one to maintain the slogan that homotopy type
-% theory is type theory plus univalence and higher inductive types.
-% %
-
-% Given any type $A$, its propositional truncation 
-% $\squash{A} : \HProp$ is generated by 
-% \begin{itemize}
-% \item a function $|\cdot|_A : A \to \squash{A}$,
-% \item for any $x,y:\squash{A}$, a path $x=y$.
-% \end{itemize}
-% % 
-% The recursion principle of $\squash{A}$  asserts that any mere
-% proposition that follows from $A$ already follows from $\squash{A}$.
-\begin{lem}
-  For any $A:\Type$ and $B:\Type_n$, if $f:A \to B$ then there is an
-  induced $g:\|A\|_n\to B$ such that $g(|a|_n)= f(a)$ for any $a:A$.
-\end{lem}
-%
-We refer the reader to~\cite[7.3]{hottbook} for more details on propositional truncation.
-
-\subsection{The (n-Connected,n-Truncated) Factorization System}
-\label{sec:epi-mono-fact}
-
-\kq{Maybe delete this subsection?}
-We now recall the presence of a factorization system in homotopy type
-theory, constituted by embeddings and surjections.
-
-A map $f:A\to B$ is an \emph{embedding} when it is $(-1)$-truncated
-(see definition in Section~\ref{ssec:hott}), and a \emph{surjection}
-when it is $(-1)$-connected, that is when every fiber of $f$ is merely
-inhabited (\ie $\|\fib f y\|$ holds for all $y$).
-
-\begin{defi}
-  Let $f:A\to B$ be a function. The {\em image} of $f$ is defined as 
-  \begin{mymath}\im(f) \defeq \sum_{y:B} \|\fib f y\|.\end{mymath}
-\end{defi}
-The canonical function $\hat f : A \to \im(f)$ is (-1)-connected,
-being the left component of an orthogonal factorization system which
-satisfies~\cite[Lemma 7.6.4]{hottbook}:
-\begin{prop}
-  A map $f:A\to B$ factors uniquely (up-to homotopy) through
-  $\im(f)$ as a (-1)-connected function followed by a (-1)-truncated
-  function.
-\end{prop}
-
-Note that this generalizes to orthogonal factorization systems
-constituted by $n$-truncated morphisms and $n$-connected morphisms, for
-every truncated level $n$. We will only use the factorization system
-at level  $-1$ in the definition of sheafification.
+\begin{mymath}t_f = \lambda
+  a,~(\fib{f}{f(a)},(a,\mathrm{idpath})).\end{mymath}
+
+% \subsection{The (n-Connected,n-Truncated) Factorization System}
+% \label{sec:epi-mono-fact}
+
+% \kq{Maybe delete this subsection?}
+% We now recall the presence of a factorization system in homotopy type
+% theory, constituted by embeddings and surjections.
+
+% A map $f:A\to B$ is an \emph{embedding} when it is $(-1)$-truncated
+% (see definition in Section~\ref{ssec:hott}), and a \emph{surjection}
+% when it is $(-1)$-connected, that is when every fiber of $f$ is merely
+% inhabited (\ie $\|\fib f y\|$ holds for all $y$).
+
+% \begin{defi}
+%   Let $f:A\to B$ be a function. The {\em image} of $f$ is defined as 
+%   \begin{mymath}\im(f) \defeq \sum_{y:B} \|\fib f y\|.\end{mymath}
+% \end{defi}
+% The canonical function $\hat f : A \to \im(f)$ is (-1)-connected,
+% being the left component of an orthogonal factorization system which
+% satisfies~\cite[Lemma 7.6.4]{hottbook}:
+% \begin{prop}
+%   A map $f:A\to B$ factors uniquely (up-to homotopy) through
+%   $\im(f)$ as a (-1)-connected function followed by a (-1)-truncated
+%   function.
+% \end{prop}
+
+% Note that this generalizes to orthogonal factorization systems
+% constituted by $n$-truncated morphisms and $n$-connected morphisms, for
+% every truncated level $n$. We will only use the factorization system
+% at level  $-1$ in the definition of sheafification.
 
@@
 
-In the definition of (a special case of) Giraud-Rezk-Lurie axiom given in
-Section~\ref{sec:giraud-ax}, the colimit of the {\em \v{C}ech nerve}
-plays a central role.
+One desired property we would like to consider concerns the colimits
+of \v{C}ech nerves (Section~\ref{sec:giraud-ax}). This section presents
+a definition of colimits in a type theoretic setting.
+% In the definition of (a special case of) Giraud-Rezk-Lurie axiom given in
+% Section~\ref{sec:giraud-ax}, the colimit of the {\em \v{C}ech nerve}
+% plays a central role.
 % 
@@
   \begin{mymath}\xymatrix{
-    C(f) := X \ar[r]^t & T_f \ar[r]^t & T_{\tilde f} \ar[r] &\cdots
+    C(f) := X \ar[r]^-t & T_f \ar[r]^-t & T_{\tilde f} \ar[r] &\cdots
   }\end{mymath}
@@
 The type of all modal $n$-types, \ie types for which $P$ holds, is
-defined as %$\Type^\modal_p$, so 
+defined as \footnote{Note that we omit the $n$
+on $\modal_n$ in $\Type_n^\modal$ to ease the reading.} %$\Type^\modal_p$, so 
 \begin{mymath}
@@
 \begin{defi}
-  For any type $A$ and $E$, $\iota : A \to E$ and $F:\Type_{n+1}$, we define
-    $\Phi_E^\iota \defeq - \circ \iota : (E \to F) \to (A \to F) $
-    as the map sending an  arrow $f:E\to F$ to its restriction $f \circ \iota$.
+  For any type $E$, $\chi : E \to \Type$ and $F:\Type_{n+1}$, we define
+    $\Phi_E^\chi : (E \to F) \to (\sum_{e:E} \chi\, e \to F) $
+    as the map sending an  arrow $f:E\to F$ to its restriction $f \circ \pi_1$.
   % \item if $A \to E$ with $n$-truncated homotopy fibers, $\Phi_E^\chi$ is the
@@
   A type $F$ in $\Type_{n+1}$ is {\em separated} if for any type $E$, and
-  all dense $n$-subobject of $\iota : A \to E$,
-  $\Phi_E^\iota$ is an embedding.%  In other words, the dotted arrow,
+  all dense $n$-subobject of $E$ classified by $\chi$,
+  $\Phi_E^\chi$ is an embedding.%  In other words, the dotted arrow,
   % if exists, is unique.
@@
 
-With topos theory point of view, it means that given a map $A\to F$,
+With topos theory point of view, it means that given a map $\sum_{e:E}
+\chi\, e \to F$,
 if there is an extension $\tilde f:E\to F$, then it is unique.
@@
   separated, and for any type $E$ and all dense $(-1)$-subobject of
-  $\iota : A \to E$, $\Phi_E^\iota$ is an
+  $E$ classified by $\chi$, $\Phi_E^\chi$ is an
   equivalence. % In other words, the dotted arrow exists and is unique.
@@
 
-In topos-theoretic words, it means that given a map $A\to F$, one can
+In topos-theoretic words, it means that given a map $\sum_{e:E}
+\chi\, e\to F$, one can
 extend it uniquely in $\tilde f:E \to F$.
@@
 \begin{prop}\label{prop:sheaf-is-sheaf}
-  $\Type_n^\modal$ is a sheaf.\footnote{Note that we omit the $n$
-on $\modal_n$ in $\Type_n^\modal$ to ease the reading.}
+  $\Type_n^\modal$ is a sheaf.
 \end{prop}
@@
   where
-  \begin{mymath}[0em]\phi : \left(\sum_{b:\left(\sum_{e:E} (\chi\, e)\right)} (f\,
-    b)\right) \to E\end{mymath}
+  \begin{mymath}[0em]\phi : \sum_{b:\left(\sum_{e:E} (\chi\, e)\right)} (f\,
+    b) \to E\end{mymath}
   defined by $\phi(x) = (x_1)_1$.
@@
 For a given induction step, the sheafification is defined in two
-steps:
-%
-\begin{enumerate}[(i)]
-\item {\em separation:} From any $T$ in $\Type_{n+1}$, we construct
-  its {\em free} separated object $\separated T$ by ``removing'' what
-  prevents $T$ from being a sheaf.
-\item {\em completion:} We add what is missing for the free 
-  separated type to be a sheaf by using closure in a sheaf.
-\end{enumerate}
+steps ; the first one construct, for any $T:\Type_{n+1}$ a separated
+type $\separated T$, and the second one construct a sheaf given any
+separated type.
+% %
+% \begin{enumerate}[(i)]
+% \item {\em separation:} From any $T$ in $\Type_{n+1}$, we construct
+%   its {\em free} separated object $\separated T$ by ``removing'' what
+%   prevents $T$ from being a sheaf.
+% \item {\em completion:} We add what is missing for the free 
+%   separated type to be a sheaf by using closure in a sheaf.
+% \end{enumerate}
 
@@
 Let $T : \Type_{n+1}$. We define $\separated T$ as the image (see
-Section~\ref{sec:epi-mono-fact}) of
+Section~\ref{ssec:hott}) of
 $\modal_n^T \circ \{\cdot\}_T$, as in
@@
 is universal among separated type below $T$. This comes from the
-following central lemma which connects $\separated T$ to the colimit of
-the \v{C}ech nerve of $\mu_T$.
+following construction which connects $\separated T$ to the colimit of
+the iterated kernel pair of $\mu_T$.
 
@@
 topos theory, where the kernel pair of $\mu_T$
-is replaced by its \v{C}ech nerve.
+is replaced by its iterated kernel pair.
 
@@
 As $A$ is separated, $\pi_1\circ f = \pi_2 \circ g$.
-For the second components, let $x:E$. Notice that for any $x:E$,
+For the second components, let $x:E$. Notice that 
 $\sum_{y:E} x = y$ has a dense $n$-subobject, $\sum_{y:\sum_{e:E} (\chi\,
@@
 }\end{mymath}
-Using the sepatation property of $B\,x$, one can show that second
+Using the separation property of $B\,x$, one can show that second
 components, transported correctly along the first components equality,
@@
 %
-Note that we also could have resolved the issue by giving explicitely
-the universe levels we wanted, like in the Coq/HoTT library\cite{hottlib}.
+% Note that we also could have resolved the issue by giving explicitely
+% the universe levels we wanted, like in the Coq/HoTT library\cite{hottlib}.
 
@@
    \\
-    \modal_{-1\phantom{n}} \defeq \lambda T:\HProp,~ \lnot\lnot T \\
-      \displaystyle{\modal_{n+1} \defeq  \lambda T:\Type_{n+1}},\\
-      \hspace{5em} \displaystyle{\sum_{u:T \to \Type_n^\modal} \modal_{-1} 
+    \modal_{-1\phantom{n}}(T) \defeq\lnot\lnot T \\
+      \displaystyle{\modal_{n+1}(T)} \defeq  
+      \displaystyle{\sum_{u:T \to \Type_n^\modal} \!\!\!\!\modal_{-1} 
       \left\|
@@
   totally satisfactory, as every type is collapsed to a truncated
-  one. But some types in homotopy type theorey are not
+  one. But some types in homotopy type theory are not
   truncated\cite[Example 8.8.6]{hottbook}.

### 2015-12-16 paper/paper_acm.tex
reduce again

@@
 
+% \usepackage[moderate]{savetrees}
 % \usepackage{mathptmx}
@@
 (T = U) \to (T \simeq U)
-\end{mymath}
+\end{mymath}%
 %
@@
 j : \Omega \rightarrow \Omega
-\end{mymath}
+\end{mymath}%
 %
@@
 \Char{\closure{A}} = j \circ \Char{A}.
-\end{mymath}
+\end{mymath}%
 %
@@
 \vspace{-0.7em}
-\end{equation}
+\end{equation}%
 %
@@
 \Omega_j = \{ P \in \E \ | \ j P  = P \}.
-\end{mymath}
+\end{mymath}%
 %
@@
 a_j : \E \rightarrow \Sh{\E}
-\end{mymath}
+\end{mymath}%
 which exhibits $\Sh{\E}$ as a reflective
@@
 In what follows, we work in the context of homotopy type theory, seen
-as Martin-Löf type theory, with univalence axiom (and thus functional
+as Martin-Löf type theory, with univalence axiom (thus functional
 extensionality) and higher inductive types.
@@
     is pointed by $c:X$, and every other point in $X$ is connected to $c$.
-  \item $\IsType {(n+1)} X = \prod_{x,y:X} \IsType n (x=y).$
+  \item $\IsType {(n+1)} X \defeq \prod_{x,y:X} \IsType n (x=y).$
   \end{itemize}
-  Then, $\Type_n = \sum_{X:\Type} \IsType n X$.
+  Then, $\Type_n \defeq \sum_{X:\Type} \IsType n X$.
 \end{defi}
@@
 %
-We also define some syntactic sugar for contractible types and mere propositions.
-
-\begin{itemize}
-\item $\IsContr \defeq \IsType {(-2)}$ and $\Contr \defeq \Type_{-2}$
-\item $\IsHProp \defeq \IsType {(-1)}$ and $\HProp \defeq \Type_{-1}$
-\end{itemize}
+When $n=-1$, we will use $\IsHProp$ and $\HProp$ instead of
+$\IsType{(-1)}$ and $\Type_{-1}$.
+% \kq{rm}
+% We also define some syntactic sugar for contractible types and mere propositions.
+
+% \begin{itemize}
+% \item $\IsContr \defeq \IsType {(-2)}$ and $\Contr \defeq \Type_{-2}$
+% \item $\IsHProp \defeq \IsType {(-1)}$ and $\HProp \defeq \Type_{-1}$
+% \end{itemize}
+
 One has a way, from any type $T$, to construct a type
@@
 \fib{f}{b} \defeq \sum_{a:A} f(a) = b.
-\end{mymath} 
+\end{mymath}%
 %
@@
  (B \to \Type_n)
-\end{mymath}
-%
- such that the diagram
-\begin{mymath}
-\xymatrix{
-  A \ar[r]^{\hspace{-1em} t_f} \ar[d]_f & \Type_n^\bullet \ar[d]^{\pi_1}\\
-  B \ar[r]_{\hspace{-1em} \chi_f} & \Type_n
-}
-\end{mymath}
-is a pullback for any $f$ with
- $n$-truncated homotopy fibers where $\Type_n^\bullet \defeq
- \sum_{A:\Type_n} A$ is the universe of pointed
-$n$-truncated types and 
-\begin{mymath}t_f = \lambda
-  a,~(\fib{f}{f(a)},(a,\mathrm{idpath})).\end{mymath}
+\end{mymath}%
+%
+ such that the usual (\cite[Theorem 4.8.4]{hottbook}) subobject classifier diagram is a pullback.
+% \begin{mymath}
+% \xymatrix{
+%   A \ar[r]^{\hspace{-1em} t_f} \ar[d]_f & \Type_n^\bullet \ar[d]^{\pi_1}\\
+%   B \ar[r]_{\hspace{-1em} \chi_f} & \Type_n
+% }
+% \end{mymath}
+% is a pullback for any $f$ with
+ % $n$-truncated homotopy fibers where $\Type_n^\bullet \defeq
+ % \sum_{A:\Type_n} A$ is the universe of pointed
+% $n$-truncated types and 
+% \begin{mymath}t_f = \lambda
+  % a,~(\fib{f}{f(a)},(a,\mathrm{idpath})).
+% \end{mymath}
 
@@
 \item a map $q_i : D_0(i) \to P$ for any
-vertex $i:G_0$, \ie \begin{mymath}[0em]q : \prod_{i:G_0} D_0(i) \to P\end{mymath}
+vertex $i:G_0$, \ie \begin{mymath}[0em]q : \prod_{i:G_0} D_0(i) \to P\end{mymath}%
 \item for any vertices $i,j:G_0$ and all edges $\phi:G_1(i,j)$, a path
   $p_{i,j}^\phi : q_j \circ D_1(\phi) = q_i$, \ie
-  \begin{mymath}[0em]p : \prod_{i,j:G_0} \prod_{\phi:G_1(i,j)} q_j \circ D_1(\phi) = q_i.\end{mymath}
+  \begin{mymath}[0em]p : \prod_{i,j:G_0} \prod_{\phi:G_1(i,j)} q_j \circ D_1(\phi) = q_i.\end{mymath}%
 \end{itemize}
@@
 \begin{mymath}\lambda f:P \to X, \left( \lambda i,~f \circ q_i\, ;\, \lambda i\, j\,
-  \phi,\ \precompose_f (p_{i, j}^\phi)) \right)\end{mymath}
-where $\precompose_f : \phi = \psi \to \phi \circ f = \psi \circ f$,
+  \phi,\, f (p_{i, j}^\phi)) \right)\end{mymath}%
+% where $\precompose_f : \phi = \psi \to \phi \circ f = \psi \circ f$,
 is an equivalence.
@@
   \item $t:~X \to T_f$
-  \item $\alpha:~\forall a\,b:X,~f a = f b \to t a = t b$
-  \item $\alpha_1:~\forall a:X,~\alpha\, a\, a\, 1 = 1$
+  \item $\alpha:~\forall a\,b:X,~f(a) = f(b) \to t(a) = t(b)$
+  \item $\alpha_1:~\forall a:X,~\alpha(a, a, 1) = 1$
   \end{itemize}
@@
   \begin{mymath}\xymatrix{
-    \sum_{a,b:X} f a = f b \dar[r]{4pt}^-{\pi_1}_-{\pi_2} & X
-  }\end{mymath}
+    \sum_{a,b:X} f(a) = f(b) \dar[r]{4pt}^-{\pi_1}_-{\pi_2} & X
+  }\end{mymath}%
   preserving the identity.
@@
     C(f) := X \ar[r]^-t & T_f \ar[r]^-t & T_{\tilde f} \ar[r] &\cdots
-  }\end{mymath}
+  }\end{mymath}%
 \end{defi}
@@
         f & \mapsto & f \circ \eta_A
-      \end{array} \right.\end{mymath}
+      \end{array} \right.\end{mymath}%
     is an equivalence.
@@
 \Type_n^\modal \ \defeq \sum_{T : \Type_n} P (T).
-\end{mymath}
+\end{mymath}%
 %
@@
 is the fiber of $\modal f$ above $\eta_Y y$:
-\begin{mymath}[-0.3em]\modal \left(\sum_{x:X}  (f x = y)\right) = \sum_{x:\modal X}
-(\modal f x = \eta_Y y),\end{mymath}
+\begin{mymath}[-0.3em]\modal \left(\sum_{x:X}  (f(x) = y)\right) = \sum_{x:\modal X}
+(\modal f(x) = \eta_Y(y)),\end{mymath}%
 where $\modal f:\modal X \to \modal Y$ is defined via the equivalence {\it (iv)} and $\eta_Y$.
@@
 \begin{mymath}[-0.3em]\xymatrix{
-  \sum_{x:X} (f x = y) \ar[r]^\eta \ar[d]_\gamma & \modal \left(\sum_{x:X}  (f x = y)\right) \ar@{=}[dl]\\
-  \sum_{x:\modal X} (\modal f x = \eta_Y y) & }\end{mymath}
+  \sum_{x:X} (f(x) = y) \ar[r]^\eta \ar[d]_\gamma & \modal \left(\sum_{x:X}  (f(x) = y)\right) \ar@{=}[dl]\\
+  \sum_{x:\modal X} (\modal f(x) = \eta_Y(y)) & }\end{mymath}%
 where $\pi_1 \circ \gamma = \eta_X \circ \pi_1$, and
@@
 
-\kq{Is the following usefull?}
-In the same way, left exact modalities preserve homotopy types.
-\begin{prop}
-  Let $k \leq n$.
-  If $P:\Type_k$, then $\modal \widehat P : \Type_k$, where $\widehat P$
-  is $P$ seen as a $n$-type.
-\end{prop}
-\begin{proof}
-  An $n$-truncated type $P$ can equivalently be described as a type for
-  which the unique map to $\one$ is with $n$-truncated fibers. Thus, the
-  property is a direct corollary of
-  Proposition~\ref{sec:defin-basic-prop} and the fact that $\modal \one =
-  \one$.
-\end{proof}
+% \kq{Is the following usefull?}
+% In the same way, left exact modalities preserve homotopy types.
+% \begin{prop}
+%   Let $k \leq n$.
+%   If $P:\Type_k$, then $\modal \widehat P : \Type_k$, where $\widehat P$
+%   is $P$ seen as a $n$-type.
+% \end{prop}
+% \begin{proof}
+%   An $n$-truncated type $P$ can equivalently be described as a type for
+%   which the unique map to $\one$ is with $n$-truncated fibers. Thus, the
+%   property is a direct corollary of
+%   Proposition~\ref{sec:defin-basic-prop} and the fact that $\modal \one =
+%   \one$.
+% \end{proof}
 
@@
   path types are $n$-modal, ie
-  \begin{mymath}\mathrm{separated}(F) = \prod_{x,y:F}~\modal_n(x=y) = (x=y).\end{mymath}
+  \begin{mymath}\mathrm{separated}(F) = \prod_{x,y:F}~\modal_n(x=y) = (x=y).\end{mymath}%
 \end{prop}
@@
   \Type_n^\modal$, such that $\phi_1 \circ \pi_1 = \phi_2 \circ
-  \pi_1$ and let $x:E$. We show $\phi_1 x = \phi_2 x$ using
+  \pi_1$ and let $x:E$. We show $\phi_1(x) = \phi_2(x)$ using
   univalence.
@@
   As $\chi$ is dense, we have a term $m_x : \modal_n(\chi\, x)$.
-  But as $\phi_2 x$ is modal, we can obtain a term $h_x : \chi\,
+  But as $\phi_2(x)$ is modal, we can obtain a term $h_x : \chi\,
   x$. 
   As $\phi_1$ and $\phi_2$ are equal on $\sum_{e:E}\chi\, e$, we
-  have an arrow $\phi_1 x \to \phi_2 x$.
-  The same method leads to an arrow $\phi_2 x \to \phi_1 x$, and we
+  have an arrow $\phi_1(x) \to \phi_2(x)$.
+  The same method leads to an arrow $\phi_2 (x) \to \phi_1 (x)$, and we
   prove they are each other inverse.
@@
 
-  We define $g$ as $g(e) = \modal_n \left( \fib \phi e \right)$,
+  We define $g$ as $g(e) = \modal_n \left( \fib \phi (e) \right)$,
   where
   \begin{mymath}[0em]\phi : \sum_{b:\left(\sum_{e:E} (\chi\, e)\right)} (f\,
-    b) \to E\end{mymath}
+    b) \to E\end{mymath}%
   defined by $\phi(x) = (x_1)_1$.
-\kq{Is it necessary to state the lemma ?}
-  Using the following properties on fibers, we can prove that it
+% \kq{Is it necessary to state the lemma ?}
+Using a modal analogous of the characterization of fibers of composition, we can prove that it
 defines an equivalence.
-  \begin{lem}
-    Let $A,B,C:\Type_n$, $f:A\to B$ and $g:B\to C$.
-    Then
-    \begin{itemize}
-    \item If $g$ is an embedding, then 
-      \begin{mymath}\prod_{b:B}\fib f b = \fib {g\circ f} {g\, b}.\end{mymath}
-    \item If all fibers of $f$ and $g$ are $n$-truncated, then
-      \begin{mymath}\prod_{c:C}\modal_n(\fib {g \circ f} c) =
-      \modal_n \left(  
-        \sum_{w:\fib g c} \modal_n (\fib f {w_1})
-      \right).\end{mymath}
-    \end{itemize}
-  \end{lem}
-
+  % \begin{lem}
+  %   Let $A,B,C:\Type_n$, $f:A\to B$ and $g:B\to C$.
+  %   Then
+  %   \begin{itemize}
+  %   \item If $g$ is an embedding, then 
+  %     \begin{mymath}\prod_{b:B}\fib f b = \fib {g\circ f} {g\, b}.\end{mymath}
+  %   \item If all fibers of $f$ and $g$ are $n$-truncated, then
+  %     \begin{mymath}\prod_{c:C}\modal_n(\fib {g \circ f} c) =
+  %     \modal_n \left(  
+  %       \sum_{w:\fib g c} \modal_n (\fib f {w_1})
+  %     \right).\end{mymath}
+  %   \end{itemize}
+  % \end{lem}
 \end{proof}
-
+% \vspace{-1em}
 Another fundamental property on sheaves is that the type of (dependent)
@@
     \pi_1$.
-    Then for any $a:A$, $(\lambda x:E,~\phi_1\, x\,
-    a)$
-    and $(\lambda x:E,~\phi_2\, x\, a)$
+    Then for any $a:A$, $(\lambda x:E,~\phi_1(x, a))$
+    and $(\lambda x:E,~\phi_2(x,a))$
     coincide on $\sum_{e:E}(\chi\, e)$, and as $B\, a$ is separated,
@@
     $f:\sum_{e:E} (\chi\, e)\to \prod_{a:A} (B\,a)$. Let $a:A$ ; the
-    map $(\lambda x,~f\,x \,a)$ is valued in the sheaf $B\, a$, so it
-    can be extended to all $E$, allowing $f$ to be extended to all $E$.
+    map $(\lambda x,~f(x,a))$ is valued in the sheaf $B\, a$, so it
+    can be extended to all $E$, allowing $f$ to be extended to all
+    $E$. \qedhere
   \end{itemize}
@@
     \modal_{-1} P \ar@{->}^{\sim}[r] & \modal_n \widehat P 
-  }\end{mymath}
+  }\end{mymath}%
 \end{cond}
@@
   \separated T \ar[r]& \left( \Type_n^\modal \right)^T
-}, \end{mymath}
+}, \end{mymath}%
 where $\{\cdot\}_T$ is the singleton map $\lambda (t:T),~\lambda
@@
 \end{array}
-\end{mymath}
+\end{mymath}%
 %
@@
   \item $\mathring \alpha:~\forall a\, b:\|X\|_{n+1},~\modal (a=b) \to
-    \mathring t a = \mathring t b$
+    \mathring t(a) = \mathring t(b)$
   \item $\mathring \alpha_1:~\forall a:\|X\|_{n+1},~
-    \mathring \alpha\, a \, a\, (\eta_{a=a} 1) = 1$
+    \mathring \alpha(a , a, \eta_{a=a} 1) = 1$
   \end{itemize}
@@
   \begin{mymath}\xymatrix{\sum_{a,b:\|X\|_{n+1}} \modal (a=b) \dar[r]{4pt}^-{\pi_1}_-{\pi_2}
-  & \|X\|_{n+1}}\end{mymath}
+  & \|X\|_{n+1}}\end{mymath}%
   preserving $\eta_{a=a} 1$.
@@
   \begin{mymath}\xymatrix{\|X\|_{n+1} \ar[r] & \|\mathring T_{X}\|_{n+1} \ar[r] & \|\mathring
-  T_{\mathring T_X} \|_{n+1} \ar[r] & \cdots} \end{mymath}
+  T_{\mathring T_X} \|_{n+1} \ar[r] & \cdots} \end{mymath}%
 \end{defi}
@@
     \label{eq:Omono}\tag{$\star$}
-    \forall a,b:A,~f a = f b \simeq \modal (a=b),
+    \forall a,b:A,~f (a) = f (b) \simeq \modal (a=b),
     \vspace{-0.7em}
@@
   then
-  \begin{mymath}\forall a,b:\|T_f\|_{n+1},~|\tilde f|_{n+1} a = |\tilde f|_{n+1} b \simeq \modal (a=b).\end{mymath}
+  \begin{mymath}\forall a,b:\|T_f\|_{n+1},~|\tilde f|_{n+1} (a) = |\tilde f|_{n+1} (b) \simeq \modal (a=b).\end{mymath}%
 \end{lem}
@@
   By induction on truncation, we need to show that 
-  \begin{mymath}\forall a,b:T_f,~\tilde f |a|_{n+1} = \tilde f |b|_{n+1} \simeq
-  \modal (|a|_{n+1}=|b|_{n+1}).\end{mymath}
-  We use the encode-decode~\cite[Section 8.9]{hottbook} method to characterize $\tilde f |a|_{n+1}
+  \begin{mymath}\forall a,b:T_f,~\tilde f (|a|_{n+1}) = \tilde f (|b|_{n+1} )\simeq
+  \modal (|a|_{n+1}=|b|_{n+1}).\end{mymath}%
+  We use the encode-decode~\cite[Section 8.9]{hottbook} method to characterize $\tilde f (|a|_{n+1})
   = x$, and the result follows. We refer to the formalization for details.
@@
   satisfying~(\ref{eq:Omono}), 
-  \begin{mymath}T_f \simeq \mathring T_B.\end{mymath}
+  \begin{mymath}T_f \simeq \mathring T_B.\end{mymath}%
   For this, we use 
@@
   & Q &
-} \end{mymath} 
+} \end{mymath}% 
 % \begin{mymath}\xymatrix @C=4em{ 
@@
   E \ar@<-2pt>[ru]_{g} \ar@<2pt>[ru]^{f}&
-}\end{mymath}
+}\end{mymath}%
 % We can restrict the previous diagram to 
@@
 %   E \ar@<-2pt>[ru]_{\pi_1\circ g} \ar@<2pt>[ru]^{\pi_1\circ f}&
-% }\end{mymath}
+% }\end{mymath}%
 As $A$ is separated, $\pi_1\circ f = \pi_2 \circ g$.
@@
   \sum_{y:E} x = y \ar@<2pt>[ur]^{\pi_2\circ f\circ \pi_1} \ar@<-2pt>[ur]_{\pi_2\circ g\circ \pi_1}&
-}\end{mymath}
+}\end{mymath}%
 Using the separation property of $B\,x$, one can show that second
@@
             (\lambda t,~\modal_n (a=t)) = u\right\|.
-\end{mymath}
+\end{mymath}%
 
@@
       (\lambda t,~\modal_{-1} (a=t)) = u\right)
-    \end{mymath}
+    \end{mymath}%
     because the type inside the truncation is already in $\HProp$.
@@
     \begin{mymath}\phi t = (\lambda t',\, \modal_{-1}(t=t') ; \eta
-    (t;\mathrm{idpath}))\end{mymath}
+    (t;\mathrm{idpath}))\end{mymath}%
     and $\psi : \modal_0T\to \modal_{-1} T$ by obtaining the
@@
     a modal proposition), and letting
-    \begin{mymath}\psi (u;x) = \eta_T a.\end{mymath}
+    \begin{mymath}\psi (u;x) = \eta_T a.\end{mymath}%
     These two maps form an equivalence (the section and retraction are
@@
         (\lambda t,~\modal_{n} (a=t)) = u\right\|.
-    \end{mymath}
+    \end{mymath}%
     It remains to prove that for every $(u,x)$ inhabiting the
@@
     every 
-    $t:T$, $u\, t = \modal_n(a=t) : \Type_n$.
+    $t:T$, $u\, t = \modal_n(a=t) : \Type_n$. \qedhere
   \end{itemize}
 \end{proof}
-
 This proves in particular that $\modal_{n+1}$ is compatible with
@@
 The last step is the left-exactness of $\modal_{n+1}$. Let $T$ be in
-$\Type_{n+1}$ such that $\Contr(\modal_{n+1} T)$.  Thanks to the just
+$\Type_{n+1}$ such that $\modal_{n+1} T$ is contractible.  Thanks to the just
 shown compatibility between $\modal_{n+1}$ and $\modal_n$ for
-$\Type_n$, left exactness means that for any $x,y : T$,
+$\Type_n$, left exactness means that for any $x,y: T$,
 $\modal_n(x=y)$ is contractible.
 
-Using a proof by univalence as we have done for proving $\modal_n (a=b) = (\mu_T a =
-\mu_T b)$ in Proposition~\ref{lem:sepiscolim}, we can show that:
+Using a proof by univalence as we have done for proving $\modal_n (a=b) = (\mu_T(a) =
+\mu_T (b))$ in Proposition~\ref{lem:sepiscolim}, we can show that:
 \begin{prop}
-  For all $a,b:T$, $\modal_n(a=b) = (\nu_T a = \nu_T b)$.
+  For all $a,b:T$, $\modal_n(a=b) \!=\! (\nu_T a = \nu_T b)$.
 \end{prop}
@@
                                     true} \end{array}\right.~:~\HProp
-                              \to \nat\end{mymath}
+                              \to \nat\end{mymath}%
 will not reduce to a numeral when applied to a undecidable
@@
 %
-After reviewing the content and some statistics about the
-formalization in Section~\ref{sec:cont-form}, we present the
-limitations of our formalization in Section~\ref{sec:limit-form}, in
-particular the issues relative to universe polymorphism.
+% After reviewing the content and some statistics about the
+% formalization in Section~\ref{sec:cont-form}, we present the
+% limitations of our formalization in Section~\ref{sec:limit-form}, in
+% particular the issues relative to universe polymorphism.
 
@@
 
-Universes are used in type theory to ensure consistency by checking
-that definitions are well-stratified according to a certain hierarchy.
-%
-Universe polymorphism~\cite{sozeau2014universe} supports generic
-definitions over universes, reusable at different levels.
-%
-Although the presence of universe polymorphism is mandatory for our
-formalization, its implementation is still too rigid to allow a
-complete formalization of our work for the following reasons.
-
-%
-If Coq handles cumulativity on $\Type$ natively, it is not
-the case for the $\Sigma$-type $\Type_n$, which require propositional
-resizing. 
-%
-This issue could be solved by adding an axiom of cumulativity
-for $\Type_n$ with an explicit management of universes. 
-%
-But as it would not have any computational content, such a solution
-would really complicate the proofs as the axiom would appear
-everywhere cumulativity is needed and it would need explicit
-annotations for universe levels everywhere in the formalization.
-%
-% Note that we also could have resolved the issue by giving explicitely
-% the universe levels we wanted, like in the Coq/HoTT library\cite{hottlib}.
-
-Another issue with universe polymorphism lies in the management of
+% Universes are used in type theory to ensure consistency by checking
+% that definitions are well-stratified according to a certain hierarchy.
+% %
+% Universe polymorphism~\cite{sozeau2014universe} supports generic
+% definitions over universes, reusable at different levels.
+% %
+% Although the presence of universe polymorphism is mandatory for our
+% formalization, its implementation is still too rigid to allow a
+% complete formalization of our work for the following reasons.
+
+% %
+% If Coq handles cumulativity on $\Type$ natively, it is not
+% the case for the $\Sigma$-type $\Type_n$, which require propositional
+% resizing. 
+% %
+% This issue could be solved by adding an axiom of cumulativity
+% for $\Type_n$ with an explicit management of universes. 
+% %
+% But as it would not have any computational content, such a solution
+% would really complicate the proofs as the axiom would appear
+% everywhere cumulativity is needed and it would need explicit
+% annotations for universe levels everywhere in the formalization.
+% %
+% % Note that we also could have resolved the issue by giving explicitely
+% % the universe levels we wanted, like in the Coq/HoTT library\cite{hottlib}.
+
+One issue with universe polymorphism lies in the management of
 recursive definitions. Indeed, the following recursive definition of
@@
 moment.
-
+%
 Note that the same increasing in the universe levels occurs in the
@@
   one. But some types in homotopy type theory are not
-  truncated\cite[Example 8.8.6]{hottbook}.
-
+  truncated\cite[Example 8.8.6]{hottbook}.q
   % At the moment, sheafification only works with truncated

### 2015-12-16 paper/paper_acm.tex
reduce again

@@
 To extend Lawvere-Tierney sheafification to homotopy type theory, the
-first thing to understand is that the construction can not be done in one
+first thing to understand is that the construction cannot be done in one
 single step anymore. It must rather be performed by induction on the
@@
 sheafification is defined for $\Type_{0}$, given a topology on
-$\Type_{-1}$. This part corresponds to sheafification for toposes. Then,
+$\Type_{-1}$. This part corresponds to sheafification for topoï. Then,
 assuming that the sheafification has been constructed above level $n$,
@@
 to construction in the Grothendieck setting, where the
-sheafification is obtained by applying the plus-construction $n+1$
+sheafification is obtained by applying the plus-construction $n\!+\!1$
 times to
@@
 have a finite homotopy level, hence this construction cannot reach
-every type.
+every type\looseness=-1.
 
@@
 theory uses the property that epimorphisms are coequalizers of their
-kernel pairs. This property does not hold anymore in higher toposes, but
-using Giraud-Rezk-Lurie axioms of $\infty$-toposes, the same property
+kernel pairs. This property does not hold anymore in higher topoï, but
+using Giraud-Rezk-Lurie axioms of $\infty$-topoï, the same property
 holds by replacing kernel pairs by \v{C}ech nerves in the statement.
@@
 %
-Finally, Section~\ref{sec:future-works} concludes on future works.
+Finally, Section~\ref{sec:future-works} concludes on future works\looseness=-1.
 
@@
 % And Section~\ref{sec:giraud-ax} defines an analogue of
-% Giraud-Rezk-Lurie axioms on higher toposes that we need in the
+% Giraud-Rezk-Lurie axioms on higher topoï that we need in the
 % definition of sheafification.
@@
 notation. In the rest of the paper, $\Type$ must be seen in an
-universe-polymorphic way.
+universe-polymorphic way\looseness=-1.
 
@@
 As noticed in \cite{hottbook}, left exact modalities correspond
-to sub-$(\infty,1)$-toposes. Indeed, modal types defines a new type
+to sub-$(\infty,1)$-topoï. Indeed, modal types defines a new type
 theory as modality preserves all type former, such as dependent
@@
   and only if $\modal \zero$ can not be inhabited in the initial type
-  theory. In that case, the modality is said to be consistent.
+  theory. In that case, the modality is said to be consistent\looseness=-1.
 \end{prop}
@@
 definition of sheaves for $(n+1)$-types and basic properties of
-sheaves that will be useful in the next section.
+sheaves that will be useful in the next section\looseness=-1.
 
@@
 % itself is not the constant function equal to $\one$, as it is the case
-% for toposes, but rather the map 
+% for topoï, but rather the map 
 % %
@@
     $\Phi_E^\chi : (E \to F) \to (\sum_{e:E} \chi\, e \to F) $
-    as the map sending an  arrow $f:E\to F$ to its restriction $f \circ \pi_1$.
+    as the map sending an  arrow $f:E\to F$ to its restriction $f \circ \pi_1$\looseness=-1.
   % \item if $A \to E$ with $n$-truncated homotopy fibers, $\Phi_E^\chi$ is the
@@
 Lawvere-Tierney topologies can still be shown to subsume Grothendieck
-toplogies for $(\infty,1)$-toposes, which is beyond the scope of this paper).
+toplogies for $(\infty,1)$-topoï, which is beyond the scope of this paper).
 %
@@
 This construction may be seen as the first induction step between
-$\HProp$ and $\Type_0$.
+$\HProp$ and $\Type_0$.\looseness=-1
 
@@
 \begin{prop}
-  For any $T:\Type_{n+1}$, $\separated T$ is separated.  
+  For any $T\!:\!\Type_{n+1}$, $\separated T$ is separated.  
 \end{prop}
@@
 are equal. The complete proof can be found in the formalization.
-This proves the following proposition
+This proves the following proposition\looseness=-1
 \begin{prop}\label{prop:sep-mod}
@@
     witness $a:T$ (which is possible because we are trying to inhabit
-    a modal proposition), and letting
-    \begin{mymath}\psi (u;x) = \eta_T a.\end{mymath}%
+    a modal proposition), and letting $\psi (u;x) = \eta_T a$.
     These two maps form an equivalence (the section and retraction are
@@
 $\Type_n$, left exactness means that for any $x,y: T$,
-$\modal_n(x=y)$ is contractible.
+$\modal_n(x=y)$ is contractible.\looseness=-1
 
@@
 that homotopy type theory provides a convenient tool to formalize some
-part of the theory of higher toposes. As a matter of comparison, this
-\TeX\xspace file contains approximatively 1800 lines.
+part of the theory of higher topoï. As a matter of comparison, this
+\TeX\xspace file contains approximatively 1800 lines\looseness=-1.
 
@@
   one. But some types in homotopy type theory are not
-  truncated\cite[Example 8.8.6]{hottbook}.q
+  truncated~\cite[Example 8.8.6]{hottbook}.
   % At the moment, sheafification only works with truncated

### 2015-12-17 paper/paper_acm.tex
reduce again

@@
 
-\category{CR-number}{subcategory}{third-level}
+\category{F}{8}{3}
 
@@
 way, we can add classical reasoning in the type theory without
-collapsing the theory to mere propositions.
+collapsing it to mere propositions.
 
@@
 
-As $\modal_{n+1} T$ is contractible, any path space of $\modal_{n+1} T$ is
+As $\modal_{n+1} T$ is contractible, path spaces of $\modal_{n+1} T$ are
 contractible, in particular $(\nu_T a=\nu_T b)$, which proves left
@@
 %
+\vspace{-0.05em}
 \[ \begin{array}{l}
@@
     \modal_{-1\phantom{n}}(T) \defeq\lnot\lnot T \\
+
       \displaystyle{\modal_{n+1}(T)} \defeq  
@@
     \end{array}
+\vspace{-0.3em}
 \]
@@
 
-\section*{Acknowledgment}
-
-We would like to thank all maintainers of the Coq/HoTT library, who
-have provided a very solid base for our formalization. 
-We also would like
-to thank Bas Spitters for helpful discussions on the subject.
+% \section*{Acknowledgment}
+
+% We would like to thank all maintainers of the Coq/HoTT library, who
+% have provided a very solid base for our formalization. 
+% We also would like
+% to thank Bas Spitters for helpful discussions on the subject.
 

### 2016-01-06 paper/paper_acm.tex
reduce again
last version

@@
 
-\documentclass[preprint,10pt,numbers]{sigplanconf}
+\documentclass[preprint,9pt,numbers]{sigplanconf}
 
@@
 \usepackage{wasysym}
-\usepackage{microtype}
+%usepackage{microtype}
 % \usepackage[subtle]{savetrees}
@@
 \DeclareMathOperator{\IsContr}{IsContr}
+\DeclareMathOperator{\IsEquiv}{IsEquiv}
 \DeclareMathOperator{\precompose}{\mathrm{precompose}}
@@
 
-\newenvironment{mymath}[1][-0.7em]{%
+\newenvironment{mymath}[1][-0em]{%
   \newcommand\mymathaux{\vspace{#1}}%
@@
 
-\category{F}{8}{3}
+\category{F}{4}{1}
 
@@
 
-Even if similar questions have been considered around the Curry-Howard
-isomorphism, there is no canonical way to extend a programming
-language such as type theory with new logical or computational
-principle while keeping consistency automatically.
-%
-For instance, much efforts have been done to provide a computational
-content to the law of excluded middle in order to define a
-constructive version of classical logic. This has lead to various
-calculi, with most notably the $\lambda \mu$-calculus of
-Parigot~\cite{parigot1993classical}, but this line of work has not
-appeared to be fruitful to define a new version of type theory with
-classical principles.
-%
-Other works have tried to extend continuation-passing-style (CPS)
-transformation to type theory, but they have been faced with the
-difficulty that the CPS transformation is incompatible with (full) dependent
-sums~\cite{barthe2002cps}, which puts emphasis on the tedious link
-between the axiom of choice and the law of excluded middle in type theory\looseness=-1.
-
-Nevertheless the axiom of choice has shown to be realizable by
-computational meaning in a classical setting by techniques turning
-around the notion of (modified) bar induction
-\cite{berardi1998computational}, Krivine's
-realizability~\cite{krivine2003dependent} and even more recently with
-restriction on elimination of dependent sums and lazy
-evaluation~\cite{herbelin2012constructive}.
-%
-
-All those works---although they provide very nice and
-sophisticated theoretical results---are not as general as the process
-of sheafification in topos theory and are not amenable to a
-direct integration into a proof assistant based on type theory such as
-Coq~\cite{coq:refman:8.4} or Agda~\cite{norell2007towards}.  
-%
-However, type theory is known to be quite closed to topos theory so
+As type theory is known to be quite close to topos theory,
 one could wonder why similar techniques have not been developed in the
@@
 two issues that this paper addresses: (i) how to lift the notion of
-Lawvere-Tierney sheafification to higher topos theory and (ii) is it
+Lawvere-Tierney sheafification to a higher dimensional setting and (ii) is it
 possible to formalize this new definition in homotopy type theory.
@@
 %
-\vspace{-0.7em}
+%\vspace{-0.7em}
 \begin{equation}\label{equ:sheaf_def}
 \Hom{\E}(E,F) \rightarrow \Hom{\E}(A,F).
-\vspace{-0.7em}
+%\vspace{-0.7em}
 \end{equation}%
@@
 \begin{mymath}
-\Omega_j = \{ P \in \E \ | \ j P  = P \}.
+\Omega_j = \{ P \in \Omega \ | \ j P  = P \}.
 \end{mymath}%
@@
 Furthermore, one can define a left adjoint to the inclusion, the
-sheafification functor
+LT-sheafification functor
 %
@@
 which exhibits $\Sh{\E}$ as a reflective
-subcategory of $\E$ (which is a particular case of localization). This
+subcategory (or reflective localization) of $\E$. This
 means that logical principles valid in $\E$ are still valid in
@@
 The main contribution of this article is the development of a
-computer-checked definition of Lawvere-Tierney sheafification in the
+computer-checked definition of Lawvere-Tierney sheafification
+(LT-sheafification) in the
 setting of homotopy type theory.
@@
 level of homotopy types. More precisely, the first layer of
-sheafification is defined for $\Type_{0}$, given a topology on
-$\Type_{-1}$. This part corresponds to sheafification for topoï. Then,
-assuming that the sheafification has been constructed above level $n$,
-one can define the sheafification for $\Type_{n+1}$.
-%
-This inductive process in the definition of sheafification is similar
+LT-sheafification is defined for $\Type_{0}$, given a topology on
+$\Type_{-1}$. This part corresponds to LT-sheafification for topoï. Then,
+assuming that the LT-sheafification has been constructed above level $n$,
+one can define the LT-sheafification for $\Type_{n+1}$.
+%
+This inductive process in the definition of LT-sheafification is similar
 to construction in the Grothendieck setting, where the
@@
 have a finite homotopy level, hence this construction cannot reach
-every type\looseness=-1.
+every type.
 
@@
 Lawvere-Tierney topology can thus be seen as a left-exact modality on
-$\Type_{-1}$ and sheafification as an inductive process that extends
+$\Type_{-1}$ and LT-sheafification as an inductive process that extends
 it as a left-exact modality on $\Type_{n}$ for any $n$.
 
-Compare to the definition of sheafification in topos theory, the main
+Compare to the definition of LT-sheafification in topos theory, the main
 change comes from the fact that in homotopy type theory, there is not
@@
 For instance, the notion of separated objects (which is a technical
-notion used in the construction of the sheafification functor), \ie
+notion used in the construction of the LT-sheafification functor), \ie
 objects for which the morphism
@@
 
-Another technical difficulty is the fact that sheafification in topos
+Another technical difficulty is the fact that LT-sheafification in topos
 theory uses the property that epimorphisms are coequalizers of their
@@
 
+\subsection{Related Works}
+\label{sec:related-works}
+
+Even if similar questions have been considered around the Curry-Howard
+isomorphism, there is no canonical way to extend a programming
+language such as type theory with new logical or computational
+principle while keeping consistency automatically.
+%
+For instance, much efforts have been done to provide a computational
+content to the law of excluded middle in order to define a
+constructive version of classical logic. This has lead to various
+calculi, with most notably the $\lambda \mu$-calculus of
+Parigot~\cite{parigot1993classical}, but this line of work has not
+appeared to be fruitful to define a new version of type theory with
+classical principles.
+%
+Other works have tried to extend continuation-passing-style (CPS)
+transformation to type theory, but they have been faced with the
+difficulty that the CPS transformation is incompatible with (full) dependent
+sums~\cite{barthe2002cps}, which puts emphasis on the tedious link
+between the axiom of choice and the law of excluded middle in type theory.
+
+Nevertheless the axiom of choice has shown to be realizable by
+computational meaning in a classical setting by techniques turning
+around the notion of (modified) bar induction
+\cite{berardi1998computational}, Krivine's
+realizability~\cite{krivine2003dependent} and even more recently with
+restriction on elimination of dependent sums and lazy
+evaluation~\cite{herbelin2012constructive}.
+%
+
+All those works---although they provide very nice and
+sophisticated theoretical results---are not as general as the process
+of LT-sheafification in topos theory and are not amenable to a
+direct integration into a proof assistant based on type theory such as
+Coq~\cite{coq:refman:8.4} or Agda~\cite{norell2007towards}.
+
 \subsection{Plan of the Paper}
@@
 Section~\ref{sec:sheafification} explains the construction of the
-sheafification functor.
+LT-sheafification functor.
 %
@@
 %
-Finally, Section~\ref{sec:future-works} concludes on future works\looseness=-1.
+Finally, Section~\ref{sec:future-works} concludes on future works.
 
@@
 theory that are central in our formalization but not specific to
-sheafification. 
+LT-sheafification. 
 % %
@@
 sum, and $\pi_1, \pi_2$ for the first and second projection of a
-dependent pair (noted $(a;b)$). We use informal mathematical language
+dependent pair (noted $(a;b)$). The identity path will be denoted
+$1$. We use informal mathematical language
 instead of type theory whenever it is possible, to ease the reading
@@
 notation. In the rest of the paper, $\Type$ must be seen in an
-universe-polymorphic way\looseness=-1.
+universe-polymorphic way.
 
@@
 because Lawvere-Tierney topologies correspond exactly to left exact
-modalities at level $-1$ and sheafification is described as an inductive
+modalities at level $-1$ and LT-sheafification is described as an inductive
 process that produces a left exact modality at level $n+1$ from a
@@
 (from \cite[Corollary 7.7.2]{hottbook})
-\item For any $A:\Type_n^\modal$ and $x,y:X$, $P(x=y)$.
+\item For any $A:\Type_n^\modal$ and $x,y:A$, $P(x=y)$.
 \item 
@@
 
-As the definition of sheafification is based on subobject classifiers
+As the definition of LT-sheafification is based on subobject classifiers
 and thus on homotopy fibers, we need to take a closer look at the way left
@@
 
-Furthermore, when $A$ and $B$ are modal and $f : A \to B$, the mere
-proposition of $f$ being an equivalence is also modal. This means that
-univalence also holds in $\Type^\modal_p$, so it induces a good notion
-of homotopy type theory.
-
-Every left exact modality induces a new homotopy type theory, but with more
+% Furthermore, when $A$ and $B$ are modal and $f : A \to B$, the mere
+% proposition of $f$ being an equivalence is also modal. This means that
+% univalence also holds in $\Type^\modal_n$, so it induces a good notion
+% of homotopy type theory.
+
+Every left exact modality induces a new type theory, but with
+more
 properties, that can make it inconsistent. We now provide a simple
@@
   and only if $\modal \zero$ can not be inhabited in the initial type
-  theory. In that case, the modality is said to be consistent\looseness=-1.
+  theory. In that case, the modality is said to be consistent.
 \end{prop}
@@
   By condition (iv) of Definition~\ref{sec:defin-basic-prop-1},
-  $\modal \zero$ is an initial object of $\Type^\modal_p$, and thus
+  $\modal \zero$ is an initial object of $\Type^\modal_n$, and thus
   corresponds to false for modal mere proposition.
   % 
-  As $\modal \one = \one$, $\Type^\modal_p$ is consistent when
+  As $\modal \one = \one$, $\Type^\modal_n$ is consistent when
   $\modal \zero \neq \one$, that is when there is no proof of
@@
 
-Another example of a left exact modality for every level $p$ is the
+Another example of a left exact modality for every level $n$ is the
 double negation modality,
@@
 In this paper, we use the double negation modality at level $-1$ only,
-and then extend it to higher homotopy types using sheafification. This
+and then extend it to higher homotopy types using LT-sheafification. This
 way, we can add classical reasoning in the type theory without
@@
 %
-and sheafification corresponds to lifting this modality to a left
+and LT-sheafification corresponds to lifting this modality to a left
 exact modality at level $0$.
@@
 any left exact modality at level $-1$ to a left exact modality at level
-$n$ using sheafification, by induction on $n$.
+$n$ using LT-sheafification, by induction on $n$.
 
@@
 definition of sheaves for $(n+1)$-types and basic properties of
-sheaves that will be useful in the next section\looseness=-1.
+sheaves that will be useful in the next section.
 
@@
     dense in $E$} if its closure is $E$, \ie
-  $\modal_n \circ \chi = \lambda e. 1$ 
+  $\modal_n \circ \chi = \lambda e, \one$ 
 \end{defi}
@@
     $\Phi_E^\chi : (E \to F) \to (\sum_{e:E} \chi\, e \to F) $
-    as the map sending an  arrow $f:E\to F$ to its restriction $f \circ \pi_1$\looseness=-1.
+    as the map sending an  arrow $f:E\to F$ to its restriction $f \circ \pi_1$.
   % \item if $A \to E$ with $n$-truncated homotopy fibers, $\Phi_E^\chi$ is the
@@
   path types are $n$-modal, ie
-  \begin{mymath}\mathrm{separated}(F) = \prod_{x,y:F}~\modal_n(x=y) = (x=y).\end{mymath}%
+  \begin{mymath}[-0.3em]\prod_{x,y:F}~\modal_n(x=y) = (x=y).\end{mymath}%
 \end{prop}
@@
 %
-For higher topos theory, in the Grotehndieck setting, sheaves seems
+For higher topos theory, in the Grothendieck setting, sheaves seems
 to correspond to the second part of our definition only (assuming that
@@
 
-A fundamental property in the construction of the sheafification
+A fundamental property in the construction of the LT-sheafification
 functor is the fact that the type of modal types at level $n$ defines
@@
 
-  We define $g$ as $g(e) = \modal_n \left( \fib \phi (e) \right)$,
+  We define $g$ as $g(e) = \modal_n \left( \fib \phi {e} \right)$,
   where
@@
 
-Note that once the first modality $\modal_{-1}$ is defined, its
+\label{sec:sheaf-left-exact}Note that once the first modality $\modal_{-1}$ is defined, its
 extension to all $\Type_n$ is automatic. This means that the new
 logical principles that can be added by the sheaf construction must be
-added in the first modality already. Sheafification is just a way
+added in the first modality already. LT-Sheafification is just a way
 to propagate those new principles to types that are not mere
@@
 
-The sheafification in topos theory described in
+The LT-sheafification in topos theory described in
 Section~\ref{sec:lawv-tiern-sheaf} works in the same way. 
@@
 This construction may be seen as the first induction step between
-$\HProp$ and $\Type_0$.\looseness=-1
+$\HProp$ and $\Type_0$.
 
@@
 
-For a given induction step, the sheafification is defined in two
+For a given induction step, the LT-sheafification is defined in two
 steps ; the first one construct, for any $T:\Type_{n+1}$ a separated
@@
 \begin{mymath}\xymatrix{
-  T \ar[r]^{\{\cdot\}_T} \ar[d]_{\mu_T} & \Type_n^T \ar[d]^{\modal_n^T} \\
+    T \ar[r]^{\{\cdot\}_T} \ar[d]_{\mu_T} & \left(\Type_n\right)^T \ar[d]^{\modal_n^T} \\
   \separated T \ar[r]& \left( \Type_n^\modal \right)^T
@@
 
-Again, this lemma is an adaptation of the sheafification process in
+Again, this lemma is an adaptation of the LT-sheafification process in
 topos theory, where the kernel pair of $\mu_T$
@@
   Let $A,S:\Type_{n+1}$, $S$ separated, and $f:A \to S$. Then if 
-  \vspace{-0.7em}
+  % \vspace{-0.7em}
   \begin{equation}
@@
     \forall a,b:A,~f (a) = f (b) \simeq \modal (a=b),
-    \vspace{-0.7em}
+    % \vspace{-0.7em}
   \end{equation}
@@
 are equal. The complete proof can be found in the formalization.
-This proves the following proposition\looseness=-1
+This proves the following proposition
 \begin{prop}\label{prop:sep-mod}
@@
 As this modality is just a step in the construction, we do not need to
-show that it is left exact, we will only do it for the sheafification
+show that it is left exact, we will only do it for the LT-sheafification
 modality.
@@
     \begin{mymath}\phi t = (\lambda t',\, \modal_{-1}(t=t') ; \eta
-    (t;\mathrm{idpath}))\end{mymath}%
+    (t;1))\end{mymath}%
     and $\psi : \modal_0T\to \modal_{-1} T$ by obtaining the
@@
 $\Type_n$, left exactness means that for any $x,y: T$,
-$\modal_n(x=y)$ is contractible.\looseness=-1
+$\modal_n(x=y)$ is contractible.
 
@@
 \begin{prop}
-  $(\modal_{\lnot\lnot})_n$, the modality obtained by sheafification
+  $(\modal_{\lnot\lnot})_n$, the modality obtained by LT-sheafification
   of the double negation modality,
@@
 ones constructed without using excluded middle.
+
+By proposition~\ref{prop:consistent}, these left-exact modalities
+induces a consistent type theory. Furthermore, the univalence remains
+true in this new type theory in the following sense:
+\begin{prop}
+  Let $n$ be a given truncation index, and $\modal$ the modality
+  associated to $n+1$ as defined in lemma~\ref{lem:type}. Then, for
+  any type $A,B:\Type_n^\modal$, the canonical arrow
+  $$\varphi~:~A = B \to A\simeq B$$
+  is an equivalence.
+\end{prop}
+\begin{proof}
+  The first thing to notice is that, if $X$ and $Y$ are modal, and
+  $f:X \to Y$, then the mere proposition $\IsEquiv f$ is also modal.
+  Therefore, it suffices to show that both $A=B$ and $A\simeq B$ are
+  modal. As $\Type_n^\modal$ is modal, $A=B$ is modal. 
+  Moreover, $(A\simeq B) \simeq \sum_{f:A\to B} \IsEquiv
+  f$. Therefore, as $A$ and $B$ are modal, $A\simeq B$ is too.
+
+  Hence, $\IsEquiv \varphi$ is modal, and then is true in the new type
+  theory, as it was in the original one.
+\end{proof}
 
@@
 part of the theory of higher topoï. As a matter of comparison, this
-\TeX\xspace file contains approximatively 1800 lines\looseness=-1.
+\TeX\xspace file contains approximatively 1800 lines.
 
@@
 recursive definitions. Indeed, the following recursive definition of
-sheafification
+LT-sheafification
 %
@@
 Thus, the induction step presented in this paper has been formalized,
-but the complete recursive sheafification can not be defined for the
+but the complete recursive LT-sheafification can not be defined for the
 moment.
@@
 %
-Using sheafification, it is possible to add new logical principles to
+Using LT-sheafification, it is possible to add new logical principles to
 homotopy type theory, such as the law of excluded middle, and this

### 2016-01-06 paper/paper_acm.tex
rework introduction

@@
 use a generalized notion of sheafification, called Lawvere-Tierney
-sheafification~\cite{tierney1972sheaf,maclanemoerdijk}, but this construction has not
+sheafification~\cite{tierney1972sheaf,maclanemoerdijk}
+(LT-sheafification), but this construction has not
 been considered yet in the setting of higher topos theory. This rises
 two issues that this paper addresses: (i) how to lift the notion of
-Lawvere-Tierney sheafification to a higher dimensional setting and (ii) is it
+LT-sheafification to a higher dimensional setting and (ii) is it
 possible to formalize this new definition in homotopy type theory.
+
+\paragraph*{Contributions.}
+
+The main contribution of this article is the development of a
+computer-checked definition of Lawvere-Tierney sheafification in the
+setting of homotopy type theory.
+%
+This provides a way to give a meaning to the propositional law of
+excluded middle, and more generally, combined with the previous work on
+forcing in type theory~\cite{jaber2012extending}, it allows to reuse
+results derived for topos theory, such as independence results.
+
+Technically, LT-sheafification in homotopy type theory allows to
+construct a model of homotopy type theory (restricted to finite homotopy
+types) with new logical principles (such as EM) from a model of
+homotopy type theory.
+
+That is, LT-sheafification can be seen as a model transformation that
+allows to justify new axioms in homotopy type theory. 
+%
+But as this transformation is internalized inside homotopy type theory
+itself, it also provides a computational content to the added axiom,
+making them more amenable to be used in practice inside a proof
+assistant.
 
@@
 \label{sec:lawv-tiern-sheaf}
+
+In order to understand the construction of LT-sheafification in homotopy
+type theory, we first briefly recall its corresponding construction in
+topos theory.
 
@@
 \subsection{Overview of the Result}
-
-The main contribution of this article is the development of a
-computer-checked definition of Lawvere-Tierney sheafification
-(LT-sheafification) in the
-setting of homotopy type theory.
-%
-This provides a way to give a meaning to the propositional law of
-excluded middle, and more generally, combined with the previous work on
-forcing in type theory~\cite{jaber2012extending}, it allows to reuse
-results derived for topos theory, such as independence results.
 
@@
 
-Even if similar questions have been considered around the Curry-Howard
-isomorphism, there is no canonical way to extend a programming
-language such as type theory with new logical or computational
-principle while keeping consistency automatically.
+Similar questions have been considered around the Curry-Howard
+isomorphism, to extend a programming language closed to type theory
+with new logical or computational principle while keeping consistency
+automatically.
 %
@@
 %
-
-All those works---although they provide very nice and
-sophisticated theoretical results---are not as general as the process
-of LT-sheafification in topos theory and are not amenable to a
-direct integration into a proof assistant based on type theory such as
-Coq~\cite{coq:refman:8.4} or Agda~\cite{norell2007towards}.
+All those works---although they provide very nice and sophisticated
+theoretical results---are not amenable to a direct integration into a
+proof assistant based on type theory such as Coq~\cite{coq:refman:8.4}
+or Agda~\cite{norell2007towards}.
+
+The work on forcing in type theory~\cite{jaber2012extending} lies in
+the same line as this paper as it also gives a computational meaning
+to a type theory enriched with new logical or computational principle.
+%
+Actually, forcing is entirely complementary to LT-sheafification as
+forcing corresponds to the presheaf construction while
+LT-sheafification corresponds to the topological transformation that
+allows to go from the presheaf to the sheaf construction.
 

### 2016-01-06 paper/paper_acm.tex
minor change

@@
 between the axiom of choice and the law of excluded middle in type theory.
-
+%
 Nevertheless the axiom of choice has shown to be realizable by
@@
 evaluation~\cite{herbelin2012constructive}.
-%
+
 All those works---although they provide very nice and sophisticated
@@
 
-The work on forcing in type theory~\cite{jaber2012extending} lies in
-the same line as this paper as it also gives a computational meaning
-to a type theory enriched with new logical or computational principle.
+The work on forcing in type theory~\cite{jaber2012extending} also
+gives a computational meaning to a type theory enriched with new
+logical or computational principle, but for other kinds of principles
+as the ones interpreted by LT-sheafification.
 %

### 2016-01-06 paper/paper_acm.tex
refactoring of introduction
rework plan of the paper

@@
 
-Nevertheless, the adaptation of the sheafification in higher topos
-theory to homotopy type theory is not straightforward because the
-construction in higher toposes is restricted to the initial Grothendieck
-setting which is still very topologically oriented, and not very
-amenable to formalization in type theory. It seems more promising to
-use a generalized notion of sheafification, called Lawvere-Tierney
-sheafification~\cite{tierney1972sheaf,maclanemoerdijk}
-(LT-sheafification), but this construction has not
-been considered yet in the setting of higher topos theory. This rises
-two issues that this paper addresses: (i) how to lift the notion of
-LT-sheafification to a higher dimensional setting and (ii) is it
-possible to formalize this new definition in homotopy type theory.
+Nevertheless, the adaptation of the sheafification process in higher
+topos theory to homotopy type theory is not straightforward because
+the construction in higher toposes is restricted to the initial
+Grothendieck setting which is still very topologically oriented, and
+not very amenable to formalization in type theory. It seems more
+promising to use a generalized notion of sheafification, called
+Lawvere-Tierney sheafification~\cite{tierney1972sheaf,maclanemoerdijk}
+(LT-sheafification), but this construction has not been considered yet
+in the setting of higher topos theory. This rises two issues that this
+paper addresses: (i) how to lift the notion of LT-sheafification to a
+higher dimensional setting and (ii) is it possible to formalize this
+new definition in homotopy type theory.
 
@@
 
+\subsection{Related Works}
+\label{sec:related-works}
+
+Similar questions have been considered around the Curry-Howard
+isomorphism, to extend a programming language closed to type theory
+with new logical or computational principle while keeping consistency
+automatically.
+%
+For instance, much efforts have been done to provide a computational
+content to the law of excluded middle in order to define a
+constructive version of classical logic. This has lead to various
+calculi, with most notably the $\lambda \mu$-calculus of
+Parigot~\cite{parigot1993classical}, but this line of work has not
+appeared to be fruitful to define a new version of type theory with
+classical principles.
+%
+Other works have tried to extend continuation-passing-style (CPS)
+transformation to type theory, but they have been faced with the
+difficulty that the CPS transformation is incompatible with (full) dependent
+sums~\cite{barthe2002cps}, which puts emphasis on the tedious link
+between the axiom of choice and the law of excluded middle in type theory.
+%
+Nevertheless the axiom of choice has shown to be realizable by
+computational meaning in a classical setting by techniques turning
+around the notion of (modified) bar induction
+\cite{berardi1998computational}, Krivine's
+realizability~\cite{krivine2003dependent} and even more recently with
+restriction on elimination of dependent sums and lazy
+evaluation~\cite{herbelin2012constructive}.
+
+All those works---although they provide very nice and sophisticated
+theoretical results---are not amenable to a direct integration into a
+proof assistant based on type theory such as Coq~\cite{coq:refman:8.4}
+or Agda~\cite{norell2007towards}.
+
+The work on forcing in type theory~\cite{jaber2012extending} also
+gives a computational meaning to a type theory enriched with new
+logical or computational principle, but for other kinds of principles
+as the ones interpreted by LT-sheafification.
+%
+Actually, forcing is entirely complementary to LT-sheafification as
+forcing corresponds to the presheaf construction while
+LT-sheafification corresponds to the topological transformation that
+allows to go from the presheaf construction to the sheaf construction.
+
+\subsection{Plan of the Paper}
+
+Section~\ref{sec:lawv-tiern-nutshell} presents an overview of the
+construction.
+% 
+Section~\ref{sec:hott} presents the foundations of homotopy type theory
+and presents a analogous of~\cite[Corollary 6.2.3.5]{lurie}. 
+%
+Section~\ref{sec:lexmod} introduces the notion of left-exact stratified
+modalities, on which this paper is based.
+%
+Section~\ref{sec:sheaves} defines Lawvere-Tierney sheaves in homotopy
+type theory.
+%
+Section~\ref{sec:sheafification} explains the construction of the
+LT-sheafification functor.
+%
+The definitions of this paper use an informal presentation of
+homotopy type theory, working entirely inside of Univalent
+Foundations. 
+ %
+Section~\ref{sec:extension-type} describes how to use this
+construction to define a modality on $\Type$ from a modality on
+$\HProp$.
+%
+Section~\ref{sec:formalization} discusses the details of the
+formalization in Coq using formal type theory, in particular issues
+concerning universe polymorphism.
+%
+Finally, Section~\ref{sec:future-works} concludes on future works.
+
+\section{Lawvere-Tierney Sheafification in a nutshell}
+\label{sec:lawv-tiern-nutshell}
+
+In order to understand the construction of LT-sheafification in homotopy
+type theory, we first briefly recall the corresponding construction in
+topos theory and then gives an overview of our construction.
+
 \subsection{Lawvere-Tierney Sheafification in Topos Theory}
 \label{sec:lawv-tiern-sheaf}
-
-In order to understand the construction of LT-sheafification in homotopy
-type theory, we first briefly recall its corresponding construction in
-topos theory.
 
@@
 extensionality) and higher inductive types.
-
-\subsection{Related Works}
-\label{sec:related-works}
-
-Similar questions have been considered around the Curry-Howard
-isomorphism, to extend a programming language closed to type theory
-with new logical or computational principle while keeping consistency
-automatically.
-%
-For instance, much efforts have been done to provide a computational
-content to the law of excluded middle in order to define a
-constructive version of classical logic. This has lead to various
-calculi, with most notably the $\lambda \mu$-calculus of
-Parigot~\cite{parigot1993classical}, but this line of work has not
-appeared to be fruitful to define a new version of type theory with
-classical principles.
-%
-Other works have tried to extend continuation-passing-style (CPS)
-transformation to type theory, but they have been faced with the
-difficulty that the CPS transformation is incompatible with (full) dependent
-sums~\cite{barthe2002cps}, which puts emphasis on the tedious link
-between the axiom of choice and the law of excluded middle in type theory.
-%
-Nevertheless the axiom of choice has shown to be realizable by
-computational meaning in a classical setting by techniques turning
-around the notion of (modified) bar induction
-\cite{berardi1998computational}, Krivine's
-realizability~\cite{krivine2003dependent} and even more recently with
-restriction on elimination of dependent sums and lazy
-evaluation~\cite{herbelin2012constructive}.
-
-All those works---although they provide very nice and sophisticated
-theoretical results---are not amenable to a direct integration into a
-proof assistant based on type theory such as Coq~\cite{coq:refman:8.4}
-or Agda~\cite{norell2007towards}.
-
-The work on forcing in type theory~\cite{jaber2012extending} also
-gives a computational meaning to a type theory enriched with new
-logical or computational principle, but for other kinds of principles
-as the ones interpreted by LT-sheafification.
-%
-Actually, forcing is entirely complementary to LT-sheafification as
-forcing corresponds to the presheaf construction while
-LT-sheafification corresponds to the topological transformation that
-allows to go from the presheaf to the sheaf construction.
-
-\subsection{Plan of the Paper}
-
-Section~\ref{sec:hott} presents the foundations of homotopy type theory
-and present a analogous of~\cite[Corollary 6.2.3.5]{lurie}. 
-%
-Section~\ref{sec:lexmod} introduces the notion of left-exact stratified
-modalities, on which this paper is based.
-%
-Section~\ref{sec:sheaves} defines Lawvere-Tierney sheaves in homotopy
-type theory.
-%
-Section~\ref{sec:sheafification} explains the construction of the
-LT-sheafification functor.
-%
-The definitions of this paper uses an informal presentation of
-homotopy type theory, working entirely inside of Univalent
-Foundations. 
- %
-Section~\ref{sec:formalization} discusses the details of the
-formalization in Coq using formal type theory, in particular issues
-concerning universe polymorphism.
-%
-Finally, Section~\ref{sec:future-works} concludes on future works.
 
@@
 
+\nt{Add something on the definition of a syntactic translation as it
+  has been done for forcing.}
+
 As a future work, we would like to improve this construction in two

### 2016-01-07 paper/paper_acm.tex
add things

--- 2016-01-06 paper/paper_acm.tex
@@
 % correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
+\hyphenation{op-tical net-works semi-conduc-tor gro-then-dieck LT-sheafi-fi-ca-tion}
 
@@
 
+\subsection{Related Works}
+\label{sec:related-works}
+
+Similar questions have been considered around the Curry-Howard
+isomorphism, to extend a programming language closed to type theory
+with new logical or computational principle while keeping consistency
+automatically.
+%
+For instance, much efforts have been done to provide a computational
+content to the law of excluded middle in order to define a
+constructive version of classical logic. This has lead to various
+calculi, with most notably the $\lambda \mu$-calculus of
+Parigot~\cite{parigot1993classical}, but this line of work has not
+appeared to be fruitful to define a new version of type theory with
+classical principles.
+%
+Other works have tried to extend continuation-passing-style (CPS)
+transformation to type theory, but they have been faced with the
+difficulty that the CPS transformation is incompatible with (full) dependent
+sums~\cite{barthe2002cps}, which puts emphasis on the tedious link
+between the axiom of choice and the law of excluded middle in type theory.
+%
+Nevertheless the axiom of choice has shown to be realizable by
+computational meaning in a classical setting by techniques turning
+around the notion of (modified) bar induction
+\cite{berardi1998computational}, Krivine's
+realizability~\cite{krivine2003dependent} and even more recently with
+restriction on elimination of dependent sums and lazy
+evaluation~\cite{herbelin2012constructive}.
+
+All those works---although they provide very nice and sophisticated
+theoretical results---are not amenable to a direct integration into a
+proof assistant based on type theory such as Coq~\cite{coq:refman:8.4}
+or Agda~\cite{norell2007towards}.
+
+The work on forcing in type theory~\cite{jaber2012extending} also
+gives a computational meaning to a type theory enriched with new
+logical or computational principle, but for other kinds of principles
+as the ones interpreted by LT-sheafification.
+%
+Actually, forcing is entirely complementary to LT-sheafification as
+forcing corresponds to the presheaf construction while
+LT-sheafification corresponds to the topological transformation that
+allows to go from the presheaf to the sheaf construction.
+
+\subsection{Plan of the Paper}
+
+Section~\ref{sec:lawv-tiern-nutshell} presents an overview of the construction. 
+% 
+Section~\ref{sec:hott} presents the foundations of homotopy type theory
+and present a analogous of~\cite[Corollary 6.2.3.5]{lurie}. 
+%
+Section~\ref{sec:lexmod} introduces the notion of left-exact stratified
+modalities, on which this paper is based.
+%
+Section~\ref{sec:sheaves} defines Lawvere-Tierney sheaves in homotopy
+type theory.
+%
+Section~\ref{sec:sheafification} explains the construction of the
+LT-sheafification functor.
+%
+The definitions of this paper uses an informal presentation of
+homotopy type theory, working entirely inside of Univalent
+Foundations. 
+ %
+Section~\ref{sec:formalization} discusses the details of the
+formalization in Coq using formal type theory, in particular issues
+concerning universe polymorphism.
+%
+Finally, Section~\ref{sec:future-works} concludes on future works.
+
+\section{Lawvere-Tierney Sheafification in a nutshell}
+\label{sec:lawv-tiern-nutshell}
+
 \subsection{Lawvere-Tierney Sheafification in Topos Theory}
@@
 it as a left-exact modality on $\Type_{n}$ for any $n$.
+Our inductive definition of these modalities will be given by
+\[ \begin{array}{l}
+   \modal : \forall \ (n : nat), \ \Type_n \to \Type_n 
+   \\
+    \modal_{-1\phantom{n}}(T) \defeq\lnot\lnot T \\
+
+      \displaystyle{\modal_{n+1}(T)} \defeq  
+      \displaystyle{\sum_{u:T \to \Type_n^\modal} \!\!\!\!\modal_{-1} 
+      \left\|
+      \sum_{a:T} u= (\lambda t,~\modal_n (a=t))
+      \right\|}
+    \end{array}
+% \vspace{-0.3em}
+\]
+where $\Type_n^\modal$ is the type of modal types for $\modal_n$.
+The rest of the paper is an inductive proof that this definition produces a
+sequence of left-exact modalities.
 
@@
 
+Such a construction can be thought in two ways:
+\begin{itemize}
+\item as a translation from one type theory into another type theory
+  satisfying new principles
+\item or as a transformation, taking a model of type theory, and
+  producing another model, where the desired principle is true. This
+  can lead to consistency results.
+\end{itemize}
+
 In what follows, we work in the context of homotopy type theory, seen
@@
 \item an {\em embedding} if $f$ is $(-1)$-truncated
-\item a {\em surjection} is every fiber of $f$ is merely inhabited
-  (\ie $\|\fib f y\|$ holds for all $y$).
+\item a {\em surjection} if every fiber of $f$ is merely inhabited
+  (i.e $\|\fib f y\|$ holds for all $y$).
 \end{itemize}
@@
 
-A fundamental property in the construction of the LT-sheafification
+A fundamental property in the construction of the LT-sheafifica\-tion
 functor is the fact that the type of modal types at level $n$ defines
@@
 \begin{prop}
-  $(\modal_{\lnot\lnot})_n$, the modality obtained by LT-sheafification
+  $(\modal_{\lnot\lnot})_n$, the modality obtained by LT-sheafifica\-tion
   of the double negation modality,
@@
 
+We can view sheafification in terms of model of type theory ; let
+$\mathfrak M$ be a model of homotopy type theory. Using sheafification
+associated to a truncation index $n$, one can build a modality
+$\modal$, which is left-exact with respect to types at most
+$n$-truncated. Thus, there is a model $\modal \mathfrak M$ of type
+theory (using results in section~\ref{sec:new-type-theory}), where
+excluded middle is true, and which is univalent when restricted to $n$-types.
+
 % \subsection{New objects}
@@
 %
-\vspace{-0.05em}
+% \vspace{-0.05em}
 \[ \begin{array}{l}
@@
     \end{array}
-\vspace{-0.3em}
+% \vspace{-0.3em}
 \]
@@
 \end{itemize}
-Moreover, in topos theory, Lawvere-Tierney subsumes Gro\-thendieck~\cite[Section~V.4]{maclanemoerdijk} in the sense that any
+Moreover, in topos theory, Lawvere-Tierney subsumes Grothen\-dieck~\cite[Section~V.4]{maclanemoerdijk} in the sense that any
 Grothen\-dieck topology gives rise to a Lawvere-Tierney topology with

### 2016-01-07 paper/paper_acm.tex

@@
 LT-sheafification corresponds to the topological transformation that
-allows to go from the presheaf to the sheaf construction.
+allows to go from the presheaf construction to the sheaf construction.
 
@@
 
-Section~\ref{sec:lawv-tiern-nutshell} presents an overview of the construction. 
+Section~\ref{sec:lawv-tiern-nutshell} presents an overview of the
+construction.
 % 
 Section~\ref{sec:hott} presents the foundations of homotopy type theory
-and present a analogous of~\cite[Corollary 6.2.3.5]{lurie}. 
+and presents a analogous of~\cite[Corollary 6.2.3.5]{lurie}. 
 %
@@
 %
-The definitions of this paper uses an informal presentation of
+The definitions of this paper use an informal presentation of
 homotopy type theory, working entirely inside of Univalent
 Foundations. 
+ %
+Section~\ref{sec:extension-type} describes how to use this
+construction to define a modality on $\Type$ from a modality on
+$\HProp$.
 %
@@
 
+In order to understand the construction of LT-sheafification in homotopy
+type theory, we first briefly recall the corresponding construction in
+topos theory and then gives an overview of our construction.
+
 \subsection{Lawvere-Tierney Sheafification in Topos Theory}
 \label{sec:lawv-tiern-sheaf}
-
-In order to understand the construction of LT-sheafification in homotopy
-type theory, we first briefly recall its corresponding construction in
-topos theory.
 

### 2016-01-08 paper/paper_acm.tex
remove LT

@@
 promising to use a generalized notion of sheafification, called
-Lawvere-Tierney sheafification~\cite{tierney1972sheaf,maclanemoerdijk}
-(LT-sheafification), but this construction has not been considered yet
+Lawvere-Tierney
+sheafification~\cite{tierney1972sheaf,maclanemoerdijk},
+ but this construction has not been considered yet
 in the setting of higher topos theory. This rises two issues that this
-paper addresses: (i) how to lift the notion of LT-sheafification to a
+paper addresses: (i) how to lift the notion of Lawvere-Tierney sheafification to a
 higher dimensional setting and (ii) is it possible to formalize this
@@
 
-Technically, LT-sheafification in homotopy type theory allows to
+Technically, Lawvere-Tienrey sheafification in homotopy type theory allows to
 construct a model of homotopy type theory (restricted to finite homotopy
@@
 
-That is, LT-sheafification can be seen as a model transformation that
+That is, it can be seen as a model transformation that
 allows to justify new axioms in homotopy type theory. 
@@
 logical or computational principle, but for other kinds of principles
-as the ones interpreted by LT-sheafification.
-%
-Actually, forcing is entirely complementary to LT-sheafification as
+as the ones interpreted by Lawvere-Tierney sheafification.
+%
+Actually, forcing is entirely complementary to Lawvere-Tierney sheafification as
 forcing corresponds to the presheaf construction while
-LT-sheafification corresponds to the topological transformation that
+Lawvere-Tierney sheafification corresponds to the topological transformation that
 allows to go from the presheaf construction to the sheaf construction.
@@
 Section~\ref{sec:sheafification} explains the construction of the
-LT-sheafification functor.
+Lawvere-Tierney sheafification functor.
 %
@@
 
-In order to understand the construction of LT-sheafification in homotopy
+In order to understand the construction of Lawvere-Tierney sheafification in homotopy
 type theory, we first briefly recall the corresponding construction in
 topos theory and then gives an overview of our construction.
+
+From now on, {\em sheafification} is to be understood as {\em
+  Lawvere-Tierney sheafification}.
 
@@
 Furthermore, one can define a left adjoint to the inclusion, the
-LT-sheafification functor
+sheafification functor
 %
@@
 level of homotopy types. More precisely, the first layer of
-LT-sheafification is defined for $\Type_{0}$, given a topology on
-$\Type_{-1}$. This part corresponds to LT-sheafification for topoï. Then,
-assuming that the LT-sheafification has been constructed above level $n$,
-one can define the LT-sheafification for $\Type_{n+1}$.
-%
-This inductive process in the definition of LT-sheafification is similar
+sheafification is defined for $\Type_{0}$, given a topology on
+$\Type_{-1}$. This part corresponds to sheafification for topoï. Then,
+assuming that the sheafification has been constructed above level $n$,
+one can define the sheafification for $\Type_{n+1}$.
+%
+This inductive process in the definition of sheafification is similar
 to construction in the Grothendieck setting, where the
@@
 Lawvere-Tierney topology can thus be seen as a left-exact modality on
-$\Type_{-1}$ and LT-sheafification as an inductive process that extends
+$\Type_{-1}$ and sheafification as an inductive process that extends
 it as a left-exact modality on $\Type_{n}$ for any $n$.
@@
 
-Compare to the definition of LT-sheafification in topos theory, the main
+Compare to the definition of sheafification in topos theory, the main
 change comes from the fact that in homotopy type theory, there is not
@@
 For instance, the notion of separated objects (which is a technical
-notion used in the construction of the LT-sheafification functor), \ie
+notion used in the construction of the sheafification functor), \ie
 objects for which the morphism
@@
 
-Another technical difficulty is the fact that LT-sheafification in topos
+Another technical difficulty is the fact that sheafification in topos
 theory uses the property that epimorphisms are coequalizers of their
@@
 \item or as a transformation, taking a model of type theory, and
-  producing another model, where the desired principle is true. This
-  can lead to consistency results.
+  producing another model, where the desired principle is true. As in
+  set theory or topos theory, it could lead to independence results.
 \end{itemize}
@@
 theory that are central in our formalization but not specific to
-LT-sheafification. 
+sheafification. 
 % %
@@
 universe-polymorphic way.
+
+Section~\ref{ssec:hott} will present homotopy levels and object
+classifiers, section~\ref{ssec:colim-homot-type} introduces a theory
+of colimits in homotopy type theory, illustrated by an important
+example in section~\ref{ssec:giraud-ax}.
 
@@
 \subsection{Colimits in Homotopy Type Theory }
-\label{sec:colim-homot-type}
+\label{ssec:colim-homot-type}
 
 One desired property we would like to consider concerns the colimits
-of \v{C}ech nerves (Section~\ref{sec:giraud-ax}). This section presents
+of \v{C}ech nerves (Section~\ref{ssec:giraud-ax}). This section presents
 a definition of colimits in a type theoretic setting.
@@
 \subsection{On Giraud-Rezk-Lurie axioms}
-\label{sec:giraud-ax}
+\label{ssec:giraud-ax}
 
@@
 because Lawvere-Tierney topologies correspond exactly to left exact
-modalities at level $-1$ and LT-sheafification is described as an inductive
+modalities at level $-1$ and sheafification is described as an inductive
 process that produces a left exact modality at level $n+1$ from a
@@
 
-As the definition of LT-sheafification is based on subobject classifiers
+As the definition of sheafification is based on subobject classifiers
 and thus on homotopy fibers, we need to take a closer look at the way left
@@
 In this paper, we use the double negation modality at level $-1$ only,
-and then extend it to higher homotopy types using LT-sheafification. This
+and then extend it to higher homotopy types using sheafification. This
 way, we can add classical reasoning in the type theory without
@@
 %
-and LT-sheafification corresponds to lifting this modality to a left
+and sheafification corresponds to lifting this modality to a left
 exact modality at level $0$.
@@
 any left exact modality at level $-1$ to a left exact modality at level
-$n$ using LT-sheafification, by induction on $n$.
+$n$ using sheafification, by induction on $n$.
 
@@
 
-The LT-sheafification in topos theory described in
+The sheafification in topos theory described in
 Section~\ref{sec:lawv-tiern-sheaf} works in the same way. 
@@
 
-For a given induction step, the LT-sheafification is defined in two
-steps ; the first one construct, for any $T:\Type_{n+1}$ a separated
-type $\separated T$, and the second one construct a sheaf given any
+For a given induction step, the sheafification is defined in two
+steps ; the first one (section~\ref{ssec:from-type-separated}) construct, for any $T:\Type_{n+1}$ a separated
+type $\separated T$, and the second one
+(section~\ref{ssec:separated-to-sheaf}) construct a sheaf given any
 separated type.
@@
 
-Again, this lemma is an adaptation of the LT-sheafification process in
+Again, this lemma is an adaptation of the sheafification process in
 topos theory, where the kernel pair of $\mu_T$
@@
 \begin{proof}[Proof of lemma~\ref{lem:sepiscolim}]
-  As $\mu_T$ is an surjection, using the theorem~\ref{cech} introduced in Section~\ref{sec:giraud-ax},
+  As $\mu_T$ is an surjection, using the theorem~\ref{cech} introduced in Section~\ref{ssec:giraud-ax},
   we know that $\separated T$ is the $(n+1)$-colimit of $\|C(\mu_T)\|_{n+1}$. 
@@
 As this modality is just a step in the construction, we do not need to
-show that it is left exact, we will only do it for the LT-sheafification
+show that it is left exact, we will only do it for the sheafification
 modality.
@@
 true in this new type theory in the following sense:
-\begin{prop}
+\begin{prop}\label{prop:univalence}
   Let $n$ be a given truncation index, and $\modal$ the modality
@@
 theory (using results in section~\ref{sec:new-type-theory}), where
-excluded middle is true, and which is univalent when restricted to $n$-types.
+excluded middle is true, and which is univalent when restricted to
+$n$-types (in the sense of proposition~\ref{prop:univalence}).
 
@@
 recursive definitions. Indeed, the following recursive definition of
-LT-sheafification
+sheafification
 %
@@
 Thus, the induction step presented in this paper has been formalized,
-but the complete recursive LT-sheafification can not be defined for the
+but the complete recursive sheafification can not be defined for the
 moment.
@@
 %
-Using LT-sheafification, it is possible to add new logical principles to
+Using sheafification, it is possible to add new logical principles to
 homotopy type theory, such as the law of excluded middle, and this

### 2016-01-08 paper/paper_acm.tex
remove LT

@@
 The answer to this question has been given recently by the advent of
-homotopy type theory~\cite{hottbook}, which is an extension of
+homotopy type theory~\cite{hottbook}: type theory is not close {\em
+  enough} to topos theory for a straightforward adaptation. 
+%
+Homotopy type theory is an extension of
 Martin-Löf type theory with principles inspired by (higher) category
@@
 %
- such that the usual (\cite[Theorem 4.8.4]{hottbook}) subobject classifier diagram is a pullback.
+ such that the usual (\cite[Theorem 4.8.4]{hottbook}) subobject
+ classifier diagram is a pullback.
+Therefore, in our construction, we will represent a subobject of a
+type $B$ with $n$-truncated homotopy fibers either as a map $f:A\to B$
+such that $\IsType n A$, either as a characteristic map $B\to \Type_n$.
 % \begin{mymath}
@@
 \begin{defi}
-  Let $E$ be a type. The {\em closure} of a subobject of $E$ with
+  Let $E$ be a type. 
+  \begin{itemize}
+
+  \item The {\em closure} of a subobject of $E$ with
   n-truncated homotopy fibers (or $n$-subobject of $E$, for short),
@@
 
-  An $n$-subobject of $E$ classified by $\chi$ is said to be {\em
+\item An $n$-subobject of $E$ classified by $\chi$ is said to be {\em
     closed in $E$} if it is equal to its closure, \ie
@@
 
-  An $n$-subobject of $E$ classified by $\chi$ is said to be {\em
+\item An $n$-subobject of $E$ classified by $\chi$ is said to be {\em
     dense in $E$} if its closure is $E$, \ie
   $\modal_n \circ \chi = \lambda e, \one$ 
+  \end{itemize}
 \end{defi}
@@
 
+As many proofs are very technical, we only give here sketches, and
+often refer to formalization for all details.
+
 \subsection{From Type to Separated Type}
@@
 }\end{mymath}%
-% We can restrict the previous diagram to 
-% \begin{mymath}\xymatrix @R=4em @C=5em{
-%   \sum_{e:E} (\chi\, e) \ar@<-2pt>[r]_{\pi_1\circ g\circ\pi_1} \ar@<2pt>[r]^{\pi_1\circ f\circ \pi_1} \ar[d]_{\mathrm{dense}}& \sum_{x :A} (Bx) \\
-%   E \ar@<-2pt>[ru]_{\pi_1\circ g} \ar@<2pt>[ru]^{\pi_1\circ f}&
-% }\end{mymath}%
-As $A$ is separated, $\pi_1\circ f = \pi_2 \circ g$.
+We can restrict the previous diagram to 
+\begin{mymath}\xymatrix @R=4em @C=5em{
+  \sum_{e:E} (\chi\, e) \ar@<-2pt>[r]_{\pi_1\circ g\circ\pi_1} \ar@<2pt>[r]^{\pi_1\circ f\circ \pi_1} \ar[d]_{\mathrm{dense}}& \sum_{x :A} (Bx) \\
+  E \ar@<-2pt>[ru]_{\pi_1\circ g} \ar@<2pt>[ru]^{\pi_1\circ f}&
+}\end{mymath}%
+and as $A$ is separated, $\pi_1\circ f = \pi_1 \circ g$.
 For the second components, let $x:E$. Notice that 
@@
 A Coq formalization based on the Coq/HoTT library~\cite{hottlib} is available at
-\url{github.com/KevinQuirin/sheafification}.
+\url{https://github.com/KevinQuirin/sheafification}.
 %
@@
 
-\nt{Add something on the definition of a syntactic translation as it
-  has been done for forcing.}
+% \nt{Add something on the definition of a syntactic translation as it
+  % has been done for forcing.}
 
@@
   than excluded middle.
+\item It would be useful to give a syntactic translation of
+  sheafification,
+  as it has been done for forcing~\cite{jaber2012extending}, to be
+  able to work in the new type theory into a proof assistant.
 \end{itemize}

### 2016-01-15 paper/paper_acm.tex
paper

@@
 \usepackage{amssymb,amsthm,amsfonts}
+\usepackage{mathbbol}
 
@@
            {\'Ecole des Mines de Nantes}
-           {kevin.quirin@inria.fr}
+           {kevin.quirin@mines-nantes.fr}
 \authorinfo{Nicolas Tabareau}
@@
 \begin{itemize}
-\item as a translation from one type theory into another type theory
-  satisfying new principles
-\item or as a transformation, taking a model of type theory, and
+\item As a translation from one type theory into another type theory
+  satisfying new principles, in the same way Gödel translation embeds
+  classical logic into  intuitionistic logic.
+\item Or as a transformation, taking a model of type theory, and
   producing another model, where the desired principle is true. As in
-  set theory or topos theory, it could lead to independence results.
+  set theory or topos theory, it leads to consistency results.
 \end{itemize}
@@
 described in~\ref{sec:examples-left-exact}.
-
 Chosing $i\geqslant 0$ is a refinement of this double negation
@@
 
-Taking $i$ as a non-fixed parameter, that would allow us to work in an
+Obviously, as truncation modalities are not left-exact~\cite[Exercise
+7.11]{hottbook}, $\modal$ isn't either. But in the following sense, when
+restricted to $i$-truncated types, it is:
+\begin{lem}
+  Let $A:\Type_i$. Then if $\modal(A)$ is contractible, for any $x,y:A$,
+  $\modal(x=y)$ is contractible.
+\end{lem}
+\begin{proof}
+  For $i$-truncated types, $\modal = \modal_i$, and $\modal_i$ is left-exact.
+\end{proof}
+
+The compatibility between the modalities $\modal_n$ and between the
+modalities $\|\cdot \|_n$ allow us to chose the truncation index as
+high as desired.
+Taking it as a non-fixed parameter allows to work in an
 universe where the new principle ({\em e.g.} mere excluded middle) is
 true for any explicit truncated type. Indeed, $i$ can be chosen
-dynamically along a proof, and thus be increased as much as needed.
+dynamically along a proof, and thus be increased as much as needed,
+without changing results for lower truncated types.
 
@@
   associated to $n+1$ as defined in lemma~\ref{lem:type}. Then, for
-  any type $A,B:\Type_n^\modal$, the canonical arrow
-  $$\varphi~:~A = B \to A\simeq B$$
-  is an equivalence.
+  any type $A,B:\Type_n^\modal$, if $\varphi$ is the canonical arrow
+  $$A = B \to A\simeq B,$$
+  then $\modal(\IsEquiv(\varphi))$ is inhabited.
 \end{prop}
@@
 % Actually, one can defined such a map for any family of $\HProp$
-% indexed by $\nat$, for example with the family $C_n \defeq \text{«\,The $n$-th Turing Machine halts\,»}$.
+% indexed by $\nat$, for example with the family $C_n \defeq
+% \text{«\,The $n$-th Turing Machine halts\,»}$.
+
+\section{Translation}
+\label{sec:translation}
+
+Let $n$ be a trunction index, and let $\modal = \modal_n \circ
+\|\cdot\|_n$ the constructed modality. We define the following
+translation
+\begin{itemize}
+\item For $\Type$
+\[
+\begin{array}{lcl}
+  \Lbrack \Type\Rbrack &\defeq& \Type_\modal
+\end{array}
+\]
+
+\item For dependent sums
+\[
+\begin{array}{lcl}
+\left[ \sum_{x:A} B \right] &\defeq&  \left( \sum_{x:\Lbrack A \Rbrack}
+                                  \Lbrack B\Rbrack ; \pi_\Sigma^{[A],[B]}
+                                \right)\\[0.5em]
+  \left[  (x;y)\right] &\defeq& ([x];[y]) \\[0.5em]
+  \left[  \pi_i t\right] &\defeq& \pi_i [t] \\[0.5em]
+\end{array}
+\]
+\item For dependent products
+\[
+\begin{array}{lcl}
+ \left[ \prod_{x:A} B \right] &\defeq& \left( \prod_{x:\Lbrack A \Rbrack}
+                                   \Lbrack B\Rbrack ; \pi_{\Pi}^{[A],[B]}
+                                  \right)\\[0.5em]
+\left[  \lambda\, x:A,~B \right] &\defeq&\lambda\,x:\Lbrack A
+                                     \Rbrack,~\Lbrack B\Rbrack
+  \\[0.5em]
+  \left[ t \, t' \right] &\defeq& [t] [t'] \\[0.5em]
+\end{array}
+\]
+\item For sums
+\[
+\begin{array}{lcl}
+\left[  A+B \right] &\defeq& \left( \modal(\Lbrack A \Rbrack + \Lbrack B
+                        \Rbrack); \pi_\modal(\Lbrack A \Rbrack + \Lbrack B
+                        \Rbrack)\right)\\[0.5em]
+\left[  \mathrm{in}_\ell t \right] &\defeq& \eta (\mathrm{in}_\ell [t]) \\[0.5em]
+\left[  \mathrm{in}_r t \right] &\defeq& \eta (\mathrm{in}_r [t]) \\[0.5em]
+\left[ \langle f ,g\rangle\right] &\defeq& \modal_{\mathrm{rec}}^{\Lbrack A\Rbrack +
+                                      \Lbrack B\Rbrack} \langle
+                                      [f],[g]\rangle\\[0.5em]
+\end{array}
+\]
+\item For paths
+\[
+\begin{array}{lcl}
+\left[  A=B \right] &\defeq& \left( \Lbrack A \Rbrack = \Lbrack B
+                             \Rbrack , \pi_=^{[A],[B]}
+                             \right)\\[0.5em]
+\left[ 1 \right] &\defeq& 1\\[0.5em]
+\end{array}
+\]
+\item For truncations ($i\leqslant n$)
+\[
+\begin{array}{lcl}
+  \left[  \|A\|_i \right] &\defeq& (\modal \| \Lbrack A\Rbrack  \|_i;
+                                   \pi_\modal(\| \Lbrack A\Rbrack
+                                   \|_i)) \\[0.5em]
+  \left[ |t|_i \right] &\defeq& \eta |[t]|_i \\[0.5em]
+  \left[ \|f\|_i \right] &\defeq& \modal_{\mathrm{rec}}^{\| \Lbrack
+                                  A\Rbrack  \|_i} \| [f] \|_i
+\end{array}
+\]
+
+\end{itemize}
+
+where $\pi_\Sigma^{A,B}$ (resp. $\pi_\Pi^{A,B}$, resp. $\pi_=^{A,B}$)
+is the proof that $\Sigma_{x:A} B$ (resp. $\prod_{x:A} B$, resp $A=B$)
+is modal as soon as $A$ and $B$ are, and $\pi_\modal^A$ a proof that
+$\modal(A)$ is always modal as described in section~\ref{sec:lexmod}.
+
+As in~\cite{jaber2012extending}, the main issue is that $\Lbrack
+U\{x/N\} \Rbrack$
+and $\Lbrack U \Rbrack\{x/[N]\}$ are equal, but not convertible. We
+therefore have to use $\mathrm{transport}$, the eliminator of the
+equality, in the conversion rule.
+
+The central theorem is
+\begin{prop}
+  Let $\Gamma$ be a valid context, $A$ a type and $t$ a term of type
+  $A$.
+  Then if $\Gamma \vdash t: A$, then $\Lbrack\Gamma\Rbrack \vdash [t]
+  : \Lbrack A\Rbrack$.
+\end{prop}
 
@@
 the project
-contains 8081 lines, and it could be reduced a bit by improving the
+contains 7914 lines, and it could be reduced a bit by improving the
 way Coq tries to rewrite and apply lemmas automatically. 
-The \texttt{coqwc} tool counts 1750 lines of specifications
-(definitions, lemmas, theorems, propositions) and 6330 lines of proof
+The \texttt{coqwc} tool counts 1611 lines of specifications
+(definitions, lemmas, theorems, propositions) and 5403 lines of proof
 script.
@@
 
-As a future work, we would like to improve this construction in two
+As a future work, we would like to improve this construction in three
 ways:

### 2016-01-17 paper/paper_acm.tex
correct proof of prop 30

@@
     Now, let define $\phi : \modal_{-1} T \to \modal_0T$ by
-    \begin{mymath}\phi t = (\lambda t',\, \modal_{-1}(t=t') ; \eta
-    (t;1))\end{mymath}%
-    and $\psi : \modal_0T\to \modal_{-1} T$ by obtaining the
+    \begin{mymath}\phi t = (\lambda t',\, (t=\eta_T (t'))
+      ;\kappa)\end{mymath}%
+    where $\kappa$ is defined by $\modal_{-1}$-induction on $t$.
+    Let $\psi : \modal_0T\to \modal_{-1} T$ by obtaining the
     witness $a:T$ (which is possible because we are trying to inhabit

### 2016-01-18 paper/paper_acm.tex
correct proof of prop 30
more on translation

@@
     Now, let define $\phi : \modal_{-1} T \to \modal_0T$ by
-    \begin{mymath}\phi t = (\lambda t',\, (t=\eta_T (t'))
+    \begin{mymath}\phi t = (\lambda t',\, \one
       ;\kappa)\end{mymath}%
-    where $\kappa$ is defined by $\modal_{-1}$-induction on $t$.
+    where $\kappa$ is defined by $\modal_{-1}$-induction on
+    $t$. Indeed, as $T$ is an $\HProp$, $(a=t) \simeq \one$. 
     Let $\psi : \modal_0T\to \modal_{-1} T$ by obtaining the
@@
 
-Unfortunately, one lose canonicity when working in $\Type^\modal$;
-indeed, for example, the function 
-\begin{mymath}\lambda P:\HProp,~\left\{\begin{array}{l} 0 \text{ if $P$ is
-                                    false} \\ 1 \text{ if $P$ is
-                                    true} \end{array}\right.~:~\HProp
-                              \to \nat\end{mymath}%
-will not reduce to a numeral when applied to a undecidable
-property. However, we keep canonicity for ``old'' terms, i.e. the
-ones constructed without using excluded middle.
-
 By proposition~\ref{prop:consistent}, these left-exact modalities
@@
 Let $n$ be a trunction index, and let $\modal = \modal_n \circ
-\|\cdot\|_n$ the constructed modality. We define the following
-translation
+\|\cdot\|_n$ the constructed modality. 
+%
+As this modality is not left exact, $\Type$ itself is not modal and we
+can only define the translation on a type thoery with only one
+universe. 
+%
+We define the translation on Martin-Löf type theory,
+but it could also be done for CIC.   
+%
 \begin{itemize}
-\item For $\Type$
+\item For types (assuming $A \neq \Type$) 
 \[
-\begin{array}{lcl}
-  \Lbrack \Type\Rbrack &\defeq& \Type_\modal
+\begin{array}{lcllcl}
+  \Lbrack \Type\Rbrack &\defeq& \Type^\modal & \qquad 
+  \Lbrack A \Rbrack &\defeq& \pi_1 \left[ A \right]
 \end{array}
@@
 
-As in~\cite{jaber2012extending}, the main issue is that $\Lbrack
-U\{x/N\} \Rbrack$
-and $\Lbrack U \Rbrack\{x/[N]\}$ are equal, but not convertible. We
+As in the forcing translation~\cite{jaber2012extending}, the main
+issue is that $\Lbrack U\{x/N\} \Rbrack$ and
+$\Lbrack U \Rbrack\{x/[N]\}$ are equal, but not convertible. We
 therefore have to use $\mathrm{transport}$, the eliminator of the
-equality, in the conversion rule.
-
-The central theorem is
-\begin{prop}
-  Let $\Gamma$ be a valid context, $A$ a type and $t$ a term of type
-  $A$.
-  Then if $\Gamma \vdash t: A$, then $\Lbrack\Gamma\Rbrack \vdash [t]
-  : \Lbrack A\Rbrack$.
-\end{prop}
+equality, in the conversion rule. This requires to use a type theory
+with explicit conversion in the syntax. We have chosen to omit this
+technical details here to simplify the presentation.
+
+% The central theorem is
+% \begin{prop}
+%   Let $\Gamma$ be a valid context, $A$ a type and $t$ a term of type
+%   $A$.
+%   Then if $\Gamma \vdash t: A$, then $\Lbrack\Gamma\Rbrack \vdash [t]
+%   : \Lbrack A\Rbrack$.
+% \end{prop}
+
+Modulo this technical point, we have the usual property that if a term
+$t$ has type $A$ in context $\Gamma$, then $[t]$ has type $\Lbrack
+A\Rbrack$ in the translated context $\Lbrack \Gamma\Rbrack$.
+%
+Furthermore, depending the modality $\modal_{-1}$ under focus, we can
+add new constants in the type theory, and give them a meaning through
+the translation.
+
+For instance, for the double negation modality, we can add the term
+$\mathop{EM}$ by defining 
+$$
+[\mathop{EM}] \defeq \lambda P, \lambda H : \lnot (P + \lnot P), H (\mathrm{in}_r (\lambda
+p : P, H (\mathrm{in}_\ell p)))
+$$
+of type 
+$$
+\Lbrack \forall P:\HProp, P + \lnot P \Rbrack \equiv \forall
+P:\Type^\modal_{-1}. \lnot (\lnot (P + \lnot P)).
+$$
+
+Unfortunately, one lose canonicity when working in $\Type^\modal$;
+indeed, for example, the translation of the function 
+% \begin{mymath}\lambda P:\HProp,~\left\{\begin{array}{l} 0 \text{ if $P$ is
+%                                     false} \\ 1 \text{ if $P$ is
+%                                     true} \end{array}\right.~:~\HProp
+%                               \to \nat\end{mymath}%
+\begin{mymath}
+\lambda P:\HProp, \langle 0 ,1\rangle (\mathop{EM} P)~:~\HProp
+                     \to \nat\end{mymath}%
+will not reduce to a numeral when applied to a given
+property. However, we keep canonicity for terms of Martin-Löf Type
+theory which do not make us of new constants.
 

### 2016-01-18 paper/paper_acm.tex
more on translation

@@
 compatible with $\modal_n$ on lower homotopy types.
-\begin{prop}
+\begin{prop} \label{prop:compatibility}
   If $T:\Type_n$, then $\modal_{n+1} \widehat T = \modal_n T$, where $\widehat T$ is $T$ seen as a
@@
 
-We can view sheafification in terms of model of type theory ; let
-$\mathfrak M$ be a model of homotopy type theory. Using sheafification
-associated to a truncation index $n$, one can build a modality
-$\modal$, which is left-exact with respect to types at most
-$n$-truncated. Thus, there is a model $\modal \mathfrak M$ of type
-theory (using results in section~\ref{sec:new-type-theory}), where
-excluded middle is true, and which is univalent when restricted to
-$n$-types (in the sense of proposition~\ref{prop:univalence}).
+We can view sheafification in terms of model of type theory but
+because of the resulting modality on $\Type$ is not left exact, we
+need to restrict ourselves to a type theory with only one universe.
+%
+Let $\mathfrak M$ be a model of homotopy type theory with one
+universe.
+%
+Using the modality $\modal$ (for any level $n$) associated to the
+sheafification, there is a model $\modal \mathfrak M$ of type theory
+with one universe (using results in
+Section~\ref{sec:new-type-theory}), where excluded middle is true, and
+which is univalent (as shown in Proposition~\ref{prop:univalence}).
+%
+This model transformation can better be described as a program
+transformation, as done in the next section.
 
@@
 
+In this section, we present how to use the sheafification modality to
+define a transformation of Matrin-Löf type theory into
+itself\footnote{We define the translation on Martin-Löf type theory,
+  but it could also be done for CIC as well.} as it has been done for
+forcing~\cite{jaber2012extending}.
+
 Let $n$ be a trunction index, and let $\modal = \modal_n \circ
@@
 %
-We define the translation on Martin-Löf type theory,
-but it could also be done for CIC.   
+To explain the translation on positive types as well, we also
+translate the sum type $A+B$, with associated constructors
+$\mathrm{in}_\ell$
+and $\mathrm{in}_r$ and eliminator $\langle \_ , \_ \rangle$.
+%
 %
@@
 \]
+\item For paths
+\[
+\begin{array}{lcl}
+\left[  A=B \right] &\defeq& \left( \Lbrack A \Rbrack = \Lbrack B
+                             \Rbrack , \pi_=^{[A],[B]}
+                             \right)\\[0.5em]
+\left[ 1 \right] &\defeq& 1\\[0.5em]
+\left[ \mathtt{transport} \ A \ P \ x \ y \ e \right] &\defeq&\mathtt{transport}
+  \ \Lbrack A \Rbrack \ [P] \ [x] \ [y] \ [e] \\[0.5em]
+\end{array}
+\]
 \item For sums
@@
 \]
-\item For paths
-\[
-\begin{array}{lcl}
-\left[  A=B \right] &\defeq& \left( \Lbrack A \Rbrack = \Lbrack B
-                             \Rbrack , \pi_=^{[A],[B]}
-                             \right)\\[0.5em]
-\left[ 1 \right] &\defeq& 1\\[0.5em]
-\end{array}
-\]
 \item For truncations ($i\leqslant n$)
@@
 $$
+where we have use an implicit rewriting with
+Proposition~\ref{prop:compatibility}.
 
@@
 property. However, we keep canonicity for terms of Martin-Löf Type
-theory which do not make us of new constants.
+theory which do not make us of new constants nor positive types. 
+%
 
@@
 As a future work, we would like to improve this construction in three
-ways:
-\begin{itemize}
-\item 
-  The extension to whole $\Type$ in lemma~\ref{lem:type} is not
+ways.
+% \begin{itemize}
+% \item 
+(i) The extension to whole $\Type$ in lemma~\ref{lem:type} is not
   totally satisfactory, as every type is collapsed to a truncated
@@
   % limit construction.
-\item 
+% \item 
+(ii) 
   We would like to have much more examples of left-exact modalities on
@@
   than excluded middle.
-\item It would be useful to give a syntactic translation of
-  sheafification,
-  as it has been done for forcing~\cite{jaber2012extending}, to be
-  able to work in the new type theory into a proof assistant.
-\end{itemize}
+% \item 
+(iii) It would be useful to improve the translation induced by the
+  sheafification modality to preserve more definitional equalities.
+  %
+  This could be done for instance by defining a higher inductive
+  isomorphic to the modality by with a better computational content. 
+  % able to work in the new type theory into a proof assistant.
+% \end{itemize}
+
 Moreover, in topos theory, Lawvere-Tierney subsumes Grothen\-dieck~\cite[Section~V.4]{maclanemoerdijk} in the sense that any
@@
 
-% \section*{Acknowledgment}
-
-% We would like to thank all maintainers of the Coq/HoTT library, who
-% have provided a very solid base for our formalization. 
-% We also would like
-% to thank Bas Spitters for helpful discussions on the subject.
+\section*{Acknowledgment}
+
+We would like to thank all maintainers of the Coq/HoTT library, who
+have provided a very solid base for our formalization. 
+%
+We also would like to thank Bas Spitters for helpful discussions on
+the subject.
+%
+This work has been funded by the CoqHoTT ERC Grant 637339.
 

### 2016-01-18 paper/paper_acm.tex
adding Section 8 in plan of the paper

@@
 $\HProp$.
+%
+%
+Section~\ref{sec:translation} defines the translation of Matin-Löf
+type theory into itself based on this modality. 
 %
@@
 In this section, we present how to use the sheafification modality to
-define a transformation of Matrin-Löf type theory into
+define a transformation of Martin-Löf type theory into
 itself\footnote{We define the translation on Martin-Löf type theory,

### 2016-01-18 paper/paper_acm.tex
typo
prop 35

--- 2016-01-18 paper/paper_acm.tex
@@
 type $B$ with $n$-truncated homotopy fibers either as a map $f:A\to B$
-such that $\IsType n A$, either as a characteristic map $B\to \Type_n$.
+such that $\IsType n \fib{f}{b}$, either as a characteristic map $B\to \Type_n$.
 % \begin{mymath}
@@
   $$A = B \to A\simeq B,$$
-  then $\modal(\IsEquiv(\varphi))$ is inhabited.
+  then $\IsEquiv(\varphi)$ is modal.
 \end{prop}
@@
 
-  Hence, $\IsEquiv \varphi$ is modal, and then is true in the new type
-  theory, as it was in the original one.
+  Hence, $\IsEquiv \varphi$ is modal.
 \end{proof}

### 2016-01-18 paper/paper_acm.tex

@@
 
-\section*{Acknowledgment}
+\section*{Acknowledgments}
 

### 2016-01-19 paper/paper_acm.tex
keywords
identity type using J
typos

@@
 % correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor gro-then-dieck LT-sheafi-fi-ca-tion}
+\hyphenation{op-tical net-works semi-conduc-tor gro-then-dieck sheafi-fi-ca-tion}
 
@@
 
-\titlebanner{banner above paper title}        % These are ignored unless
-\preprintfooter{short description of paper}   % 'preprint' option specified.
+\titlebanner{}        % These are ignored unless
+\preprintfooter{}   % 'preprint' option specified.
 
@@
 % you may leave them out
-\terms
-term1, term2
+% \terms
+% term1, term2
 
 \keywords
-keyword1, keyword2
+sheaf, homotopy type theory, modalities
 
@@
   Let $D$ be a diagram, $m$ a truncation index, and
-  $P:\Type_{m}$. 
+  $P:\Type_{m}$ a colimit of $D$. 
   Then, if $\|D\|_m$ is the same diagram as $D$, but where every type
@@
 The Giraud-Rezk-Lurie axioms are the $\infty$-version of Giraud's
-axioms that characterize a topos. Namely, there are 4 axioms on a
+axioms that characterize a topos. Namely, there are four axioms on a
 $(\infty,1)$-category that have been shown to be equivalent to
@@
 process that produces a left exact modality at level $n+1$ from a
-level exact modality at level $n$.
+left exact modality at level $n$.
 
@@
 
-A fundamental property in the construction of the LT-sheafifica\-tion
+A fundamental property in the construction of the sheafifica\-tion
 functor is the fact that the type of modal types at level $n$ defines
@@
 logical principles that can be added by the sheaf construction must be
-added in the first modality already. LT-Sheafification is just a way
+added in the first modality already. Sheafification is just a way
 to propagate those new principles to types that are not mere
@@
 \begin{prop}
-  $(\modal_{\lnot\lnot})_n$, the modality obtained by LT-sheafifica\-tion
+  $(\modal_{\lnot\lnot})_n$, the modality obtained by sheafifica\-tion
   of the double negation modality,
@@
 hypothesis to a classical setting, which is where the continuum hypothesis is
-really meaningful.  However, we haven't work out the details and left
+really meaningful.  However, we haven't worked out the details and left
 this for future work.
@@
 As this modality is not left exact, $\Type$ itself is not modal and we
-can only define the translation on a type thoery with only one
+can only define the translation on a type theory with only one
 universe.
@@
 $$
-where we have use an implicit rewriting with
+where we have uses an implicit rewriting with
 Proposition~\ref{prop:compatibility}.
@@
 \begin{mymath}
-\lambda P:\HProp, \langle 0 ,1\rangle (\mathop{EM} P)~:~\HProp
+\lambda P:\HProp, \langle \lambda x,\,0 ,\lambda x,\,1\rangle (\mathop{EM} P)~:~\HProp
                      \to \nat\end{mymath}%
@@
 part of the theory of higher topoï. As a matter of comparison, this
-\TeX\xspace file contains approximatively 1800 lines.
+\TeX\xspace file contains approximatively 2000 lines.
 
@@
   %
-  This could be done for instance by defining a higher inductive
-  isomorphic to the modality by with a better computational content. 
+  This could be done for instance by defining a higher inductive type
+  isomorphic to the modality but with a better computational content. 
   % able to work in the new type theory into a proof assistant.

### 2016-01-19 paper/paper_acm.tex
typos
pass on intro

@@
 %
-which transports equalities to equivalences is an equivalence.
+which transports equalities between types to equivalences is an equivalence.
 %
@@
 
-Technically, Lawvere-Tienrey sheafification in homotopy type theory allows to
-construct a model of homotopy type theory (restricted to finite homotopy
-types) with new logical principles (such as EM) from a model of
+Technically, Lawvere-Tienrey sheafification in homotopy type theory
+allows to construct a model of homotopy type theory (with only one
+universe) with new logical principles (such as EM) from a model of
 homotopy type theory.
@@
 Similar questions have been considered around the Curry-Howard
-isomorphism, to extend a programming language closed to type theory
+isomorphism, to extend a programming language close to type theory
 with new logical or computational principle while keeping consistency
@@
 %
-Actually, forcing is entirely complementary to Lawvere-Tierney sheafification as
-forcing corresponds to the presheaf construction while
+Actually, this version of forcing is entirely complementary to Lawvere-Tierney sheafification as
+it corresponds to the presheaf construction while
 Lawvere-Tierney sheafification corresponds to the topological transformation that
@@
 Section~\ref{sec:hott} presents the foundations of homotopy type theory
-and presents a analogous of~\cite[Corollary 6.2.3.5]{lurie}. 
+and an analogue of a theorem in higher toposes that is crucial
+in our construction~\cite[Corollary 6.2.3.5]{lurie}. 
 %
@@
 %
+We note $J$ the eliminator of the identity type. 
+%
 %
@@
 \left[ 1 \right] &\defeq& 1\\[0.5em]
-\left[ \mathtt{transport} \ A \ P \ x \ y \ e \right] &\defeq&\mathtt{transport}
-  \ \Lbrack A \Rbrack \ [P] \ [x] \ [y] \ [e] \\[0.5em]
+\left[ J \right] &\defeq& J \\[0.5em]
 \end{array}
@@
 $$
-where we have uses an implicit rewriting with
+where we have used an implicit rewriting with
 Proposition~\ref{prop:compatibility}.

### 2016-01-19 paper/paper_acm.tex
j instead of not not

@@
 This inductive process in the definition of sheafification is similar
-to construction in the Grothendieck setting, where the
+to the construction in the Grothendieck setting, where the
 sheafification is obtained by applying the plus-construction $n\!+\!1$
@@
 left-exact modality introduced in~\cite[Chapter~7]{hottbook}. A
-Lawvere-Tierney topology can thus be seen as a left-exact modality on
+Lawvere-Tierney topology $j$ can thus be seen as a left-exact modality on
 $\Type_{-1}$ and sheafification as an inductive process that extends
@@
    \\
-    \modal_{-1\phantom{n}}(T) \defeq\lnot\lnot T \\
+    \modal_{-1\phantom{n}}(T) \defeq \mathop{j} T \\
 

### 2016-01-19 paper/paper_acm.tex
pass on section 2

@@
 using Giraud-Rezk-Lurie axioms of $\infty$-topoï, the same property
-holds by replacing kernel pairs by \v{C}ech nerves in the statement.
+holds by replacing kernel pairs by \v{C}ech nerves in the statement
+(Section~\ref{ssec:giraud-ax} presents the analogue theorem in
+homotopy type theory).
 

### 2016-01-19 paper/paper_acm.tex
fixing notation of the iterated kernel pair

@@
 In~\ref{ssec:from-type-separated}, we will need to know how colimits
-behave with respect to truncations. An answer is given by the following:
+behave with respect to truncations. An answer is given by the
+following lemma.
 
@@
 In~\cite{boulier}, the authors propose an analogue of this property~:
-they give, for any map $f$, a diagram $D_f$ whose colimit is $\im(f)$.
+they give, for any map $f$, a diagram $C(f)$ whose colimit is $\im(f)$.
 
@@
 
-Then, the desired diagram is the mapping telescope of the iterations
+Then, the considered diagram $C(f)$  is the mapping telescope of the iterations
 of $T$.
@@
 \begin{thm}[Colimit of $C(f)$~\cite{boulier}]\label{cech}
-  For any surjection $f : X \to Y$, the colimit of $C(f)$ is $Y$.
+  For any morphism $f : X \to Y$, the colimit of $C(f)$ is $\im(f)$,
+  the image of $f$.
 \end{thm}

### 2016-01-19 paper/paper_acm.tex
typo

--- 2016-01-19 paper/paper_acm.tex
@@
 
-Technically, Lawvere-Tienrey sheafification in homotopy type theory
+Technically, Lawvere-Tierney sheafification in homotopy type theory
 allows to construct a model of homotopy type theory (with only one

### 2016-01-19 paper/paper_acm.tex
adding ref to formalization

--- 2016-01-19 paper/paper_acm.tex
@@
 our presentation (also the setting using truncated types provides
-minor changes), but they can be found in our Coq formalization.
+minor changes), but they can be found in our Coq formalization (see
+Section~\ref{sec:formalization}).
 

### 2016-01-19 paper/paper_acm.tex
typo
correction

--- 2016-01-19 paper/paper_acm.tex
@@
 
-\begin{prop}
+\begin{prop}\label{prop:mod_prop}
 
@@
 \item For any $A:\Type_n^\modal$ and $x,y:A$, $P(x=y)$.
+\item For any $A,B:\Type_n^\modal$, $P(A=B)$.
 \item 
@@
   Let $n$ be a given truncation index, and $\modal$ the modality
-  associated to $n+1$ as defined in lemma~\ref{lem:type}. Then, for
+  associated to $n$ as defined in lemma~\ref{lem:type}. Then, for
   any type $A,B:\Type_n^\modal$, if $\varphi$ is the canonical arrow
@@
   Therefore, it suffices to show that both $A=B$ and $A\simeq B$ are
-  modal. As $\Type_n^\modal$ is modal, $A=B$ is modal. 
+  modal. By proposition~\ref{prop:mod_prop}, $A=B$ is modal. 
   Moreover, $(A\simeq B) \simeq \sum_{f:A\to B} \IsEquiv

### 2016-01-19 paper/paper_acm.tex

@@
 to sub-$(\infty,1)$-topoï. Indeed, modal types defines a new type
-theory as modality preserves all type former, such as dependent
+theory as a modality preserves all type former, such as dependent
 products or dependent sums.

### 2016-01-19 paper/paper_acm.tex
adding brackets to ease the reading
typo
repetition and typos

@@
   path types are $n$-modal, ie
-  \begin{mymath}[-0.3em]\prod_{x,y:F}~\modal_n(x=y) = (x=y).\end{mymath}%
+  \begin{mymath}[-0.3em]\prod_{x,y:F}~\left( \modal_n(x=y) \right) = (x=y).\end{mymath}%
 \end{prop}
@@
 
-In topos-theoretic words, it means that given a map $\sum_{e:E}
+In topos-theoretic words, it means that given a map $f : \sum_{e:E}
 \chi\, e\to F$, one can
-extend it uniquely in $\tilde f:E \to F$.
+extend it uniquely to $\tilde f:E \to F$.
 
@@
 For a given induction step, the sheafification is defined in two
-steps ; the first one (section~\ref{ssec:from-type-separated}) construct, for any $T:\Type_{n+1}$ a separated
+phases; the first one (section~\ref{ssec:from-type-separated}) constructs, for any $T:\Type_{n+1}$ a separated
 type $\separated T$, and the second one
-(section~\ref{ssec:separated-to-sheaf}) construct a sheaf given any
+(section~\ref{ssec:separated-to-sheaf}) constructs a sheaf given any
 separated type.

### 2016-01-19 paper/paper_acm.tex
adding _{\lnot \lnot}

@@
 
-In the previous section, we defined a (countably) infinite family of
+In the previous section, we have defined a (countably) infinite family of
 modalities $\Type_i \to \Type_i$. One can extend them to whole
@@
 %
-Using the modality $\modal$ (for any level $n$) associated to the
-sheafification, there is a model $\modal \mathfrak M$ of type theory
+Using the modality $\modal_{\lnot \lnot}$ (for any level $n$) associated to the
+sheafification, there is a model $\modal_{\lnot \lnot} \mathfrak M$ of type theory
 with one universe (using results in

### 2016-01-19 paper/paper_acm.tex
correction of the translation
transport -> J
pass on conclusion

@@
                                   \right)\\[0.5em]
-\left[  \lambda\, x:A,~B \right] &\defeq&\lambda\,x:\Lbrack A
-                                     \Rbrack,~\Lbrack B\Rbrack
+\left[  \lambda\, x:A,~M \right] &\defeq&\lambda\,x:\Lbrack A
+                                     \Rbrack,~[ B ]
   \\[0.5em]
@@
 \begin{array}{lcl}
-\left[  A=B \right] &\defeq& \left( \Lbrack A \Rbrack = \Lbrack B
-                             \Rbrack , \pi_=^{[A],[B]}
+\left[  A=B \right] &\defeq& \left( [A] = [ B] , \pi_=^{[A],[B]}
                              \right)\\[0.5em]
@@
 $\Lbrack U \Rbrack\{x/[N]\}$ are equal, but not convertible. We
-therefore have to use $\mathrm{transport}$, the eliminator of the
+therefore have to use $J$, the eliminator of the
 equality, in the conversion rule. This requires to use a type theory
@@
 with new principles, leveraging Lawvere-Tierney sheafification to
-higher topos theory. 
+homotopy type theory. 
 %
@@
 (ii) 
-  We would like to have much more examples of left-exact modalities on
+  We would like to have more examples of left-exact modalities on
   $\HProp$, in order to extend type theory with different principles

### 2016-04-05 paper/paper_lipics.tex
Lipics

--- 2016-01-19 paper/paper_acm.tex
@@
-%-----------------------------------------------------------------------------
-%
-%               Template for sigplanconf LaTeX Class
-%
-% Name:         sigplanconf-template.tex
-%
-% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
-%               file for SIGPLAN conference proceedings.
-%
-% Guide:        Refer to "Author's Guide to the ACM SIGPLAN Class,"
-%               sigplanconf-guide.pdf
-%
-% Author:       Paul C. Anagnostopoulos
-%               Windfall Software
-%               978 371-2316
-%               paul@windfall.com
-%
-% Created:      15 February 2005
-%
-%-----------------------------------------------------------------------------
-
-\documentclass[preprint,9pt,numbers]{sigplanconf}
-
-% The following \documentclass options may be useful:
-
-% preprint      Remove this option only once the paper is in final form.
-% 10pt          To set in 10-point type instead of 9-point.
-% 11pt          To set in 11-point type instead of 9-point.
-% numbers       To obtain numeric citation style instead of author/year.
-
+\documentclass[a4paper,UKenglish]{lipics-v2016}
+%This is a template for producing LIPIcs articles. 
+%See lipics-manual.pdf for further information.
+%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
+%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
+% for section-numbered lemmas etc., use "numberwithinsect"
 \usepackage{amsmath}
@@
 \usepackage{xspace}
-\usepackage[colorlinks=true]{hyperref}
+%\usepackage[colorlinks=true]{hyperref}
 
@@
 \usepackage{mathbbol}
+\usepackage{microtype}%if unwanted, comment out or use option "draft"
+
+%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory
+
+\bibliographystyle{plainurl}% the recommended bibstyle
+
+% Author macros::begin %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\title{A Sample Article for the LIPIcs series\footnote{This work was partially supported by someone.}}
+%\titlerunning{A Sample LIPIcs Article} %optional, in case that the title is too long; the running title should fit into the top page column
+\title{Lawvere-Tierney Sheafification in Homotopy Type Theory}
+
+%% Please provide for each author the \author and \affil macro, even when authors have the same affiliation, i.e. for each author there needs to be the  \author and \affil macros
+\author[1]{Kevin Quirin}
+\author[2]{Nicolas Tabareau}
+\affil[1]{École des Mines de Nantes, France\\
+  \texttt{kevin.quirin@mines-nantes.fr}}
+\affil[2]{Inria, France\\
+  \texttt{nicolas.tabareau@inria.fr}}
+\authorrunning{K. Quirin and N. Tabareau} %mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et. al.'
+
+\Copyright{Kevin Quirin and Nicolas Tabareau}
+%mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/
+
+\subjclass{F.4.1 Mathematical Logic and Formal Languages}% mandatory: Please choose ACM 1998 classifications from http://www.acm.org/about/class/ccs98-html .
+E.g., cite as "F.1.1 Models of Computation".
+\keywords{sheaf, homotopy type theory, modalities}% mandatory: Please provide 1-5 keywords
+% Author macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\EventEditors{John Q. Open and Joan R. Acces}
+\EventNoEds{2}
+\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
+\EventShortTitle{CVIT 2016}
+\EventAcronym{CVIT}
+\EventYear{2016}
+\EventDate{December 24--27, 2016}
+\EventLocation{Little Whinging, United Kingdom}
+\EventLogo{}
+\SeriesVolume{42}
+\ArticleNo{23}
+% Editor-only macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
-
 % correct bad hyphenation here
@@
 \usepackage{wasysym}
-%usepackage{microtype}
-% \usepackage[subtle]{savetrees}
-
 \usepackage[all]{xy}
@@
 
-\special{papersize=8.5in,11in}
-\setlength{\pdfpageheight}{\paperheight}
-\setlength{\pdfpagewidth}{\paperwidth}
-
-\conferenceinfo{LICS '16}{July 5--8, 2016, New York City, USA}
-\copyrightyear{2016}
-\copyrightdata{978-1-nnnn-nnnn-n/yy/mm}
-\copyrightdoi{nnnnnnn.nnnnnnn}
-
-% Uncomment the publication rights you want to use.
-%\publicationrights{transferred}
-%\publicationrights{licensed}     % this is the default
-%\publicationrights{author-pays}
-
-\titlebanner{}        % These are ignored unless
-\preprintfooter{}   % 'preprint' option specified.
-
-\title{Lawvere-Tierney Sheafification\\ in Homotopy Type Theory}
-% \subtitle{Subtitle Text, if any}
-
-\authorinfo{Kevin Quirin}
-           {\'Ecole des Mines de Nantes}
-           {kevin.quirin@mines-nantes.fr}
-\authorinfo{Nicolas Tabareau}
-           {Inria}
-           {nicolas.tabareau@inria.fr}
-
 \maketitle
@@
 
-\category{F}{4}{1}
+%\category{F.4.1}{MATHEMATICAL LOGIC AND FORMAL LANGUAGES}{Mathematical Logic}
+% \category{F}{4}{1}
 
@@
 
-\keywords
-sheaf, homotopy type theory, modalities
+% \keywords
+% sheaf, homotopy type theory, modalities
 
@@
 Our inductive definition of these modalities will be given by
-\[ \begin{array}{l}
+\begin{mymath} \begin{array}{l}
    \modal : \forall \ (n : nat), \ \Type_n \to \Type_n 
@@
 % \vspace{-0.3em}
-\]
+\end{mymath}
 where $\Type_n^\modal$ is the type of modal types for $\modal_n$.
@@
 an $n$-type is simply a type
-for which equality becomes trivial up to level $n$. 
+for which equality becomes trivial above level $n$. 
 %
@@
 
-Another example of a left exact modality for every level $n$ is the
+Another example of a modality for every level $n$ is the
 double negation modality,
@@
 It follows from homotopy fibers preservation that any $n$-subobject of
-a type seen as a $n$-subobject of its closure is closed.
+a type seen as a $n$-subobject of its closure is dense.
 
@@
 %
-Let $A:\Type_{n+1}$ be a sheaf and $B:A \to \Type_{n+1}$ be a sheaf
-family. We want to show that $\sum_{x:A} (Bx)$ is separated. Let $E$
+Let $A:\Type_{n+1}$ be a separated type and $B:A \to \Type_{n+1}$ be a
+family of separated types. We want to show that $\sum_{x:A} (Bx)$ is separated. Let $E$
 be a type, and $\sum_{e:E} (\chi\,e)$ a dense $n$-subobject of E.
@@
 \begin{mymath}\xymatrix @R=4em @C=4em{
-  \sum_{e:E} (\chi\, e) \ar@<-2pt>[r]_{g\circ\pi_1} \ar@<2pt>[r]^{f\circ \pi_1} \ar[d]_{\mathrm{dense}}& \sum_{x :A} (Bx) \\
+  \sum_{e:E} (\chi\, e) \ar@<-2pt>[r]_{g\circ\pi_1} \ar@<2pt>[r]^{f\circ \pi_1} \ar[d]_{\mathrm{dense}}& \sum_{x :A} (B\,x) \\
   E \ar@<-2pt>[ru]_{g} \ar@<2pt>[ru]^{f}&
@@
 \begin{mymath}\xymatrix @R=4em @C=5em{
-  \sum_{e:E} (\chi\, e) \ar@<-2pt>[r]_{\pi_1\circ g\circ\pi_1} \ar@<2pt>[r]^{\pi_1\circ f\circ \pi_1} \ar[d]_{\mathrm{dense}}& \sum_{x :A} (Bx) \\
+  \sum_{e:E} (\chi\, e) \ar@<-2pt>[r]_{\pi_1\circ g\circ\pi_1} \ar@<2pt>[r]^{\pi_1\circ f\circ \pi_1} \ar[d]_{\mathrm{dense}}& A \\
   E \ar@<-2pt>[ru]_{\pi_1\circ g} \ar@<2pt>[ru]^{\pi_1\circ f}&
@@
   any type $A,B:\Type_n^\modal$, if $\varphi$ is the canonical arrow
-  $$A = B \to A\simeq B,$$
+  \begin{mymath}A = B \to A\simeq B,\end{mymath}
   then $\IsEquiv(\varphi)$ is modal.
@@
 \item For types (assuming $A \neq \Type$) 
-\[
+\begin{mymath}
 \begin{array}{lcllcl}
@@
 \end{array}
-\]
+\end{mymath}
 
 \item For dependent sums
-\[
+\begin{mymath}
 \begin{array}{lcl}
@@
                                   \Lbrack B\Rbrack ; \pi_\Sigma^{[A],[B]}
-                                \right)\\[0.5em]
-  \left[  (x;y)\right] &\defeq& ([x];[y]) \\[0.5em]
-  \left[  \pi_i t\right] &\defeq& \pi_i [t] \\[0.5em]
+                                \right)\\begin{mymath}0.5em]
+  \left[  (x;y)\right] &\defeq& ([x];[y]) \\begin{mymath}0.5em]
+  \left[  \pi_i t\right] &\defeq& \pi_i [t] \\begin{mymath}0.5em]
 \end{array}
-\]
+\end{mymath}
 \item For dependent products
-\[
+\begin{mymath}
 \begin{array}{lcl}
@@
                                    \Lbrack B\Rbrack ; \pi_{\Pi}^{[A],[B]}
-                                  \right)\\[0.5em]
+                                  \right)\\begin{mymath}0.5em]
 \left[  \lambda\, x:A,~M \right] &\defeq&\lambda\,x:\Lbrack A
-                                     \Rbrack,~[ B ]
-  \\[0.5em]
-  \left[ t \, t' \right] &\defeq& [t] [t'] \\[0.5em]
+                                     \Rbrack,~[ M ]
+  \\begin{mymath}0.5em]
+  \left[ t \, t' \right] &\defeq& [t] [t'] \\begin{mymath}0.5em]
 \end{array}
-\]
+\end{mymath}
 \item For paths
-\[
+\begin{mymath}
 \begin{array}{lcl}
 \left[  A=B \right] &\defeq& \left( [A] = [ B] , \pi_=^{[A],[B]}
-                             \right)\\[0.5em]
-\left[ 1 \right] &\defeq& 1\\[0.5em]
-\left[ J \right] &\defeq& J \\[0.5em]
+                             \right)\\begin{mymath}0.5em]
+\left[ 1 \right] &\defeq& 1\\begin{mymath}0.5em]
+\left[ J \right] &\defeq& J \\begin{mymath}0.5em]
 \end{array}
-\]
+\end{mymath}
 \item For sums
-\[
+\begin{mymath}
 \begin{array}{lcl}
@@
                         \Rbrack); \pi_\modal(\Lbrack A \Rbrack + \Lbrack B
-                        \Rbrack)\right)\\[0.5em]
-\left[  \mathrm{in}_\ell t \right] &\defeq& \eta (\mathrm{in}_\ell [t]) \\[0.5em]
-\left[  \mathrm{in}_r t \right] &\defeq& \eta (\mathrm{in}_r [t]) \\[0.5em]
+                        \Rbrack)\right)\\begin{mymath}0.5em]
+\left[  \mathrm{in}_\ell t \right] &\defeq& \eta (\mathrm{in}_\ell [t]) \\begin{mymath}0.5em]
+\left[  \mathrm{in}_r t \right] &\defeq& \eta (\mathrm{in}_r [t]) \\begin{mymath}0.5em]
 \left[ \langle f ,g\rangle\right] &\defeq& \modal_{\mathrm{rec}}^{\Lbrack A\Rbrack +
                                       \Lbrack B\Rbrack} \langle
-                                      [f],[g]\rangle\\[0.5em]
+                                      [f],[g]\rangle\\begin{mymath}0.5em]
 \end{array}
-\]
+\end{mymath}
 \item For truncations ($i\leqslant n$)
-\[
+\begin{mymath}
 \begin{array}{lcl}
@@
                                    \pi_\modal(\| \Lbrack A\Rbrack
-                                   \|_i)) \\[0.5em]
-  \left[ |t|_i \right] &\defeq& \eta |[t]|_i \\[0.5em]
+                                   \|_i)) \\begin{mymath}0.5em]
+  \left[ |t|_i \right] &\defeq& \eta |[t]|_i \\begin{mymath}0.5em]
   \left[ \|f\|_i \right] &\defeq& \modal_{\mathrm{rec}}^{\| \Lbrack
@@
 \end{array}
-\]
+\end{mymath}
 
@@
 $\mathop{EM}$ by defining 
-$$
+\[
 [\mathop{EM}] \defeq \lambda P, \lambda H : \lnot (P + \lnot P), H (\mathrm{in}_r (\lambda
 p : P, H (\mathrm{in}_\ell p)))
-$$
+\]
 of type 
-$$
+\begin{mymath}
 \Lbrack \forall P:\HProp, P + \lnot P \Rbrack \equiv \forall
 P:\Type^\modal_{-1}. \lnot (\lnot (P + \lnot P)).
-$$
+\end{mymath}
 where we have used an implicit rewriting with
@@
 % \vspace{-0.05em}
-\[ \begin{array}{l}
+\begin{mymath} \begin{array}{l}
    \modal : \forall \ (n : nat), \ \Type_n \to \Type_n 
@@
 % \vspace{-0.3em}
-\]
+\end{mymath}
 %
@@
 
-\bibliographystyle{abbrvnat}
+%\bibliographystyle{abbrvnat}
 \bibliography{sheaf_lics}
@@
 \end{document}
+
+% \begin{document}
+
+% \maketitle
+
+% \begin{abstract}
+% Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Praesent convallis orci arcu, eu mollis dolor.
+Aliquam eleifend suscipit lacinia.
+Maecenas quam mi, porta ut lacinia sed, convallis ac dui.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Suspendisse potenti.
+%  \end{abstract}
+
+% \section{Typesetting instructions -- please read carefully}
+% Please comply with the following instructions when preparing your article for a LIPIcs proceedings volume. 
+% \begin{itemize}
+% \item Use pdflatex and an up-to-date LaTeX system.
+% \item Use further LaTeX packages only if required. Avoid usage of packages like \verb+enumitem+, \verb+enumerate+, \verb+cleverref+. Keep it simple, i.e. use as few additional packages as possible.
+% \item Add custom made macros carefully and only those which are needed in the article (i.e., do not simply add your convolute of macros collected over the years).
+% \item Do not use a different main font. For example, the usage of the \verb+times+-package is forbidden.
+% \item Provide full author names (especially with regard to the first name) in the \verb+\author+ macro and in the \verb+\Copyright+ macro.
+% \item Fill out the \verb+\subjclass+ and \verb+\keywords+ macros. For the \verb+\subjclass+, please refer to the ACM classification at \url{http://www.acm.org/about/class/ccs98-html}.
+% \item Take care of suitable linebreaks and pagebreaks. No overfull \verb+\hboxes+ should occur in the warnings log.
+% \item Provide suitable graphics of at least 300dpi (preferrably in pdf format).
+% \item Use the provided sectioning macros: \verb+\section+, \verb+\subsection+, \verb+\subsection*+, \verb+\paragraph+, \verb+\subparagraph*+, ...
+``Self-made'' sectioning commands (for example, \verb+\noindent{\bf My+ \verb+subparagraph.}+ will be removed and replaced by standard LIPIcs style sectioning commands.
+% \item Do not alter the spacing of the  \verb+lipics-v2016.cls+ style file. Such modifications will be removed.
+% \item Do not use conditional structures to include/exclude content. Instead, please provide only the content that should be published -- in one file -- and nothing else.
+% \item Remove all comments, especially avoid commenting large text blocks and using \verb+\iffalse+ $\ldots$ \verb+\fi+ constructions.
+% \item Keep the standard style (\verb+plainurl+) for the bibliography as provided by the\linebreak \verb+lipics-v2016.cls+ style file.
+% \item Use BibTex and provide exactly one BibTex file for your article.
+The BibTex file should contain only entries that are referenced in the article.
+Please make sure that there are no errors and warnings with the referenced BibTex entries.
+% \item Use a spellchecker to get rid of typos.
+% \item A manual for the LIPIcs style is available at \url{http://drops.dagstuhl.de/styles/lipics/lipics-v2016-authors/lipics-v2016-manual.pdf}.
+% \end{itemize}
+
+% \section{Lorem ipsum dolor sit amet}
+
+% Lorem ipsum dolor sit amet, consectetur adipiscing elit \cite{DBLP:journals/cacm/Knuth74}.
+Praesent convallis orci arcu, eu mollis dolor.
+Aliquam eleifend suscipit lacinia.
+Maecenas quam mi, porta ut lacinia sed, convallis ac dui.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Suspendisse potenti.
+Donec eget odio et magna ullamcorper vehicula ut vitae libero.
+Maecenas lectus nulla, auctor nec varius ac, ultricies et turpis.
+Pellentesque id ante erat.
+In hac habitasse platea dictumst.
+Curabitur a scelerisque odio.
+Pellentesque elit risus, posuere quis elementum at, pellentesque ut diam.
+Quisque aliquam libero id mi imperdiet quis convallis turpis eleifend.
+
+% \begin{lemma}[Lorem ipsum]
+% \label{lemma:lorem}
+% Vestibulum sodales dolor et dui cursus iaculis.
+Nullam ullamcorper purus vel turpis lobortis eu tempus lorem semper.
+Proin facilisis gravida rutrum.
+Etiam sed sollicitudin lorem.
+Proin pellentesque risus at elit hendrerit pharetra.
+Integer at turpis varius libero rhoncus fermentum vitae vitae metus.
+% \end{lemma}
+
+% \begin{proof}
+% Cras purus lorem, pulvinar et fermentum sagittis, suscipit quis magna.
+% \end{proof}
+
+% \begin{theorem}[Curabitur pulvinar, \cite{DBLP:books/mk/GrayR93}]
+% \label{theorem:curabitur}
+% Nam liber tempor cum soluta nobis eleifend option congue nihil imperdiet doming id quod mazim placerat facer possim assum.
+Lorem ipsum dolor sit amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna aliquam erat volutpat.
+% \end{theorem}
+
+% \subsection{Curabitur dictum felis id sapien}
+
+% Curabitur dictum felis id sapien mollis ut venenatis tortor feugiat.
+Curabitur sed velit diam.
+Integer aliquam, nunc ac egestas lacinia, nibh est vehicula nibh, ac auctor velit tellus non arcu.
+Vestibulum lacinia ipsum vitae nisi ultrices eget gravida turpis laoreet.
+Duis rutrum dapibus ornare.
+Nulla vehicula vulputate iaculis.
+Proin a consequat neque.
+Donec ut rutrum urna.
+Morbi scelerisque turpis sed elit sagittis eu scelerisque quam condimentum.
+Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
+Aenean nec faucibus leo.
+Cras ut nisl odio, non tincidunt lorem.
+Integer purus ligula, venenatis et convallis lacinia, scelerisque at erat.
+Fusce risus libero, convallis at fermentum in, dignissim sed sem.
+Ut dapibus orci vitae nisl viverra nec adipiscing tortor condimentum \cite{DBLP:journals/cacm/Dijkstra68a}.
+Donec non suscipit lorem.
+Nam sit amet enim vitae nisl accumsan pretium.
+
+% \begin{lstlisting}[caption={Useless code},label=list:8-6,captionpos=t,float,abovecaptionskip=-\medskipamount]
+% for i:=maxint to 0 do 
+% begin 
+%     j:=square(root(i));
+% end;
+% \end{lstlisting}
+
+% \subsection{Proin ac fermentum augue}
+
+% Proin ac fermentum augue.
+Nullam bibendum enim sollicitudin tellus egestas lacinia euismod orci mollis.
+Nulla facilisi.
+Vivamus volutpat venenatis sapien, vitae feugiat arcu fringilla ac.
+Mauris sapien tortor, sagittis eget auctor at, vulputate pharetra magna.
+Sed congue, dui nec vulputate convallis, sem nunc adipiscing dui, vel venenatis mauris sem in dui.
+Praesent a pretium quam.
+Mauris non mauris sit amet eros rutrum aliquam id ut sapien.
+Nulla aliquet fringilla sagittis.
+Pellentesque eu metus posuere nunc tincidunt dignissim in tempor dolor.
+Nulla cursus aliquet enim.
+Cras sapien risus, accumsan eu cursus ut, commodo vel velit.
+Praesent aliquet consectetur ligula, vitae iaculis ligula interdum vel.
+Integer faucibus faucibus felis.
+
+% \begin{itemize}
+% \item Ut vitae diam augue. 
+% \item Integer lacus ante, pellentesque sed sollicitudin et, pulvinar adipiscing sem. 
+% \item Maecenas facilisis, leo quis tincidunt egestas, magna ipsum condimentum orci, vitae facilisis nibh turpis et elit. 
+% \end{itemize}
+
+% \section{Pellentesque quis tortor}
+
+% Nec urna malesuada sollicitudin.
+Nulla facilisi.
+Vivamus aliquam tempus ligula eget ornare.
+Praesent eget magna ut turpis mattis cursus.
+Aliquam vel condimentum orci.
+Nunc congue, libero in gravida convallis \cite{DBLP:conf/focs/HopcroftPV75}, orci nibh sodales quam, id egestas felis mi nec nisi.
+Suspendisse tincidunt, est ac vestibulum posuere, justo odio bibendum urna, rutrum bibendum dolor sem nec tellus.
+
+% \begin{lemma} [Quisque blandit tempus nunc]
+% Sed interdum nisl pretium non.
+Mauris sodales consequat risus vel consectetur.
+Aliquam erat volutpat.
+Nunc sed sapien ligula.
+Proin faucibus sapien luctus nisl feugiat convallis faucibus elit cursus.
+Nunc vestibulum nunc ac massa pretium pharetra.
+Nulla facilisis turpis id augue venenatis blandit.
+Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus.
+% \end{lemma}
+
+% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
+
+% \subparagraph*{Acknowledgements.}
+
+% I want to thank \dots
+
+% \appendix
+% \section{Morbi eros magna}
+
+% Morbi eros magna, vestibulum non posuere non, porta eu quam.
+Maecenas vitae orci risus, eget imperdiet mauris.
+Donec massa mauris, pellentesque vel lobortis eu, molestie ac turpis.
+Sed condimentum convallis dolor, a dignissim est ultrices eu.
+Donec consectetur volutpat eros, et ornare dui ultricies id.
+Vivamus eu augue eget dolor euismod ultrices et sit amet nisi.
+Vivamus malesuada leo ac leo ullamcorper tempor.
+Donec justo mi, tempor vitae aliquet non, faucibus eu lacus.
+Donec dictum gravida neque, non porta turpis imperdiet eget.
+Curabitur quis euismod ligula.
+
+% %%
+% %% Bibliography
+% %%
+
+% %% Either use bibtex (recommended),
+
+% \bibliography{lipics-v2016-sample-article}
+
+% %% .. or use the thebibliography environment explicitely
+
+% \end{document}

### 2016-04-05 paper/paper_lipics_shorten.tex
shorten version

--- 2016-04-05 paper/paper_lipics.tex
@@
   \vspace{#1}%
-  \begin{equation*}%
+  $%
   }{ %
@@
     % \vspace{#1}
-  \end{equation*}}
+  $ }
 
@@
 \end{prop}
-\begin{proof}
-  By condition (iv) of Definition~\ref{sec:defin-basic-prop-1},
-  $\modal \zero$ is an initial object of $\Type^\modal_n$, and thus
-  corresponds to false for modal mere proposition.
-  % 
-  As $\modal \one = \one$, $\Type^\modal_n$ is consistent when
-  $\modal \zero \neq \one$, that is when there is no proof of
-  $\modal \zero$.
-\end{proof}
 
@@
 \end{prop}
-\begin{proof}
-  First, we prove that $\Type_n^\modal$ is separated. Let $E:\Type$
-  and $\chi:E \to \Type_n$, dense in $E$. Let $\phi_1,\phi_2:E \to
-  \Type_n^\modal$, such that $\phi_1 \circ \pi_1 = \phi_2 \circ
-  \pi_1$ and let $x:E$. We show $\phi_1(x) = \phi_2(x)$ using
-  univalence.
-
-  As $\chi$ is dense, we have a term $m_x : \modal_n(\chi\, x)$.
-  But as $\phi_2(x)$ is modal, we can obtain a term $h_x : \chi\,
-  x$. 
-  As $\phi_1$ and $\phi_2$ are equal on $\sum_{e:E}\chi\, e$, we
-  have an arrow $\phi_1(x) \to \phi_2(x)$.
-  The same method leads to an arrow $\phi_2 (x) \to \phi_1 (x)$, and we
-  prove they are each other inverse.
-
-  Now, we prove that $\Type_n^\modal$ is a sheaf. Let $E:\Type$ and
-  $\chi:E \to \HProp$, dense in $E$. Let $f:\sum_{e:E} \chi\, e \to
-  \Type_n^\modal$. We want to extend $f$ to $E \to \Type_n^\modal$.
-
-  We define $g$ as $g(e) = \modal_n \left( \fib \phi {e} \right)$,
-  where
-  \begin{mymath}[0em]\phi : \sum_{b:\left(\sum_{e:E} (\chi\, e)\right)} (f\,
-    b) \to E\end{mymath}%
-  defined by $\phi(x) = (x_1)_1$.
-% \kq{Is it necessary to state the lemma ?}
-Using a modal analogous of the characterization of fibers of composition, we can prove that it
-defines an equivalence.
-  % \begin{lem}
-  %   Let $A,B,C:\Type_n$, $f:A\to B$ and $g:B\to C$.
-  %   Then
-  %   \begin{itemize}
-  %   \item If $g$ is an embedding, then 
-  %     \begin{mymath}\prod_{b:B}\fib f b = \fib {g\circ f} {g\, b}.\end{mymath}
-  %   \item If all fibers of $f$ and $g$ are $n$-truncated, then
-  %     \begin{mymath}\prod_{c:C}\modal_n(\fib {g \circ f} c) =
-  %     \modal_n \left(  
-  %       \sum_{w:\fib g c} \modal_n (\fib f {w_1})
-  %     \right).\end{mymath}
-  %   \end{itemize}
-  % \end{lem}
-\end{proof}
 % \vspace{-1em}
@@
 \end{prop}
-\begin{proof}
-  \begin{itemize}
-  \item {\em Separation:} Let $E:\Type$, $\chi:E \to \Type_n$ dense
-    and $\phi_1,\phi_2:E\to \prod_{a:A} (B\, a)$ equal on
-    $\sum_{e:E}(\chi\, e)$ \ie $\phi_1\circ \pi_1 = \phi_2\circ
-    \pi_1$.
-    Then for any $a:A$, $(\lambda x:E,~\phi_1(x, a))$
-    and $(\lambda x:E,~\phi_2(x,a))$
-    coincide on $\sum_{e:E}(\chi\, e)$, and as $B\, a$ is separated,
-    they coincide also on all $E$.
-  \item {\em Sheaf:} Let $E:\Type$, $\chi:E\to \HProp$ dense and
-    $f:\sum_{e:E} (\chi\, e)\to \prod_{a:A} (B\,a)$. Let $a:A$ ; the
-    map $(\lambda x,~f(x,a))$ is valued in the sheaf $B\, a$, so it
-    can be extended to all $E$, allowing $f$ to be extended to all
-    $E$. \qedhere
-  \end{itemize}
-%  \nt{add a short proof here}
-\end{proof}
 
@@
 
-\begin{proof}
-We use the following lemma:
-\begin{lem}
-  A $(n+1)$-truncated type $T$ with an embedding $f : T \to U$
-  into a separated $(n+1)$-truncated type $U$ is itself separated.
-\end{lem}
-As $\separated T$ embeds in $\left( \Type_n^\modal \right)^T$, we only
-have to show that the latter is separated. But it is the case because
-$\Type_n^\modal$ is a sheaf (by Proposition~\ref{prop:sheaf-is-sheaf})
-and a function type is a sheaf as soon
-as its codomain is a sheaf (by Proposition~\ref{prop:sheaf-forall}).
-\end{proof}
-
 To show that $\separated$ is a modality, we need to show that $\separated T$
@@
 \end{lem}
-
-\begin{proof}[Sketch of proof]
-  By induction on truncation, we need to show that 
-  \begin{mymath}\forall a,b:T_f,~\tilde f (|a|_{n+1}) = \tilde f (|b|_{n+1} )\simeq
-  \modal (|a|_{n+1}=|b|_{n+1}).\end{mymath}%
-  We use the encode-decode~\cite[Section 8.9]{hottbook} method to characterize $\tilde f (|a|_{n+1})
-  = x$, and the result follows. We refer to the formalization for details.
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:sepiscolim}]
-  As $\mu_T$ is an surjection, using the theorem~\ref{cech} introduced in Section~\ref{ssec:giraud-ax},
-  we know that $\separated T$ is the $(n+1)$-colimit of $\|C(\mu_T)\|_{n+1}$. 
-  % 
-  Thus the result follows from the fact that $\|C(\mu_T)\|_{n+1} =
-  \mathring T$,
-  which we now show.
-
-  First, we show that $\mu_T$ satisfies condition~(\ref{eq:Omono}),
-  \ie{} $\forall a,b:T$, $\modal_n (a=b) = (\mu_T a =
-  \mu_T b)$. By univalence, we want arrows in both ways, forming an
-  equivalence.
-  \begin{itemize}
-  \item Suppose $p : (\mu_T a = \mu_T b)$. Then projecting $p$ along
-    first components yields $q : \prod_{t:T} \modal_n(a=t) = \modal_n (b=t)
-    $.
-    Taking for example $t=b$, we deduce $\modal_n (a=b) = \modal_n(b=b)$,
-    and the latter is inhabited by $\eta_{b=b} 1$.
-  \item Suppose now $p : \modal_n(a=b)$. Let $\iota$ be the first
-    projection from $\separated T \to (T \to \Type_n^\modal)$. $\iota$ is
-    an embedding, thus it suffices to prove $\iota (\mu_T a) = \iota
-    (\mu_T b)$, \ie $\prod_{t:T}\modal_n (a=t) = \modal_n (b=t)$. The latter
-    remains true by univalence.
-  \end{itemize}
-  The fact that these two form an equivalence is technical, we refer to
-  the formalization for an explicit proof.
-
-  Then, by induction, we need to show that for any $A,B:\Type_{n+1}$ such
-  that $\varphi:~A \simeq B$, with an arrow $f:A \to \separated T$
-  satisfying~(\ref{eq:Omono}), 
-  \begin{mymath}T_f \simeq \mathring T_B.\end{mymath}%
-  For this, we use 
-  \begin{align*}
-    T_f &= \mathrm{Coeq}_1 \left( 
-          \xymatrix{
-          \sum_{a,b:A} f a = f b \dar[r]{4pt}^-{\pi_1}_-{\pi_2} & A
-  }
-                                                                  \right)
-    \\
-    \mathring T_B &= \mathrm{Coeq}_1 \left( 
-                    \xymatrix{\sum_{a,b:B} \modal (a=b) \dar[r]{4pt}^-{\pi_1}_-{\pi_2}
-                                                                & B}
-                                                                  \right)
-  \end{align*}
-
-  By~(\ref{eq:Omono}), $\sum_{a,b:A} f a = f b \simeq \sum_{a,b:A} \modal
-  (a=b)$, and by $\varphi$, $\sum_{a,b:A} f a = f b \simeq
-  \sum_{a,b:B} \modal ( a = b)$.
-
-Thus, $T_f \simeq \mathring T$.
-\end{proof}
 
@@
 \end{prop}
-\begin{proof}
-  Let $T,Q:\Type_{n+1}$ such that $Q$ is a sheaf. Let $f:T\to Q$.
-  Because $Q$ is a sheaf, it is in particular separated;
-  % 
-  thus we can extend $f$ to $\separated f:\separated T\to Q$.
-
-  But as $\modal_{n+1} T$ is the closure of $\separated T$, $\separated T$ is dense
-  into $\modal_{n+1} T$, so the sheaf property of $Q$ allows to extend
-  $\separated f$ to $\modal_{n+1} f:\modal_{n+1} T \to Q$.
-
-  As all these steps are universal, the composition is.
-\end{proof}
 
@@
 \end{prop}
-\begin{proof}
-  We prove it by induction on $n$:
-  \begin{itemize}
-  \item For $n=-1$: Let $T:\HProp$. Then
-    \begin{mymath}
-      \modal_{0} \widehat T = \sum_{u:T \to \Type_n^\modal} \modal_{-1}\left( \sum_{a:T} 
-      (\lambda t,~\modal_{-1} (a=t)) = u\right)
-    \end{mymath}%
-    because the type inside the truncation is already in $\HProp$.
-    Now, let define $\phi : \modal_{-1} T \to \modal_0T$ by
-    \begin{mymath}\phi t = (\lambda t',\, \one
-      ;\kappa)\end{mymath}%
-    where $\kappa$ is defined by $\modal_{-1}$-induction on
-    $t$. Indeed, as $T$ is an $\HProp$, $(a=t) \simeq \one$. 
-    Let $\psi : \modal_0T\to \modal_{-1} T$ by obtaining the
-    witness $a:T$ (which is possible because we are trying to inhabit
-    a modal proposition), and letting $\psi (u;x) = \eta_T a$.
-    These two maps form an equivalence (the section and retraction are
-    trivial because the equivalence is between mere propositions).
-  \item Suppose now that $\modal_{n+1}$ is compatible with all $\modal_k$ on
-    lower homotopy types. Let $\modal_{n+2}$ be as above, and let
-    $T:\Type_{n+1}$. Then, as $\modal_{n+1}$ is compatible with $\modal_{n}$, and
-    $(a=t)$ is in $\Type_n$,
-    \begin{mymath}
-      \modal_{n+2} \widehat T= \hspace{-1em} \sum_{u:T \to
-        \Type_{n+1}^\modal} 
-      \hspace{-1em} \modal_{-1}\left\| \sum_{a:T} 
-        (\lambda t,~\modal_{n} (a=t)) = u\right\|.
-    \end{mymath}%
-    It remains to prove that for every $(u,x)$ inhabiting the
-    $\Sigma$-type above, $u$ is in $T\to\Type_n^\modal$, \ie that for
-    every $t:T$, $\IsType n (u\, t)$.  But for any truncation index
-    $p$,
-    the type $\IsType p X:\HProp$ is a sheaf as soon as $X$ is, so we can get rid
-    of $\modal_{-1}$ and of the truncation, which tells us that for
-    every 
-    $t:T$, $u\, t = \modal_n(a=t) : \Type_n$. \qedhere
-  \end{itemize}
-\end{proof}
+
 This proves in particular that $\modal_{n+1}$ is compatible with
@@
 \end{lem}
-\begin{proof}
-  For $i$-truncated types, $\modal = \modal_i$, and $\modal_i$ is left-exact.
-\end{proof}
 
@@
 \end{prop}
-\begin{proof}
-  The first thing to notice is that, if $X$ and $Y$ are modal, and
-  $f:X \to Y$, then the mere proposition $\IsEquiv f$ is also modal.
-  Therefore, it suffices to show that both $A=B$ and $A\simeq B$ are
-  modal. By proposition~\ref{prop:mod_prop}, $A=B$ is modal. 
-  Moreover, $(A\simeq B) \simeq \sum_{f:A\to B} \IsEquiv
-  f$. Therefore, as $A$ and $B$ are modal, $A\simeq B$ is too.
-
-  Hence, $\IsEquiv \varphi$ is modal.
-\end{proof}
 
@@
 % \text{«\,The $n$-th Turing Machine halts\,»}$.
-
-\section{Translation}
-\label{sec:translation}
-
-In this section, we present how to use the sheafification modality to
-define a transformation of Martin-Löf type theory into
-itself\footnote{We define the translation on Martin-Löf type theory,
-  but it could also be done for CIC as well.} as it has been done for
-forcing~\cite{jaber2012extending}.
-
-Let $n$ be a trunction index, and let $\modal = \modal_n \circ
-\|\cdot\|_n$ the constructed modality. 
-%
-As this modality is not left exact, $\Type$ itself is not modal and we
-can only define the translation on a type theory with only one
-universe.
-%
-To explain the translation on positive types as well, we also
-translate the sum type $A+B$, with associated constructors
-$\mathrm{in}_\ell$
-and $\mathrm{in}_r$ and eliminator $\langle \_ , \_ \rangle$.
-%
-We note $J$ the eliminator of the identity type. 
-%
-%
-\begin{itemize}
-\item For types (assuming $A \neq \Type$) 
-\begin{mymath}
-\begin{array}{lcllcl}
-  \Lbrack \Type\Rbrack &\defeq& \Type^\modal & \qquad 
-  \Lbrack A \Rbrack &\defeq& \pi_1 \left[ A \right]
-\end{array}
-\end{mymath}
-
-\item For dependent sums
-\begin{mymath}
-\begin{array}{lcl}
-\left[ \sum_{x:A} B \right] &\defeq&  \left( \sum_{x:\Lbrack A \Rbrack}
-                                  \Lbrack B\Rbrack ; \pi_\Sigma^{[A],[B]}
-                                \right)\\begin{mymath}0.5em]
-  \left[  (x;y)\right] &\defeq& ([x];[y]) \\begin{mymath}0.5em]
-  \left[  \pi_i t\right] &\defeq& \pi_i [t] \\begin{mymath}0.5em]
-\end{array}
-\end{mymath}
-\item For dependent products
-\begin{mymath}
-\begin{array}{lcl}
- \left[ \prod_{x:A} B \right] &\defeq& \left( \prod_{x:\Lbrack A \Rbrack}
-                                   \Lbrack B\Rbrack ; \pi_{\Pi}^{[A],[B]}
-                                  \right)\\begin{mymath}0.5em]
-\left[  \lambda\, x:A,~M \right] &\defeq&\lambda\,x:\Lbrack A
-                                     \Rbrack,~[ M ]
-  \\begin{mymath}0.5em]
-  \left[ t \, t' \right] &\defeq& [t] [t'] \\begin{mymath}0.5em]
-\end{array}
-\end{mymath}
-\item For paths
-\begin{mymath}
-\begin{array}{lcl}
-\left[  A=B \right] &\defeq& \left( [A] = [ B] , \pi_=^{[A],[B]}
-                             \right)\\begin{mymath}0.5em]
-\left[ 1 \right] &\defeq& 1\\begin{mymath}0.5em]
-\left[ J \right] &\defeq& J \\begin{mymath}0.5em]
-\end{array}
-\end{mymath}
-\item For sums
-\begin{mymath}
-\begin{array}{lcl}
-\left[  A+B \right] &\defeq& \left( \modal(\Lbrack A \Rbrack + \Lbrack B
-                        \Rbrack); \pi_\modal(\Lbrack A \Rbrack + \Lbrack B
-                        \Rbrack)\right)\\begin{mymath}0.5em]
-\left[  \mathrm{in}_\ell t \right] &\defeq& \eta (\mathrm{in}_\ell [t]) \\begin{mymath}0.5em]
-\left[  \mathrm{in}_r t \right] &\defeq& \eta (\mathrm{in}_r [t]) \\begin{mymath}0.5em]
-\left[ \langle f ,g\rangle\right] &\defeq& \modal_{\mathrm{rec}}^{\Lbrack A\Rbrack +
-                                      \Lbrack B\Rbrack} \langle
-                                      [f],[g]\rangle\\begin{mymath}0.5em]
-\end{array}
-\end{mymath}
-\item For truncations ($i\leqslant n$)
-\begin{mymath}
-\begin{array}{lcl}
-  \left[  \|A\|_i \right] &\defeq& (\modal \| \Lbrack A\Rbrack  \|_i;
-                                   \pi_\modal(\| \Lbrack A\Rbrack
-                                   \|_i)) \\begin{mymath}0.5em]
-  \left[ |t|_i \right] &\defeq& \eta |[t]|_i \\begin{mymath}0.5em]
-  \left[ \|f\|_i \right] &\defeq& \modal_{\mathrm{rec}}^{\| \Lbrack
-                                  A\Rbrack  \|_i} \| [f] \|_i
-\end{array}
-\end{mymath}
-
-\end{itemize}
-
-where $\pi_\Sigma^{A,B}$ (resp. $\pi_\Pi^{A,B}$, resp. $\pi_=^{A,B}$)
-is the proof that $\Sigma_{x:A} B$ (resp. $\prod_{x:A} B$, resp $A=B$)
-is modal as soon as $A$ and $B$ are, and $\pi_\modal^A$ a proof that
-$\modal(A)$ is always modal as described in section~\ref{sec:lexmod}.
-
-As in the forcing translation~\cite{jaber2012extending}, the main
-issue is that $\Lbrack U\{x/N\} \Rbrack$ and
-$\Lbrack U \Rbrack\{x/[N]\}$ are equal, but not convertible. We
-therefore have to use $J$, the eliminator of the
-equality, in the conversion rule. This requires to use a type theory
-with explicit conversion in the syntax. We have chosen to omit this
-technical details here to simplify the presentation.
-
-% The central theorem is
-% \begin{prop}
-%   Let $\Gamma$ be a valid context, $A$ a type and $t$ a term of type
-%   $A$.
-%   Then if $\Gamma \vdash t: A$, then $\Lbrack\Gamma\Rbrack \vdash [t]
-%   : \Lbrack A\Rbrack$.
-% \end{prop}
-
-Modulo this technical point, we have the usual property that if a term
-$t$ has type $A$ in context $\Gamma$, then $[t]$ has type $\Lbrack
-A\Rbrack$ in the translated context $\Lbrack \Gamma\Rbrack$.
-%
-Furthermore, depending the modality $\modal_{-1}$ under focus, we can
-add new constants in the type theory, and give them a meaning through
-the translation.
-
-For instance, for the double negation modality, we can add the term
-$\mathop{EM}$ by defining 
-\[
-[\mathop{EM}] \defeq \lambda P, \lambda H : \lnot (P + \lnot P), H (\mathrm{in}_r (\lambda
-p : P, H (\mathrm{in}_\ell p)))
-\]
-of type 
-\begin{mymath}
-\Lbrack \forall P:\HProp, P + \lnot P \Rbrack \equiv \forall
-P:\Type^\modal_{-1}. \lnot (\lnot (P + \lnot P)).
-\end{mymath}
-where we have used an implicit rewriting with
-Proposition~\ref{prop:compatibility}.
-
-Unfortunately, one lose canonicity when working in $\Type^\modal$;
-indeed, for example, the translation of the function 
-% \begin{mymath}\lambda P:\HProp,~\left\{\begin{array}{l} 0 \text{ if $P$ is
-%                                     false} \\ 1 \text{ if $P$ is
-%                                     true} \end{array}\right.~:~\HProp
-%                               \to \nat\end{mymath}%
-\begin{mymath}
-\lambda P:\HProp, \langle \lambda x,\,0 ,\lambda x,\,1\rangle (\mathop{EM} P)~:~\HProp
-                     \to \nat\end{mymath}%
-will not reduce to a numeral when applied to a given
-property. However, we keep canonicity for terms of Martin-Löf Type
-theory which do not make us of new constants nor positive types. 
-%
 

### 2016-04-05 paper/paper_lipics.tex
Reduce

@@
 says that for any types $T$ and $U$, the canonical map 
-\begin{mymath}
-(T = U) \to (T \simeq U)
-\end{mymath}%
+$(T = U) \to (T \simeq U)$
 %
@@
 %
-Section~\ref{sec:translation} defines the translation of Matin-Löf
+Section~\ref{ssec:translation} defines the translation of Matin-Löf
 type theory into itself based on this modality. 
@@
 type theory, we first briefly recall the corresponding construction in
-topos theory and then gives an overview of our construction.
-
+topos theory and then give an overview of our construction.
 From now on, {\em sheafification} is to be understood as {\em
@@
 %
-\begin{mymath}
+$
 j : \Omega \rightarrow \Omega
-\end{mymath}%
+$
 %
@@
 corresponds to the subobject of $E$ whose characteristic function is 
-%
-\begin{mymath}
+$
 \Char{\closure{A}} = j \circ \Char{A}.
-\end{mymath}%
-%
+$
 A subobject $A$ of $E$ is said to
@@
 sheaves, is again a topos, with classifying object
-%
-\begin{mymath}
+$
 \Omega_j = \{ P \in \Omega \ | \ j P  = P \}.
-\end{mymath}%
-%
+$
 Thus, in case of the double negation, the resulting topos is boolean  
@@
 sheafification functor
-%
-\begin{mymath}
+$
 a_j : \E \rightarrow \Sh{\E}
-\end{mymath}%
+$
 which exhibits $\Sh{\E}$ as a reflective
@@
 Our inductive definition of these modalities will be given by
-\begin{mymath} \begin{array}{l}
+\begin{equation}\label{eq:induction} \begin{array}{l}
    \modal : \forall \ (n : nat), \ \Type_n \to \Type_n 
@@
 % \vspace{-0.3em}
-\end{mymath}
+\end{equation}
 where $\Type_n^\modal$ is the type of modal types for $\modal_n$.
@@
 \item a map $q_i : D_0(i) \to P$ for any
-vertex $i:G_0$, \ie \begin{mymath}[0em]q : \prod_{i:G_0} D_0(i) \to P\end{mymath}%
+vertex $i:G_0$, \ie $q : \prod_{i:G_0} D_0(i) \to P$
 \item for any vertices $i,j:G_0$ and all edges $\phi:G_1(i,j)$, a path
   $p_{i,j}^\phi : q_j \circ D_1(\phi) = q_i$, \ie
-  \begin{mymath}[0em]p : \prod_{i,j:G_0} \prod_{\phi:G_1(i,j)} q_j \circ D_1(\phi) = q_i.\end{mymath}%
+  $p : \prod_{i,j:G_0} \prod_{\phi:G_1(i,j)} q_j \circ D_1(\phi) = q_i.$
 \end{itemize}
 
-Then $P$ is the {\em colimit} of $D$ if for any other $X:\Type$, the
-map
-\begin{mymath}\lambda f:P \to X, \left( \lambda i,~f \circ q_i\, ;\, \lambda i\, j\,
-  \phi,\, f (p_{i, j}^\phi)) \right)\end{mymath}%
-% where $\precompose_f : \phi = \psi \to \phi \circ f = \psi \circ f$,
-is an equivalence.
+Then $P$ is the {\em colimit} of $D$ if for any other $X:\Type$, 
+\begin{mymath}[-0.5em]\IsEquiv\left(\lambda f:P \to X, \left( \lambda i,~f \circ q_i\, ;\, \lambda i\, j\,
+  \phi,\, f (p_{i, j}^\phi)) \right)\right).\end{mymath}
 \end{defi}
@@
   We view $T_f$ as the coequalizer of
-  \begin{mymath}\xymatrix{
+  $\xymatrix{
     \sum_{a,b:X} f(a) = f(b) \dar[r]{4pt}^-{\pi_1}_-{\pi_2} & X
-  }\end{mymath}%
+  }$
   preserving the identity.
-
   We call $\tilde f$ the map $T_f \to Y$ given by induction.
@@
   is given by the diagram
-  \begin{mymath}\xymatrix{
+  $\xymatrix{
     C(f) := X \ar[r]^-t & T_f \ar[r]^-t & T_{\tilde f} \ar[r] &\cdots
-  }\end{mymath}%
+  }$
 \end{defi}
@@
   For any  $A:\Type$ and $B:A \to \Type_n^\modal$,  $P\left(\prod_{a:A} (B\,
-  a)\right)$. \\
+  a)\right)$. 
 (from \cite[Lemma 7.7.2]{hottbook})
 \item For any $A,B:\Type_n$,
-  $\modal (A\times B) = (\modal A) \times (\modal B)$. \\
+  $\modal (A\times B) = (\modal A) \times (\modal B)$. 
 (from \cite[Corollary 7.7.2]{hottbook})
@@
   $g:\prod_{a:A} B(\eta_A(a))$, there exists a map $f:\prod_{z:\modal
-    A} (B\, z)$ such that for all $a:A$, $f(\eta_A(a)) = g(a)$. \\
+    A} (B\, z)$ such that for all $a:A$, $f(\eta_A(a)) = g(a)$. 
  (from \cite[Theorem
@@
 
-Moreover the following diagram commute
-\begin{mymath}[-0.3em]\xymatrix{
-  \sum_{x:X} (f(x) = y) \ar[r]^\eta \ar[d]_\gamma & \modal \left(\sum_{x:X}  (f(x) = y)\right) \ar@{=}[dl]\\
-  \sum_{x:\modal X} (\modal f(x) = \eta_Y(y)) & }\end{mymath}%
-where $\pi_1 \circ \gamma = \eta_X \circ \pi_1$, and
-$\pi_2 \circ \gamma$ is given by the modalisation of
-paths.
+% Moreover the following diagram commute
+% \begin{mymath}[-0.3em]\xymatrix{
+%   \sum_{x:X} (f(x) = y) \ar[r]^\eta \ar[d]_\gamma & \modal \left(\sum_{x:X}  (f(x) = y)\right) \ar@{=}[dl]\\
+%   \sum_{x:\modal X} (\modal f(x) = \eta_Y(y)) & }\end{mymath}%
+% where $\pi_1 \circ \gamma = \eta_X \circ \pi_1$, and
+% $\pi_2 \circ \gamma$ is given by the modalisation of
+% paths.
 \end{prop}
@@
 \end{prop}
+
+% \vspace{-1em}
+Another fundamental property on sheaves is that the type of (dependent)
+functions is a sheaf as soon as its codomain is a sheaf.
+
+\begin{prop}\label{prop:sheaf-forall}
+  If $A:\Type_{n+1}$ and $B:A \to \Type_{n+1}$ such that for any
+  $a:A$, $(B~a)$ is a sheaf, then $\prod_{a:A}B~a$ is a sheaf.
+\end{prop}
+
+\section{Sheafification of Left Exact Modalities}
+\label{sec:sheafification}
+
+% We mimic the construction in~\cite{maclanemoerdijk}.
+
+From any left exact modality $\modal_{-1}$ on $\HProp$, we define by
+induction on the truncation index $n$, a
+left exact modality $\modal_n$ at level $n$ that extends $\modal_{-1}$ in the
+sense that the following property holds.
+%
+% We have to be careful: if we want to propagate the properties of the
+% first modality $\modal_{-1}$, all the higher modalities $\modal_n$ must be compatible
+% with $\modal_{-1}$, \ie we want the property
+\begin{cond}\label{cond:hprop}
+  For any mere proposition $P$ (where $\widehat P$ is $P$ seen as a
+  $\Type_n$),  $\modal_n \widehat P = \modal_{-1} P$ and the
+  following coherence diagram commutes
+  \begin{mymath}\xymatrix{
+    P \ar@{->}^{\sim}[r] \ar[d]_{\eta_{-1}} & \widehat P \ar[d]^{\eta_n} \\
+    \modal_{-1} P \ar@{->}^{\sim}[r] & \modal_n \widehat P 
+  }\end{mymath}%
+\end{cond}
+
+\label{sec:sheaf-left-exact}Note that once the first modality $\modal_{-1}$ is defined, its
+extension to all $\Type_n$ is automatic. This means that the new
+logical principles that can be added by the sheaf construction must be
+added in the first modality already. Sheafification is just a way
+to propagate those new principles to types that are not mere
+propositions.
+
+%   extend the new principles it gives to
+% every $\Type_n,\,n\geqslant 0$.  Actually, if $n_0$ is a fixed
+% truncation index, and $\modal$ a left exact modality on $\Type_{n_0}$,
+% then we can in the same way extend the properties of the modality to
+% any $\Type_n$, for $n > n_0$.
+
+The sheafification in topos theory described in
+Section~\ref{sec:lawv-tiern-sheaf} works in the same way. 
+%
+From a left exact modality on the internal logic (a Lawvere-Tierney
+topology), a new left exact modality is defined on the whole topos.
+%
+This construction may be seen as the first induction step between
+$\HProp$ and $\Type_0$.
+
+% \kq{c'est une vision plutôt grothendieck en fait. La separation ``enlève''
+%   ce qui empêche à $T$ d'être un faisceau, et la sheafification
+%   rajoute ce qui manque pour être un faisceau. Mais c'était plutôt mal
+%   dit, effectivement}
+
+For a given induction step, the sheafification is defined in two
+phases; the first one (section~\ref{ssec:from-type-separated}) constructs, for any $T:\Type_{n+1}$ a separated
+type $\separated T$, and the second one
+(section~\ref{ssec:separated-to-sheaf}) constructs a sheaf given any
+separated type.
+% %
+% \begin{enumerate}[(i)]
+% \item {\em separation:} From any $T$ in $\Type_{n+1}$, we construct
+%   its {\em free} separated object $\separated T$ by ``removing'' what
+%   prevents $T$ from being a sheaf.
+% \item {\em completion:} We add what is missing for the free 
+%   separated type to be a sheaf by using closure in a sheaf.
+% \end{enumerate}
+
+% \subsection{For mere propositions}
+% \label{ssec:h-propositions}
+
+% For the case $n=-1$, one can take any left exact modality on
+% $\HProp$. Here we took the $\lnot\lnot$ modality:
+% \begin{mymath}\forall P:\HProp,~\modal_j P = \lnot\lnot P.\end{mymath}
+% One can easily show that it defines a left exact modality.
+
+% The $\lnot\lnot$-modality allows to work with a propositional law of
+% excluded middle.
+
+In the rest of the section, we suppose given a truncation index
+$n\geqslant -1$, and a left exact modality $\modal_n$ on $\Type_n$,
+compatible with a left exact modality $\modal_{-1}$ on $\HProp$ in the
+sense of condition~\ref{cond:hprop}.
+
+As many proofs are very technical, we only give here sketches, and
+often refer to formalization for all details.
+
+\subsection{From Type to Separated Type}
+\label{ssec:from-type-separated}
+
+Let $T : \Type_{n+1}$. We define $\separated T$ as the image (see
+Section~\ref{ssec:hott}) of
+$\modal_n^T \circ \{\cdot\}_T$, as in
+\begin{mymath}\xymatrix{
+    T \ar[r]^{\{\cdot\}_T} \ar[d]_{\mu_T} & \left(\Type_n\right)^T \ar[d]^{\modal_n^T} \\
+  \separated T \ar[r]& \left( \Type_n^\modal \right)^T
+}, \end{mymath}%
+where $\{\cdot\}_T$ is the singleton map $\lambda (t:T),~\lambda
+(t':T),~t=t'$. 
+%
+$\separated T$ can be given explicitly by
+%
+\begin{mymath}
+\begin{array}{rcl}
+\separated T &\defeq & \im (\lambda~t:T,~\lambda~ t',~ \modal_n (t = t')) \\
+          & \defeq & \sum\limits_{u:T \to \Type_n^\modal} \left\| \sum_{a:T} 
+            (\lambda t,~\modal_n (a=t)) = u\right\|.
+\end{array}
+\end{mymath}%
+%
+This corresponds to the free separated object used in the topos-theoretic construction, but using $\Type_n^\modal$ instead of the
+$j$-subobject classifier $\Omega_j$.
+%
+\begin{prop}
+  For any $T\!:\!\Type_{n+1}$, $\separated T$ is separated.  
+\end{prop}
+
 \begin{proof}
+We use the following lemma:
+\begin{lem}
+  A $(n+1)$-truncated type $T$ with an embedding $f : T \to U$
+  into a separated $(n+1)$-truncated type $U$ is itself separated.
+\end{lem}
+As $\separated T$ embeds in $\left( \Type_n^\modal \right)^T$, we only
+have to show that the latter is separated. But it is the case because
+$\Type_n^\modal$ is a sheaf (by Proposition~\ref{prop:sheaf-is-sheaf})
+and a function type is a sheaf as soon
+as its codomain is a sheaf (by Proposition~\ref{prop:sheaf-forall}).
+\end{proof}
+
+To show that $\separated$ is a modality, we need to show that $\separated T$
+is universal among separated type below $T$. This comes from the
+following construction which connects $\separated T$ to the colimit of
+the iterated kernel pair of $\mu_T$.
+
+\begin{defi}
+  Let $X:\Type$. Let $\mathring T_X$ be the higher inductive type
+  generated by
+  \begin{itemize}
+  \item $\mathring t:~\|X\|_{n+1} \to \mathring T_X$
+  \item $\mathring \alpha:~\forall a\, b:\|X\|_{n+1},~\modal (a=b) \to
+    \mathring t(a) = \mathring t(b)$
+  \item $\mathring \alpha_1:~\forall a:\|X\|_{n+1},~
+    \mathring \alpha(a , a, \eta_{a=a} 1) = 1$
+  \end{itemize}
+\end{defi}
+We view $\mathring T$ as the coequalizer of
+$\xymatrix{\sum_{a,b:\|X\|_{n+1}} \modal (a=b) \dar[r]{4pt}^-{\pi_1}_-{\pi_2}
+  & \|X\|_{n+1}}$
+preserving $\eta_{a=a} 1$.
+
+We consider the diagram $\mathring T :=
+\xymatrix{\|X\|_{n+1} \ar[r] & \|\mathring T_{X}\|_{n+1} \ar[r] & \|\mathring
+  T_{\mathring T_X} \|_{n+1} \ar[r] & \cdots}$
+
+We already know that $\separated T$ is the $(n+1)$-colimit of
+$\|C(\mu_T)\|_{n+1}$. We can show, with an encode-decode method, that
+a embedding-like property is preserved by $\widetilde\cdot$: 
+\begin{equation}\label{eq:Omono}
+  \mathrm{if}~\forall a,b:A,~f (a) = f (b) \simeq \modal (a=b),
+\end{equation} 
+\vspace{-2em}
+\begin{mymath}
+  \mathrm{then}~\forall a,b:\|T_f\|_{n+1},~|\tilde f|_{n+1} (a) = |\tilde
+  f|_{n+1} (b) \simeq \modal (a=b)
+\end{mymath}
+This can be used to show that the two diagrams $\|C(\mu_T)\|_{n+1}$ and
+$\mathring T$ are equivalent, and thus
+\begin{lem}\label{lem:sepiscolim}
+  Let $T:\Type_{n+1}$. Then $\separated T$ is the $(n+1)$-colimit of the
+  diagram $\mathring T$.
+\end{lem}
+
+Again, this lemma is an adaptation of the sheafification process in
+topos theory, where the kernel pair of $\mu_T$
+is replaced by its iterated kernel pair.
+
+% We know that $\separated T$ is the $(n+1)$-colimit of the iterated
+% kernel pair of $\mu_T$.
+
+% We will need the following lemma:
+
+% \begin{lem}
+%   Let $A,S:\Type_{n+1}$, $S$ separated, and $f:A \to S$. Then if 
+%   % \vspace{-0.7em}
+%   \begin{equation}
+%     \label{eq:Omono}\tag{$\star$}
+%     \forall a,b:A,~f (a) = f (b) \simeq \modal (a=b),
+%     % \vspace{-0.7em}
+%   \end{equation}
+%   then
+%   \begin{mymath}\forall a,b:\|T_f\|_{n+1},~|\tilde f|_{n+1} (a) = |\tilde f|_{n+1} (b) \simeq \modal (a=b).\end{mymath}%
+% \end{lem}
+
+% \begin{proof}[Sketch of proof]
+%   By induction on truncation, we need to show that 
+%   \begin{mymath}\forall a,b:T_f,~\tilde f (|a|_{n+1}) = \tilde f (|b|_{n+1} )\simeq
+%   \modal (|a|_{n+1}=|b|_{n+1}).\end{mymath}%
+%   We use the encode-decode~\cite[Section 8.9]{hottbook} method to characterize $\tilde f (|a|_{n+1})
+%   = x$, and the result follows. We refer to the formalization for details.
+% \end{proof}
+
+Now, let $Q$ be any separated $\Type_{n+1}$, and $f:X \to Q$. Then the
+following diagram commutes
+\begin{mymath}[-0.5em]\xymatrix{
+\|X\|_{n+1} \ar[r] \ar[rd] & \|\mathring T_{X}\|_{n+1} \ar[r] \ar[d] & \|\mathring
+  T_{\mathring T_X} \|_{n+1} \ar[ld]\cdots \\
+  & Q &
+} \end{mymath}% 
+% \begin{mymath}\xymatrix @C=4em{ 
+%     \cdots \overline{\Delta_3} \tar[r]{4pt} \ar[rd]_{f\circ \pi_1} & \overline{\Delta_2}
+%     \dar[r]{2pt} \ar[d]^{f\circ \pi_1} &
+%     \overline{\Delta_1} = T \ar[ld]^{f\circ \pi_1}\\
+%     & Q &
+%   }\end{mymath}
+But we know (lemma~\ref{lem:sepiscolim}) that $\separated T$ is the
+$(n+1)$-colimit of the diagram $\mathring T$, thus there is an universal
+arrow $\separated T \to Q$.
+%
+This is enough to prove that $(\separated,\mu)$ defines a reflective
+subuniverse on $\Type_{n+1}$.
+We can now state
+\begin{prop}\label{prop:sep-mod}
+  $(\separated,\mu)$ defines a modality on $\Type_{n+1}$.
+\end{prop}
+
+As this modality is just a step in the construction, we do not need to
+show that it is left exact, we will only do it for the sheafification
+modality.
+
+\subsection{From Separated Type to Sheaf}
+\label{ssec:separated-to-sheaf}
+
+%\nt{put the definition of $\modal_{n+1}$ upfront}
+For any $T$ in $\Type_{n+1}$, 
+$\modal_{n+1}T$ is defined as the closure of $\separated T$,
+seen as a subobject of $T \to \Type_n^\modal$. 
+%
+$\modal_{n+1}T$ can be given explicitly by
+\begin{mymath}
+\modal_{n+1} T \ \defeq \sum_{u:T \to \Type_n^\modal} \modal_{-1}\left\| \sum_{a:T} 
+            (\lambda t,~\modal_n (a=t)) = u\right\|.
+\end{mymath}%
+
+To prove that $\modal_{n+1} T$ is a sheaf for any $T:\Type_{n+1}$, we
+use the following lemma.
+\begin{lem}
+  Let $X:\Type_{n+1}$ and $U$ be a sheaf. If $X$ embeds
+  in $U$, and is closed in $U$, then $X$ is a sheaf.
+\end{lem}
+
+As $T\to \Type_n^\modal$ is a sheaf, and $\modal_{n+1}T$ is closed in
+$T\to \Type_n^\modal$, $\modal_{n+1}T$ is a sheaf. We now prove that
+it forms a modality.
+
+\begin{prop}\label{prop:Sn-modality}
+  $(\modal_{n+1},\nu)$ defines a modality.
+\end{prop}
+
+It remains to show that $\modal_{n+1}$ is left exact and is compatible
+with $\modal_{-1}$. To do that, we need to extend the notion of
+compatibility and show that actually every modality $\modal_{n+1}$ is
+compatible with $\modal_n$ on lower homotopy types.
+\begin{prop} \label{prop:compatibility}
+  If $T:\Type_n$, then $\modal_{n+1} \widehat T = \modal_n T$, where $\widehat T$ is $T$ seen as a
+  $\Type_{n+1}$.
+\end{prop}
+
+This proves in particular that $\modal_{n+1}$ is compatible with
+$\modal_{-1}$ in the sense of condition~\ref{cond:hprop}.
+
+The last step is the left-exactness of $\modal_{n+1}$. Let $T$ be in
+$\Type_{n+1}$ such that $\modal_{n+1} T$ is contractible.  Thanks to the just
+shown compatibility between $\modal_{n+1}$ and $\modal_n$ for
+$\Type_n$, left exactness means that for any $x,y: T$,
+$\modal_n(x=y)$ is contractible.
+
+Using a proof by univalence as we have done for proving $\modal_n (a=b) = (\mu_T(a) =
+\mu_T (b))$ in Proposition~\ref{lem:sepiscolim}, we can show that:
+\begin{prop}
+  For all $a,b:T$, $\modal_n(a=b) \!=\! (\nu_T a = \nu_T b)$.
+\end{prop}
+
+As $\modal_{n+1} T$ is contractible, path spaces of $\modal_{n+1} T$ are
+contractible, in particular $(\nu_T a=\nu_T b)$, which proves left
+exactness.
+
+\subsection{Summary}
+\label{ssec:summary}
+
+Starting from any left-exact modality $\modal_{-1}$ on $\HProp$, we
+have defined for any truncation level $n$, a new left-exact modality
+$\modal_n$ on $\Type_n$, which corresponds to $\modal_{-1}$ when
+restricted to $\HProp$.
+
+When $\modal_{-1}$ is consistent (in the sense of
+proposition~\ref{prop:consistent}), 
+$\modal_{n}\zero=\modal_{-1}\zero$ is also not inhabited, hence the homotopy type theory induced by
+$\modal_n$ is consistent. 
+%
+In particular, the modality induced by the double negation modality on
+$\HProp$ is consistent.
+
+In topos theory, the topos of Lawvere-Tierney sheaves for the double
+negation topology is a boolean topos. In homotopy type theory, this
+result can be expressed as:
+
+\begin{prop}
+  $(\modal_{\lnot\lnot})_n$, the modality obtained by sheafifica\-tion
+  of the double negation modality,
+  %
+  induces a type theory where the propositional excluded middle law holds.
+\end{prop}
+
+Combined with forcing in type theory~\cite{jaber2012extending}, it
+should be possible to lift the proof of independence of the continuum
+hypothesis to a classical setting, which is where the continuum hypothesis is
+really meaningful.  However, we haven't worked out the details and left
+this for future work.
+
+% \nt{Recap what has been done, and give the final result about EM with
+%   the dense topology as a corollary.}
+
+\section{Extension to Type}
+\label{sec:extension-type}
+
+In the previous section, we have defined a (countably) infinite family of
+modalities $\Type_i \to \Type_i$. One can extend them to whole
+$\Type$ by composing with truncation:
+
+\begin{lem}\label{lem:type}
+  Let $\modal_i:\Type_i \to \Type_i$ be a modality. Then $\modal
+  \defeq \modal_i
+  \circ \|\cdot\|_i : \Type \to \Type$ is a modality in the sense
+  of~\cite[Section 7.7]{hottbook}
+\end{lem}
+
+If $\modal_{-1}$ is the double negation modality on $\HProp$ and
+$i=-1$, $\modal$ is exactly the double negation modality on $\Type$
+described in~\ref{sec:examples-left-exact}.
+Chosing $i\geqslant 0$ is a refinement of this double negation
+modality on $\Type$: it will collapse every type to a $\Type_i$,
+instead of an $\HProp$.
+
+Obviously, as truncation modalities are not left-exact~\cite[Exercise
+7.11]{hottbook}, $\modal$ isn't either. But when
+restricted to $i$-truncated types, $\modal = \modal_i$, so:
+\begin{lem}
+  Let $A:\Type_i$. Then if $\modal(A)$ is contractible, for any $x,y:A$,
+  $\modal(x=y)$ is contractible.
+\end{lem}
+
+The compatibility between the modalities $\modal_n$ and between the
+modalities $\|\cdot \|_n$ allow us to chose the truncation index as
+high as desired.
+Taking it as a non-fixed parameter allows to work in an
+universe where the new principle ({\em e.g.} mere excluded middle) is
+true for any explicit truncated type. Indeed, $i$ can be chosen
+dynamically along a proof, and thus be increased as much as needed,
+without changing results for lower truncated types.
+
+By proposition~\ref{prop:consistent}, these left-exact modalities
+induces a consistent type theory. Furthermore, the univalence remains
+true in this new type theory in the following sense:
+\begin{prop}\label{prop:univalence}
+  Let $n$ be a given truncation index, and $\modal$ the modality
+  associated to $n$ as defined in lemma~\ref{lem:type}. Then, for
+  any type $A,B:\Type_n^\modal$, if $\varphi$ is the canonical arrow
+  $A = B \to A\simeq B$,
+  then $\IsEquiv(\varphi)$ is modal.
+\end{prop}
+
+We can view sheafification in terms of model of type theory but
+because of the resulting modality on $\Type$ is not left exact, we
+need to restrict ourselves to a type theory with only one universe.
+%
+Let $\mathfrak M$ be a model of homotopy type theory with one
+universe.
+%
+Using the modality $\modal_{\lnot \lnot}$ (for any level $n$) associated to the
+sheafification, there is a model $\modal_{\lnot \lnot} \mathfrak M$ of type theory
+with one universe (using results in
+Section~\ref{sec:new-type-theory}), where excluded middle is true, and
+which is univalent (as shown in Proposition~\ref{prop:univalence}).
+%
+This model transformation can better be described as a program
+transformation, as done in the next section.
+
+% \subsection{New objects}
+% \label{sec:new-objects}
+
+% Working in $\Type^\modal$, one can consider new objects. For example,
+% let $c:nat \to \nat$ implementing the Collatz function. Then, for any
+% $n:\nat$, the type \begin{mymath}C_n \defeq \sum_{k:\nat}\left( c^k(n) = 1 \land k\text{ is the
+%   smallest such $\nat$}\right)\end{mymath} is an $\HProp$.
+% Thus, for any $n$, one have $C_n + \lnot C_n$, and one can {\em
+%   constructively} build a map $\varphi:\nat \to \two$ such that
+% $\varphi(n) = \True$ if and only if $n$ satisfies the Collatz
+% conjecture.
+% Actually, one can defined such a map for any family of $\HProp$
+% indexed by $\nat$, for example with the family $C_n \defeq
+% \text{«\,The $n$-th Turing Machine halts\,»}$.
+
+\section{Formalization} 
+\label{sec:formalization}
+
+A Coq formalization based on the Coq/HoTT library~\cite{hottlib} is available at
+\url{https://github.com/KevinQuirin/sheafification}.
+%
+% After reviewing the content and some statistics about the
+% formalization in Section~\ref{sec:cont-form}, we present the
+% limitations of our formalization in Section~\ref{sec:limit-form}, in
+% particular the issues relative to universe polymorphism.
+
+\subsection{Content of the formalization}
+\label{sec:cont-form}
+
+% We provide a more detailed insight of the structure of our formalization:
+% \begin{itemize} 
+% \item Colimits and iterated kernel pairs are formalized in
+% \texttt{Limit}, \texttt{T.v}, \texttt{OT.v}v \texttt{OT\_Tf.v}, \texttt{T\_telescope.v},\\ \texttt{Tf\_Omono\_sep.v}.% (552 lines).
+% \item
+% Reflective subuniverses and modalities are formalized in\\
+% \texttt{reflective\_subuniverse.v}, \texttt{modalities.v}. % (1053 lines).
+% \item 
+% %
+%   The definition of the dense topology as a left exact modality on
+%   $\HProp$ is given in \texttt{sheaf\_base\_case.v}. % (186 lines).
+% \item
+% Section~\ref{sec:sheaves} is formalized in
+% \texttt{sheaf\_def\_and\_thm.v}. % (1029 lines).
+% \item
+% Section~\ref{sec:sheafification} is formalized in
+% \texttt{sheaf\_induction.v}. % (2340 lines).
+% \end{itemize}
+
+Overall, % with other files containing technical lemmas,
+the project
+contains 7914 lines, and it could be reduced a bit by improving the
+way Coq tries to rewrite and apply lemmas automatically. 
+The \texttt{coqwc} tool counts 1611 lines of specifications
+(definitions, lemmas, theorems, propositions) and 5403 lines of proof
+script.
+%
+This constitutes a significant amount of work but the part dedicated
+to sheaves and sheafification is only 2200 lines of proof script,
+which seems quite reasonable and encouraging, because it suggests
+that homotopy type theory provides a convenient tool to formalize some
+part of the theory of higher topoï. As a matter of comparison, this
+\TeX\xspace file contains approximatively 2000 lines.
+
+% In our formalization, some things are admitted, either because they
+% are just an adaptation of a proposition already formalized
+% in~\cite{hottbook}, either because they are true, but very technical:
+% \begin{itemize}
+% \item in \texttt{modalities.v}, an equality is admitted in the proof
+%   of proposition~\ref{sec:defin-basic-prop}.
+% \item in \texttt{nat\_lemmas.v}, the proof that $\IsHProp (n\leqslant
+%   m)$ is admitted.
+% \item in \texttt{sheaf\_induction.v}, in the proof that
+%   $C(\mu_T)=(C\id)$ (see lemma~\ref{lem:sepiscolim}), we admit the proof
+%   of equality of projections.
+% \item in \texttt{sheaf\_induction.v}, we admit some very technical
+%   proofs in proofs that $\square$ and $\star$ are modalities.
+% \item in \texttt{sheaf\_induction.v}, cumulativity is an axiom,
+%   because of universes issues.
+% \end{itemize}
+
+\subsection{Limitations of the formalization}
+\label{sec:limit-form}
+
+In the formalization, we had to use the \texttt{type-in-type} option, to handle
+the universe issues we faced.
+
+% Universes are used in type theory to ensure consistency by checking
+% that definitions are well-stratified according to a certain hierarchy.
+% %
+% Universe polymorphism~\cite{sozeau2014universe} supports generic
+% definitions over universes, reusable at different levels.
+% %
+% Although the presence of universe polymorphism is mandatory for our
+% formalization, its implementation is still too rigid to allow a
+% complete formalization of our work for the following reasons.
+
+% %
+% If Coq handles cumulativity on $\Type$ natively, it is not
+% the case for the $\Sigma$-type $\Type_n$, which require propositional
+% resizing. 
+% %
+% This issue could be solved by adding an axiom of cumulativity
+% for $\Type_n$ with an explicit management of universes. 
+% %
+% But as it would not have any computational content, such a solution
+% would really complicate the proofs as the axiom would appear
+% everywhere cumulativity is needed and it would need explicit
+% annotations for universe levels everywhere in the formalization.
+% %
+% % Note that we also could have resolved the issue by giving explicitely
+% % the universe levels we wanted, like in the Coq/HoTT library\cite{hottlib}.
+
+One issue with universe polymorphism lies in the management of
+recursive definitions. Indeed, the recursive definition of
+sheafification~\ref{eq:induction} is not allowed. 
+This is because Coq forces the universe of the first $\Type_n$
+occurring in the definition to be the same for every $n$, whereas the
+universe of the first $\Type_{n+1}$ occurring in $\modal_{n+1}$ should be at
+least one level higher as the one of $\Type_n$ occurring in
+$\modal_{n}$ because of the use of $\Sigma$-type over
+$T \to \Type_n^\modal$ and equality on the return type of $\modal_n$.
+%
+% \nt{explain why it is an issue.}
+%
+Thus, the induction step presented in this paper has been formalized,
+but the complete recursive sheafification can not be defined for the
+moment.
+%
+Note that the same increasing in the universe levels occurs in the
+Rezk completion for categories~\cite{rezk}. In the definition of the
+completion, they use the Yoneda embedding and representable functors,
+which is similar to our use of characteristic functions.
+
+%
+This restriction in our formalization may be solved by
+generalizing the management of universe polymorphism for recursive definition
+%
+or by the use of general ``resizing axiom'' which is still under
+discussion in the community.
+
+\section{Conclusion and Future Works}
+\label{sec:future-works}
+
+In this paper, we have defined a way to extend homotopy type theory
+with new principles, leveraging Lawvere-Tierney sheafification to
+homotopy type theory. 
+%
+Using sheafification, it is possible to add new logical principles to
+homotopy type theory, such as the law of excluded middle, and this
+without axiom.
+%
+Beyond this result, our work is part of a more general line of work
+which aim to illustrate that homotopy type theory is a very promising
+theory for formalizing mathematics inside a proof assistant.
+
+% \nt{Add something on the definition of a syntactic translation as it
+  % has been done for forcing.}
+
+As a future work, we would like to improve this construction in three
+ways.
+% \begin{itemize}
+% \item 
+(i) The extension to whole $\Type$ in lemma~\ref{lem:type} is not
+  totally satisfactory, as every type is collapsed to a truncated
+  one. But some types in homotopy type theory are not
+  truncated~\cite[Example 8.8.6]{hottbook}.
+  % At the moment, sheafification only works with truncated
+  % types. This is fine from a programming point of view, 
+  % but many important types in homotopy theory are not
+  % truncated, \eg the circle $\mathbb S^2$.  Thus, we plan to extend
+  % the sheafification to all types by using a
+  % limit construction.
+% \item 
+(ii) 
+  We would like to have more examples of left-exact modalities on
+  $\HProp$, in order to extend type theory with different principles
+  than excluded middle.
+% \item 
+(iii) It would be useful to improve the translation induced by the
+  sheafification modality to preserve more definitional equalities.
+  %
+  This could be done for instance by defining a higher inductive type
+  isomorphic to the modality but with a better computational content. 
+  % able to work in the new type theory into a proof assistant.
+% \end{itemize}
+
+Moreover, in topos theory, Lawvere-Tierney subsumes Grothen\-dieck~\cite[Section~V.4]{maclanemoerdijk} in the sense that any
+Grothen\-dieck topology gives rise to a Lawvere-Tierney topology with
+the same notion of sheaves. Higher Lawvere-Tierney sheaves are
+presented here, and higher Grothendieck sheaves have been defined
+in~\cite{lurie}. It should be interesting to check if the subsumption
+remains true in higher topos theory.
+
+% \section*{Acknowledgments}
+
+% We would like to thank all maintainers of the Coq/HoTT library, who
+% have provided a very solid base for our formalization. 
+% %
+% We also would like to thank Bas Spitters for helpful discussions on
+% the subject.
+% %
+% This work has been funded by the CoqHoTT ERC Grant 637339.
+
+%\bibliographystyle{abbrvnat}
+\bibliography{sheaf_lics}
+
+\newpage
+\section*{Appendix}
+
+\begin{proof}[Proof of Proposition~\ref{prop:sheaf-is-sheaf}]
   First, we prove that $\Type_n^\modal$ is separated. Let $E:\Type$
@@
 \end{proof}
-% \vspace{-1em}
-Another fundamental property on sheaves is that the type of (dependent)
-functions is a sheaf as soon as its codomain is a sheaf.
-
-\begin{prop}\label{prop:sheaf-forall}
-  If $A:\Type_{n+1}$ and $B:A \to \Type_{n+1}$ such that for any
-  $a:A$, $(B~a)$ is a sheaf, then $\prod_{a:A}B~a$ is a sheaf.
-\end{prop}
-\begin{proof}
+
+\begin{proof}[Proof of Proposition~\ref{prop:sheaf-forall}]
   \begin{itemize}
@@
 %  \nt{add a short proof here}
-\end{proof}
-
-\section{Sheafification of Left Exact Modalities}
-\label{sec:sheafification}
-
-% We mimic the construction in~\cite{maclanemoerdijk}.
-
-From any left exact modality $\modal_{-1}$ on $\HProp$, we define by
-induction on the truncation index $n$, a
-left exact modality $\modal_n$ at level $n$ that extends $\modal_{-1}$ in the
-sense that the following property holds.
-%
-% We have to be careful: if we want to propagate the properties of the
-% first modality $\modal_{-1}$, all the higher modalities $\modal_n$ must be compatible
-% with $\modal_{-1}$, \ie we want the property
-\begin{cond}\label{cond:hprop}
-  For any mere proposition $P$ (where $\widehat P$ is $P$ seen as a
-  $\Type_n$),  $\modal_n \widehat P = \modal_{-1} P$ and the
-  following coherence diagram commutes
-  \begin{mymath}\xymatrix{
-    P \ar@{->}^{\sim}[r] \ar[d]_{\eta_{-1}} & \widehat P \ar[d]^{\eta_n} \\
-    \modal_{-1} P \ar@{->}^{\sim}[r] & \modal_n \widehat P 
-  }\end{mymath}%
-\end{cond}
-
-\label{sec:sheaf-left-exact}Note that once the first modality $\modal_{-1}$ is defined, its
-extension to all $\Type_n$ is automatic. This means that the new
-logical principles that can be added by the sheaf construction must be
-added in the first modality already. Sheafification is just a way
-to propagate those new principles to types that are not mere
-propositions.
-
-%   extend the new principles it gives to
-% every $\Type_n,\,n\geqslant 0$.  Actually, if $n_0$ is a fixed
-% truncation index, and $\modal$ a left exact modality on $\Type_{n_0}$,
-% then we can in the same way extend the properties of the modality to
-% any $\Type_n$, for $n > n_0$.
-
-The sheafification in topos theory described in
-Section~\ref{sec:lawv-tiern-sheaf} works in the same way. 
-%
-From a left exact modality on the internal logic (a Lawvere-Tierney
-topology), a new left exact modality is defined on the whole topos.
-%
-This construction may be seen as the first induction step between
-$\HProp$ and $\Type_0$.
-
-% \kq{c'est une vision plutôt grothendieck en fait. La separation ``enlève''
-%   ce qui empêche à $T$ d'être un faisceau, et la sheafification
-%   rajoute ce qui manque pour être un faisceau. Mais c'était plutôt mal
-%   dit, effectivement}
-
-For a given induction step, the sheafification is defined in two
-phases; the first one (section~\ref{ssec:from-type-separated}) constructs, for any $T:\Type_{n+1}$ a separated
-type $\separated T$, and the second one
-(section~\ref{ssec:separated-to-sheaf}) constructs a sheaf given any
-separated type.
-% %
-% \begin{enumerate}[(i)]
-% \item {\em separation:} From any $T$ in $\Type_{n+1}$, we construct
-%   its {\em free} separated object $\separated T$ by ``removing'' what
-%   prevents $T$ from being a sheaf.
-% \item {\em completion:} We add what is missing for the free 
-%   separated type to be a sheaf by using closure in a sheaf.
-% \end{enumerate}
-
-% \subsection{For mere propositions}
-% \label{ssec:h-propositions}
-
-% For the case $n=-1$, one can take any left exact modality on
-% $\HProp$. Here we took the $\lnot\lnot$ modality:
-% \begin{mymath}\forall P:\HProp,~\modal_j P = \lnot\lnot P.\end{mymath}
-% One can easily show that it defines a left exact modality.
-
-% The $\lnot\lnot$-modality allows to work with a propositional law of
-% excluded middle.
-
-In the rest of the section, we suppose given a truncation index
-$n\geqslant -1$, and a left exact modality $\modal_n$ on $\Type_n$,
-compatible with a left exact modality $\modal_{-1}$ on $\HProp$ in the
-sense of condition~\ref{cond:hprop}.
-
-As many proofs are very technical, we only give here sketches, and
-often refer to formalization for all details.
-
-\subsection{From Type to Separated Type}
-\label{ssec:from-type-separated}
-
-Let $T : \Type_{n+1}$. We define $\separated T$ as the image (see
-Section~\ref{ssec:hott}) of
-$\modal_n^T \circ \{\cdot\}_T$, as in
-\begin{mymath}\xymatrix{
-    T \ar[r]^{\{\cdot\}_T} \ar[d]_{\mu_T} & \left(\Type_n\right)^T \ar[d]^{\modal_n^T} \\
-  \separated T \ar[r]& \left( \Type_n^\modal \right)^T
-}, \end{mymath}%
-where $\{\cdot\}_T$ is the singleton map $\lambda (t:T),~\lambda
-(t':T),~t=t'$. 
-%
-$\separated T$ can be given explicitly by
-%
-\begin{mymath}
-\begin{array}{rcl}
-\separated T &\defeq & \im (\lambda~t:T,~\lambda~ t',~ \modal_n (t = t')) \\
-          & \defeq & \sum\limits_{u:T \to \Type_n^\modal} \left\| \sum_{a:T} 
-            (\lambda t,~\modal_n (a=t)) = u\right\|.
-\end{array}
-\end{mymath}%
-%
-This corresponds to the free separated object used in the topos-theoretic construction, but using $\Type_n^\modal$ instead of the
-$j$-subobject classifier $\Omega_j$.
-%
-\begin{prop}
-  For any $T\!:\!\Type_{n+1}$, $\separated T$ is separated.  
-\end{prop}
-
-\begin{proof}
-We use the following lemma:
-\begin{lem}
-  A $(n+1)$-truncated type $T$ with an embedding $f : T \to U$
-  into a separated $(n+1)$-truncated type $U$ is itself separated.
-\end{lem}
-As $\separated T$ embeds in $\left( \Type_n^\modal \right)^T$, we only
-have to show that the latter is separated. But it is the case because
-$\Type_n^\modal$ is a sheaf (by Proposition~\ref{prop:sheaf-is-sheaf})
-and a function type is a sheaf as soon
-as its codomain is a sheaf (by Proposition~\ref{prop:sheaf-forall}).
-\end{proof}
-
-To show that $\separated$ is a modality, we need to show that $\separated T$
-is universal among separated type below $T$. This comes from the
-following construction which connects $\separated T$ to the colimit of
-the iterated kernel pair of $\mu_T$.
-
-\begin{defi}
-  Let $X:\Type$. Let $\mathring T_X$ be the higher inductive type
-  generated by
-  \begin{itemize}
-  \item $\mathring t:~\|X\|_{n+1} \to \mathring T_X$
-  \item $\mathring \alpha:~\forall a\, b:\|X\|_{n+1},~\modal (a=b) \to
-    \mathring t(a) = \mathring t(b)$
-  \item $\mathring \alpha_1:~\forall a:\|X\|_{n+1},~
-    \mathring \alpha(a , a, \eta_{a=a} 1) = 1$
-  \end{itemize}
-
-  We view $\mathring T$ as the coequalizer of
-  \begin{mymath}\xymatrix{\sum_{a,b:\|X\|_{n+1}} \modal (a=b) \dar[r]{4pt}^-{\pi_1}_-{\pi_2}
-  & \|X\|_{n+1}}\end{mymath}%
-  preserving $\eta_{a=a} 1$.
-
-  We consider the diagram $\mathring T$ :
-  \begin{mymath}\xymatrix{\|X\|_{n+1} \ar[r] & \|\mathring T_{X}\|_{n+1} \ar[r] & \|\mathring
-  T_{\mathring T_X} \|_{n+1} \ar[r] & \cdots} \end{mymath}%
-\end{defi}
-
-\begin{lem}\label{lem:sepiscolim}
-  Let $T:\Type_{n+1}$. Then $\separated T$ is the $(n+1)$-colimit of the
-  diagram $\mathring T$.
-\end{lem}
-
-Again, this lemma is an adaptation of the sheafification process in
-topos theory, where the kernel pair of $\mu_T$
-is replaced by its iterated kernel pair.
-
-We will need the following lemma:
-
-\begin{lem}
-  Let $A,S:\Type_{n+1}$, $S$ separated, and $f:A \to S$. Then if 
-  % \vspace{-0.7em}
-  \begin{equation}
-    \label{eq:Omono}\tag{$\star$}
-    \forall a,b:A,~f (a) = f (b) \simeq \modal (a=b),
-    % \vspace{-0.7em}
-  \end{equation}
-  then
-  \begin{mymath}\forall a,b:\|T_f\|_{n+1},~|\tilde f|_{n+1} (a) = |\tilde f|_{n+1} (b) \simeq \modal (a=b).\end{mymath}%
-\end{lem}
-
-\begin{proof}[Sketch of proof]
-  By induction on truncation, we need to show that 
-  \begin{mymath}\forall a,b:T_f,~\tilde f (|a|_{n+1}) = \tilde f (|b|_{n+1} )\simeq
-  \modal (|a|_{n+1}=|b|_{n+1}).\end{mymath}%
-  We use the encode-decode~\cite[Section 8.9]{hottbook} method to characterize $\tilde f (|a|_{n+1})
-  = x$, and the result follows. We refer to the formalization for details.
 \end{proof}
@@
 
-Now, let $Q$ be any separated $\Type_{n+1}$, and $f:X \to Q$. Then the
-following diagram commutes
-
-\begin{mymath}\xymatrix{
-\|X\|_{n+1} \ar[r] \ar[rd] & \|\mathring T_{X}\|_{n+1} \ar[r] \ar[d] & \|\mathring
-  T_{\mathring T_X} \|_{n+1} \ar[ld]\cdots \\
-  & Q &
-} \end{mymath}% 
-% \begin{mymath}\xymatrix @C=4em{ 
-%     \cdots \overline{\Delta_3} \tar[r]{4pt} \ar[rd]_{f\circ \pi_1} & \overline{\Delta_2}
-%     \dar[r]{2pt} \ar[d]^{f\circ \pi_1} &
-%     \overline{\Delta_1} = T \ar[ld]^{f\circ \pi_1}\\
-%     & Q &
-%   }\end{mymath}
-But we know (lemma~\ref{lem:sepiscolim}) that $\separated T$ is the
-$(n+1)$-colimit of the diagram $\mathring T$, thus there is an universal
-arrow $\separated T \to Q$.
-%
-This is enough to state the following proposition.
-\begin{prop}\label{prop:sep-subu}
-  $(\separated,\mu)$ defines a reflective subuniverse on $\Type_{n+1}$.
-\end{prop}
-
-To show that $\separated$ is a modality, it remains to show point
-{\it (v)} of the definition.
-%
+\begin{proof}[Proof of Proposition~\ref{prop:sep-mod}]
+  Only point {\it (v)} of the definition remains to be checked.
+
   Let $A:\Type_{n+1}$ be a separated type and $B:A \to \Type_{n+1}$ be a
@@
 are equal. The complete proof can be found in the formalization.
-This proves the following proposition
-\begin{prop}\label{prop:sep-mod}
-  $(\separated,\mu)$ defines a modality on $\Type_{n+1}$.
-\end{prop}
-
-As this modality is just a step in the construction, we do not need to
-show that it is left exact, we will only do it for the sheafification
-modality.
-
-\subsection{From Separated Type to Sheaf}
-\label{ssec:separated-to-sheaf}
-
-%\nt{put the definition of $\modal_{n+1}$ upfront}
-For any $T$ in $\Type_{n+1}$, 
-$\modal_{n+1}T$ is defined as the closure of $\separated T$,
-seen as a subobject of $T \to \Type_n^\modal$. 
-%
-$\modal_{n+1}T$ can be given explicitly by
-\begin{mymath}
-\modal_{n+1} T \ \defeq \sum_{u:T \to \Type_n^\modal} \modal_{-1}\left\| \sum_{a:T} 
-            (\lambda t,~\modal_n (a=t)) = u\right\|.
-\end{mymath}%
-
-To prove that $\modal_{n+1} T$ is a sheaf for any $T:\Type_{n+1}$, we
-use the following lemma.
-\begin{lem}
-  Let $X:\Type_{n+1}$ and $U$ be a sheaf. If $X$ embeds
-  in $U$, and is closed in $U$, then $X$ is a sheaf.
-\end{lem}
-
-As $T\to \Type_n^\modal$ is a sheaf, and $\modal_{n+1}T$ is closed in
-$T\to \Type_n^\modal$, $\modal_{n+1}T$ is a sheaf. We now prove that
-it forms a reflective subuniverse.
-
-\begin{prop}
-  $(\modal_{n+1},\nu)$ defines a reflective subuniverse.
-\end{prop}
-\begin{proof}
+\end{proof}
+
+\begin{proof}[Proof of Proposition~\ref{prop:Sn-modality}]
   Let $T,Q:\Type_{n+1}$ such that $Q$ is a sheaf. Let $f:T\to Q$.
@@
   As all these steps are universal, the composition is.
+
+  Using the same technique as in proposition~\ref{prop:sep-mod}, we
+  can prove that $\modal_{n+1}$ preserves dependent sums.
 \end{proof}
 
-Using the same technique as in proposition~\ref{prop:sep-mod}, we have
-\begin{prop}
-  $(\modal_{n+1},\nu)$ defines a modality.
-\end{prop}
-
-It remains to show that $\modal_{n+1}$ is left exact and is compatible
-with $\modal_{-1}$. To do that, we need to extend the notion of
-compatibility and show that actually every modality $\modal_{n+1}$ is
-compatible with $\modal_n$ on lower homotopy types.
-\begin{prop} \label{prop:compatibility}
-  If $T:\Type_n$, then $\modal_{n+1} \widehat T = \modal_n T$, where $\widehat T$ is $T$ seen as a
-  $\Type_{n+1}$.
-\end{prop}
-\begin{proof}
+\begin{proof}[Proof of Proposition~\ref{prop:compatibility}]
   We prove it by induction on $n$:
@@
     Now, let define $\phi : \modal_{-1} T \to \modal_0T$ by
-    \begin{mymath}\phi t = (\lambda t',\, \one
-      ;\kappa)\end{mymath}%
+    $\phi(t) = (\lambda t',\, \one
+      ;\kappa)$
     where $\kappa$ is defined by $\modal_{-1}$-induction on
@@
 \end{proof}
-This proves in particular that $\modal_{n+1}$ is compatible with
-$\modal_{-1}$ in the sense of condition~\ref{cond:hprop}.
-
-The last step is the left-exactness of $\modal_{n+1}$. Let $T$ be in
-$\Type_{n+1}$ such that $\modal_{n+1} T$ is contractible.  Thanks to the just
-shown compatibility between $\modal_{n+1}$ and $\modal_n$ for
-$\Type_n$, left exactness means that for any $x,y: T$,
-$\modal_n(x=y)$ is contractible.
-
-Using a proof by univalence as we have done for proving $\modal_n (a=b) = (\mu_T(a) =
-\mu_T (b))$ in Proposition~\ref{lem:sepiscolim}, we can show that:
-\begin{prop}
-  For all $a,b:T$, $\modal_n(a=b) \!=\! (\nu_T a = \nu_T b)$.
-\end{prop}
-
-As $\modal_{n+1} T$ is contractible, path spaces of $\modal_{n+1} T$ are
-contractible, in particular $(\nu_T a=\nu_T b)$, which proves left
-exactness.
-
-\subsection{Summary}
-\label{ssec:summary}
-
-Starting from any left-exact modality $\modal_{-1}$ on $\HProp$, we
-have defined for any truncation level $n$, a new left-exact modality
-$\modal_n$ on $\Type_n$, which corresponds to $\modal_{-1}$ when
-restricted to $\HProp$.
-
-When $\modal_{-1}$ is consistent (in the sense of
-proposition~\ref{prop:consistent}), 
-$\modal_{n}\zero=\modal_{-1}\zero$ is also not inhabited, hence the homotopy type theory induced by
-$\modal_n$ is consistent. 
-%
-In particular, the modality induced by the double negation modality on
-$\HProp$ is consistent.
-
-In topos theory, the topos of Lawvere-Tierney sheaves for the double
-negation topology is a boolean topos. In homotopy type theory, this
-result can be expressed as:
-
-\begin{prop}
-  $(\modal_{\lnot\lnot})_n$, the modality obtained by sheafifica\-tion
-  of the double negation modality,
-  %
-  induces a type theory where the propositional excluded middle law holds.
-\end{prop}
-
-Combined with forcing in type theory~\cite{jaber2012extending}, it
-should be possible to lift the proof of independence of the continuum
-hypothesis to a classical setting, which is where the continuum hypothesis is
-really meaningful.  However, we haven't worked out the details and left
-this for future work.
-
-% \nt{Recap what has been done, and give the final result about EM with
-%   the dense topology as a corollary.}
-
-\section{Extension to Type}
-\label{sec:extension-type}
-
-In the previous section, we have defined a (countably) infinite family of
-modalities $\Type_i \to \Type_i$. One can extend them to whole
-$\Type$ by composing with truncation:
-
-\begin{lem}\label{lem:type}
-  Let $\modal_i:\Type_i \to \Type_i$ be a modality. Then $\modal
-  \defeq \modal_i
-  \circ \|\cdot\|_i : \Type \to \Type$ is a modality in the sense
-  of~\cite[Section 7.7]{hottbook}
-\end{lem}
-
-If $\modal_{-1}$ is the double negation modality on $\HProp$ and
-$i=-1$, $\modal$ is exactly the double negation modality on $\Type$
-described in~\ref{sec:examples-left-exact}.
-Chosing $i\geqslant 0$ is a refinement of this double negation
-modality on $\Type$: it will collapse every type to a $\Type_i$,
-instead of an $\HProp$.
-
-Obviously, as truncation modalities are not left-exact~\cite[Exercise
-7.11]{hottbook}, $\modal$ isn't either. But in the following sense, when
-restricted to $i$-truncated types, it is:
-\begin{lem}
-  Let $A:\Type_i$. Then if $\modal(A)$ is contractible, for any $x,y:A$,
-  $\modal(x=y)$ is contractible.
-\end{lem}
-\begin{proof}
-  For $i$-truncated types, $\modal = \modal_i$, and $\modal_i$ is left-exact.
-\end{proof}
-
-The compatibility between the modalities $\modal_n$ and between the
-modalities $\|\cdot \|_n$ allow us to chose the truncation index as
-high as desired.
-Taking it as a non-fixed parameter allows to work in an
-universe where the new principle ({\em e.g.} mere excluded middle) is
-true for any explicit truncated type. Indeed, $i$ can be chosen
-dynamically along a proof, and thus be increased as much as needed,
-without changing results for lower truncated types.
-
-By proposition~\ref{prop:consistent}, these left-exact modalities
-induces a consistent type theory. Furthermore, the univalence remains
-true in this new type theory in the following sense:
-\begin{prop}\label{prop:univalence}
-  Let $n$ be a given truncation index, and $\modal$ the modality
-  associated to $n$ as defined in lemma~\ref{lem:type}. Then, for
-  any type $A,B:\Type_n^\modal$, if $\varphi$ is the canonical arrow
-  \begin{mymath}A = B \to A\simeq B,\end{mymath}
-  then $\IsEquiv(\varphi)$ is modal.
-\end{prop}
-\begin{proof}
+
+\begin{proof}[Proof of Proposition~\ref{prop:univalence}]
   The first thing to notice is that, if $X$ and $Y$ are modal, and
@@
 
-We can view sheafification in terms of model of type theory but
-because of the resulting modality on $\Type$ is not left exact, we
-need to restrict ourselves to a type theory with only one universe.
-%
-Let $\mathfrak M$ be a model of homotopy type theory with one
-universe.
-%
-Using the modality $\modal_{\lnot \lnot}$ (for any level $n$) associated to the
-sheafification, there is a model $\modal_{\lnot \lnot} \mathfrak M$ of type theory
-with one universe (using results in
-Section~\ref{sec:new-type-theory}), where excluded middle is true, and
-which is univalent (as shown in Proposition~\ref{prop:univalence}).
-%
-This model transformation can better be described as a program
-transformation, as done in the next section.
-
-% \subsection{New objects}
-% \label{sec:new-objects}
-
-% Working in $\Type^\modal$, one can consider new objects. For example,
-% let $c:nat \to \nat$ implementing the Collatz function. Then, for any
-% $n:\nat$, the type \begin{mymath}C_n \defeq \sum_{k:\nat}\left( c^k(n) = 1 \land k\text{ is the
-%   smallest such $\nat$}\right)\end{mymath} is an $\HProp$.
-% Thus, for any $n$, one have $C_n + \lnot C_n$, and one can {\em
-%   constructively} build a map $\varphi:\nat \to \two$ such that
-% $\varphi(n) = \True$ if and only if $n$ satisfies the Collatz
-% conjecture.
-% Actually, one can defined such a map for any family of $\HProp$
-% indexed by $\nat$, for example with the family $C_n \defeq
-% \text{«\,The $n$-th Turing Machine halts\,»}$.
-
-\section{Translation}
-\label{sec:translation}
+\subsection{Translation}
+\label{ssec:translation}
 
@@
                                   \Lbrack B\Rbrack ; \pi_\Sigma^{[A],[B]}
-                                \right)\\begin{mymath}0.5em]
-  \left[  (x;y)\right] &\defeq& ([x];[y]) \\begin{mymath}0.5em]
-  \left[  \pi_i t\right] &\defeq& \pi_i [t] \\begin{mymath}0.5em]
+                                \right)\\
+  \left[  (x;y)\right] &\defeq& ([x];[y]) \\
+  \left[  \pi_i t\right] &\defeq& \pi_i [t] \\
 \end{array}
@@
                                    \Lbrack B\Rbrack ; \pi_{\Pi}^{[A],[B]}
-                                  \right)\\begin{mymath}0.5em]
+                                  \right)\\
 \left[  \lambda\, x:A,~M \right] &\defeq&\lambda\,x:\Lbrack A
                                      \Rbrack,~[ M ]
-  \\begin{mymath}0.5em]
-  \left[ t \, t' \right] &\defeq& [t] [t'] \\begin{mymath}0.5em]
+  \\
+  \left[ t \, t' \right] &\defeq& [t] [t'] \\
 \end{array}
@@
 \left[  A=B \right] &\defeq& \left( [A] = [ B] , \pi_=^{[A],[B]}
-                             \right)\\begin{mymath}0.5em]
-\left[ 1 \right] &\defeq& 1\\begin{mymath}0.5em]
-\left[ J \right] &\defeq& J \\begin{mymath}0.5em]
+                             \right)\\
+\left[ 1 \right] &\defeq& 1\\
+\left[ J \right] &\defeq& J \\
 \end{array}
@@
                         \Rbrack); \pi_\modal(\Lbrack A \Rbrack + \Lbrack B
-                        \Rbrack)\right)\\begin{mymath}0.5em]
-\left[  \mathrm{in}_\ell t \right] &\defeq& \eta (\mathrm{in}_\ell [t]) \\begin{mymath}0.5em]
-\left[  \mathrm{in}_r t \right] &\defeq& \eta (\mathrm{in}_r [t]) \\begin{mymath}0.5em]
+                        \Rbrack)\right)\\
+\left[  \mathrm{in}_\ell t \right] &\defeq& \eta (\mathrm{in}_\ell [t]) \\
+\left[  \mathrm{in}_r t \right] &\defeq& \eta (\mathrm{in}_r [t]) \\
 \left[ \langle f ,g\rangle\right] &\defeq& \modal_{\mathrm{rec}}^{\Lbrack A\Rbrack +
                                       \Lbrack B\Rbrack} \langle
-                                      [f],[g]\rangle\\begin{mymath}0.5em]
+                                      [f],[g]\rangle\\
 \end{array}
@@
                                    \pi_\modal(\| \Lbrack A\Rbrack
-                                   \|_i)) \\begin{mymath}0.5em]
-  \left[ |t|_i \right] &\defeq& \eta |[t]|_i \\begin{mymath}0.5em]
+                                   \|_i)) \\
+  \left[ |t|_i \right] &\defeq& \eta |[t]|_i \\
   \left[ \|f\|_i \right] &\defeq& \modal_{\mathrm{rec}}^{\| \Lbrack
@@
 %
-
-\section{Formalization} 
-\label{sec:formalization}
-
-A Coq formalization based on the Coq/HoTT library~\cite{hottlib} is available at
-\url{https://github.com/KevinQuirin/sheafification}.
-%
-% After reviewing the content and some statistics about the
-% formalization in Section~\ref{sec:cont-form}, we present the
-% limitations of our formalization in Section~\ref{sec:limit-form}, in
-% particular the issues relative to universe polymorphism.
-
-\subsection{Content of the formalization}
-\label{sec:cont-form}
-
-% We provide a more detailed insight of the structure of our formalization:
-% \begin{itemize} 
-% \item Colimits and iterated kernel pairs are formalized in
-% \texttt{Limit}, \texttt{T.v}, \texttt{OT.v}v \texttt{OT\_Tf.v}, \texttt{T\_telescope.v},\\ \texttt{Tf\_Omono\_sep.v}.% (552 lines).
-% \item
-% Reflective subuniverses and modalities are formalized in\\
-% \texttt{reflective\_subuniverse.v}, \texttt{modalities.v}. % (1053 lines).
-% \item 
-% %
-%   The definition of the dense topology as a left exact modality on
-%   $\HProp$ is given in \texttt{sheaf\_base\_case.v}. % (186 lines).
-% \item
-% Section~\ref{sec:sheaves} is formalized in
-% \texttt{sheaf\_def\_and\_thm.v}. % (1029 lines).
-% \item
-% Section~\ref{sec:sheafification} is formalized in
-% \texttt{sheaf\_induction.v}. % (2340 lines).
-% \end{itemize}
-
-Overall, % with other files containing technical lemmas,
-the project
-contains 7914 lines, and it could be reduced a bit by improving the
-way Coq tries to rewrite and apply lemmas automatically. 
-The \texttt{coqwc} tool counts 1611 lines of specifications
-(definitions, lemmas, theorems, propositions) and 5403 lines of proof
-script.
-%
-This constitutes a significant amount of work but the part dedicated
-to sheaves and sheafification is only 2200 lines of proof script,
-which seems quite reasonable and encouraging, because it suggests
-that homotopy type theory provides a convenient tool to formalize some
-part of the theory of higher topoï. As a matter of comparison, this
-\TeX\xspace file contains approximatively 2000 lines.
-
-% In our formalization, some things are admitted, either because they
-% are just an adaptation of a proposition already formalized
-% in~\cite{hottbook}, either because they are true, but very technical:
-% \begin{itemize}
-% \item in \texttt{modalities.v}, an equality is admitted in the proof
-%   of proposition~\ref{sec:defin-basic-prop}.
-% \item in \texttt{nat\_lemmas.v}, the proof that $\IsHProp (n\leqslant
-%   m)$ is admitted.
-% \item in \texttt{sheaf\_induction.v}, in the proof that
-%   $C(\mu_T)=(C\id)$ (see lemma~\ref{lem:sepiscolim}), we admit the proof
-%   of equality of projections.
-% \item in \texttt{sheaf\_induction.v}, we admit some very technical
-%   proofs in proofs that $\square$ and $\star$ are modalities.
-% \item in \texttt{sheaf\_induction.v}, cumulativity is an axiom,
-%   because of universes issues.
-% \end{itemize}
-
-\subsection{Limitations of the formalization}
-\label{sec:limit-form}
-
-In the formalization, we had to use the \texttt{type-in-type} option, to handle
-the universe issues we faced.
-
-% Universes are used in type theory to ensure consistency by checking
-% that definitions are well-stratified according to a certain hierarchy.
-% %
-% Universe polymorphism~\cite{sozeau2014universe} supports generic
-% definitions over universes, reusable at different levels.
-% %
-% Although the presence of universe polymorphism is mandatory for our
-% formalization, its implementation is still too rigid to allow a
-% complete formalization of our work for the following reasons.
-
-% %
-% If Coq handles cumulativity on $\Type$ natively, it is not
-% the case for the $\Sigma$-type $\Type_n$, which require propositional
-% resizing. 
-% %
-% This issue could be solved by adding an axiom of cumulativity
-% for $\Type_n$ with an explicit management of universes. 
-% %
-% But as it would not have any computational content, such a solution
-% would really complicate the proofs as the axiom would appear
-% everywhere cumulativity is needed and it would need explicit
-% annotations for universe levels everywhere in the formalization.
-% %
-% % Note that we also could have resolved the issue by giving explicitely
-% % the universe levels we wanted, like in the Coq/HoTT library\cite{hottlib}.
-
-One issue with universe polymorphism lies in the management of
-recursive definitions. Indeed, the following recursive definition of
-sheafification
-%
-% \vspace{-0.05em}
-\begin{mymath} \begin{array}{l}
-   \modal : \forall \ (n : nat), \ \Type_n \to \Type_n 
-   \\
-    \modal_{-1\phantom{n}}(T) \defeq\lnot\lnot T \\
-
-      \displaystyle{\modal_{n+1}(T)} \defeq  
-      \displaystyle{\sum_{u:T \to \Type_n^\modal} \!\!\!\!\modal_{-1} 
-      \left\|
-      \sum_{a:T} u= (\lambda t,~\modal_n (a=t))
-      \right\|}
-    \end{array}
-% \vspace{-0.3em}
-\end{mymath}
-%
-is not allowed. 
-%
-This is because Coq forces the universe of the first $\Type_n$
-occurring in the definition to be the same for every $n$, whereas the
-universe of the first $\Type_{n+1}$ occurring in $\modal_{n+1}$ should be at
-least one level higher as the one of $\Type_n$ occurring in
-$\modal_{n}$ because of the use of $\Sigma$-type over
-$T \to \Type_n^\modal$ and equality on the return type of $\modal_n$.
-%
-% \nt{explain why it is an issue.}
-%
-Thus, the induction step presented in this paper has been formalized,
-but the complete recursive sheafification can not be defined for the
-moment.
-%
-Note that the same increasing in the universe levels occurs in the
-Rezk completion for categories~\cite{rezk}. In the definition of the
-completion, they use the Yoneda embedding and representable functors,
-which is similar to our use of characteristic functions.
-
-%
-This restriction in our formalization may be solved by
-generalizing the management of universe polymorphism for recursive definition
-%
-or by the use of general ``resizing axiom'' which is still under
-discussion in the community.
-
-\section{Conclusion and Future Works}
-\label{sec:future-works}
-
-In this paper, we have defined a way to extend homotopy type theory
-with new principles, leveraging Lawvere-Tierney sheafification to
-homotopy type theory. 
-%
-Using sheafification, it is possible to add new logical principles to
-homotopy type theory, such as the law of excluded middle, and this
-without axiom.
-%
-Beyond this result, our work is part of a more general line of work
-which aim to illustrate that homotopy type theory is a very promising
-theory for formalizing mathematics inside a proof assistant.
-
-% \nt{Add something on the definition of a syntactic translation as it
-  % has been done for forcing.}
-
-As a future work, we would like to improve this construction in three
-ways.
-% \begin{itemize}
-% \item 
-(i) The extension to whole $\Type$ in lemma~\ref{lem:type} is not
-  totally satisfactory, as every type is collapsed to a truncated
-  one. But some types in homotopy type theory are not
-  truncated~\cite[Example 8.8.6]{hottbook}.
-  % At the moment, sheafification only works with truncated
-  % types. This is fine from a programming point of view, 
-  % but many important types in homotopy theory are not
-  % truncated, \eg the circle $\mathbb S^2$.  Thus, we plan to extend
-  % the sheafification to all types by using a
-  % limit construction.
-% \item 
-(ii) 
-  We would like to have more examples of left-exact modalities on
-  $\HProp$, in order to extend type theory with different principles
-  than excluded middle.
-% \item 
-(iii) It would be useful to improve the translation induced by the
-  sheafification modality to preserve more definitional equalities.
-  %
-  This could be done for instance by defining a higher inductive type
-  isomorphic to the modality but with a better computational content. 
-  % able to work in the new type theory into a proof assistant.
-% \end{itemize}
-
-Moreover, in topos theory, Lawvere-Tierney subsumes Grothen\-dieck~\cite[Section~V.4]{maclanemoerdijk} in the sense that any
-Grothen\-dieck topology gives rise to a Lawvere-Tierney topology with
-the same notion of sheaves. Higher Lawvere-Tierney sheaves are
-presented here, and higher Grothendieck sheaves have been defined
-in~\cite{lurie}. It should be interesting to check if the subsumption
-remains true in higher topos theory.
-
-\section*{Acknowledgments}
-
-We would like to thank all maintainers of the Coq/HoTT library, who
-have provided a very solid base for our formalization. 
-%
-We also would like to thank Bas Spitters for helpful discussions on
-the subject.
-%
-This work has been funded by the CoqHoTT ERC Grant 637339.
-
-%\bibliographystyle{abbrvnat}
-\bibliography{sheaf_lics}
-
 \end{document}

### 2016-04-05 paper/paper_lipics.tex
Reduce

@@
 
-\subsection{Plan of the Paper}
-
-Section~\ref{sec:lawv-tiern-nutshell} presents an overview of the
-construction.
-% 
-Section~\ref{sec:hott} presents the foundations of homotopy type theory
-and an analogue of a theorem in higher toposes that is crucial
-in our construction~\cite[Corollary 6.2.3.5]{lurie}. 
-%
-Section~\ref{sec:lexmod} introduces the notion of left-exact stratified
-modalities, on which this paper is based.
-%
-Section~\ref{sec:sheaves} defines Lawvere-Tierney sheaves in homotopy
-type theory.
-%
-Section~\ref{sec:sheafification} explains the construction of the
-Lawvere-Tierney sheafification functor.
-%
-The definitions of this paper use an informal presentation of
-homotopy type theory, working entirely inside of Univalent
-Foundations. 
- %
-Section~\ref{sec:extension-type} describes how to use this
-construction to define a modality on $\Type$ from a modality on
-$\HProp$.
-%
-%
-Section~\ref{ssec:translation} defines the translation of Matin-Löf
-type theory into itself based on this modality. 
-%
-Section~\ref{sec:formalization} discusses the details of the
-formalization in Coq using formal type theory, in particular issues
-concerning universe polymorphism.
-%
-Finally, Section~\ref{sec:future-works} concludes on future works.
+% \subsection{Plan of the Paper}
+
+% Section~\ref{sec:lawv-tiern-nutshell} presents an overview of the
+% construction.
+% % 
+% Section~\ref{sec:hott} presents the foundations of homotopy type theory
+% and an analogue of a theorem in higher toposes that is crucial
+% in our construction~\cite[Corollary 6.2.3.5]{lurie}. 
+% %
+% Section~\ref{sec:lexmod} introduces the notion of left-exact stratified
+% modalities, on which this paper is based.
+% %
+% Section~\ref{sec:sheaves} defines Lawvere-Tierney sheaves in homotopy
+% type theory.
+% %
+% Section~\ref{sec:sheafification} explains the construction of the
+% Lawvere-Tierney sheafification functor.
+% %
+% The definitions of this paper use an informal presentation of
+% homotopy type theory, working entirely inside of Univalent
+% Foundations. 
+%  %
+% Section~\ref{sec:extension-type} describes how to use this
+% construction to define a modality on $\Type$ from a modality on
+% $\HProp$.
+% %
+% %
+% Section~\ref{ssec:translation} defines the translation of Matin-Löf
+% type theory into itself based on this modality. 
+% %
+% Section~\ref{sec:formalization} discusses the details of the
+% formalization in Coq using formal type theory, in particular issues
+% concerning universe polymorphism.
+% %
+% Finally, Section~\ref{sec:future-works} concludes on future works.
 
@@
 
-One has a way, from any type $T$, to construct a type
-$\|T\|_n:\Type_n$ as the HIT generated by
+From any type $T$, the type
+$\|T\|_n:\Type_n$ can be constructed as the HIT generated by
 \begin{itemize}
@@
 %
-The homotopy fiber $\fib{f}{b}$ of a function $f$ at element $b$ is
+The homotopy fiber of a function $f$ at element $b$ is
 defined as 
-\begin{mymath}[-0.3em]
+$
 \fib{f}{b} \defeq \sum_{a:A} f(a) = b.
-\end{mymath}%
+$
 %
@@
 %
- such that the usual (\cite[Theorem 4.8.4]{hottbook}) subobject
- classifier diagram is a pullback.
+ such that the usual subobject
+ classifier diagram (\cite[Theorem 4.8.4]{hottbook}) is a pullback.
 Therefore, in our construction, we will represent a subobject of a
 type $B$ with $n$-truncated homotopy fibers either as a map $f:A\to B$
-such that $\IsType n \fib{f}{b}$, either as a characteristic map $B\to \Type_n$.
+such that $\IsType n \fib{f}{b}$, either as a map $B\to \Type_n$.
 % \begin{mymath}
@@
   $P:\Type_{m}$ a colimit of $D$. 
-  Then, if $\|D\|_m$ is the same diagram as $D$, but where every type
+  Then, if $\|D\|_m$ is the same diagram as $D$ where every type
   is $m$-truncated, $P$ is a $m$-colimit\footnote{$P$ is a $m$-colimit
@@
 \begin{mymath}
-\begin{array}{rcl}
-\separated T &\defeq & \im (\lambda~t:T,~\lambda~ t',~ \modal_n (t = t')) \\
-          & \defeq & \sum\limits_{u:T \to \Type_n^\modal} \left\| \sum_{a:T} 
-            (\lambda t,~\modal_n (a=t)) = u\right\|.
-\end{array}
+%\begin{array}{rclcl}
+\separated T \defeq  \im (\lambda~t:T,~\lambda~ t',~ \modal_n (t = t')).
+ %          \left( =  \sum\limits_{u:T \to \Type_n^\modal} \left\| \sum_{a:T} 
+  %          (\lambda t,~\modal_n (a=t)) = u\right\|\right).
+%\end{array}
 \end{mymath}%
@@
 
-Now, let $Q$ be any separated $\Type_{n+1}$, and $f:X \to Q$. Then the
-following diagram commutes
+Now, let $Q$ be any separated $\Type_{n+1}$, and $f:X \to Q$. Then this
+ diagram commutes
 \begin{mymath}[-0.5em]\xymatrix{
-\|X\|_{n+1} \ar[r] \ar[rd] & \|\mathring T_{X}\|_{n+1} \ar[r] \ar[d] & \|\mathring
+\|X\|_{n+1} \ar[r] \ar[rd]_{|f|_{n+1}} & \|\mathring T_{X}\|_{n+1} \ar[r] \ar[d] & \|\mathring
   T_{\mathring T_X} \|_{n+1} \ar[ld]\cdots \\
@@
 This model transformation can better be described as a program
-transformation, as done in the next section.
+transformation, as done in the appendix (section~\ref{ssec:translation}).
 
@@
 recursive definitions. Indeed, the recursive definition of
-sheafification~\ref{eq:induction} is not allowed. 
+sheafification~(\ref{eq:induction}) is not allowed. 
 This is because Coq forces the universe of the first $\Type_n$
@@
 % \item 
-(iii) It would be useful to improve the translation induced by the
-  sheafification modality to preserve more definitional equalities.
-  %
-  This could be done for instance by defining a higher inductive type
-  isomorphic to the modality but with a better computational content. 
+% (iii) It would be useful to improve the translation induced by the
+%   sheafification modality to preserve more definitional equalities.
+%   %
+%   This could be done for instance by defining a higher inductive type
+%   isomorphic to the modality but with a better computational content. 
   % able to work in the new type theory into a proof assistant.

### 2016-04-06 paper/paper_lipics.tex
foo

@@
 discussion in the community.
+%
+We note that for this solution to work, we need to have a slighlty
+weaker definition of modalities than the one in~\cite{hottlib}. We
+usually ask the reflector $\modal$ to map a universe $\Type^i$ to
+itself, while here, we need $\modal T$, when $T:\Type^i$ to live in a
+possibly higher universe $\Type^j$. For modalities on whole $\Type$,
+we checked that all properties are still satisfied (see
+\url{https://github.com/KevinQuirin/HoTT/extended_modalities/} for the
+formalization).
 

### 2016-04-06 paper/paper_lipics.tex
finish

--- 2016-04-05 paper/paper_lipics.tex
@@
 
-\let\oldthebibliography=\thebibliography
-  \let\endoldthebibliography=\endthebibliography
-  \renewenvironment{thebibliography}[1]{%
-    \begin{oldthebibliography}{#1}%
-      \setlength{\parskip}{0ex}%
-      \setlength{\itemsep}{0ex}%
-  }%
-  {%
-    \end{oldthebibliography}%
-  }
+% \let\oldthebibliography=\thebibliography
+%   \let\endoldthebibliography=\endthebibliography
+%   \renewenvironment{thebibliography}[1]{%
+%     \begin{oldthebibliography}{#1}%
+%       \setlength{\parskip}{0ex}%
+%       \setlength{\itemsep}{0ex}%
+%   }%
+%   {%
+%     \end{oldthebibliography}%
+%   }
 
@@
 dense subobjects. This idea is formalized by saying that for every
-dense subobject $A$ of $E$, the following canonical map is an
-isomorphism
-%
-%\vspace{-0.7em}
-\begin{equation}\label{equ:sheaf_def}
-\Hom{\E}(E,F) \rightarrow \Hom{\E}(A,F).
-%\vspace{-0.7em}
-\end{equation}%
+dense subobject $A$ of $E$, the canonical map $\Hom{\E}(E,F) \rightarrow \Hom{\E}(A,F)$ is an
+isomorphism.
+%
+%\begin{equation}\label{equ:sheaf_def}
+%\Hom{\E}(E,F) \rightarrow \Hom{\E}(A,F).
+%\end{equation}%
 %
@@
     \end{array}
-% \vspace{-0.3em}
 \end{equation}
@@
 
-For instance, the notion of separated objects (which is a technical
-notion used in the construction of the sheafification functor), \ie
-objects for which the morphism
-(\ref{equ:sheaf_def}) is just an embedding, is generalized in two
+For instance, in the notion of separated objects defined in section~\ref{sec:sheafification},
+%(which is a technical
+%notion used in the construction of the sheafification functor), \ie
+%objects for which the morphism
+%(\ref{equ:sheaf_def}) is just an embedding, 
+subobjects are generalized in two
 different ways.  
@@
 The subobjects to be considered are given by $n$-truncated homotopy fibers,
-while the morphism (\ref{equ:sheaf_def}) is still required to be an
+while the morphism $\Hom{\E}(E,F) \rightarrow \Hom{\E}(A,F)$ is still required to be an
 embedding.
@@
 %
-\begin{mymath}[0em]
+\begin{mymath}[-0em]
   \chi : \sum_{A:\Type} \sum_{f:A \to B} \prod_{b\in B}
@@
 Then $P$ is the {\em colimit} of $D$ if for any other $X:\Type$, 
-\begin{mymath}[-0.5em]\IsEquiv\left(\lambda f:P \to X, \left( \lambda i,~f \circ q_i\, ;\, \lambda i\, j\,
+\begin{mymath}[-0em]\IsEquiv\left(\lambda f:P \to X, \left( \lambda i,~f \circ q_i\, ;\, \lambda i\, j\,
   \phi,\, f (p_{i, j}^\phi)) \right)\right).\end{mymath}
@@
   \item For every $n$-truncated types $A$ and $B$, if $P(B)$ then
-    \begin{mymath}\left\{
+    \begin{mymath}[-0em]\left\{
       \begin{array}{rcl}
@@
 on $\modal_n$ in $\Type_n^\modal$ to ease the reading.} %$\Type^\modal_p$, so 
-\begin{mymath}
+\begin{mymath}[-0em]
 \Type_n^\modal \ \defeq \sum_{T : \Type_n} P (T).
@@
 is the fiber of $\modal f$ above $\eta_Y y$:
-\begin{mymath}[-0.3em]\modal \left(\sum_{x:X}  (f(x) = y)\right) = \sum_{x:\modal X}
+\begin{mymath}[-0em]\modal \left(\sum_{x:X}  (f(x) = y)\right) = \sum_{x:\modal X}
 (\modal f(x) = \eta_Y(y)),\end{mymath}%
@@
 
+For space reason, the main ideas of the proofs are given in the
+appendix, and we refer the reader to the formalization for all details.
+
 In this section, we suppose given a truncation index $n\geqslant -1$,
@@
 if there is an extension $\tilde f:E\to F$, then it is unique.
-Actually, there is a simpler characterisation:
-
-\begin{prop}
-  A type $F$ is $\Type_{n+1}$ is separated if, and only if all its
-  path types are $n$-modal, ie
-  \begin{mymath}[-0.3em]\prod_{x,y:F}~\left( \modal_n(x=y) \right) = (x=y).\end{mymath}%
-\end{prop}
+Equivalently, it can be checked that a type is separated if, and only if all its path types
+are $n$-modal.
+
+% \begin{prop}
+%   A type $F$ is $\Type_{n+1}$ is separated if, and only if all its
+%   path types are $n$-modal, ie
+%   \begin{mymath}[-0.3em]\prod_{x,y:F}~\left( \modal_n(x=y) \right) = (x=y).\end{mymath}%
+% \end{prop}
 
@@
   For any mere proposition $P$ (where $\widehat P$ is $P$ seen as a
-  $\Type_n$),  $\modal_n \widehat P = \modal_{-1} P$ and the
+  $\Type_n$),  $\modal_n \widehat P \simeq \modal_{-1} P$ and the
   following coherence diagram commutes
-  \begin{mymath}\xymatrix{
+  \begin{mymath}[-0em]\xymatrix{
     P \ar@{->}^{\sim}[r] \ar[d]_{\eta_{-1}} & \widehat P \ar[d]^{\eta_n} \\
@@
 
-As many proofs are very technical, we only give here sketches, and
-often refer to formalization for all details.
-
 \subsection{From Type to Separated Type}
@@
 $\modal_n^T \circ \{\cdot\}_T$, as in
-\begin{mymath}\xymatrix{
+\begin{mymath}[-0em]\xymatrix{
     T \ar[r]^{\{\cdot\}_T} \ar[d]_{\mu_T} & \left(\Type_n\right)^T \ar[d]^{\modal_n^T} \\
@@
 %
-\begin{mymath}
+\begin{mymath}[-0em]
 %\begin{array}{rclcl}
@@
 
-Now, let $Q$ be any separated $\Type_{n+1}$, and $f:X \to Q$. Then this
- diagram commutes
-\begin{mymath}[-0.5em]\xymatrix{
-\|X\|_{n+1} \ar[r] \ar[rd]_{|f|_{n+1}} & \|\mathring T_{X}\|_{n+1} \ar[r] \ar[d] & \|\mathring
-  T_{\mathring T_X} \|_{n+1} \ar[ld]\cdots \\
-  & Q &
-} \end{mymath}% 
+Now, let $Q$ be any separated $\Type_{n+1}$, and $f:X \to Q$. 
+Then $|f|_{n+1}$ can be extended to build a cocone over $\mathring
+T$ into $Q$.
+
+% Then this
+%  diagram commutes
+% \begin{mymath}[-0.5em]\xymatrix{
+% \|X\|_{n+1} \ar[r] \ar[rd]_{|f|_{n+1}} & \|\mathring T_{X}\|_{n+1} \ar[r] \ar[d] & \|\mathring
+%   T_{\mathring T_X} \|_{n+1} \ar[ld]\cdots \\
+%   & Q &
+% } \end{mymath}% 
 % \begin{mymath}\xymatrix @C=4em{ 
@@
 subuniverse on $\Type_{n+1}$.
-We can now state
+As $(\separated,\mu)$ is a modality and closure is universal, we can state
 \begin{prop}\label{prop:sep-mod}
@@
 $\modal_{n+1}T$ can be given explicitly by
-\begin{mymath}
+\begin{mymath}[-0em]
 \modal_{n+1} T \ \defeq \sum_{u:T \to \Type_n^\modal} \modal_{-1}\left\| \sum_{a:T} 
@@
 As $T\to \Type_n^\modal$ is a sheaf, and $\modal_{n+1}T$ is closed in
-$T\to \Type_n^\modal$, $\modal_{n+1}T$ is a sheaf. We now prove that
-it forms a modality.
+$T\to \Type_n^\modal$, $\modal_{n+1}T$ is a sheaf. We can now state:
 
@@
 A Coq formalization based on the Coq/HoTT library~\cite{hottlib} is available at
-\url{https://github.com/KevinQuirin/sheafification}.
+\url{https://github.com/KevinQuirin/sheafification}. This development
+contains 7914 lines, with approximatively 20\% lines of
+definitions, which seems quite reasonable and encouraging, because it
+suggests that homotopy type theory provides a convenient tool to
+formalize some part of the theory of higher topoï.
 %
@@
 
-\subsection{Content of the formalization}
-\label{sec:cont-form}
-
-% We provide a more detailed insight of the structure of our formalization:
-% \begin{itemize} 
-% \item Colimits and iterated kernel pairs are formalized in
-% \texttt{Limit}, \texttt{T.v}, \texttt{OT.v}v \texttt{OT\_Tf.v}, \texttt{T\_telescope.v},\\ \texttt{Tf\_Omono\_sep.v}.% (552 lines).
-% \item
-% Reflective subuniverses and modalities are formalized in\\
-% \texttt{reflective\_subuniverse.v}, \texttt{modalities.v}. % (1053 lines).
-% \item 
-% %
-%   The definition of the dense topology as a left exact modality on
-%   $\HProp$ is given in \texttt{sheaf\_base\_case.v}. % (186 lines).
-% \item
-% Section~\ref{sec:sheaves} is formalized in
-% \texttt{sheaf\_def\_and\_thm.v}. % (1029 lines).
-% \item
-% Section~\ref{sec:sheafification} is formalized in
-% \texttt{sheaf\_induction.v}. % (2340 lines).
-% \end{itemize}
-
-Overall, % with other files containing technical lemmas,
-the project
-contains 7914 lines, and it could be reduced a bit by improving the
-way Coq tries to rewrite and apply lemmas automatically. 
-The \texttt{coqwc} tool counts 1611 lines of specifications
-(definitions, lemmas, theorems, propositions) and 5403 lines of proof
-script.
-%
-This constitutes a significant amount of work but the part dedicated
-to sheaves and sheafification is only 2200 lines of proof script,
-which seems quite reasonable and encouraging, because it suggests
-that homotopy type theory provides a convenient tool to formalize some
-part of the theory of higher topoï. As a matter of comparison, this
-\TeX\xspace file contains approximatively 2000 lines.
-
-% In our formalization, some things are admitted, either because they
-% are just an adaptation of a proposition already formalized
-% in~\cite{hottbook}, either because they are true, but very technical:
-% \begin{itemize}
-% \item in \texttt{modalities.v}, an equality is admitted in the proof
-%   of proposition~\ref{sec:defin-basic-prop}.
-% \item in \texttt{nat\_lemmas.v}, the proof that $\IsHProp (n\leqslant
-%   m)$ is admitted.
-% \item in \texttt{sheaf\_induction.v}, in the proof that
-%   $C(\mu_T)=(C\id)$ (see lemma~\ref{lem:sepiscolim}), we admit the proof
-%   of equality of projections.
-% \item in \texttt{sheaf\_induction.v}, we admit some very technical
-%   proofs in proofs that $\square$ and $\star$ are modalities.
-% \item in \texttt{sheaf\_induction.v}, cumulativity is an axiom,
-%   because of universes issues.
-% \end{itemize}
-
-\subsection{Limitations of the formalization}
-\label{sec:limit-form}
-
-In the formalization, we had to use the \texttt{type-in-type} option, to handle
+% \subsection{Content of the formalization}
+% \label{sec:cont-form}
+
+% % We provide a more detailed insight of the structure of our formalization:
+% % \begin{itemize} 
+% % \item Colimits and iterated kernel pairs are formalized in
+% % \texttt{Limit}, \texttt{T.v}, \texttt{OT.v}v \texttt{OT\_Tf.v}, \texttt{T\_telescope.v},\\ \texttt{Tf\_Omono\_sep.v}.% (552 lines).
+% % \item
+% % Reflective subuniverses and modalities are formalized in\\
+% % \texttt{reflective\_subuniverse.v}, \texttt{modalities.v}. % (1053 lines).
+% % \item 
+% % %
+% %   The definition of the dense topology as a left exact modality on
+% %   $\HProp$ is given in \texttt{sheaf\_base\_case.v}. % (186 lines).
+% % \item
+% % Section~\ref{sec:sheaves} is formalized in
+% % \texttt{sheaf\_def\_and\_thm.v}. % (1029 lines).
+% % \item
+% % Section~\ref{sec:sheafification} is formalized in
+% % \texttt{sheaf\_induction.v}. % (2340 lines).
+% % \end{itemize}
+
+% Overall, % with other files containing technical lemmas,
+% the project
+% contains 7914 lines, and it could be reduced a bit by improving the
+% way Coq tries to rewrite and apply lemmas automatically. 
+% The \texttt{coqwc} tool counts 1611 lines of specifications
+% (definitions, lemmas, theorems, propositions) and 5403 lines of proof
+% script.
+% %
+% This constitutes a significant amount of work but the part dedicated
+% to sheaves and sheafification is only 2200 lines of proof script,
+% which seems quite reasonable and encouraging, because it suggests
+% that homotopy type theory provides a convenient tool to formalize some
+% part of the theory of higher topoï. As a matter of comparison, this
+% \TeX\xspace file contains approximatively 2000 lines.
+
+% % In our formalization, some things are admitted, either because they
+% % are just an adaptation of a proposition already formalized
+% % in~\cite{hottbook}, either because they are true, but very technical:
+% % \begin{itemize}
+% % \item in \texttt{modalities.v}, an equality is admitted in the proof
+% %   of proposition~\ref{sec:defin-basic-prop}.
+% % \item in \texttt{nat\_lemmas.v}, the proof that $\IsHProp (n\leqslant
+% %   m)$ is admitted.
+% % \item in \texttt{sheaf\_induction.v}, in the proof that
+% %   $C(\mu_T)=(C\id)$ (see lemma~\ref{lem:sepiscolim}), we admit the proof
+% %   of equality of projections.
+% % \item in \texttt{sheaf\_induction.v}, we admit some very technical
+% %   proofs in proofs that $\square$ and $\star$ are modalities.
+% % \item in \texttt{sheaf\_induction.v}, cumulativity is an axiom,
+% %   because of universes issues.
+% % \end{itemize}
+
+\paragraph*{On the use of universe polymorphism}
+%\subsection{Limitations of the formalization}
+%\label{sec:limit-form}
+
+The development uses the new universe polymorphism feature of
+Coq~\cite{sozeau2014universe}. However, it still lacks some features, and
+in the formalization, we had to use the \texttt{type-in-type} option, to handle
 the universe issues we faced.
@@
 %
-or by the use of general ``resizing axiom'' which is still under
+and by the use of general ``resizing axiom'' which is still under
 discussion in the community.
+%
+We note that for this solution to work, we need to have a more general 
+definition of modalities than the one in~\cite{hottlib}, which ask the
+reflector $\modal$ to map a universe $\Type^i$ to
+itself. Here, we relax this constraint by allowing the reflector to
+map a universe $\Type^i$ to a
+possibly higher universe $\Type^j$. 
+We have checked that the library~\cite{hottlib} still works with this
+generalized definition (see
+\url{https://github.com/KevinQuirin/HoTT/tree/extended_modalities/} for the
+formalization).
 
@@
 % \end{itemize}
-
-Moreover, in topos theory, Lawvere-Tierney subsumes Grothen\-dieck~\cite[Section~V.4]{maclanemoerdijk} in the sense that any
+%
+(iii) In topos theory, Lawvere-Tierney subsumes Grothen\-dieck~\cite[Section~V.4]{maclanemoerdijk} in the sense that any
 Grothen\-dieck topology gives rise to a Lawvere-Tierney topology with
@@
 %\bibliographystyle{abbrvnat}
+\newpage
 \bibliography{sheaf_lics}
@@
 \newpage
-\section*{Appendix}
+\appendix
+\section{Proofs}
 
@@
 
-\subsection{Translation}
+\section{Translation}
 \label{ssec:translation}

### 2016-04-07 paper/paper_lipics.tex
minor corrections

@@
 
-In the rest of the section, we suppose given a truncation index
+In the rest of this section, we suppose given a truncation index
 $n\geqslant -1$, and a left exact modality $\modal_n$ on $\Type_n$,
@@
 
+\nt{explain why it is acceptable}
+
 % Universes are used in type theory to ensure consistency by checking
@@
 %
-and by the use of general ``resizing axiom'' which is still under
+and by the use of general ``resizing axioms'' which are still under
 discussion in the community.
@@
 We have checked that the library~\cite{hottlib} still works with this
-generalized definition (see
+generalized definition after minor modifications (see
 \url{https://github.com/KevinQuirin/HoTT/tree/extended_modalities/} for the

### 2016-04-07 paper/paper_lipics.tex
finish

@@
 \subsection{Overview of the Result}
+\label{ssec:overview}
 
@@
 in the formalization, we had to use the \texttt{type-in-type} option, to handle
-the universe issues we faced.
-
-\nt{explain why it is acceptable}
+the universe issues we faced. If not totally satisfactory, this option
+is a convenient way to write the Coq code, the following
+universes issues being checked on paper, like in the library
+Unimath\footnote{\url{https://github.com/UniMath}, {Vladimir
+    Voevodsky, Benedikt Ahrens, Daniel Grayson and others}}.
 
@@
 \section{Proofs}
+As said in section~\ref{sec:sheafification}, some proofs were ommited
+in the paper, due to space reason. The reader can find the key points
+here, and see the complete proofs in the Coq formalization, at
+\url{https://github.com/KevinQuirin/sheafification}.
 
@@
 
-In this section, we present how to use the sheafification modality to
-define a transformation of Martin-Löf type theory into
-itself\footnote{We define the translation on Martin-Löf type theory,
-  but it could also be done for CIC as well.} as it has been done for
-forcing~\cite{jaber2012extending}.
+In section~\ref{ssec:overview}, we announce that sheafification can be
+seen a translation of type theories\footnote{We define the translation on Martin-Löf type theory,
+  but it could also be done for CIC as well.}, as it has been done for
+forcing~\cite{jaber2012extending}. Actually, this is not a specific
+property of sheafification, but something shared by all modalities. We
+hence don't explicit it in the paper, but we do in this appendix.
 

### 2016-04-08 paper/paper_lipics.tex
minor changes

@@
 in the formalization, we had to use the \texttt{type-in-type} option, to handle
-the universe issues we faced. If not totally satisfactory, this option
-is a convenient way to write the Coq code, the following
-universes issues being checked on paper, like in the library
-Unimath\footnote{\url{https://github.com/UniMath}, {Vladimir
-    Voevodsky, Benedikt Ahrens, Daniel Grayson and others}}.
+the universe issues we faced. If not entirely satisfactory, this option
+is a convenient way to check the whole construction in Coq, but for
+universes.
+%
+The following universe issues have being checked on paper, like in the
+Unimath
+library.\footnote{\url{https://github.com/UniMath}, {Vladimir Voevodsky, Benedikt Ahrens, Daniel Grayson and others}}
 
@@
 %
-This restriction in our formalization may be solved by
+So universe issues in our formalization may be solved by
 generalizing the management of universe polymorphism for recursive definition
 %
-and by the use of general ``resizing axioms'' which are still under
+and also by using of general ``resizing axioms'' which are still under
 discussion in the community.
@@
 Beyond this result, our work is part of a more general line of work
-which aim to illustrate that homotopy type theory is a very promising
+which aims to illustrate that homotopy type theory is a very promising
 theory for formalizing mathematics inside a proof assistant.

### 2016-04-08 paper/paper_lipics.tex
foo

--- 2016-04-07 paper/paper_lipics.tex
@@
 \begin{defi}
-  Let $f:X \to Y$ be a map. The kernel pair $T_f$ of $f$ is the higher inductive type given
+  Let $f:X \to Y$ be a map. The coequalizer $T_f$ of the kernel pair of $f$ is the higher inductive type given
   by
@@
 \begin{defi}
-  Let $f$ be a map from $X$ to $Y$. Then the iterated kernel pair of
-  $f$ $C(f)$
+  Let $f$ be a map from $X$ to $Y$. Then the {\em iterated kernel pair} of
+  $f$ 
   is given by the diagram
@@
 property of sheafification, but something shared by all modalities. We
-hence don't explicit it in the paper, but we do in this appendix.
+hence don't explicit it in the paper, but only here in this appendix.
 
